### 2012-09-28 paper/cpp/listing_02_prereq_ptrvec.hpp.tex
sync

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}boost}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}container}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}vector.hpp\PYZgt{}}
\PY{k}{template} \PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{T}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{vec\PYZus{}t} \PY{o}{:} \PY{n}{boost}\PY{o}{:}\PY{o}{:}\PY{n}{ptr\PYZus{}vector}\PY{o}{\PYZlt{}}\PY{n}{T}\PY{o}{\PYZgt{}}
\PY{p}{\PYZob{}}
  \PY{k}{const} \PY{n}{T} \PY{o}{\PYZam{}}\PY{k}{operator}\PY{p}{[}\PY{p}{]}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{i}\PY{p}{)} \PY{k}{const}
  \PY{p}{\PYZob{}}   
    \PY{k}{return} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{at}\PY{p}{(}
      \PY{p}{(}\PY{n}{i} \PY{o}{+} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}
    \PY{p}{)}\PY{p}{;} 
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\end{Verbatim}

### 2012-09-28 paper/cpp/listings.hpp

@@ -1,6 +1,6 @@
-#include "listing_00.hpp"
-#include "listing_01.hpp"
-#include "listing_02.hpp"
+#include "listing_00_prereq_blitz.hpp"
+#include "listing_01_prereq_declret.hpp"
+#include "listing_02_prereq_ptrvec.hpp"
 #include "listing_03.hpp"
 #include "listing_04.hpp"
 #include "listing_05.hpp"

### 2012-09-28 paper/pyt/listings.py
adding a first test for Python; Cyclic boundary conditions; more general version of Shift

@@ -1,0 +1,133 @@
+#listing00
+
+import numpy 
+
+#listing01
+
+class Shift():
+  def __init__(self, plus, mnus):
+    self.plus = plus
+    self.mnus = mnus
+  def __radd__(self, arg): 
+    return type(arg)(
+      arg.start + self.plus, 
+      arg.stop  + self.plus
+    )
+  def __rsub__(self, arg): 
+    return type(arg)(
+      arg.start - self.mnus, 
+      arg.stop  - self.mnus
+    )
+
+one = Shift(1,1) 
+hlf = Shift(1,0)
+two = Shift(2,2)
+
+#listing02
+
+def f(psi_l, psi_r, C): 
+  return numpy.where(C >= 0, psi_l * C, psi_r * C)
+
+#listing03
+
+def donorcell_1D(psi_arg, C_arg, i, j, d): 
+  psi = psi_arg.swapaxes(0,d)
+  C = C_arg.swapaxes(0,d)
+  return (
+    f(psi[i,   j], psi[i+one, j], C[i+hlf, j]) - 
+    f(psi[i-one, j], psi[i,   j], C[i-hlf, j]) 
+  ).swapaxes(0,d)
+
+#listing04
+
+def donorcell_2D(psi, n, C, i, j):
+  print "psi,c shape", psi[n].shape, C[0].shape 
+  print "psi.swap, c.swap shape", psi[n].swapaxes(0,1).shape, C[1].swapaxes(0,1).shape
+  psi[n+1][i,j] = (psi[n][i,j] 
+    - donorcell_1D(psi[n], C[0], i, j, 0)
+    - donorcell_1D(psi[n], C[1], j, i, 1)
+  ) # loopa po wymiarach???
+
+#listing05
+
+class Mpdata(object):
+  def __init__(self, n_iters):
+    self.n_steps = n_iters
+    self.n_halos = n_iters
+
+  def op_2D(self, psi, n, C, i, j, step):
+    donorcell_2D(psi, n, C, i, j)
+
+#listing06
+
+class Cyclic(object):
+  def __init__(self, d, i, hlo): #TODO: i & j not needed in principle...
+    self.d = d
+    self.left_halo = slice(i.start - hlo, i.start)
+    self.rght_edge = slice(i.stop - hlo, i.stop)
+    self.rght_halo = slice(i.stop, i.stop + hlo)
[…]

### 2012-09-28 paper/pyt/test.py

@@ -1,109 +1,0 @@
-# -- listing 2.0
-import numpy #TODO: as np?
-
-# -- listing 2.3
-
-class Shift():
-  # TODO? only if isinstance(slice)?
-  def __init__(self,plus,mnus):
-    self.plus = plus
-    self.mnus = mnus
-  def __radd__(self, a):
-    return slice(a.start + self.plus, a.stop + self.plus)
-  def __rsub__(self, a):
-    return slice(a.start - self.mnus, a.stop - self.mnus)
-
-
-one = Shift(1,1) 
-hlf = Shift(1,0)
-two = Shift(2,2)
-
-# -- listing 2.6
-
-def f(psi_l, psi_r, C):
-  return numpy.where(C >= 0, psi_l * C, psi_r * C)
-
-# -- listing 2.7
-
-def donorcell_1D(psi_arg, C_arg, i, j, d):
-  psi = psi_arg.swapaxes(0,d)
-  C = C_arg.swapaxes(0,d)
-  return (
-    f(psi[i,   j], psi[i+one, j], C[i+hlf, j]) - 
-    f(psi[i-one, j], psi[i,   j], C[i-hlf, j])
-  ).swapaxes(0,d)
-
-# -- listing 2.8
-
-def donorcell_2D(psi, n, C, i, j):
-  print "psi,c shape", psi[n].shape, C[0].shape 
-  print "psi.swap, c.swap shape", psi[n].swapaxes(0,1).shape, C[1].swapaxes(0,1).shape
-  psi[n+1][i,j] = (psi[n][i,j] 
-    - donorcell_1D(psi[n], C[0], i, j, 0)
-    - donorcell_1D(psi[n], C[1], j, i, 1)
-  ) # loopa po wymiarach???
-
-# -- listing 2.13
-
-class Mpdata(object):
-  def __init__(self, n_iters):
-    self.n_steps = n_iters
-    self.n_halos = n_iters
-
-  def op_2D(self, psi, n, C, i, j, step):
-    donorcell_2D(psi, n, C, i, j)
-
-# -- listing 2.15
-
-class Solver_2D(object):
-  def __init__(self, adv, nx, ny):
-    self.adv = adv
-    self.n = 0
-    self.psi = []
-    self.C = []
-    self.i = slice(adv.n_halos, nx + adv.n_halos)
-    self.j = slice(adv.n_halos, ny + adv.n_halos)
-
[…]

### 2012-09-28 paper/pyt/testing.py

@@ -1,15 +1,14 @@
 """Unit test for test.py"""
-import test 
+from listings import *
 import unittest
 import numpy
 
 class KnownValues(unittest.TestCase):
-    #sprawdzic pierwsza wartosc!!!
     knownValue = [[10, 5, 1, 0.5, 0.5,
                    numpy.array(
-        [[ 1.,   1.,   0.,   0.,   0. ],
-         [ 1.,   1.,   0.,   0.,   0. ],
-         [ 1.,   1.,   0.,   0.,   0. ],
+        [[ 1.,   1.,   1.,   0.,   0. ],
+         [ 1.,   1.,   1.,   0.,   0. ],
+         [ 0.,   0.,   0.,   0.,   0. ],
          [ 0.,   0.,   0.,   0.,   0. ],
          [ 0.,   0.,   0.,   0.,   0. ],
          [ 0.,   0.,   0.,   0.,   0. ],
@@ -30,7 +29,7 @@ class KnownValues(unittest.TestCase):
          [ 0.,   0.,   0.,   0.,   0. ],
          [ 0.,   0.,   0.,   0.,   0. ]]
         )],
-                  [10, 5, 6, 1., 0.,
+                  [10, 5, 6, 0., 1.,
                    numpy.array(
         [[ 1.,   1.,   0.,   0.,   0. ],
          [ 1.,   1.,   0.,   0.,   0. ],
@@ -55,10 +54,10 @@ class KnownValues(unittest.TestCase):
          [ 0.,   0.,   0.,   0.,   0. ],
          [ 0.,   0.,   0.,   0.,   0. ]]
         )],
-                  [10, 5, 1, -1., 0.,
+                  [10, 5, 1, 0., -1.,
                    numpy.array(
-        [[ 1.,   1.,   0.,   0.,  0. ],
-         [ 1.,   1.,   0.,   0.,  0. ],
+        [[ 1.,   1.,   0.,   0.,   0. ],
+         [ 1.,   1.,   0.,   0.,   0. ],
          [ 1.,   1.,   0.,   0.,   0. ],
          [ 0.,   0.,   0.,   0.,   0. ],
          [ 0.,   0.,   0.,   0.,   0. ],
@@ -69,8 +68,8 @@ class KnownValues(unittest.TestCase):
          [ 0.,   0.,   0.,   0.,   0. ]]
         ),
                    numpy.array(
[…]

### 2012-09-29 paper/cpp/listing_00_prereq_blitz.hpp
moving all code into listngs.* files

@@ -1,3 +1,0 @@
-#include <blitz/array.h>
-using arr_t = blitz::Array<double, 2>;
-using rng_t = blitz::Range;

### 2012-09-29 paper/cpp/listing_01_prereq_declret.hpp

@@ -1,3 +1,0 @@
-#define return_macro(expr) \
-  -> decltype(safeToReturn(expr)) \
-{ return safeToReturn(expr); } 

### 2012-09-29 paper/cpp/listing_02_prereq_ptrvec.hpp

@@ -1,11 +1,0 @@
-#include <boost/ptr_container/ptr_vector.hpp>
-template <class T>
-struct vec_t : boost::ptr_vector<T>
-{
-  const T &operator[](const int i) const
-  {   
-    return this->at(
-      (i + this->size()) % this->size()
-    ); 
-  }
-};

### 2012-09-29 paper/cpp/listing_03.hpp

@@ -1,13 +1,0 @@
-struct hlf_t {} h;
-
-inline rng_t operator+(
-  const rng_t &i, const hlf_t &
-) { 
-  return i + 1; 
-} 
-
-inline rng_t operator-(
-  const rng_t &i, const hlf_t &
-) { 
-  return i; 
-}

### 2012-09-29 paper/cpp/listing_04.hpp

@@ -1,9 +1,0 @@
-template <class n_t>
-inline rng_t operator^(
-  const rng_t &r, const n_t &n) 
-{ 
-  return rng_t(
-    (r - n).first(), 
-    (r + n).last()
-  ); 
-} 

### 2012-09-29 paper/cpp/listing_05.hpp

@@ -1,13 +1,0 @@
-struct pi_ij : blitz::RectDomain<2> 
-{ 
-  pi_ij(const rng_t &i, const rng_t &j) :
-    RectDomain<2>(blitz::TinyVector<rng_t, 2>(i,j)) 
-  {}  
-};
-
-struct pi_ji : blitz::RectDomain<2> 
-{ 
-  pi_ji(const rng_t &j, const rng_t &i) :
-    RectDomain<2>(blitz::TinyVector<rng_t, 2>(i,j)) 
-  {}  
-}; 

### 2012-09-29 paper/cpp/listing_06.hpp

@@ -1,9 +1,0 @@
-template <class T1, class T2, class T3> 
-inline auto F(
-  const T1 &psi_l, 
-  const T2 &psi_r, 
-  const T3 &C
-) return_macro(
-  .5 * (C + abs(C)) * psi_l + 
-  .5 * (C - abs(C)) * psi_r
-)

### 2012-09-29 paper/cpp/listing_07.hpp

@@ -1,10 +1,0 @@
-template <class pi>  
-inline auto donorcell_1D( 
-  const arr_t &psi, 
-  const arr_t &C, 
-  const rng_t &i, 
-  const rng_t &j
-) return_macro(
-  F(psi(pi(i,  j)), psi(pi(i+1,j)), C(pi(i+h,j))) -
-  F(psi(pi(i-1,j)), psi(pi(i,  j)), C(pi(i-h,j)))
-)

### 2012-09-29 paper/cpp/listing_08.hpp

@@ -1,9 +1,0 @@
-inline void donorcell_2D(
-  const vec_t<arr_t> &psi, const int n,
-  const vec_t<arr_t> &C, 
-  const rng_t &i, const rng_t &j
-) { 
-  psi[n+1](i,j) = psi[n](i,j)
-    - donorcell_1D<pi_ij>(psi[n], C[0], i, j)
-    - donorcell_1D<pi_ji>(psi[n], C[1], j, i); 
-}

### 2012-09-29 paper/cpp/listing_09.hpp

@@ -1,6 +1,0 @@
-template <class nom_t, class den_t>
-static inline auto frac(
-  const nom_t &nom, const den_t &den
-) return_macro(
-  where(den > 0, nom / den, 0)
-) 

### 2012-09-29 paper/cpp/listing_10.hpp

@@ -1,10 +1,0 @@
-template <class pi>
-inline auto A(
-  const arr_t &psi,
-  const rng_t &i, const rng_t &j
-) return_macro(
-  frac(
-    psi(pi(i+1, j)) - psi(pi(i,j)),
-    psi(pi(i+1, j)) + psi(pi(i,j))
-  ) 
-) 

### 2012-09-29 paper/cpp/listing_11.hpp

@@ -1,12 +1,0 @@
-template <class pi>
-inline auto B(
-  const arr_t &psi, 
-  const rng_t &i, const rng_t &j
-) return_macro(
- .5 * frac(
-    psi(pi(i+1, j+1)) + psi(pi(i, j+1)) -
-    psi(pi(i+1, j-1)) - psi(pi(i, j-1)),
-    psi(pi(i+1, j+1)) + psi(pi(i, j+1)) +
-    psi(pi(i+1, j-1)) + psi(pi(i, j-1))
-  )
-)

### 2012-09-29 paper/cpp/listing_12.hpp

@@ -1,16 +1,0 @@
-template <class pi>
-inline auto antidiff_2D(
-  const arr_t &psi, 
-  const rng_t &i, const rng_t &j,
-  const vec_t<arr_t> &C, const int d
-) return_macro(
-  abs(C[d](pi(i+h, j)))
-  * A<pi>(psi, i, j)
-  * (1 - abs(C[d](pi(i+h, j))))
-  - C[d](pi(i+h, j)) 
-  * B<pi>(psi, i, j)
-  * .25 * (
-    C[d+1](pi(i+1, j+h)) + C[d+1](pi(i, j+h)) +
-    C[d+1](pi(i+1, j-h)) + C[d+1](pi(i, j-h)) 
-  ) 
-) 

### 2012-09-29 paper/cpp/listing_13.hpp

@@ -1,50 +1,0 @@
-template <int n_iters>
-struct mpdata 
-{
-  // member fields
-  vec_t<arr_t> tmp0, tmp1;
-  static const int n_steps = n_iters;
-  static const int n_halos = n_iters;
-
-  // ctor
-  mpdata(const rng_t &i, const rng_t &j) 
-  {
-    const int hlo = n_halos;
-    tmp0.push_back(new arr_t(i ^ h ^ hlo, j ^ hlo));
-    tmp0.push_back(new arr_t(i ^ hlo, j ^ h ^ hlo));
-    if (n_iters < 2) return;
-    tmp1.push_back(new arr_t(i ^ h ^ hlo, j ^ hlo));
-    tmp1.push_back(new arr_t(i ^ hlo, j ^ h ^ hlo));
-  }
-
-  // the 2D advection operator method
-  inline void op_2D( 
-    const vec_t<arr_t> &psi, const int n,
-    const vec_t<arr_t> &C, 
-    const rng_t &i, const rng_t &j,
-    const int step
-  ) {
-    // chosing input/output for antidiff. velocity
-    const vec_t<arr_t> 
-      &C_unco = (step == 1) ? C : 
-        (step % 2) ? tmp0 : tmp1,
-      &C_corr = (step  % 2) ? tmp1 : tmp0;
-    
-    // calculating the antidiffusive velocities
-    if (step > 0) {
-      const int hlo = n_steps - 1 - step;
-      enum {x,y};
-      C_corr[x](i ^ h ^ hlo, j ^ hlo) =
-        antidiff_2D<pi_ij>(
-          psi[n], i ^ h ^ hlo, j ^ hlo, C_unco, x
-        );
-      C_corr[y](i ^ hlo, j ^ h ^ hlo) =
-        antidiff_2D<pi_ji>(
-          psi[n], j ^ h ^ hlo, i ^ hlo, C_unco, y
-        );
-    }
-
-    // performing a donor-cell step with C or C_corr
-    donorcell_2D(psi, n, step==0 ? C : C_corr, i, j);
-  }
-};

### 2012-09-29 paper/cpp/listing_14.hpp

@@ -1,26 +1,0 @@
-template <class pi>
-struct cyclic
-{
-  // (could-be-private) member fields
-  pi left_halo, rght_edge, rght_halo, left_edge;
-
-  // ctor with member-field initialisation 
-  cyclic(const rng_t &i, const rng_t &j, const int hlo) :
-    //     ( i.first  ...   i.last         ), (  j ) 
-    left_halo(
-      rng_t(i.first()-hlo,  i.first()-1    ), j^hlo),
-    rght_edge(
-      rng_t(i.last()-hlo+1, i.last()       ), j^hlo),
-    rght_halo(
-      rng_t(i.last()+1,     i.last()+hlo   ), j^hlo),
-    left_edge(
-      rng_t(i.first(),      i.first()+hlo-1), j^hlo)
-  {} 
-
-  // public method
-  void fill_halos(const arr_t &psi)
-  {
-    psi(left_halo) = psi(rght_edge);     
-    psi(rght_halo) = psi(left_edge);     
-  }
-};

### 2012-09-29 paper/cpp/listing_15.hpp

@@ -1,38 +1,0 @@
-template <class adv_t, class bcx_t, class bcy_t>
-struct solver_2D
-{
-  vec_t<arr_t> psi, C;
-  int n;
-  rng_t i, j;
-  adv_t adv;
-  bcx_t bcx;
-  bcy_t bcy;
-
-  solver_2D(int nx, int ny) :
-    n(0), i(0, nx-1), j(0, ny-1), 
-    adv(i, j), 
-    bcx(i, j, adv_t::n_halos), 
-    bcy(j, i, adv_t::n_halos)
-  {
-    const int hlo = adv.n_halos;
-    for (int l = 0; l < 2; ++l) 
-      psi.push_back(new arr_t(i ^ hlo, j ^ hlo));
-    C.push_back(new arr_t(i ^ h ^ hlo, j ^ hlo));
-    C.push_back(new arr_t(i ^ hlo, j ^ h ^ hlo));
-  }
-  
-  arr_t state() {return psi[n](i,j).reindex({0,0});}
-  arr_t Cx() { return C[0](i ^ h, j); } // TODO: bcd!
-  arr_t Cy() { return C[1](i, j ^ h); } // TODO: bcd!
-
-  void solve(const int nt) {
-    for (int t = 0; t < nt; ++t) {
-      for (int s = 0; s < adv.n_steps; ++s) {
-        bcx.fill_halos(psi[n]);
-        bcy.fill_halos(psi[n]);
-        adv.op_2D(psi, n, C, i, j, s);
-        n = (n + 1) % 2;
-      }
-    }
-  }
-};

### 2012-09-29 paper/cpp/listings.hpp

@@ -1,16 +1,257 @@
-#include "listing_00_prereq_blitz.hpp"
-#include "listing_01_prereq_declret.hpp"
-#include "listing_02_prereq_ptrvec.hpp"
-#include "listing_03.hpp"
-#include "listing_04.hpp"
-#include "listing_05.hpp"
-#include "listing_06.hpp"
-#include "listing_07.hpp"
-#include "listing_08.hpp"
-#include "listing_09.hpp"
-#include "listing_10.hpp"
-#include "listing_11.hpp"
-#include "listing_12.hpp"
-#include "listing_13.hpp"
-#include "listing_14.hpp"
-#include "listing_15.hpp"
+//listing00
+#include <blitz/array.h>
+using arr_t = blitz::Array<double, 2>;
+using rng_t = blitz::Range;
+using idx_t = blitz::RectDomain<2>;
+using rngvec_t = blitz::TinyVector<rng_t, 2>;
+//listing01
+#define return_macro(expr) \
+  -> decltype(safeToReturn(expr)) \
+{ return safeToReturn(expr); } 
+//listing02
+#include <boost/ptr_container/ptr_vector.hpp>
+template <class T>
+struct ptrvec_t : boost::ptr_vector<T>
+{
+  const T &operator[](const int i) const
+  {   
+    return this->at(
+      (i + this->size()) % this->size()
+    ); 
+  }
+};
+//listing03
+struct hlf_t {} h;
+
+inline rng_t operator+(
+  const rng_t &i, const hlf_t &
+) { 
+  return i + 1; 
+} 
+
+inline rng_t operator-(
+  const rng_t &i, const hlf_t &
+) { 
+  return i; 
+}
+//listing04
+template <class n_t>
+inline rng_t operator^(
+  const rng_t &r, const n_t &n) 
+{ 
+  return rng_t(
+    (r - n).first(), 
+    (r + n).last()
+  ); 
+} 
+//listing05
+struct pi_ij : idx_t
+{ 
+  pi_ij(const rng_t &i, const rng_t &j) :
+    idx_t(rngvec_t(i,j)) 
+  {}  
+};
+
+struct pi_ji : idx_t
+{ 
+  pi_ji(const rng_t &j, const rng_t &i) :
+    idx_t(rngvec_t(i,j)) 
+  {}  
+}; 
[…]

### 2012-09-29 paper/cpp/plot.cpp

@@ -1,3 +1,4 @@
+//listing16
 #include "listings.hpp"
 #define GNUPLOT_ENABLE_BLITZ
 #include <gnuplot-iostream/gnuplot-iostream.h>
@@ -59,3 +60,4 @@ int main()
   gp << "splot '-' binary" << fmt << " notitle\n";
   gp.sendBinary(slv3.state().copy());
 }
+//listing17

### 2012-09-29 paper/for/test.f

@@ -1,3 +1,4 @@
+!listing08
 program test
   use solver_2D_m
   use mpdata_m
@@ -35,3 +36,4 @@ program test
     tmp(0:2,0:3) = 1
   end subroutine
 end program
+!listing09

### 2012-09-29 paper/pyt/listings.py

@@ -1,9 +1,6 @@
 #listing00
-
 import numpy 
-
 #listing01
-
 class Shift():
   def __init__(self, plus, mnus):
     self.plus = plus
@@ -18,18 +15,14 @@ class Shift():
       arg.start - self.mnus, 
       arg.stop  - self.mnus
     )
-
+#listing02
 one = Shift(1,1) 
 hlf = Shift(1,0)
 two = Shift(2,2)
-
-#listing02
-
+#listing03
 def f(psi_l, psi_r, C): 
   return numpy.where(C >= 0, psi_l * C, psi_r * C)
-
-#listing03
-
+#listing04
 def donorcell_1D(psi_arg, C_arg, i, j, d): 
   psi = psi_arg.swapaxes(0,d)
   C = C_arg.swapaxes(0,d)
@@ -37,19 +30,13 @@ def donorcell_1D(psi_arg, C_arg, i, j, d):
     f(psi[i,   j], psi[i+one, j], C[i+hlf, j]) - 
     f(psi[i-one, j], psi[i,   j], C[i-hlf, j]) 
   ).swapaxes(0,d)
-
-#listing04
-
+#listing05
 def donorcell_2D(psi, n, C, i, j):
-  print "psi,c shape", psi[n].shape, C[0].shape 
-  print "psi.swap, c.swap shape", psi[n].swapaxes(0,1).shape, C[1].swapaxes(0,1).shape
   psi[n+1][i,j] = (psi[n][i,j] 
     - donorcell_1D(psi[n], C[0], i, j, 0)
     - donorcell_1D(psi[n], C[1], j, i, 1)
   ) # loopa po wymiarach???
-
-#listing05
-
+#listing06
 class Mpdata(object):
   def __init__(self, n_iters):
     self.n_steps = n_iters
@@ -57,9 +44,7 @@ class Mpdata(object):
 
   def op_2D(self, psi, n, C, i, j, step):
     donorcell_2D(psi, n, C, i, j)
-
-#listing06
-
+#listing07
 class Cyclic(object):
   def __init__(self, d, i, hlo): #TODO: i & j not needed in principle...
     self.d = d
@@ -72,9 +57,7 @@ class Cyclic(object):
     psi = psi_arg.swapaxes(0, self.d) # TODO: does it work in 3D?
     psi[self.left_halo] = psi[self.rght_edge]
     psi[self.rght_halo] = psi[self.left_edge]
-
-#
-
+#listing08
[…]

### 2012-09-29 paper/for/listings.f
adding missing listings.f

@@ -1,0 +1,274 @@
+!listing00
+module arr_m
+  implicit none
+  type :: arr_t
+    real, dimension(:,:), pointer :: X
+  end type
+end module
+!listing01
+module arakawa_c_m
+  implicit none
+
+  type :: half_t
+  end type
+
+  type(half_t) :: h
+
+  interface operator (+)
+    module procedure ph
+  end interface
+
+  interface operator (-)
+    module procedure mh
+  end interface
+
+  ! TODO
+  !interface operator (^)
+  !  module procedure ext
+  !end interface
+
+  contains
+
+  elemental function ph(i, h)
+    integer, intent(in) :: i
+    type(half_t), intent(in) :: h
+    integer :: ph
+    ph = i + 1
+  end function
+
+  elemental function mh(i, h)
+    integer, intent(in) :: i 
+    type(half_t), intent(in) :: h
+    integer :: mh
+    mh = i
+  end function
+end module
+!listing02
+module adv_m
+  use arr_m
+  implicit none
+
+  type, abstract :: adv_t
+    integer :: n_steps, n_halos
+    contains 
+    procedure(op_2D_i), deferred :: op_2D
+  end type
+
+  abstract interface
+    subroutine op_2D_i(this, psi, n, C, i, j, s)
+      import :: adv_t
+      import :: arr_t
+      implicit none
+      class(adv_t) :: this
+      class(arr_t) :: psi(:), C(:)
+      integer, intent(in) :: n, s
+      integer, intent(in), dimension(:) :: i, j
+    end subroutine
+  end interface
+end module
+!listing03
+module mpdata_m
+  use adv_m
+  implicit none
+  
+  type, extends(adv_t) :: mpdata_t
+    contains
+    procedure :: ctor => mpdata_ctor
+    procedure :: op_2D => mpdata_op_2D
+  end type
+
+  contains
+
+  subroutine mpdata_ctor(this)
+    implicit none
+    class(mpdata_t) :: this
+    this%n_steps = 2
+    this%n_halos = 2
+  end subroutine
+
+  subroutine mpdata_op_2D(this, psi, n, C, i, j, s)
+    use arakawa_c_m
+    implicit none
+    class(mpdata_t) :: this
+    class(arr_t) :: psi(:), C(:)
+    integer, intent(in) :: n, s
+    integer, intent(in), dimension(:) :: i, j
+
+    psi(n+1)%X(i,j) = psi(n)%X(i,j) - (    &
+      F(                                   &
[…]

### 2012-09-29 paper/pyt/testing.py
adding known value

@@ -103,7 +103,14 @@ class KnownValues(unittest.TestCase):
          [ 0.,   0.,   0.,   0.,   0. ],
          [ 1.,   1.,   0.,   0.,   0. ],
          [ 1.,   1.,   0.,   0.,   0. ]]
-        )]       
+        )],
+                  [3,3,1,0.5,0.5,
+                   numpy.array([[0, 0, 0],
+                                [0, 1, 0],
+                                [0, 0, 0]]),
+                   numpy.array([[0, 0,   0],
+                                [0, 0.,  0.5],
+                                [0, 0.5 ,0]])]
                   ]
 
     def testMpdata(self):

### 2012-09-29 paper/cpp/listings.hpp
adding test_mpdata.cpp; switching from pi_ij, pi_ji to pi<0> and pi<1>

@@ -45,16 +45,20 @@ inline rng_t operator^(
   ); 
 } 
 //listing05
-struct pi_ij : idx_t
+template <int d> struct pi;
+
+template <>
+struct pi<0> : idx_t
 { 
-  pi_ij(const rng_t &i, const rng_t &j) :
+  pi(const rng_t &i, const rng_t &j) :
     idx_t(rngvec_t(i,j)) 
   {}  
 };
 
-struct pi_ji : idx_t
+template <>
+struct pi<1> : idx_t
 { 
-  pi_ji(const rng_t &j, const rng_t &i) :
+  pi(const rng_t &j, const rng_t &i) :
     idx_t(rngvec_t(i,j)) 
   {}  
 }; 
@@ -69,15 +73,23 @@ inline auto F(
   .5 * (C - abs(C)) * psi_r
 )
 //listing07
-template <class pi>  
+template <int d>  
 inline auto donorcell_1D( 
   const arr_t &psi, 
   const arr_t &C, 
   const rng_t &i, 
   const rng_t &j
 ) return_macro(
-  F(psi(pi(i,  j)), psi(pi(i+1,j)), C(pi(i+h,j))) -
-  F(psi(pi(i-1,j)), psi(pi(i,  j)), C(pi(i-h,j)))
+  F(
+    psi(pi<d>(i,  j)),  
+    psi(pi<d>(i+1,j)), 
+    C(pi<d>(i+h,j))
+  ) -
+  F(
+    psi(pi<d>(i-1,j)), 
+    psi(pi<d>(i,  j)), 
+    C(pi<d>(i-h,j))
+  )
 )
 //listing08
 inline void donorcell_2D(
@@ -86,8 +98,8 @@ inline void donorcell_2D(
   const rng_t &i, const rng_t &j
 ) { 
   psi[n+1](i,j) = psi[n](i,j)
-    - donorcell_1D<pi_ij>(psi[n], C[0], i, j)
-    - donorcell_1D<pi_ji>(psi[n], C[1], j, i); 
+    - donorcell_1D<0>(psi[n], C[0], i, j)
+    - donorcell_1D<1>(psi[n], C[1], j, i); 
 }
 //listing09
 template <class nom_t, class den_t>
@@ -97,45 +109,47 @@ static inline auto frac(
   where(den > 0, nom / den, 0)
 ) 
 //listing10
-template <class pi>
+template <int d>
 inline auto A(
[…]

### 2012-09-29 paper/cpp/plot.cpp

@@ -17,11 +17,11 @@ int main()
 {
   int nx = 32, ny = 32, nt = 200;
 
-  solver_2D<mpdata<1>, cyclic<pi_ij>, cyclic<pi_ji>> 
+  solver_2D<mpdata<1>, cyclic<0>, cyclic<1>> 
     slv1(nx, ny);
-  solver_2D<mpdata<2>, cyclic<pi_ij>, cyclic<pi_ji>> 
+  solver_2D<mpdata<2>, cyclic<0>, cyclic<1>> 
     slv2(nx, ny); 
-  solver_2D<mpdata<3>, cyclic<pi_ij>, cyclic<pi_ji>> 
+  solver_2D<mpdata<3>, cyclic<0>, cyclic<1>> 
     slv3(nx, ny); 
 
   Gnuplot gp;

### 2012-09-29 paper/cpp/test_mpdata.cpp

@@ -1,0 +1,16 @@
+#include "listings.hpp"
+#define GNUPLOT_ENABLE_BLITZ
+
+int main()
+{
+  int nx = 3, ny = 3, nt = 1;
+
+  solver_2D<mpdata<2>, cyclic<0>, cyclic<1>> slv(nx, ny); 
+
+  slv.state() = 0; 
+  slv.state()(rng_t(1,1), rng_t(1,1)) = 1;  
+  slv.Cx() = .5; 
+  slv.Cy() = .5;
+  slv.solve(nt);
+  std::cerr << slv.state() << blitz::sum(slv.state()) << std::endl;
+}

### 2012-09-29 paper/for/listing_00.f

@@ -1,6 +1,0 @@
-module arr_m
-  implicit none
-  type :: arr_t
-    real, dimension(:,:), pointer :: X
-  end type
-end module

### 2012-09-29 paper/for/listing_01.f

@@ -1,37 +1,0 @@
-module arakawa_c_m
-  implicit none
-
-  type :: half_t
-  end type
-
-  type(half_t) :: h
-
-  interface operator (+)
-    module procedure ph
-  end interface
-
-  interface operator (-)
-    module procedure mh
-  end interface
-
-  ! TODO
-  !interface operator (^)
-  !  module procedure ext
-  !end interface
-
-  contains
-
-  elemental function ph(i, h)
-    integer, intent(in) :: i
-    type(half_t), intent(in) :: h
-    integer :: ph
-    ph = i + 1
-  end function
-
-  elemental function mh(i, h)
-    integer, intent(in) :: i 
-    type(half_t), intent(in) :: h
-    integer :: mh
-    mh = i
-  end function
-end module

### 2012-09-29 paper/for/listing_02.f

@@ -1,22 +1,0 @@
-module adv_m
-  use arr_m
-  implicit none
-
-  type, abstract :: adv_t
-    integer :: n_steps, n_halos
-    contains 
-    procedure(op_2D_i), deferred :: op_2D
-  end type
-
-  abstract interface
-    subroutine op_2D_i(this, psi, n, C, i, j, s)
-      import :: adv_t
-      import :: arr_t
-      implicit none
-      class(adv_t) :: this
-      class(arr_t) :: psi(:), C(:)
-      integer, intent(in) :: n, s
-      integer, intent(in), dimension(:) :: i, j
-    end subroutine
-  end interface
-end module

### 2012-09-29 paper/for/listing_03.f

@@ -1,55 +1,0 @@
-module mpdata_m
-  use adv_m
-  implicit none
-  
-  type, extends(adv_t) :: mpdata_t
-    contains
-    procedure :: ctor => mpdata_ctor
-    procedure :: op_2D => mpdata_op_2D
-  end type
-
-  contains
-
-  subroutine mpdata_ctor(this)
-    implicit none
-    class(mpdata_t) :: this
-    this%n_steps = 2
-    this%n_halos = 2
-  end subroutine
-
-  subroutine mpdata_op_2D(this, psi, n, C, i, j, s)
-    use arakawa_c_m
-    implicit none
-    class(mpdata_t) :: this
-    class(arr_t) :: psi(:), C(:)
-    integer, intent(in) :: n, s
-    integer, intent(in), dimension(:) :: i, j
-
-    psi(n+1)%X(i,j) = psi(n)%X(i,j) - (    &
-      F(                                   &
-        psi(n)%X(i,  j), psi(n)%X(i+1,j),  &
-        C(0)%X(i+h,j)                      &
-      ) -                                  &
-      F(                                   &
-        psi(n)%X(i-1,j), psi(n)%X(i,  j),  &
-        C(0)%X(i-h,j)                      &
-      )                                    &
-    ) - (                                  &
-      F(                                   &
-        psi(n)%X(i,j  ), psi(n)%X(i,j+1),  &
-        C(1)%X(i,j+h)                      &
-      ) -                                  &
-      F(                                   &
-        psi(n)%X(i,j-1), psi(n)%X(i,j  ),  &
-        C(1)%X(i,j-h)                      &
-      )                                    &
-    )   
-
-    contains 
-    elemental function F(psi_l, psi_r, C)
-      implicit none
-      real :: F
-      real, intent(in) :: psi_l, psi_r, C
-    end function
-  end subroutine
-end module

### 2012-09-29 paper/for/listing_04.f

@@ -1,19 +1,0 @@
-module bcd_m
-  use arr_m
-  implicit none
-
-  type, abstract :: bcd_t
-    contains
-    procedure(fill_halos_i), deferred :: fill_halos
-  end type
- 
-  abstract interface 
-    subroutine fill_halos_i(this, psi)
-      import :: arr_t
-      import :: bcd_t
-      implicit none
-      class(bcd_t) :: this
-      real, dimension(:,:), pointer :: psi
-    end subroutine
-  end interface
-end module

### 2012-09-29 paper/for/listing_05.f

@@ -1,25 +1,0 @@
-module cyclic_m
-  use bcd_m
-  implicit none
-  
-  type, extends(bcd_t) :: cyclic_t
-    contains
-    procedure :: ctor => cyclic_ctor 
-    procedure :: fill_halos => cyclic_fill_halos
-  end type
-
-  contains
-
-  subroutine cyclic_ctor(this)
-    implicit none
-    class(cyclic_t) :: this
-  end subroutine
-
-  subroutine cyclic_fill_halos(this, psi)
-    implicit none
-    class(cyclic_t) :: this
-    real, dimension(:,:), pointer :: psi
-    !psi() = psi()
-    !psi() = psi()
-  end subroutine
-end module

### 2012-09-29 paper/for/listing_06.f

@@ -1,102 +1,0 @@
-module solver_2D_m
-  use arr_m
-  use adv_m
-  use bcd_m
-  implicit none
-
-  type :: solver_2D_t
-    class(arr_t), allocatable :: psi(:), C(:)
-    class(adv_t), pointer :: adv
-    class(bcd_t), pointer :: bcx, bcy
-    integer,  dimension(:), allocatable :: i, j
-    contains
-    procedure :: ctor  => solver_2D_ctor 
-    procedure :: solve => solver_2D_solve
-    procedure :: state => solver_2D_state
-    procedure :: dtor  => solver_2D_dtor 
-  end type 
-
-  contains
-
-  subroutine solver_2D_ctor(this, nx,ny,adv,bcx,bcy)
-    use arakawa_c_m
-    implicit none
-    class(solver_2D_t) :: this
-    integer, intent(in) :: nx, ny
-    class(adv_t), pointer :: adv
-    class(bcd_t), pointer :: bcx, bcy
-
-    allocate(this%i(0:nx-1), this%j(0:ny-1))
-
-    block
-      integer :: i, hlo
-      hlo = adv%n_halos
-      
-      allocate(this%psi(0:1))
-      do i=0, 1
-        allocate(this%psi(i)%X(        &
-          -hlo - h : nx - 1 + hlo + h, &
-          -hlo     : ny - 1 + hlo      &
-        )) 
-      end do
-
-      allocate(this%C(0:1))
-      do i=0, 1
-        allocate(this%C(i)%X(          &
-          -hlo     : nx - 1 + hlo,     &
-          -hlo - h : ny - 1 + hlo + h  &
-        ))   
-      end do
-    end block
-
-    this%adv => adv
-    this%bcx => bcx
-    this%bcy => bcy
-  end subroutine
-
-  subroutine solver_2D_dtor(this)
-    implicit none
-    class(solver_2D_t) :: this
-
-    block
-      integer :: i
-      do i=0,1
-        deallocate(this%psi(i)%X)   
-      end do
-      do i=0,1
-        deallocate(this%C(i)%X)   
-      end do
-    end block
-    deallocate(this%i, this%j, this%psi)
-  end subroutine
-  
-  subroutine solver_2D_solve(this, nt)
-    implicit none
-    class(solver_2D_t) :: this
-    integer, intent(in) :: nt
-
-    block
-      integer :: t, s, n 
-      n = 0
-      s = 1
-      do t = 1, nt
-        !do s = 1, adv%n_steps
[…]

### 2012-09-29 paper/pyt/listings.py

@@ -77,7 +77,7 @@ class Solver_2D(object):
       self.psi.append(numpy.empty((nx + 2 * hlo, ny + 2 * hlo))) 
     self.C.append(numpy.empty((nx + 1 + 2 * hlo, ny + 2 * hlo)))
     self.C.append(numpy.empty((nx + 2 * hlo, ny + 1 + 2 * hlo)))
-#listing09
+
   def solve(self, nt):
     for t in range(nt):
       for s in range(self.adv.n_steps):
@@ -93,22 +93,4 @@ class Solver_2D(object):
     return self.C[0]
   def Cy(self):
     return self.C[1]
-#listing10
-if __name__ == '__main__':
-  nx, ny = 10, 5
-
-  slv1 = Solver_2D(Mpdata(1), Cyclic, Cyclic, nx, ny)
-
-  slv1.state()[:] = 0
-  slv1.state()[0:2,0:3] = 1
-  slv1.Cx()[:] = .5
-  slv1.Cy()[:] = .5
-  
-  print slv1.state()
-  
-  nt = 1
-  slv1.solve(nt)
-  print slv1.state()
-
-  #slv2 = Solver_2D(Mpdata(2), Cyclic, Cyclic, nx, ny)
-  #slv2.solve(nt)
+#listing09

### 2012-09-30 paper/pyt/testing.py
adding new known value with Cx/Cy array; does not work

@@ -110,7 +110,46 @@ class KnownValues(unittest.TestCase):
                                 [0, 0, 0]]),
                    numpy.array([[0, 0,   0],
                                 [0, 0.,  0.5],
-                                [0, 0.5 ,0]])]
+                                [0, 0.5 ,0]])],
+                   [3,3,1,
+                    numpy.array([[0.,0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0.],
+                                 [0.,0., 0.5, 0., 0.],
+                                 [0.,0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0.],]),
+                    numpy.array([[0.,0., 0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0., 0.],
+                                 [0.,0., 0., 0.5, 0., 0.],
+                                 [0.,0., 0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0., 0.],]),
+                                 
+                    numpy.array([[0, 0, 0],
+                                [0, 1, 0],
+                                [0, 0, 0]]),
+                    numpy.array([[0, 0,   0],
+                                [0, 0.,  0.5],
+                                [0, 0.5 ,0]])],
+    
+                   [3,3,1,
+                    numpy.array([[0.,0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0.],
+                                 [0.,0.1, 0.5, 0.4, 0.],
+                                 [0.,0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0.],]),
+                    numpy.array([[0.,0., 0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0., 0.],
+                                 [0.,0., 0.3, 0.5, -0.1, 0.],
+                                 [0.,0., 0., 0., 0., 0.],
+                                 [0.,0., 0., 0., 0., 0.],]),
+                                 
+                    numpy.array([[0, 0, 0],
+                                [1, 1, 1],
+                                [0, 0, 0]]),
+                    numpy.array([[0, 0,   0],
+                                [0.6, 0.4,  1.],
+                                [0.1, 0.5 ,0.4]])]
                   ]
 
     def testMpdata(self):
@@ -121,6 +160,7 @@ class KnownValues(unittest.TestCase):
              slv.Cy()[:] = cy
[…]

### 2012-10-01 paper/for/listings.f
using tuples instead of lists to store a sequence of psi and C

@@ -58,7 +58,6 @@ module adv_m
     subroutine op_2D_i(this, psi, n, C, i, j, s)
       import :: adv_t
       import :: arr_t
-      implicit none
       class(adv_t) :: this
       class(arr_t) :: psi(:), C(:)
       integer, intent(in) :: n, s
@@ -80,7 +79,6 @@ module mpdata_m
   contains
 
   subroutine mpdata_ctor(this)
-    implicit none
     class(mpdata_t) :: this
     this%n_steps = 2
     this%n_halos = 2
@@ -88,7 +86,6 @@ module mpdata_m
 
   subroutine mpdata_op_2D(this, psi, n, C, i, j, s)
     use arakawa_c_m
-    implicit none
     class(mpdata_t) :: this
     class(arr_t) :: psi(:), C(:)
     integer, intent(in) :: n, s
@@ -116,7 +113,6 @@ module mpdata_m
 
     contains 
     elemental function F(psi_l, psi_r, C)
-      implicit none
       real :: F
       real, intent(in) :: psi_l, psi_r, C
     end function
@@ -136,7 +132,6 @@ module bcd_m
     subroutine fill_halos_i(this, psi)
       import :: arr_t
       import :: bcd_t
-      implicit none
       class(bcd_t) :: this
       real, dimension(:,:), pointer :: psi
     end subroutine
@@ -156,12 +151,10 @@ module cyclic_m
   contains
 
   subroutine cyclic_ctor(this)
-    implicit none
     class(cyclic_t) :: this
   end subroutine
 
   subroutine cyclic_fill_halos(this, psi)
-    implicit none
     class(cyclic_t) :: this
     real, dimension(:,:), pointer :: psi
     !psi() = psi()
@@ -184,14 +177,13 @@ module solver_2D_m
     procedure :: ctor  => solver_2D_ctor 
     procedure :: solve => solver_2D_solve
     procedure :: state => solver_2D_state
-    procedure :: dtor  => solver_2D_dtor 
+    procedure :: dtor  => solver_2D_dtor !TODO: final
   end type 
 
   contains
 
   subroutine solver_2D_ctor(this, nx,ny,adv,bcx,bcy)
     use arakawa_c_m
-    implicit none
     class(solver_2D_t) :: this
     integer, intent(in) :: nx, ny
     class(adv_t), pointer :: adv
@@ -226,7 +218,6 @@ module solver_2D_m
   end subroutine
 
[…]

### 2012-10-01 paper/pyt/listings.py

@@ -62,8 +62,6 @@ class Solver_2D(object):
   def __init__(self, adv, bcx, bcy, nx, ny):
     self.adv = adv
     self.n = 0
-    self.psi = []
-    self.C = []
 
     hlo = adv.n_halos
 
@@ -73,10 +71,14 @@ class Solver_2D(object):
     self.bcx = bcx(0, self.i, hlo)
     self.bcy = bcy(1, self.j, hlo)
 
-    for i in range(2):
-      self.psi.append(numpy.empty((nx + 2 * hlo, ny + 2 * hlo))) 
-    self.C.append(numpy.empty((nx + 1 + 2 * hlo, ny + 2 * hlo)))
-    self.C.append(numpy.empty((nx + 2 * hlo, ny + 1 + 2 * hlo)))
+    self.psi = (
+      numpy.empty((nx + 2 * hlo, ny + 2 * hlo)),
+      numpy.empty((nx + 2 * hlo, ny + 2 * hlo)) 
+    )
+    self.C = (
+      numpy.empty((nx + 1 + 2 * hlo, ny + 2 * hlo)),
+      numpy.empty((nx + 2 * hlo, ny + 1 + 2 * hlo))
+    )
 
   def solve(self, nt):
     for t in range(nt):

### 2013-01-08 paper/pyt/simple_test.py
removing some unneeded files from the pyt directory

@@ -1,18 +1,0 @@
-from listings import *
-import numpy 
-
-nx, ny = 3, 3
-Cx, Cy = 0, 1
-psi_in = numpy.array([[0,0,0],[0,1,0],[0,0,0]])
-
-slv = Mpdata_2D(2, Cyclic, Cyclic, nx, ny)
-slv.state()[:] = psi_in
-slv.courant(0)[:] = Cx
-slv.courant(1)[:] = Cy
-slv.solve(1)
-
-print slv.state()
-
-#if (abs(slv.state() - read_file(fout, nx, ny)) >= .5 * pow(10, -dec)).any(): 
-#  print slv.state()
-#  raise Exception()

### 2013-01-08 paper/pyt/testing_numpy_zmianahalo.py

@@ -1,105 +1,0 @@
-"""Unit test for listing.py using numpy function for testing"""
-from listings_zmianahalo import *
-import numpy
-import numpy.testing
-import pdb
-
-knownValue = [[3,3,1,0.5,0.5,
-               numpy.array([[0., 0, 0],
-                            [0,  1, 0],
-                            [0,  0, 0]]),
-               numpy.array([[0.,  0,   0],
-                            [0,   0.,  0.5],
-                            [0,   0.5, 0]]),
-               numpy.array([[0, 0,   0],
-                            [0, 0.,  0.5],
-                            [0, 0.5, 0]])],
-              [3,3,1,0.5,0.5,
-               numpy.array([[0., 0, 0],
-                            [0,  0, 0],
-                            [0,  0, 1]]),
-               numpy.array([[0.,  0.,  0.5],
-                            [0.,  0.,  0.],
-                            [0.5, 0.,  0]]),
-               numpy.array([[0,   0.,  0.5],
-                            [0.,  0.,  0.],
-                            [0.5, 0.,  0]])],
-              [3,3,4,1.,0.,
-               numpy.array([[0., 0, 0],
-                            [0,  0, 0],
-                            [0,  1, 0]]),
-               numpy.array([[0.,  1., 0],
-                            [0,   0., 0.],
-                            [0,   0., 0]]),
-               numpy.array([[0, 1,  0],
-                            [0, 0., 0.],
-                            [0, 0., 0]])],
-               [3,3,3,0.,1.,
-               numpy.array([[0., 0, 0],
-                            [0,  1, 0],
-                            [0,  0, 0]]),
-               numpy.array([[0.,  0., 0],
-                            [0,   1., 0.],
-                            [0,   0., 0]]),
-               numpy.array([[0, 0.,  0],
-                            [0, 1., 0.],
-                            [0, 0., 0]])],
-              #[3,3,1,0.2,0.2,
-              # numpy.array([[0., 0, 0],
-              #              [0,  1, 0],
-              #              [0,  0, 0]]),
-              # numpy.array([[0, 0,   0],
-              #              [0, 0.6, 0.2],
-              #              [0, 0.2, 0.]]),
-              # numpy.array([[0, 0,     0],
[…]

### 2013-01-10 paper/pyt/test.py
adding an assert into test.py to assure that the file to be loaded contains enough data (thanks to Davide Del Vento for pointing the issue out)

@@ -8,6 +8,7 @@ def read_file(fname, nx, ny):
     for line in f:
       tmp[x,:] = numpy.fromstring(line, dtype=real_t, sep='\t')
       x += 1
+  assert(x == nx)
   return tmp
 
 def main():

### 2013-01-10 paper/doc/paper.tex
adding paper draft LaTeX source to the repo (paper/doc)

\documentclass[final,5p,times,twocolumn]{elsarticle}

% input encoding assuring proper handling of Polish characters (surnames!)
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\input{formulae/preamble}

% for Arakawa-C figure
\usepackage{tikz}
\usetikzlibrary{calc}

% for conpact itemize etc
\usepackage{enumitem}
\setitemize{noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}

% for source code listings
\usepackage{fancyvrb}
\fvset{
  frame=single,
  fontsize=\scriptsize,
  fontfamily=courier,
  framerule=.2mm,
  framesep=.7mm,
  numbers=left,
  xleftmargin=4mm,
  numbersep=.25mm,
  firstnumber=last
}
\input{pygments}

\newcounter{lstnocpp}
\newcounter{lstnopyt}
\newcounter{lstnofor}

\setcounter{lstnocpp}{-1}
\setcounter{lstnopyt}{-1}
\setcounter{lstnofor}{-1}

\newcounter{linenocpp}
\newcounter{linenopyt}
\newcounter{linenofor}

\setcounter{linenocpp}{0}
\setcounter{linenopyt}{0}
\setcounter{linenofor}{0}

\newcommand*\FancyVerbStartString{}
\newcommand*\FancyVerbStopString{}

\newcommand{\codecpp}[4]{%
  \addtocounter{lstnocpp}{1}%
  \renewcommand*\FancyVerbStartString{\PY{c+c1}{//#2}}%
  \renewcommand*\FancyVerbStopString{\PY{c+c1}{//#3}}%
  \setcounter{FancyVerbLine}{\thelinenocpp}%
  \fvset{label={listing~C.\thelstnocpp~(C++)},rulecolor=\color{black},stepnumber=#4}%
  \input{#1}%
  \setcounter{linenocpp}{\value{FancyVerbLine}}%
}
\newcommand{\codepyt}[4]{%
  \addtocounter{lstnopyt}{1}%
  \renewcommand*\FancyVerbStartString{\PY{c}{\PYZsh{}#2}}
  \renewcommand*\FancyVerbStopString{\PY{c}{\PYZsh{}#3}}
  \setcounter{FancyVerbLine}{\thelinenopyt}%
  \fvset{label={listing~P.\thelstnopyt~(Python)},rulecolor=\color{blue},stepnumber=#4}%
  \input{#1}%
  \setcounter{linenopyt}{\value{FancyVerbLine}}%
}
\newcommand{\codefor}[4]{%
  \addtocounter{lstnofor}{1}%
  \renewcommand*\FancyVerbStartString{\PY{c}{!#2}}
  \renewcommand*\FancyVerbStopString{\PY{c}{!#3}}
  \setcounter{FancyVerbLine}{\thelinenofor}%
  \fvset{label={listing~F.\thelstnofor~(Fortran)},rulecolor=\color{red},stepnumber=#4}%
  \input{#1}%
  \setcounter{linenofor}{\value{FancyVerbLine}}%
}

\newcommand{\prog}[1]{{\rm\bf#1}}
\newcommand{\url}[1]{{#1}}

%\journal{Computer Physics Communications}
\journal{arXiv. Paper presented at the American Meteorological Society 2013 Annual Meeting in Austin, Texas, USA}

\begin{document}
  \begin{frontmatter}

    \title{
      Object-oriented implementations of the MPDATA advection equation solver in~C++,~Python~and~Fortran
    }

    \author[1]{Sylwester Arabas}
    \author[1]{Dorota Jarecka}
    \author[1]{Anna Jaruga}
    \author[2]{Maciej Fijałkowski}

    \address[1]{Institute of Geophysics, Faculty of Physics, University of Warsaw}
    \address[2]{PyPy Team}

    \begin{abstract}
        Three object-oriented implementations of a prototype solver of the advection equation are introduced.
        Presented programs are based on Blitz++ (C++), NumPy (Python), and Fortran's built-in array containers.
        The solvers include an implementation of the Multidimensional Positive-Definite 
          Advective Transport Algorithm (MPDATA).
        The introduced codes exemplify how the application of 
          object-oriented programming (OOP) techniques allows to reproduce the mathematical notation 
          used in the literature within the program code.
        The introduced codes serve as a basis for discussion on the tradeoffs of the programming language choice.
        The main angles of comparison are code brevity and syntax clarity
          (and hence maintainability and auditability) as well as performance.
        In case of Python, a significant performance gain is observed when switching from the standard 
          interpreter (CPython) to the PyPy implementation of Python.
        Entire source code of all three implementations is embedded in the text and is licensed
          under the terms of the GNU GPL license. 
    \end{abstract}

    % up to six in Comp. Phys. Comm.
    \begin{keyword}
      object-oriented programming, advection equation, MPDATA, C++, Fortran, Python
    \end{keyword}

  \end{frontmatter}

  \tableofcontents

  \section{Introduction}

  Object oriented programming (OOP) {\em ''has become recognised as the almost unique successful 
    paradigm for creating complex software''} \citep[][Sec.~1.3]{Press_et_al_2007}.
  It is intriguing that while the quoted statement comes from the very book subtitled 
   {\em The Art of Scientific Computing}, hardly any (if not none) of currently operational 
    weather and climate prediction systems - flagship examples of complex scientific software - 
    makes extensive use of OOP techniques\footnote{
  Fortran have been the language of choice in oceanic \citep{Griffies_et_al_2000}, 
    weather-prediction \citep{Sundberg_2009} and Earth system \citep{Legutke_2012} modelling, 
    and none of its 20-century editions were object-oriented languages \citep[see e.g.][for discussion]{Norton_et_al_2007}.}.
  Application of OOP techniques in development of numerical modelling software may help to
    {\bf (i)} maintain modularity and separation of program logic layers (e.g. separation of
      a dynamical core, parallelisation mechanisms, data input/output, error handling and
      the description of physical processes); and
    {\bf (ii)} {\bf shorten and simplify the source code and improve its readability by reproducing within 
      the program logic the mathematical abstractions and notation used in the literature}.
  The first application is attainable, yet arguably cumbersome, with procedural programming.
  The latter, virtually impossible to obtain with procedural programming, is the focus of this paper.
  It also enables the compiler or library authors to relieve the user (i.e. scientific programmer)
    from hand-coding (premature) optimisations, a practice long recognised as having {\em a strong negative impact when debugging
    and maintenance are considered} \citep{Knuth_1974}.

  MPDATA \citep{Smolarkiewicz_1984} stands for Multidimensional Positive Definite Advective Transport Algorithm and is
    an example of a numerical procedure used in weather, climate and ocean simulation systems
    \citep[e.g.][respectively]{Ziemianski_et_al_2011,Abiodun_et_al_2011,Ezer_et_al_2002}.
  MPDATA is a solver for systems of equations of the following form:
  \begin{equation}\label{eq:adv}
    \input{formulae/adv}
  \end{equation}
  that describe transport of quantity $\psi$ by fluid flow with velocity $\vec{v}$.
  Quoting Numerical Recipes once more, development of methods to numerically solve such problems 
    {\em ''is an art as much as a science''} \citep[][Sec.~20.1]{Press_et_al_2007},
    and MPDATA is an example of the state-of-the art in this field.
  MPDATA is designed to accurately solve equation (\ref{eq:adv}) in arbitrary
    number of dimensions assuring positive-definiteness of scalar field $\psi$ 
    and incurring small numerical diffusion
    \citep[see][for a recent review of MPDATA-based techniques]{Smolarkiewicz_2006}.

  In this paper we introduce and discuss object-oriented implementations of an MPDATA-based 
    two-dimensional (2D) advection equation solver written in C++11 (\citeauthor{ISO_14882} 14882:2011), 
    Python \citep{Rossum_and_Drake_2011} and Fortran 2008 (\citeauthor{ISO_1539-1} 1539-1:2010). 
  In the following section we introduce the three implementations
    briefly describing the algorithm itself and
    discussing where and how the OOP techniques may be applied in its implementation.
  The syntax and nomenclature of OOP techniques are used without introduction,
    for an overview of OOP in context of C++, Python and Fortran, consult for example
    \citep[][Part~II]{Stroustrup_2000}, \citep[][Chapter~5]{Pilgrim_2004} and
    \citep[][Chapter~11]{Markus_2012}, respectively.
  The third section covers performance evaluation of the three implementations.
  The fourth section covers discussion of the tradeoffs of the programming language choice.
  The fifth section closes the article with a brief summary.

  Throughout the paper the three implementations are presented by discussing 
    source code listings which cover entire program code.
  Subsections \ref{sec:array}-\ref{sec:solver} describe all three implementations,
    while subsequent sections \ref{sec:cyclic}-\ref{sec:example} cover discussion of C++ code 
    only.
  The relevant parts of Python and Fortran codes do not differ significantly, and for readability reasons 
    are presented in the \ref{app:P} and the \ref{app:F}, respectively.

  The entire code is licensed under the terms of the GNU General Public License license version 3 \citep{GPLv3}.

  All listings include line numbers printed to the left of the source code, with separate numbering for
    C++ (listings prefixed with C, black frame),
  \codecpp{code/cpp/listings.hpp}{listing00}{listing01}{1}
    Python (listings prefixed with P, blue frame) and
  \codepyt{code/pyt/listings.py}{listing00}{listing01}{1}
    Fortran (listings prefixed with F, red frame).
  \codefor{code/for/listings.f}{listing00}{listing01}{1}
  Programming language constructs when inlined in the text are 
    typeset in bold, e.g. \prog{GOTO 2}.

  \section{Implementation}\label{sec:impl}

  Double precision floating-point format is used in all three implementations.
  The codes begins with the following definitions:
  \codecpp{code/cpp/listings.hpp}{listing01}{listing02}{1}
  \codepyt{code/pyt/listings.py}{listing01}{listing02}{1}
  \codefor{code/for/listings.f}{listing01}{listing02}{1}
  which provide a convenient way of switching to different precision.

  All codes are strucutred in a way allowing compilation of the code 
    in exactly the same order as presented in the text within one source file,
    hence every Fortran listing contains definition of a separate module.

  \subsection{Array containers}\label{sec:array}

  Solution of equation (\ref{eq:adv}) using MPDATA implies discretisation of the $\psi$ and 
    the $\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}$ (Courant number) fields onto a grid - an array of points.
  Among the three considered languages only Fortran is equipped with built-in
    array handling facilities of practical use in high-performance computing.

  Presented C++ implementation of MPDATA is built upon the Blitz++ library. %TODO: cite something
  Blitz offers object-oriented representation of n-dimensional arrays,
    and array-valued mathematical expressions.
  In particular, it offers loop-free notation for array arithmetics
    that does not incur creation of intermediate temporary objects, 
    a crucial feature for achieving high-performance in array manipulation
    \citep[][]{Veldhuizen_et_al_1997}.
  Blitz++ is a header-only library\footnote{unless debugging mode is used} -- to use it, it is enough to include the appropriate header file,
    and optionally expose the required classes to the present namespace:
  \codecpp{code/cpp/listings.hpp}{listing02}{listing03}{1}
  Here, \prog{arr\_t}, \prog{rng\_t} and \prog{idx\_t} will serve as alias identifiers for the
    2D double precision version of the Blitz++ \prog{Array} 
    container, for the \prog{Range} class representing a sequence of integers numbering 
    array elements, and for the 2D version of the \prog{RectDomain} class representing 
    a vector of \prog{Range} objects defining a 2D slab of an array, respectively.

  Arguably, the power of Blitz++ comes from the possibility to express array expressions as objects.
  In particular, it is possible to define a function that returns an array expression;
    i.e. not the resultant array, but an object representing a ,,recipe'' defining the operations
    to be performed on the arguments.
  As a consequence, the return types of such functions become unintelligible, and the \prog{auto} 
    return type declaration from the C++11 standard allows to simplify the code significantly,
    even more if used through the following preprocessor macro:
  \codecpp{code/cpp/listings.hpp}{listing03}{listing04}{1}
    in which a call to \prog{blitz::safeToReturn()} function is also included ensuring
    all arrays involved in the expression being returned continue to exist in the
    caller scope.
  For example, definition of a function returning its array-valued argument doubled reads:
    \prog{auto~f(arr\_t~x)~return\_macro(2*x)}.
  This is the only preprocessor macro defined herein.

  For the Python implementation of MPDATA the NumPy package is used.
  In order to make the code compatible with both the standard CPython
    as well as the alternative PyPy implementation of Python \citep[][]{Bolz_et_al_2011},
    the Python implementation includes the following sequence of \prog{import} statements:
  \codepyt{code/pyt/listings.py}{listing02}{listing03}{1}
  First, the PyPy's built-in NumPy implementation named \prog{numpypy} is imported if applicable (i.e. if running PyPy), 
    and the lazy evaluation mode is turned on through the \prog{set\_invalidation(False)} call.
  PyPy's lazy evaluation obtained with the help of a just-in-time compiler enables it to achieve
    an analogous to Blitz++ temporary-array-free handling of array-valued expressions 
    (see discussion in section~\ref{sec:perf}).
  Second, the NumPy package is instructed to ignore any floating-point errors, if such an option
    is available in the interpreter (it is not supported in PyPy as of version 1.9).
  The above lines conclude all code modifications that needed to be added in order to run
    the code with PyPy.

  Fortran features built-in multidimensional array handling mechanisms,
    hence there is no need for using an external package as with C++ and Python.
  Fortran array-handling features are not object-oriented, however.
  In particular, it is not possible to alter the default behaviour through
    inheritance.

  \subsection{Array container sequences}\label{sec:sequence}

  While discretisation in space of the scalar field $\psi(x,y)$ into its $\psi_{[i,j]}$ 
    grid representation requires floating-point array containers (cf. section \ref{sec:array}), 
    discretisation in time requires a container class for storing
    sequences of such arrays, i.e. \{$\psi_{[i,j]}^{[n]}$, $\psi_{[i,j]}^{[n+1]}$\}.
  Similarly the vector field $\vec{C}$ is in fact a \{$C^{[x]}$, $C^{[y]}$\} sequence.

  Using an additional array dimension to represent the sequence elements is not considered for two reasons.
  First, the arrays constituting the sequence elements may have different sizes
      (e.g.  $C^{[x]}$ and $C^{[y]}$ on the Arakawa-C grid discussed in section~\ref{sec:grid}).
  Second, the order of dimensions would need to be different for different languages to assure that
      the contiguous dimension is not used for time levels.

  In the C++ implementation the Boost\footnote{
    Boost is a free and open-source collection of peer-reviewed C++ libraries available at \url{http://www.boost.org/}.
    Several parts of Boost have been integrated into or inspired new additions to the C++ standard.
  } \prog{ptr\_vector} class is used to represent sequences of Blitz++ arrays 
    and at the same time to handle automatic freeing of dynamically allocated memory.
  For clearer ,,translation'' of mathematical formul\ae~into C++ code, the
    \prog{ptr\_vector} class is further customised by defining a derived structure whose element-access \prog{[~]} 
    operator is overloaded with a modulo-variant:
  \codecpp{code/cpp/listings.hpp}{listing04}{listing05}{1}
  Consequently the last element of any such sequence may be accessed at index \prog{-1}, the last but one at \prog{-2}, 
    and so on.

  In the Python implementation the built-in \prog{tuple} type is used to store sequences of NumPy arrays.
  Employment of negative indices for handling from-the-end addressing of elements
    is a built-in feature of all sequence containers in Python.

  Fortran does not feature any built-in sequence container capable of storing arrays, 
    hence a custom \prog{arrvec\_t} type is introduced:
  \codefor{code/for/listings.f}{listing02}{listing03}{1}
  The \prog{arr\_t} and \prog{arrptr\_t} helper types are defined solely for the purpose 
    of overcoming the Fortran's limitation of not allowing to declare arrays of arrays.
  The \prog{ctor()} method initialises the container for a given number of elements \prog{n}.
  The \prog{init()} method initialises the \prog{n}-th element of the container with 
    an newly allocated 2D array spanning indices \prog{i(1)}:\prog{i(2)}, 
    and \prog{j(1)}:\prog{j(2)} in the first, and last dimensions respectively\footnote{When an array
    is passed as function argument in Fortran its base is locally set to unity, regardless of the setting
    at the caller scope, hence the first elements of \prog{i} and \prog{j} are always \prog{i(1)} and \prog{j(1)}.}.
  The \prog{dtor()} method serves as a destructor\footnote{as of time of writing there is no
    open-source Fortran compiler supporting destructor-like FINAL methods defined in the Fortran 2003 standard} 
    and deallocates all the memory allocated
    through previous calls to \prog{init}.
  Access to the arrays stored in the container is provided through the \prog{at} member field.
  Assuming \prog{psi} is an instance of \prog{arrptr\_t}, the \prog{(i,j)} element of the \prog{n}-th array in \prog{psi}
    may be accessed with \prog{psi\%at( n )\%p\%a( i, j )}.

%  TODO: mention why module procedures have to be prefixed with module name in Fortran
  \subsection{Staggered grid}\label{sec:grid}

  \begin{figure}[h!]
  \center
  \begin{tikzpicture}
    \coordinate (Origin)   at (0,0);
    \coordinate (XAxisMin) at (-2.5,0);
    \coordinate (XAxisMax) at (3,0);
    \coordinate (YAxisMin) at (0,-2.5);
    \coordinate (YAxisMax) at (0,3);
    \draw [thin, gray,-latex] (XAxisMin) -- (XAxisMax);% Draw x axis
    \draw [thin, gray,-latex] (YAxisMin) -- (YAxisMax);% Draw y axis

    \clip (-2.5,-2.5) rectangle (2.5cm,2.5cm); % Clips the picture...
    \draw[style=help lines,dashed] (-14,-14) grid[step=2cm] (14,14);
          % Draws a grid in the new coordinates.
          %\filldraw[fill=gray, fill opacity=0.3, draw=black] (0,0) rectangle (2,2);
              % Puts the shaded rectangle
    \foreach \x in {-7,-6,...,7}{% Two indices running over each
      \foreach \y in {-7,-6,...,7}{% node on the grid we have drawn 
        \node[draw,circle,inner sep=2pt,fill] at (2*\x,2*\y) {};
            % Places a dot at those points
      }
    }
    \draw [black] (0,0) -- (0,0) node [below right] {$\!\psi_{[i,j]}$};
    \draw [black] (-2,0) -- (-2,0) node [below right] {$\!\psi_{[i-1,j]}$};
    \draw [black] (0,2) -- (0,2) node [below right] {$\!\psi_{[i,j+1]}$};

    \draw [ultra thick,-latex,red] (.8,0) -- (1.2,0) node [above right] {$\!\!\!\!\!\!\!\!\!\!C^{[x]}_{[i+\phlf,j]}$};
    \draw [ultra thick,-latex,red] (-1.2,0) -- (-.8,0) node [above right] {$\!\!\!\!\!\!\!\!\!\!C^{[x]}_{[i\mhlf,j]}$};
    \draw [ultra thick,-latex,red] (.8,2) -- (1.2,2) node [above] {};
    \draw [ultra thick,-latex,red] (-1.2,2) -- (-.8,2) node [above] {};
    \draw [ultra thick,-latex,red] (.8,-2) -- (1.2,-2) node [above] {};
    \draw [ultra thick,-latex,red] (-1.2,-2) -- (-.8,-2) node [above] {};

    \draw [ultra thick,-latex,red] (0,.8) -- (0,1.2) node [above] {};
    \draw [ultra thick,-latex,red] (0,-1.2) -- (0,-.8) node [below right] {$C^{[y]}_{[i,j\mhlf]}$};
    \draw [ultra thick,-latex,red] (-2,.8) -- (-2,1.2) node [above] {};
    \draw [ultra thick,-latex,red] (-2,-1.2) -- (-2,-.8) node [above] {};
    \draw [ultra thick,-latex,red] (2,.8) -- (2,1.2) node [above] {};
    \draw [ultra thick,-latex,red] (2,-1.2) -- (2,-.8) node [above] {};
  \end{tikzpicture}
  \caption{\label{fig:grid}
    A schematic of the Arakawa-C grid.
  }
  \end{figure}
  The so-called Arakawa-C staggered grid \citep{Arakawa_and_Lamb_1977} depicted in Figure~\ref{fig:grid}.
    is a natural choice for MPDATA.
  MPDATA is a finite-difference approximation to the flux-form of the equation~(\ref{eq:adv}) 
    resulting in a finite-volume scheme.
  Using the Arakawa-C grid allows to evaluate the fluxes at 
    the grid cell (finite volume) edges without interpolation.

  As a consequence, the discretised representations of the $\psi$ scalar field, and each 
    component of the ${\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}}$ vector field 
    in eq.~(\ref{eq:adv}) are defined over different grid point locations.
  In mathematical notation this can be indicated by usage of fractional indices, e.g.
    $C^{[x]}_{[i\mhlf,j]}$, $C^{[x]}_{[i+\phlf,j]}$, $C^{[y]}_{[i,j\mhlf]}$ and $C^{[y]}_{[i,j+\phlf]}$
    to depict the grid values of the ${\vec{C}}$ vector components surrounding $\psi_{[i,j]}$.
  However, fractional indexing does not have a built-in counterpart in any of the
    employed programming languages.
  A desired syntax would translate:
  \begin{itemize}
    \item{\prog{$i-\phlf$} ~~ $\leadsto$ ~~ \prog{$i-1$}} 
    \item{\prog{$i+\phlf$} ~~ $\leadsto$ ~~ \prog{$i$}.}
  \end{itemize}
  OOP offers a convenient way to implement such notation
    by overloading the \prog{+} and \prog{-} operators for objects representing array indices.

%TODO: explain why h=(-1,0) is needed with Blitz and Python (in order not to lose memory)

  In the C++ implementation first a global instance \prog{h} of an empty structure 
    \prog{hlf\_t} is defined, and then the plus and minus operators for \prog{hlf\_t} and \prog{rng\_t} are overloaded:
  \codecpp{code/cpp/listings.hpp}{listing05}{listing06}{1}
  This way, the arrays representing vector field components can be indexed using
    \prog{(i+h,j)}, \prog{(i-h,j)} etc. where \prog{h}~represents the half.

  In NumPy in order to prevent copying of array data during slicing one needs to operate on the
    so-called array views.
  Array views are obtained when indexing the arrays with objects of the Python's
    built-it \prog{slice} type (or tuples of such objects in case of multi-dimensional arrays).
  Python forbids overloading of operators of built-in types such as \prog{slices}, 
    and does not define addition/subtraction operators for \prog{slice}/\prog{int} pairs.
  Consequently, a custom logic has to be defined not only for fractional indexing,
    but also for shifting the slices by integer intervals ($i\pm1$).
  It is implemented here by declaring a \prog{Shift} class with the adequate operator overloads:
  \codepyt{code/pyt/listings.py}{listing03}{listing04}{1}
    and two instances of it to represent unity and half
    in expressions like \prog{i+one}, \prog{i+hlf}, where \prog{i} is an instance of \prog{slice}
    \footnote{\label{fnt:slice}one could argue that not using an own implementation of a slice-representing class
    in NumPy is a design flaw -- being able to modify behaviour of a hypothetical numpy.slice class 
    through inheritance would allow to implement the same behaviour as obtained in listing P.3 without the need to represent 
    the unity as a separate class}:
  \codepyt{code/pyt/listings.py}{listing04}{listing05}{1}

  In Fortran, similarly as in C++ and Python, the logic related to employment of the Arakawa-C grid
    allowing employment of fractional indices in array indexing is obtained through
    definition and instantiation of an object representing the half, and having appropriate
    operator overloads:
  \codefor{code/for/listings.f}{listing03}{listing04}{1}

  \subsection{Halo regions}

  The MPDATA formul\ae~defining $\psi^{[n+1]}_{[i,j]}$ as a function of $\psi^{[n]}_{[i,j]}$
    (discussed in the following sections) feature terms such as $\psi_{[i-1,j-1]}$.
  One way of assuring validity of these formul\ae~on the edges of the domain (e.g. for i=0) 
    is to introduce the so-called halo region surrounding the domain.
  The method of populating the halo region with data depends on the boundary condition type.
  Employment of the halo-region logic implies repeated usage of array range 
    extensions in the code such as $i \leadsto i \pm halo$.

  An \prog{ext()} function is defined in all three implementation, in order to simplify 
    coding of array range extensions:
% TODO: works for int 1 and slice
  \codecpp{code/cpp/listings.hpp}{listing06}{listing07}{1}
  \codepyt{code/pyt/listings.py}{listing05}{listing06}{1}
  \codefor{code/for/listings.f}{listing04}{listing05}{1}
  Consequently, a range depicted by $i\pm1/2$ may be expressed in the code as \prog{ext(i, h)}.

%TODO: {discuss how halo region relates to parallel programming}

  \subsection{Array index permutations}\label{sec:pi}
  Hereinafter, the $\pi_{a,b}^{d}$ symbol is used to denote a  
    cyclic permutation of the order $d$ of the set $\{a,b\}$.
  It is used to generalise the MPDATA formul\ae~into multiple dimensions
    using the following notation:
  \begin{equation*}\label{eq:pi}
    \input{formulae/pi}
  \end{equation*}

  Blitz++ ships with the \prog{RectDomain} class (aliased here as \prog{idx\_t}) 
    for specifying array ranges in multiple dimensions.
  The $\pi$ permutation is implemented in C++ as a function \prog{pi()}
    returning an instance of \prog{idx\_t} constructed with a permutation
    of the \prog{i} and \prog{j} arguments of \prog{pi()}, with permutation order
    dependant on the template parameter \prog{d}:
  \codecpp{code/cpp/listings.hpp}{listing07}{listing08}{1}

  NumPy uses tuples of slices for addressing multi-dimensional arrays with a
    single object, hence the following definition of function \prog{pi()} suffices to represent $\pi$:
  \codepyt{code/pyt/listings.py}{listing06}{listing07}{1}

  In the Fortran implementation an analogous behaviour is obtained by introducing
    a function \prog{pi()} returning a pointer the array elements specified
    by \prog{i} and \prog{j} interpreted as (i,j) or
    (j,i) depending on the value of the argument \prog{d}.
  In addition to \prog{pi()}, a helper \prog{span()} function returning the 
    length of one of the vectors passed as argument is defined:
  \codefor{code/for/listings.f}{listing05}{listing06}{1}

%  TODO: templates -> branchless code by design, while F relies on opti

  \subsection{A prototype solver}\label{sec:solver}

  The tasks to be handled by a prototype advection equation solver proposed herein are:
  \begin{description}
    \item[(i)]{storing arrays representing the $\psi$ and $\vec{C}$ fields and any required housekeeping data,}
    \item[(ii)]{allocating and deallocating the required memory,}
    \item[(iii)]{providing access to the solver state, and}
    \item[(iv)]{performing the integration by invoking the advection-operator and
      boundary-condition handling routines.}
  \end{description}
  In the following C++ definition of the \prog{solver} structure, task (i) is represented with the definition of the structure
    member fields; task (ii) is split between the \prog{solver}'s constructor and the destructors of \prog{arrvec\_t};
    task (iii) is handled by the accessor methods; task (iv) is handled within the \prog{solve} method:
  \codecpp{code/cpp/listings.hpp}{listing08}{listing09}{1}
  The \prog{solver} structure is an abstract definition (containing a pure virtual method) 
     requiring its ancestors to implement at least
     the \prog{advop()} method which is expected to fill \prog{psi[n+1]} with the values of \prog{psi[n]} advected
     within one timestep.
  The two template parameters \prog{bcx\_t} and \prog{bcy\_t} allow the solver to operate with 
    any kind of boundary condition structures that fulfil the requirements implied by the calls to the 
    methods of \prog{bcx} and \prog{bcy}, respectively.

  The sizes of the arrays representing the two time-levels $\psi^{[n]}$ and $\psi^{[n+1]}$ of $\psi$ 
    are defined by the domain size ({\em nx}$~\times$~{\em ny}) plus the halo region.
  The size of the halo region is an argument of the constructor.

  The arrays representing the $C^{[x]}$ and $C^{[y]}$ components of $\vec{C}$,
    require {\em nx+1}~$\times$~{\em ny} 
    and {\em nx}$~\times$~{\em ny+1} elements, respectively
    (being laid out on the Arakawa-C staggered grid).

%  TODO: clarify what cycle() does

  Python definition of the \prog{Solver} class follows closely the C++ structure definition:
  \codepyt{code/pyt/listings.py}{listing07}{listing08}{1}
  The key difference stems from the fact that unlike Blitz++, NumPy does not allow an array
    to have arbitrary index base -- in NumPy the first element is always addressed with 0.
  Consequently, while in C++ (and Fortran) the computational domain is chosen to start at (i=0, j=0)
    and hence a part of the halo region to have negative indices, in Python the halo region starts %TODO: footnote o roznych zakresach dla MPI
    at (0,0).
  However, since the whole halo logic is hidden within the solver, such details are not exposed to the
    user.
  The \prog{bcx} and \prog{bcy} boundary-condition specifications are passed to the solver through
    constructor-like \prog{\_\_init\_\_()} method as opposed to template parameters in C++.

  The C++ and Python prototype solver implementations introduced above
    allowed the solvers to operate with any boundary condition
    objects provided the requirements imposed by calls to these objects' methods were fulfilled 
    (at compile-time in case of C++, and at run-time in the case of Python).
  In order to obtain an analogous behaviour with Fortran, it is required to
    define, prior to definition of a solver type, 
    an abstract type with deferred procedures having abstract interfaces 
    \citep[see Table 2.1 in][for a summary of approximate correspondence of OOP 
      nomenclature between Fortran and C++]{Rouson_et_al_2012}:
  \codefor{code/for/listings.f}{listing06}{listing07}{1}
  Having defined the abstract type for boundary-condition objects, 
    a definition of a solver class following closely the C++ and Python counterparts may be provided:
  \codefor{code/for/listings.f}{listing07}{listing08}{1}

  As a side note, it is worth mentioning that the proposed design is readily extendable 
    to handle concurrent computations using domain decomposition.
  The proposed prototype solver is basically ready to be applied 
    in a distributed memory set-up (e.g. using MPI)
    since all necessary communication logic could be placed in the boundary-condition objects
    not requiring any changes to the solver itself.
  Shared-memory parallelisation (e.g. OpenMP) would require minimal changes to the solver structure
    (the \prog{advop()} method would need to be called in a concurrently executed loop),
    leaving the boundary condition- and advection operator-representing logic basically untouched.

  Finally, it has to be underlined that the proposed prototype solver is designed
    just as a minimal example to present the concepts described in the paper.
%  It is by no means an optimal or versatile solution.

  \subsection{Periodic boundaries (C++)}\label{sec:cyclic}

  The solver definition described in section~\ref{sec:solver} required a given
    boundary condition object to implement a \prog{fill\_halos()} method.
  An implementation of periodic boundary conditions in C++ is provided in the following listing:
  \codecpp{code/cpp/listings.hpp}{listing09}{listing10}{1}

  As hinted by the member field names, the \prog{fill\_halos()} methods
    fill the left/right halo regions with data from the right/left edges of the domain.
  Thanks to employment of the function \prog{pi()} described in section~\ref{sec:pi}
    the same code may be applied in any dimension (here being a template parameter).

  Listings P.8 and F.8 contain the Python and Fortran counterparts to listing C.9.

  \subsection{Donor-cell formul\ae~(C++)}\label{sec:donor}
  MPDATA is an iterative algorithm in which each iteration takes the form of the 
    so-called donor-cell formula which itself is a first-order advection scheme.
  The set of functions related to the donor-cell scheme is encapsulated within
    the \prog{donorcell} namespace in C++:
  \codecpp{code/cpp/listings.hpp}{listing10}{listing11}{1}

  MPDATA and donor-cell are explicit forward-in-time algorithms -- they allow to predict $\psi^{[n+1]}$ as a
    function of $\psi^{[n]}$ where $n$ and $n+1$ denote two adjacent time levels.
  The donor-cell scheme may be written as \citep[eq. 2 in][]{Smolarkiewicz_1984}:
  \begin{equation}\label{eq:donor}
    \input{formulae/donor}
  \end{equation}
  where $N$ is the number of dimensions, 
    and F is the so-called flux function \citep[][eq.~3]{Smolarkiewicz_1984}:
  \begin{equation}\label{eq:donor:F}
    \input{formulae/donor-F}
  \end{equation}

  The flux function takes the following form in C++:
  \codecpp{code/cpp/listings.hpp}{listing11}{listing12}{1}

  Equation~\ref{eq:donor} is split into the terms under the summation 
    (effectively the 1-dimensional donor-cell formula):
  \codecpp{code/cpp/listings.hpp}{listing12}{listing13}{1}

  and the actual two-dimensional donor-cell formula:
  \codecpp{code/cpp/listings.hpp}{listing13}{listing14}{1}

  Finally, the C++ namespace is ended with:
  \codecpp{code/cpp/listings.hpp}{listing14}{listing15}{1}

  Listings P.9-P11 and F.9-F.13 contain the Python and Fortran counterparts to listings C.12-C.15.

  \subsection{A donor-cell solver (C++)}\label{sec:donorcell_solver}

  As mentioned in the previous section, the donor-cell formula
    constitutes an advection scheme, hence we may use it to create
    a \prog{solver\_donorcell} implementation of the abstract \prog{solver} class:
  \codecpp{code/cpp/listings.hpp}{listing15}{listing16}{1}

  Listings P.12 and F.14 contain the Python and Fortran counterparts to listing C.16.

  \subsection{MPDATA formul\ae~(C++)}\label{sec:mpdata}

  MPDATA introduces corrective steps to the algorithm defined by equation~\ref{eq:donor} and \ref{eq:donor:F}.
  Each corrective step has the form of equation~\ref{eq:donor} with the Courant numbers replaced
    with those defined by the so-called antidiffusive velocities, and with $\psi$ fields from 
    the previous iterations.
  The Courant number field corresponding to the basic MPDATA antidiffusive velocity takes the following 
    form \citep[eqs~13,~14~in][]{Smolarkiewicz_1984}:
  \begin{equation}\label{eq:antidiff1}
    \input{formulae/antidiff1}
  \end{equation}
  where
  \begin{equation}\label{eq:antidiff2}
    \input{formulae/antidiff2}
  \end{equation}
  and where for positive-definite $\psi$, the $A$ and $B$ terms take the 
  following form\footnote{
    Since $\psi\ge0$, $|A|\le1$ and $|B|\le1$. 
    See \citet[Sec. 4.2]{Smolarkiewicz_2006} for description of adaptation of the 
    formul\ae~for advection of fields of variable sign
  }:
  \begin{equation}\label{eq:A}
    \input{formulae/A}
  \end{equation}
  \begin{equation}\label{eq:B}
    \input{formulae/B}
  \end{equation}

  The set of functions related to MPDATA is encapsulated within
    the \prog{mpdata} namespace in C++:
  \codecpp{code/cpp/listings.hpp}{listing16}{listing17}{1}

  If the denominator in equations \ref{eq:A} or \ref{eq:B} equals zero for a given {\em i} and {\em j}, 
    the corresponding $A_{[i,j]}$ and $B_{[i,j]}$ are set to zero what may be conveniently 
    represented with the \prog{where} construct (available in all three considered languages):
  \codecpp{code/cpp/listings.hpp}{listing17}{listing18}{1}

  The $A$ term defined in equation \ref{eq:A} takes the following form:
  \codecpp{code/cpp/listings.hpp}{listing18}{listing19}{1}

  The $B$ term defined in equation \ref{eq:B} takes the following form:
  \codecpp{code/cpp/listings.hpp}{listing19}{listing20}{1}

  Equation \ref{eq:antidiff2} takes the following form:
  \codecpp{code/cpp/listings.hpp}{listing20}{listing21}{1}

  Equation \ref{eq:antidiff1} take the following form:
  \codecpp{code/cpp/listings.hpp}{listing21}{listing22}{1}

  Finally, the C++ namespace is ended with:
  \codecpp{code/cpp/listings.hpp}{listing22}{listing23}{1}

  Listings P.13-PP.17 and F.15-F.21 contain the Python and Fortran counterparts to listing C.16-C.22.

  \subsection{An MPDATA solver (C++)}\label{sec:mpdatasolver}

  An MPDATA solver may be now constructed with the following definition in C++:
  \codecpp{code/cpp/listings.hpp}{listing23}{listing24}{1}

  \subsection{Usage example (C++)}\label{sec:example}

  The following listing provides an example of how the MPDATA solver
    defined in section~\ref{sec:mpdatasolver} may be used together
    with the cyclic boundary conditions defined in section~\ref{sec:cyclic}.
  In the example a Gaussian signal is advected in a 2D domain 
    defined over a grid of 24$\times$24 cells.
  The program first plots the initial condition, then performs
    the integration for 75 timesteps with three different 
    settings of the number of iterations used in MPDATA.
  The signal shape at the end of each simulation is plotted as well.
  Plotting is done with the help of the gnuplot-iostream library.
  The resultant plot is presented herein as Figure~\ref{fig:mpdata}

  \codecpp{code/cpp/plot.cpp}{listing24}{listing25}{1}

  \begin{figure}
    \pgfimage[height=.95\textheight]{code/cpp/figure}
    \caption{\label{fig:mpdata}
      Plot generated by the program given in listing~C.24.
      See section~\ref{sec:example} for details.
      %TODO: describe the test, point out adv., numerical diffusion, periodic boundaries, not much diff between 2 and 7
    }
  \end{figure}

  \section{Performance evaluation}\label{sec:perf}

  \begin{figure}
    \center
    \includegraphics[height=.45\textwidth,angle=-90]{plot-mem-host=eyrie-cpu=GenuineIntel-6-15-bogomips=3190}
    \caption{\label{fig:mem}
      Memory consumption statistics for the test runs described in Section~\ref{sec:perf}
        plotted as a function of grid size.
      Peak resident set size (rss) values reported by the GNU time utility are normalised by the size of
        data that needs to be allocated in the program to store all declared grid-sized arrays.
      Asymptotic values reached at the largest grid sizes are indicative 
        of temporary storage requirements.
    }
  \end{figure}
  \begin{figure}
    \center
    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=eyrie-cpu=GenuineIntel-6-15-bogomips=3190}
    \caption{\label{fig:cpu-eyrie}
      Execution time statistics for the test runs described in Section~\ref{sec:perf}
        plotted as a function of grid size.
      Values of the total user mode cpu time reported by the GNU time utility are
        and normalised by the grid size ($nx=ny$) and the number of timesteps $nt=2^{24}/(nx \cdot ny)$.
      Before normalisation the time reported for an $nt=0$ run for a corresponding
        domain size is subtracted from the values.
      Both the $nt=0$ and $nt=2^{24}/()nx \cdot ny$ runs are repeated three times and
        the shortest times are taken into account only.
      Results obtained with an Intel\textsuperscript{\textregistered} 
        Core\textsuperscript{\texttrademark} 2 Duo 1.6 GHz processor.
    }
  \end{figure}
  \begin{figure}
    \center
    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=skua-cpu=AuthenticAMD-16-10-bogomips=5620}
    \caption{\label{fig:cpu-skua}
      Same as Fig.~\ref{fig:cpu-eyrie} for an AMD Phenom\textsuperscript{\texttrademark} II 800 MHz processor.
    }
  \end{figure}

  The three introduced implementations of MPDATA were tested with the following set-ups 
    employing free and open-source tools:
  \begin{description}
    \item[C++:]{~
      \begin{itemize}
        \item{GCC g++ 4.8.0\footnote{\label{fnt:gcc-snapshot}GNU Compiler Collection trunk revision 193662 packaged in the Debian's gcc-snapshot\_20121120-1}
          with Blitz++ 0.10}
      \end{itemize}
    }
    \item[Python:]{~
      \begin{itemize}
        \item{CPython 2.7.3 with NumPy 1.6.2}
        \item{PyPy 1.9.0 with built-in NumPy implementation}
      \end{itemize}
    }
    \item[Fortran:]{~
      \begin{itemize}
        \item{GCC gfortran 4.8.0\textsuperscript{\ref{fnt:gcc-snapshot}}}
      \end{itemize}
    }
  \end{description}
  All performance tests were run on a Debian GNU/Linux system with the above-listed software obtained 
    via binary packages from the Debian's ''sid'' collection (most recent package versions at the time of writing).
  The tests were performed on two 64-bit machines equipped with
    an AMD Phenom\textsuperscript{\texttrademark} II X6 1055T (800 MHz, ca. 5600 BogoMIPS)
    and an Intel\textsuperscript{\textregistered} Core\textsuperscript{\texttrademark}~2~Duo P7500 (1.6 GHz, ca. 3200 BogoMIPS)
    processors.

  For both C++ and Fortran the GCC compilers were invoked with the \prog{-Ofast} and \prog{-march=native} 
    options imposing, among other optimisations, automatic vectorisation and finite-only arithmetics 
    (i.e.  non-standard-compliant treatment of NaN and Inf values).
  GCC compilation did result in vectorised code for C++ only (vectorised loops within the Blitz++ headers, 
    as reported by the messages triggered by the \prog{-ftree-vectorizer-verbose} option).
%  At least a part of possible auto-vectorisations was not done with the Fortran implementation 
%    due to employment of non-contiguous pointers (in particular when referring to the grid elements 
%    excluding the halo region).

  The Python code was tested with the standard Python implementation CPython and with PyPy.
  PyPy is an alternative implementation of Python featuring a just-in-time compiler. 
  PyPy includes a partial reimplementation of NumPy that compiles NumPy expressions into native assembler.
  Thanks to employment of lazy evaluation of array expressions (cf. Sect. \ref{sec:array})
    PyPy allows to eliminate the use of temporary matrices for storing intermediate results,
    and to perform multiple operations on the arrays within a single array index traversal.
  Consequently, PyPy allows to overcome the same performance-limiting factors as those addressed by Blitz++, although 
    the underlying mechanisms are different.
  In contrast to other solutions for improving performance of NumPy-based codes such as
    Cython or numexpr, PyPy does not require
    any modifications to the code and allows to preserve its object-oriented structure.
  PyPy may thus serve as a drop-in replacement for CPython ready to be used with 
    previously-developed codes.

  The same set of tests was run with all four set-ups.
  Each test set consisted of 16 program runs.
  The test programs are analogous to the example code presented in section~\ref{sec:example}.
  The tests were run with different grid sizes ranging from 64$\times$64 to 2048$\times$2048.
  The Gaussian impulse was advected for $nt=2^{24}/(nx\cdot ny)$ timesteps, 
    in order to assure comparable timing accuracy for all grid sizes.
  Three MPDATA iterations were used (i.e. two corrective steps).
  The initial condition was loaded from a text file, and the final values were compared at the end of the test
    with values loaded from another text file assuring the same results were obtained with all four set-ups.
  The tests were run multiple times; program startup, data loading, and output verification times were
    subtracted from the reported values (see caption of Figure~\ref{fig:cpu-eyrie} for details).

  Figure \ref{fig:mem} presents a plot of the peak memory use\footnote{the resident set size (rss)
    as reported by GNU time (version 1.7-24)} as a function of grid size.
  The plotted values are normalised by the nominal size of all data arrays used in the program
    (i.e. two (nx+2)$\times$(ny+2) arrays representing the two time levels of $\psi$, 
     a (nx+1)$\times$(ny+2) array representing the $C^{[x]}$ component of the Courant number field,
     a (nx+2)$\times$(ny+1) array representing the $C^{[y]}$ component, 
     and two pairs of arrays of the size of $C^{[x]}$ and $C^{[y]}$ for storing the 
     antidiffusive velocities, all composed of 8-byte double-precision floating point numbers).
  Plotted statistics reveal a notable memory footprint of the Python interpreter itself
    for both CPython and PyPy, losing its significance for domains larger than 1024$\times$1024.
  The roughly asymptotic values reached in all four set-ups for grid sizes larger that 1024$\times$1024
    are indicative of the amount of temporary memory used for array manipulation.
  PyPy- and Blitz++-based set-ups consume notably less memory than Fortran and CPython confirming 
    the effectiveness of the just-in-time compilation (PyPy), and the expression-templates (Blitz++) techniques
    for elimination of temporary storage during array operations.

  The CPU time statistics presented in Figures \ref{fig:cpu-eyrie} and \ref{fig:cpu-skua} reveal
    notable differences between results obtained with the two different processors,
    however the results agree qualitatively leading to the following observations
    (where by referring to language names, only the results obtained with the herein considered
     program codes, and software/hardware configurations are meant):
  \begin{itemize}
    \item{Fortran gives shortest execution times for any domain size;}
    \item{C++ execution times are less than twice those of Fortran for grids larger than 
      256$\times$256;}
    \item{CPython requires from around 4 to almost 10 times more CPU time than Fortran depending on the grid size;}
    \item{PyPy execution times are in most cases closer to C++ than CPython.}
  \end{itemize}
  The support for OOP features in gfortran, the NumPy support in PyPy, and the auto-vectorisation
    mechanisms in GCC are still in active development and hence the performance with some of the set-ups may 
    likely change with newer versions of these packages.

  It is worth mentioning, that even though the three implementations are equally structured,
    the three considered languages have some inherent differences influencing the execution times.
  Notably, while Fortran and Blitz++ offer runtime array-bounds and array-shape checks as options
    not intended for use in production binaries, NumPy performs them always.
  Additionally, the C++ and Fortran set-ups may, in principle, benefit from GCC's auto-vectorisation
    features which do not have yet counterparts in CPython or PyPy.
  Finally, Fortran uses different ordering for storing array elements in memory, but since
    all tests were carried out using square grids, this should not have had any impact on the
    performance (both Blitz++ and NumPy support Fortran's column-major ordering as well, 
    however this feature is still missing from PyPy's built-in NumPy implementation as of PyPy 1.9)

  The authors do expect some performance gain could readily
    be obtained by introducing into the codes what is referred to as
    {\em premature optimisation} -- code rearrangements aimed solely
    at the purpose of increasing performance.
  These were avoided intentionally as they degrade code readability,
    should in principle be handled by the compilers,
    and are generally advised to be avoided \citep[e.g.][section 3.12]{bib_CERNcpp}.

  \section{Discussion on the tradeoffs of language choice}

  One of the aims of this paper is to exemplify the applicability of OOP features of the three considered
    programming languages (or language-library pairs) for representing
    what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012} within the source code
    of scientific software.
  Presented benchmark tests, although quite simplistic, together with the experience gained 
    from the development of equally-structured codes in three different
    languages provide a basis for discussion on the tradeoffs of the programming language.
  The discussion concerns in principle the development of finite-difference solvers for 
    partial differential equations, but is likely applicable to same extent to
    development of scientific software in general.
  A partly objective and partly subjective summary of pros and cons of C++, Python and Fortran,
    as observed during the development of the presented codes, in presented in the four following subsections.

  \subsection{OOP for blackboard abstractions}

  It was shown in section~\ref{sec:impl} that C++/Blitz++, Python/NumPy and Fortran
    provide comparable functionalities in terms of matching the blackboard abstractions
    within the program code.
  Taking into account solely the part of code representing particular formul\ae~
    (e.g. listings C.21, P.17, F.20 and equation \ref{eq:antidiff1}) all three
    languages allow to match (or surpass) \LaTeX~in its brevity of formula translation syntax.
  All three languages were shown to be capable of providing mechanisms to compactly represent such abstractions as:
  \begin{itemize}
    \item{loop-free array arithmetics}
    \item{definitions of functions returning array-valued expressions}
    \item{permutations of array indices allowing dimension-independent definitions
      of functions (see e.g. listings C.12 and C.13, P.10 and P.11, F11 and F.12)}
    \item{fractional indexing of arrays corresponding to employment of a staggered grid}
    \item{associating identifiers with slabs of arrays (e.g. for representing 
      a psi[left\_halo]=psi[rght\_edge] assignment)}
  \end{itemize}
  Three issues specific to Fortran were observed which 
    resulted in employment in some places of a more repetitive or cumbersome syntax than in C++ or Python:
  \begin{itemize}
    \item{Fortran does not feature a mechanism allowing to reuse a single piece of code (algorithm)
      with different data types (compare e.g. listings C.6, P.5 and F.4) such as
      templates in C++ and the so-called duck typing in Python;}
    \item{Fortran does not allow function calls to appear on the left hand side
      of assignment (see e.g. how the \prog{tmp} pointers were used as a workaround in the \prog{cyclic\_fill\_halos}
      method in listing F.8)}
    \item{Fortran's built-in array-handling features are not object-oriented,
      and hence not easily extendable via inheritance, but also subject to
      some inherent limitations such as lack of support for arrays of arrays 
      (cf. Sect. \ref{sec:sequence})}
  \end{itemize}
  Interestingly, the limitation in extendability via inheritance was found to
    exist partially in NumPy as well (see footnote \ref{fnt:slice}).
  The lack of a counterpart in Fortran to the C++ template mechanism was identified
    \citep{Cary_et_al_1997}
    as one of the key deficiencies of Fortran when compared with C++ in context 
    of applicability to object-oriented scientific programming.

  \subsection{Performance}

  The timing and memory usage statistics presented in figures \ref{fig:mem}-\ref{fig:cpu-skua}
    reveal that no single language/library/compiler set-up 
    corresponded to both shortest execution time and smallest memory footprint.
  Moreover, the shortest execution times of Fortran correspond to largest memory use.

  One may consider performance measures addressing not only the program efficiency but also 
    the factors influencing the development and maintenance time/cost.
  Taking into account such measures as code length or coding time,
    the Python environment clearly gains significantly.
  Presented Python code is shorter than the C++ and Fortran counterparts,
    and is simpler in terms of syntax and usage (see discussion below).

  Employment of the PyPy drop-in replacement for the standard Python implementation brings 
    Python's performance significantly closer to those of C++ and Fortran, in some
    cases making it the least memory consuming set-up.
  Python has already been the language of choice for scientific software projects having code clarity 
    or ease of use as the first requirement \citep[see e.g.][]{Barnes_and_Jones_2011}.
  PyPy's capability to improve performance of unmodified Python code may actually 
    make Python a favourable choice even if high performance is crucial, especially
    if a performance-to-development cost measure is to be considered.

  \subsection{Ease of use and abuse}

  Using the number of lines of code or the number of distinct language keywords
    needed to implement the MPDATA-based
    solver presented in section \ref{sec:impl} as measures of syntax 
    brevity, Python clearly surpasses its rivals.
  Python was developed with emphasis on code readability and object-orientation.
  Arguably, taking it to the extreme - Python uses line indentation to define 
    blocks of code and treats even single integers as objects.
  As a consequence Python is easy to learn and easy to teach.
  It is also much harder to abuse Python than C++ or Fortran
    (for instance with \prog{goto} statements, employment of the preprocessor,
    or the implicit typing in Fortran).

  Python simplifies also the building and dissemination of software,
    especially if multiple architectures and operating systems are targeted.
  However, CMake is a tool that allows to efficiently automate builds, tests and 
    packaging of C++ and Fortran programs.

  Python is also definitely easiest to debug among the three languages.
  There exist great debugging tools for C++, but the debugging and development is 
    often hindered by indecipherable compiler messages flooded with lengthy type names
    stemming from employment of templates.
  OOP with Fortran is a relatively new concept, and the support for the Fortran OOP
    features among free and open source compilers, debuggers and other programming
    aids in still not mature.

  A factor contributing to the ease of use of the array-handling 
    facilities is the clarity of rules determining their performance.
  With both Fortran and Python, the memory footprint caused by employment
    of temporary objects in array arithmetics is dependant on compiler optimisations.
  In contrast, Blitz++ ensures temporary-array-free computations by design.
  Similarly, the C++ template mechanism and such features as the C++11
    \prog{constexpr} construct allow the programmer to enforce evaluation
    at compile time\footnote{Useful in context of scientific computing for example to perform
    compile-time dimensional analysis of the physically-meaningful expressions within
    the code with no runtime performance penalty \citet{Schabel_et_al_2008}}.

  \subsection{Added values}

  The size of the programmers' community of a given language 
    influence the availability of trained personnel, 
    the availability of reusable software components,
    the maturity of compilers and tools, 
    and the volume and coverage of online and printed information resources.
  The OOP features were introduced to Fortran only recently, and have not gained
    wide popularity among users \citep{Worth_2008}\footnote{an anecdotal yet significant
    example being the incomplete support for syntax-highlighting of modern Fortran in Vim and Emacs editors}.
  Fortran is no longer routinely taught at the universities \citep{Kendall_et_al_2008},
    in contrast to C++ and Python.
  An example confirming the arguably ceasing popularity of Fortran in academia 
    is the discontinuation of all but C++ printed editions of the ''Numerical Recipes'' 
    series of Press et al.
  Fortran is a domain-specific language while Python and C++ are general-purpose languages
    with disproportionately larger users' communities.

  Yet still, in the case of C++ and Python it is also the size of the particular
    numerical-library users' community that is of importance.
  Blitz++ is one of several packages that offer high-performance object-oriented
    array manipulation functionality with C++ (and is not necessarily optimal for every
    purpose \citep{Iglberger_et_al_2012}).
  In contrast, the NumPy package became a de facto standard solution for Python.
  Consequently, numerous Python libraries adopted NumPy but
    there are apparently very few C++ libraries offering Blitz++ support out of the box
    (the gnuplot-iostream used in listing C.24 being a much-appreciated counterexample).
  However, Blitz++ allows to instantiate its arrays using preexisting data what
    allows to interface with virtually any library, although requiring additional
    coding.

  The availability and quality of object-oriented 
    libraries differs among the three considered languages.
  The built-in standard libraries of Python and C++ are richer than
    those of Fortran and offer versatile data types, collections of
    algorithms and facilities for interaction with host operating system.
%  External Fortran libraries are mostly limited to scientific applications, 
%    while there exist collections of general-purpose C++ and Python libraries.
%  Interfacing any of the three considered languages from
%    another one does not pose a problem, though.
  In the authors' experience, the small popularity of OOP techniques among
    Fortran users is reflected in the library designs.
  What makes correct use of external libraries more difficult with Fortran
    is the lack of standard exception handling mechanisms in Fortran,
    long and {\em much requested by the numerical community} \citep[][Foreword]{Press_et_al_1996}.

  Finally, the three languages differ as well with regard to availability of 
    mechanisms for leveraging shared-memory parallelisation (e.g. with multi-core processors).
  GCC supports OpenMP with Fortran and C++, as well as the C++11 built-in support for threads.
  The CPython and PyPy implementations of Python don't offer any
    built-in solution for multithreading. 
  However, there are third-party party libraries such as numexpr which can employ multiple cores from
    a single program.
  There exist object-oriented libraries for both C++ and Python to 
    transparently use graphics processing units (GPU) such as Thrust \citep{Thrust}
    and Theano \citep{Bergstra_et_al_2010}.

  \section{Summary and outlook}

  Three object-oriented implementations of a prototype solver 
    for the advection equation were introduced.
  The solvers are based on MPDATA - an algorithm of particular applicability
    in geophysical fluid dynamics \citep{Smolarkiewicz_2006}.
  The three implementations follow the same structure but are implemented
    in three different languages (or language/library pairs):
  \begin{itemize}
    \item{C++ with Blitz++,}
    \item{Python with NumPy and}
    \item{Fortran}
  \end{itemize}
  It was the aim of this study to assess the capabilities of these languages
    for object-oriented scientific programming in context of 
    development of finite-difference partial differential equation solvers.
  The emphasis was on the capabilities for reproduction within 
    the program code of the mathematical abstractions used for defining the algorithms
    in journal papers or lecture blackboards.

  Presented programs were developed making use of such recent
    developments as support for C++11 and Fortran 2008 in GCC, and
    the NumPy support in the PyPy implementation of Python.
  The key conclusion is that the considered language/library/compiler
    set-ups offer unprecedented possibilities for using OOP to compactly 
    and faithfully represent within the program source code 
    the mathematical abstractions employed when describing 
    numerical algorithms in literature.
  The readability and brevity of such code may contribute to its 
    auditability, indispensable for credible and reproducible research in computational science 
    \citep{Post_et_al_2005, Merali_et_al_2010, Stodden_et_al_2012}.
  Employment of OOP helps as well to keep complex programs maintainable, avoiding accumulation of the technical/code 
    debt \citep{Buschmann_2011} that besets scientific software in such domains as climate modelling
    \citep{Freeman_et_al_2010}.
  The fact that all considered standards are open and the employed 
    tools implementing them are free and open-source
    is of importance as well \citep{Anel_2011}.

  The performance evaluation revealed that:
    \begin{itemize}
      \item{the Fortran set-up offered shortest execution times,}
      \item{it took the C++ set-up less than twice longer to compute than Fortran,}
      \item{C++ and PyPy set-ups offered significantly smaller memory consumption 
        than Fortran and CPython,}
      \item{the PyPy set-up was roughly twice slower than C++ and up to twice faster than CPython.}
    \end{itemize}
  The three equally-structured implementations required ca. 200, 300, and 550 lines of code
    in Python, C++ and Fortran, respectively.

  In addition to the source code presented within the text,
    a set of tests and build-/test-automation scripts
    allowing to reproduce the analysis and plots presented in section~\ref{sec:perf} are all 
    available for download from the project repository\footnote{git repository at \url{http://github.com/slayoo/mpdata/}},
    and are released under the GNU GPL license \citep{GPLv3}.

  The OOP design enhances the possibilities to reuse and extend the presented code.
  Development is underway of an object-oriented C++ library featuring concepts presented herein,
    supporting integration in one to three dimensions, handling systems of equations with source terms, 
    providing miscellaneous options of MPDATA and several parallel processing approaches.

  \section*{Acknowledgements}
  {\small
    \noindent 
    Thanks are due Piotr Smolarkiewicz and Hanna Pawłowska.
    \noindent
    Authors would like to thank Tobias Burnus, Julian Cummings, Patrik Jonsson and
      Arjen Markus for their feedback on questions posted to Blitz++ and gfortran
      mailing lists.

    \noindent
    SA, AJ and DJ acknowledge funding from the Polish National Science Centre
      (project no. 2011/01/N/ST10/01483).

    \noindent
    Part of the work was carried out during a visit of SA to the National
      Center for Atmospheric Research (NCAR) in Boulder, Colorado, USA.
    NCAR is operated by the University Corporation for Atmospheric Research.
    The visit was funded by the Foundation for Polish Science (START programme).

    \noindent
    Development of NumPy support in PyPy was led by Alex Gaynor, Matti Picus and MF.
  }

  \appendix

  \setcounter{section}{15} 
  \section{Python code for sections \ref{sec:cyclic}--\ref{sec:mpdatasolver}}\label{app:P}

  \subsection*{{\bf Periodic Boundaries} (cf. Sect. \ref{sec:cyclic})}
  \noindent 
  \codepyt{code/pyt/listings.py}{listing08}{listing09}{1}

  \subsection*{{\bf Donor-cell formul\ae}~(cf. Sect. \ref{sec:donor})}
  \codepyt{code/pyt/listings.py}{listing09}{listing10}{1}
  \codepyt{code/pyt/listings.py}{listing10}{listing11}{1}
  \codepyt{code/pyt/listings.py}{listing11}{listing12}{1}

  \subsection*{{\bf Donor-cell solver} (cf. Sect. \ref{sec:donorcell_solver})}
  \codepyt{code/pyt/listings.py}{listing12}{listing13}{1}

  \subsection*{{\bf MPDATA formul\ae}~(cf. Sect. \ref{sec:mpdata})}
  \codepyt{code/pyt/listings.py}{listing13}{listing14}{1}
  \codepyt{code/pyt/listings.py}{listing14}{listing15}{1}
  \codepyt{code/pyt/listings.py}{listing15}{listing16}{1}
  \codepyt{code/pyt/listings.py}{listing16}{listing17}{1}
  \codepyt{code/pyt/listings.py}{listing17}{listing18}{1}

  \subsection*{{\bf An MPDATA solver} (cf. Sect. \ref{sec:mpdatasolver})}
  \codepyt{code/pyt/listings.py}{listing18}{listing19}{1}

  \setcounter{section}{5} 
  \section{Fortran code for sections \ref{sec:cyclic}--\ref{sec:mpdatasolver}}\label{app:F}

  \subsection*{{\bf Periodic boundaries} (cf. Sect. \ref{sec:cyclic})}
  \codefor{code/for/listings.f}{listing08}{listing09}{1}

  \subsection*{{\bf Donor-cell formul\ae}~(cf. Sect. \ref{sec:donor})}
  \codefor{code/for/listings.f}{listing09}{listing10}{1}
  \codefor{code/for/listings.f}{listing10}{listing11}{1}
  \codefor{code/for/listings.f}{listing11}{listing12}{1}
  \codefor{code/for/listings.f}{listing12}{listing13}{1}
  \codefor{code/for/listings.f}{listing13}{listing14}{1}

  \subsection*{{\bf Donor-cell solver} (cf. Sect. \ref{sec:donorcell_solver})}
  \codefor{code/for/listings.f}{listing14}{listing15}{1}

  \subsection*{{\bf MPDATA formul\ae}~(cf. Sect. \ref{sec:mpdata})}
  \codefor{code/for/listings.f}{listing15}{listing16}{1}
  \codefor{code/for/listings.f}{listing16}{listing17}{1}
  \codefor{code/for/listings.f}{listing17}{listing18}{1}
  \codefor{code/for/listings.f}{listing18}{listing19}{1}
  \codefor{code/for/listings.f}{listing19}{listing20}{1}
  \codefor{code/for/listings.f}{listing20}{listing21}{1}
  \codefor{code/for/listings.f}{listing21}{listing22}{1}

  \subsection*{{\bf An MPDATA solver} (cf. Sect. \ref{sec:mpdatasolver})}
  \codefor{code/for/listings.f}{listing22}{listing23}{1}

  \bibliographystyle{model1-num-names}
  \bibliography{paper}

\end{document}

### 2013-01-10 paper/doc/paper.tex
first typo corrections to the arxiv v1

@@
   Double precision floating-point format is used in all three implementations.
-  The codes begins with the following definitions:
+  The codes begin with the following definitions:
   \codecpp{code/cpp/listings.hpp}{listing01}{listing02}{1}
@@
     features among free and open source compilers, debuggers and other programming
-    aids in still not mature.
+    aids is still not mature.
 
@@
   GCC supports OpenMP with Fortran and C++, as well as the C++11 built-in support for threads.
-  The CPython and PyPy implementations of Python don't offer any
+  The CPython and PyPy implementations of Python do not offer any
     built-in solution for multithreading. 

### 2013-01-16 paper/for/listings.f
switching from pointers to allocatables where applicable, compacting arrvec by using associated() instead of the inited array (thanks to Arjen Markus for his suggestions to do it)

@@ -12,7 +12,7 @@ module arrvec_m
   implicit none
 
   type :: arr_t
-    real(real_t), pointer, contiguous :: a(:,:)
+    real(real_t), allocatable :: a(:,:)
   end type
 
   type :: arrptr_t
@@ -20,8 +20,8 @@ module arrvec_m
   end type
 
   type :: arrvec_t
-    class(arrptr_t), pointer :: at(:)
-    logical, pointer, contiguous :: inited(:)
+    class(arrptr_t), allocatable :: at(:)
+    integer :: length
     contains
     procedure :: ctor => arrvec_ctor
     procedure :: init => arrvec_init
@@ -33,26 +33,27 @@ module arrvec_m
   subroutine arrvec_ctor(this, n)
     class(arrvec_t) :: this
     integer, intent(in) :: n
+
+    this%length = n
     allocate(this%at( -n : n-1 ))
-    allocate(this%inited( 0 : n-1 ))
-    this%inited = .false.
   end subroutine
 
   subroutine arrvec_init(this, n, i, j)
     class(arrvec_t) :: this
     integer, intent(in) :: n
     integer, intent(in) :: i(2), j(2)
+
     allocate(this%at(n)%p)
     allocate(this%at(n)%p%a( i(1) : i(2), j(1) : j(2) ))
-    this%inited(n) = .true.
-    this%at(n - size(this%inited))%p => this%at(n)%p
+    this%at(n - this%length)%p => this%at(n)%p
   end subroutine
 
   subroutine arrvec_dtor(this)
     class(arrvec_t) :: this
+
     integer :: i
-    do i = 0, size(this%inited) - 1
-      if (this%inited(i)) then
+    do i = 0, this%length - 1
+      if (associated(this%at(i)%p)) then
         deallocate(this%at(i)%p%a)
         deallocate(this%at(i)%p)
       end if
@@ -128,7 +129,7 @@ module pi_m
   contains
   function pi(d, arr, i, j) result(return)
     integer, intent(in) :: d
-    real(real_t), intent(in), pointer :: arr(:,:)
+    real(real_t), allocatable, target :: arr(:,:)
     real(real_t), pointer :: return(:,:)
     integer, intent(in) :: i(2), j(2)
     select case (d) 
@@ -166,7 +167,7 @@ module bcd_m
     subroutine bcd_fill_halos(this, a, j)
       import :: bcd_t, real_t
       class(bcd_t ) :: this
[…]

### 2013-01-16 paper/doc/paper.tex

@@
 
-  All codes are strucutred in a way allowing compilation of the code 
+  All codes are structured in a way allowing compilation of the code 
     in exactly the same order as presented in the text within one source file,

### 2013-01-17 paper/doc/paper.tex
paper.tex: ancestor -> descendant (thanks Arjen)

@@
   The \prog{solver} structure is an abstract definition (containing a pure virtual method) 
-     requiring its ancestors to implement at least
+     requiring its descendants to implement at least
      the \prog{advop()} method which is expected to fill \prog{psi[n+1]} with the values of \prog{psi[n]} advected

### 2013-01-20 paper/tests/timing/filter.sh
adding cache plots

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+egrep "^time:| of all " Testing/Temporary/LastTest.log \
+  | awk -f filter1.awk | tr -s " " | tr -s "%" | tr "%" "0" \
+  | awk -f filter2.awk \
+  | awk -f filter3.awk 
+

### 2013-03-05 paper/pyt/numba/listings_noslice.py
adding test_pythononly.py - a simple way to test python version; adding listings_noslice.py - a python version without using a slice as an argument of class method (no shift class, ext changed)

@@ -1,0 +1,244 @@
+#listing00
+# code licensed under the terms of GNU GPL v3
+# copyright holder: University of Warsaw
+#listing01
+real_t = 'float64'
+#listing02
+try:
+  import numpypy
+  from _numpypy.pypy import set_invalidation
+  set_invalidation(False)
+except ImportError:
+  pass
+import numpy
+try: 
+  numpy.seterr(all='ignore')
+except AttributeError:
+  pass
+import pdb
+#listing03
+def sl_add(sl, int_num):
+  return slice(sl.start + int_num, sl.stop + int_num)
+
+#listing05
+def ext(r, lft, rgh):
+  return slice(
+    r.start - lft, 
+    r.stop + rgh
+  )
+#listing06
+def pi(d, *idx): 
+  return (idx[d], idx[d-1]) 
+#listing07
+class Solver(object):
+  # ctor-like method
+  def __init__(self, bcx, bcy, nx, ny, hlo):
+    self.n = 0
+    self.hlo = hlo
+    self.i = slice(hlo, nx + hlo)
+    self.j = slice(hlo, ny + hlo)
+
+    self.bcx = bcx(0, self.i.start, self.i.stop, hlo)
+    self.bcy = bcy(1, self.j.start, self.j.stop, hlo)
+
+    self.psi = (
+      numpy.empty((
+        ext(self.i, self.hlo, self.hlo).stop, 
+        ext(self.j, self.hlo, self.hlo).stop
+      ), real_t),
+      numpy.empty((
+        ext(self.i, self.hlo, self.hlo).stop, 
+        ext(self.j, self.hlo, self.hlo).stop
+      ), real_t)
+    )
+
+    self.C = (
+      numpy.empty((
+        ext(self.i, 1, 0).stop, 
+        ext(self.j, self.hlo, self.hlo).stop
+      ), real_t),
+      numpy.empty((
+        ext(self.i, self.hlo, self.hlo).stop, 
+        ext(self.j, 1, 0).stop
+      ), real_t)
+    )
+
+  # accessor methods
+  def state(self):
+    return self.psi[self.n][self.i, self.j]
+
+  # helper methods invoked by solve()
+  def courant(self,d):
+    return self.C[d][:]
+
+  def cycle(self):
+    self.n  = (self.n + 1) % 2 - 2
+
[…]

### 2013-03-05 paper/pyt/numba/listings_noslice.py
adding import numba and autojit for almost all functions, pi was changed (numba didn't like f(*arg)); autojit doesn't work frac(numba: Comparing objects is not yet supported)

@@ -15,20 +15,46 @@ try:
   numpy.seterr(all='ignore')
 except AttributeError:
   pass
+
+try:
+    from numba import jit, float64, int_, void, autojit
+    float64_1d=float64[:]
+    float64_2d=float64[:,:]
+    print "Numba or Numbapro imported"
+except ImportError:
+    def void(*args):
+        def dont_decorate(fn):
+            return fn
+        return dont_decorate
+    float64 = void
+    float64_1d = void
+    float64_2d = void
+    int_ = void
+    def jit(cl):
+        return cl
+    autojit = jit
+    print "Not using Numba nor Numbapro"
+
 import pdb
+
 #listing03
+@autojit
 def sl_add(sl, int_num):
   return slice(sl.start + int_num, sl.stop + int_num)
 
 #listing05
+@autojit
 def ext(r, lft, rgh):
   return slice(
     r.start - lft, 
     r.stop + rgh
   )
 #listing06
-def pi(d, *idx): 
+@autojit
+def pi(d, idx_0, idx_1):
+  idx = [idx_0, idx_1]
   return (idx[d], idx[d-1]) 
+
 #listing07
 class Solver(object):
   # ctor-like method
@@ -111,13 +137,16 @@ class Cyclic(object):
     )
 
 #listing09
+@autojit
 def f(psi_l, psi_r, C):
   #pdb.set_trace()
   return (
     (C + abs(C)) * psi_l + 
     (C - abs(C)) * psi_r
   ) / 2
+
 #listing10
+@autojit
 def donorcell(d, psi, C, i, j):
   #pdb.set_trace()
   return (
@@ -132,13 +161,16 @@ def donorcell(d, psi, C, i, j):
         C[pi(d, sl_add(i,-1), j)]
     ) 
   )
+
 #listing11
+@autojit
 def donorcell_op(psi, n, C, i, j):
   #pdb.set_trace()
   psi[n+1][i,j] = (psi[n][i,j] 
     - donorcell(0, psi[n], C[0], i, j)
     - donorcell(1, psi[n], C[1], j, i)
   )
+
 #listing12
 class Donorcell(Solver):
   def __init__(self, bcx, bcy, nx, ny):
@@ -149,17 +181,23 @@ class Donorcell(Solver):
[…]

### 2013-03-07 paper/for/listings.f
fortran: one more deallocate eliminated with use of allocatables

@@ -187,7 +187,7 @@ module solver_m
   implicit none
 
   type, abstract :: solver_t
-    class(arrvec_t), pointer :: psi, C
+    class(arrvec_t), allocatable :: psi, C
     integer :: n, hlo
     integer :: i(2), j(2) 
     class(bcd_t), pointer :: bcx, bcy
@@ -202,7 +202,7 @@ module solver_m
   abstract interface
     subroutine solver_advop(this)
       import solver_t
-      class(solver_t) :: this
+      class(solver_t), target :: this
     end subroutine
   end interface
 
@@ -251,7 +251,6 @@ module solver_m
     class(solver_t) :: this
     call this%psi%dtor()
     call this%C%dtor()
-    deallocate(this%psi, this%C)
   end subroutine
   
   function solver_state(this) result (return)
@@ -369,7 +368,8 @@ module donorcell_m
   end function
 !listing12
   subroutine donorcell_op(psi, n, C, i, j)  
-    class(arrvec_t), pointer :: psi, C
+    class(arrvec_t), allocatable :: psi
+    class(arrvec_t), pointer :: C
     integer, intent(in) :: n
     integer, intent(in) :: i(2), j(2) 
     
@@ -404,9 +404,11 @@ module solver_donorcell_m
   end subroutine
 
   subroutine donorcell_advop(this)
-    class(donorcell_t) :: this
+    class(donorcell_t), target :: this
+    class(arrvec_t), pointer :: C
+    C => this%C
     call donorcell_op(                                 &
-      this%psi, this%n, this%C, this%i, this%j         &
+      this%psi, this%n, C, this%i, this%j         &
     )
   end subroutine
 
@@ -539,7 +541,7 @@ module solver_mpdata_m
   end subroutine
 
   subroutine mpdata_advop(this)
-    class(mpdata_t) :: this
+    class(mpdata_t), target :: this
     integer :: step
 
     associate (i => this%i, j => this%j, im => this%im,&
@@ -548,7 +550,11 @@ module solver_mpdata_m
     )
       do step=0, this%n_iters-1
         if (step == 0) then
-          call donorcell_op(psi, n, this%C, i, j)
+          block
+            class(arrvec_t), pointer :: C
+            C => this%C
+            call donorcell_op(psi, n, C, i, j)
+          end block
         else
           call this%cycle()
           call bcx%fill_halos(                    &

### 2013-03-07 paper/tests/timing/inout.sh
fixing shebang in inout.sh

@@ -1,4 +1,4 @@
-#/bin/bash
+#!/bin/bash
 Cx=.2
 Cy=.5
 it=3

### 2013-03-07 paper/for/listings.f
eliminating destructors through usage of allocatables

@@ -20,12 +20,12 @@ module arrvec_m
   end type
 
   type :: arrvec_t
+    class(arr_t), allocatable :: arrs(:)
     class(arrptr_t), allocatable :: at(:)
     integer :: length
     contains
     procedure :: ctor => arrvec_ctor
     procedure :: init => arrvec_init
-    procedure :: dtor => arrvec_dtor
   end type
 
   contains
@@ -36,29 +36,17 @@ module arrvec_m
 
     this%length = n
     allocate(this%at( -n : n-1 ))
+    allocate(this%arrs( 0 : n-1 ))
   end subroutine
 
   subroutine arrvec_init(this, n, i, j)
-    class(arrvec_t) :: this
+    class(arrvec_t), target :: this
     integer, intent(in) :: n
     integer, intent(in) :: i(2), j(2)
 
-    allocate(this%at(n)%p)
-    allocate(this%at(n)%p%a( i(1) : i(2), j(1) : j(2) ))
-    this%at(n - this%length)%p => this%at(n)%p
-  end subroutine
-
-  subroutine arrvec_dtor(this)
-    class(arrvec_t) :: this
-
-    integer :: i
-    do i = 0, this%length - 1
-      if (associated(this%at(i)%p)) then
-        deallocate(this%at(i)%p%a)
-        deallocate(this%at(i)%p)
-      end if
-    end do
-    deallocate(this%at)
+    allocate(this%arrs(n)%a( i(1) : i(2), j(1) : j(2) ))
+    this%at(n)%p => this%arrs(n)
+    this%at(n - this%length)%p => this%arrs(n)
   end subroutine
 end module
 !listing03
@@ -247,12 +235,6 @@ module solver_m
     )
   end subroutine
 
-  subroutine solver_dtor(this)
-    class(solver_t) :: this
-    call this%psi%dtor()
-    call this%C%dtor()
-  end subroutine
-  
   function solver_state(this) result (return)
     class(solver_t) :: this
     real(real_t), pointer :: return(:,:)
@@ -390,7 +372,6 @@ module solver_donorcell_m
   type, extends(solver_t) :: donorcell_t
     contains
     procedure :: ctor => donorcell_ctor
-    procedure :: dtor => donorcell_dtor
     procedure :: advop => donorcell_advop
   end type
 
@@ -410,11 +391,6 @@ module solver_donorcell_m
     call donorcell_op(                                 &
[…]

### 2013-03-07 paper/for/test_arrvec.f

@@ -1,12 +1,11 @@
 program test_arrvec
   use arrvec_m
-  class(arrvec_t), pointer :: psi
-  integer, pointer, contiguous :: i(:), j(:)
-  integer :: c, nx = 10
+  class(arrvec_t), allocatable :: psi
+  integer, dimension(2) :: i, j
+  integer :: c, nx = 10, ny = 10
 
-  allocate(i(nx), j(nx))
-  i = (/ (c, c=0, nx-1) /)
-  j = (/ (c, c=-2, nx-3) /)
+  i = (/ 0, nx - 1 /)
+  j = (/ 0, ny - 1 /)
 
   allocate(psi)
   call psi%ctor(2)
@@ -18,9 +17,4 @@ program test_arrvec
   psi%at(0)%p%a(1,1) = 10
   print*, psi%at(0)%p%a(1,1)
   print*, psi%at(-2)%p%a(1,1)
-
-  call psi%dtor
-  deallocate(psi)
-  deallocate(i, j)
 end
-

### 2013-03-07 paper/for/test.f
eliminating destructors through usage of allocatables

@@ -62,8 +62,6 @@ program test
         stop 1
       end if
     end block
-
-    call slv%dtor()
   end block
 
   contains

### 2013-03-07 paper/doc/paper.tex
some paper.tex changes

@@
   Using an additional array dimension to represent the sequence elements is not considered for two reasons.
-  First, the arrays constituting the sequence elements may have different sizes
-      (e.g.  $C^{[x]}$ and $C^{[y]}$ on the Arakawa-C grid discussed in section~\ref{sec:grid}).
+  First, the $C^{[x]}$ and $C^{[y]}$ arrays constituting the sequence have different sizes
+    (see discussion of the Arakawa-C grid in section~\ref{sec:grid}).
   Second, the order of dimensions would need to be different for different languages to assure that
@@
   \end{figure}
-  The so-called Arakawa-C staggered grid \citep{Arakawa_and_Lamb_1977} depicted in Figure~\ref{fig:grid}.
+  The so-called Arakawa-C staggered grid \citep{Arakawa_and_Lamb_1977} depicted in Figure~\ref{fig:grid}
     is a natural choice for MPDATA.
-  MPDATA is a finite-difference approximation to the flux-form of the equation~(\ref{eq:adv}) 
-    resulting in a finite-volume scheme.
-  Using the Arakawa-C grid allows to evaluate the fluxes at 
-    the grid cell (finite volume) edges without interpolation.
-
+%  MPDATA is a finite-difference approximation to the flux-form of the equation~(\ref{eq:adv}) 
+%    resulting in a finite-volume scheme.
+%  Using the Arakawa-C grid allows to evaluate the fluxes at 
+%    the grid cell (finite volume) edges without interpolation.
   As a consequence, the discretised representations of the $\psi$ scalar field, and each 
@@
   Python forbids overloading of operators of built-in types such as \prog{slices}, 
-    and does not define addition/subtraction operators for \prog{slice}/\prog{int} pairs.
+    and does not define addition/subtraction operators for \prog{slice} and \prog{int} pairs.
   Consequently, a custom logic has to be defined not only for fractional indexing,
@@
 
-  In Fortran, similarly as in C++ and Python, the logic related to employment of the Arakawa-C grid
-    allowing employment of fractional indices in array indexing is obtained through
+  In Fortran fractional array indexing is obtained through
     definition and instantiation of an object representing the half, and having appropriate
@@
   The $\pi$ permutation is implemented in C++ as a function \prog{pi()}
-    returning an instance of \prog{idx\_t} constructed with a permutation
-    of the \prog{i} and \prog{j} arguments of \prog{pi()}, with permutation order
-    dependant on the template parameter \prog{d}:
+    returning an instance of \prog{idx\_t}.
+  In order to ensure compile-time evaluation, the permutation order is passed 
+    via the template parameter \prog{d}:
   \codecpp{code/cpp/listings.hpp}{listing07}{listing08}{1}
+  (note the different order of \prog{i} and \prog{j} arguments in the two template specialisations).
 
@@
 
-  In the Fortran implementation an analogous behaviour is obtained by introducing
-    a function \prog{pi()} returning a pointer the array elements specified
+  In the Fortran implementation \prog{pi()} returns a pointer to the array elements specified
     by \prog{i} and \prog{j} interpreted as (i,j) or
@@
 
-  The authors do expect some performance gain could readily
-    be obtained by introducing into the codes what is referred to as
-    {\em premature optimisation} -- code rearrangements aimed solely
-    at the purpose of increasing performance.
+  The authors do expect some performance gain could 
+    be obtained by introducing into the codes some ''manual'' optimisations -- 
+    code rearrangements aimed solely at the purpose of increasing performance.
   These were avoided intentionally as they degrade code readability,
@@
     especially if multiple architectures and operating systems are targeted.
-  However, CMake is a tool that allows to efficiently automate builds, tests and 
-    packaging of C++ and Fortran programs.
+  However, CMake\footnote{CMake is a family of open-source, cross-platform
+    tools automating building, testing and packaging of C/C++/Fortran software,
+    see http://cmake.org/ for details}
+    is a tool (or arguably the tool) that allows to efficiently automate 
+    builds, tests and packaging of C++ and Fortran programs.
 
@@
     \end{itemize}
-  The three equally-structured implementations required ca. 200, 300, and 550 lines of code
+  The three equally-structured implementations required ca. 200, 300, and 550 lines of code %TODO update line numbers
     in Python, C++ and Fortran, respectively.

### 2013-03-08 paper/doc/paper.tex
suggestions of changes in the part 4 and 5

@@
 
-  One of the aims of this paper is to exemplify the applicability of OOP features of the three considered
-    programming languages (or language-library pairs) for representing
-    what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012} within the source code
-    of scientific software.
+  One of the aim of this paper is to exemplify OOP features of the three
+    programming languages (or language-library pairs)  trying to represent
+    what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012} within the
+     scientific software.
   Presented benchmark tests, although quite simplistic, together with the experience gained 
-    from the development of equally-structured codes in three different
-    languages provide a basis for discussion on the tradeoffs of the programming language.
+    from the development of equally structured codes in three different
+    languages provide a basis for discussion of the tradeoffs of the programming language.
   The discussion concerns in principle the development of finite-difference solvers for 
-    partial differential equations, but is likely applicable to same extent to
-    development of scientific software in general.
+    partial differential equations, but is likely applicable the scientific software in general.
   A partly objective and partly subjective summary of pros and cons of C++, Python and Fortran,
-    as observed during the development of the presented codes, in presented in the four following subsections.
+    as observed during the development of the presented codes, is presented in the four following subsections.
 
@@
   \begin{itemize}
-    \item{loop-free array arithmetics}
-    \item{definitions of functions returning array-valued expressions}
+    \item{loop-free array arithmetics;}
+    \item{definitions of functions returning array-valued expressions;}
     \item{permutations of array indices allowing dimension-independent definitions
-      of functions (see e.g. listings C.12 and C.13, P.10 and P.11, F11 and F.12)}
-    \item{fractional indexing of arrays corresponding to employment of a staggered grid}
+      of functions (see e.g. listings C.12 and C.13, P.10 and P.11, F11 and F.12);}
+    \item{fractional indexing of arrays corresponding to employment of a staggered grid;}
     \item{associating identifiers with slabs of arrays (e.g. for representing 
-      a psi[left\_halo]=psi[rght\_edge] assignment)}
+      a psi[left\_halo]=psi[rght\_edge] assignment).} (DOROTA: w zasadzie tylko przyklad rozumiem)
   \end{itemize}
-  Three issues specific to Fortran were observed which 
-    resulted in employment in some places of a more repetitive or cumbersome syntax than in C++ or Python:
+  Three issues specific to Fortran that 
+    resulted in employment of a more repetitive or cumbersome syntax than in C++ or Python
+    were observed:
   \begin{itemize}
@@
       some inherent limitations such as lack of support for arrays of arrays
+      (DOROTA: nie podoba mi sie to 'but also', moze wystarczy tylko pierwsza czesc punktu?)
       (cf. Sect. \ref{sec:sequence})}
@@
     or ease of use as the first requirement \citep[see e.g.][]{Barnes_and_Jones_2011}.
-  PyPy's capability to improve performance of unmodified Python code may actually 
-    make Python a favourable choice even if high performance is crucial, especially
+  PyPy's capability to improve performance of unmodified Python code may 
+    make Python a favourable choice even if high performance is important, especially
     if a performance-to-development cost measure is to be considered.
@@
 
-  Python simplifies also the building and dissemination of software,
+(DOROTA: zastanawiam sie, czy tego nie wyrzucic)
+  Python simplifies also building and shearing of software,
     especially if multiple architectures and operating systems are targeted.
@@
   There exist great debugging tools for C++, but the debugging and development is 
-    often hindered by indecipherable compiler messages flooded with lengthy type names
-    stemming from employment of templates.
+    often hindered by indecipherable compiler messages.
   OOP with Fortran is a relatively new concept, and the support for the Fortran OOP
     features among free and open source compilers, debuggers and other programming
-    aids is still not mature.
+    aids are still not mature.
 
@@
   With both Fortran and Python, the memory footprint caused by employment
-    of temporary objects in array arithmetics is dependant on compiler optimisations.
+    of temporary objects in array arithmetics is dependent on compiler optimisations.
   In contrast, Blitz++ ensures temporary-array-free computations by design.
@@
     influence the availability of trained personnel, 
-    the availability of reusable software components,
-    the maturity of compilers and tools, 
-    and the volume and coverage of online and printed information resources.
+   reusable software components and information resources.
+    It also affect the maturity and quality of compilers and tools. 
+  Fortran is a domain-specific language while Python and C++ are general-purpose languages
+    with disproportionately larger users' communities.
   The OOP features were introduced to Fortran only recently, and have not gained
@@
     example being the incomplete support for syntax-highlighting of modern Fortran in Vim and Emacs editors}.
-  Fortran is no longer routinely taught at the universities \citep{Kendall_et_al_2008},
+  Fortran is almost no longer taught at the universities \citep{Kendall_et_al_2008},
     in contrast to C++ and Python.
-  An example confirming the arguably ceasing popularity of Fortran in academia 
-    is the discontinuation of all but C++ printed editions of the ''Numerical Recipes'' 
+  An example ofdecreasing popularity of Fortran in academia 
+    is the discontinuation of Fortran printed editions of the ''Numerical Recipes'' 
     series of Press et al.
-  Fortran is a domain-specific language while Python and C++ are general-purpose languages
-    with disproportionately larger users' communities.
-
-  Yet still, in the case of C++ and Python it is also the size of the particular
+
+  In the case of C++ and Python it is also the size of the particular
     numerical-library users' community that is of importance.
@@
     (the gnuplot-iostream used in listing C.24 being a much-appreciated counterexample).
+ ( DOROTA: nie rozumiem nastepnego zdania)
   However, Blitz++ allows to instantiate its arrays using preexisting data what
@@
 
+(DOROTA: a moze skrocic to i przeniesc do 4.1?)
   The availability and quality of object-oriented 
@@
 
+(DOROTA: nastepny akapit zaznaczylabym Fijalowi do uwaznego przejrzenia, 
+Ja nie rozumime tej czesci o numexpr itp)
   Finally, the three languages differ as well with regard to availability of 
@@
     for the advection equation were introduced.
-  The solvers are based on MPDATA - an algorithm of particular applicability
-    in geophysical fluid dynamics \citep{Smolarkiewicz_2006}.
-  The three implementations follow the same structure but are implemented
-    in three different languages (or language/library pairs):
+  The solvers are based on the MPDATA algorithm \citep{Smolarkiewicz_2006}.
+  All implementations follow the same structure but are implemented
+    in three different language and library pairs:
   \begin{itemize}
-    \item{C++ with Blitz++,}
-    \item{Python with NumPy and}
-    \item{Fortran}
+    \item{C++ with Blitz++;}
+    \item{Python with NumPy;}
+    \item{Fortran.}
   \end{itemize}
-  It was the aim of this study to assess the capabilities of these languages
-    for object-oriented scientific programming in context of 
-    development of finite-difference partial differential equation solvers.
-  The emphasis was on the capabilities for reproduction within 
-    the program code of the mathematical abstractions used for defining the algorithms
-    in journal papers or lecture blackboards.
+(DOROTA: przelozylam to i delikatnie zmienilam, ale nadal nie rozuemi czesci 'employed tools...')
+  The fact that all considered standards are open and the employed
+    tools implementing them are free and open-source
+    is certainly a great advantage \citep{Anel_2011}.
+
+  Finite-difference partial differential equation solvers were example used to study 
+    the capabilities of these languages
+    for object-oriented scientific programming.
+  The emphasis was on the capabilities for reproduction 
+    of the mathematical abstractions used for explaining the algorithms
+    in journal papers within the program code.
 
@@
     the NumPy support in the PyPy implementation of Python.
-  The key conclusion is that the considered language/library/compiler
-    set-ups offer unprecedented possibilities for using OOP to compactly 
-    and faithfully represent within the program source code 
-    the mathematical abstractions employed when describing 
-    numerical algorithms in literature.
+  The key conclusion is that all considered language/library/compiler
+    set-ups offer possibilities for using OOP to compactly 
+     represent the mathematical abstractions within the program  code. 
+(OROTA: dwa nastepne zdania bym przeformulowala, ale chce sie upewnic, co wlasciwie chcemy tu powiedziec)
   The readability and brevity of such code may contribute to its 
@@
     \citep{Freeman_et_al_2010}.
-  The fact that all considered standards are open and the employed 
-    tools implementing them are free and open-source
-    is of importance as well \citep{Anel_2011}.
 
@@
 
+(DOROTA: to trzeba zmienic, nie wiem co dokladnie chciales powiedziec.)
   The OOP design enhances the possibilities to reuse and extend the presented code.

### 2013-03-12 paper/doc/paper.tex
new plots, mention of clang, usage example listing for Python and Fortran

--- 2013-03-07 paper/doc/paper.tex
@@
     \center
-    \includegraphics[height=.45\textwidth,angle=-90]{plot-mem-host=eyrie-cpu=GenuineIntel-6-15-bogomips=3190}
+    \includegraphics[height=.45\textwidth,angle=-90]{plot-mem-host=laptok-cpu=GenuineIntel-6-42-bogomips=3190}
     \caption{\label{fig:mem}
@@
     \center
-    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=eyrie-cpu=GenuineIntel-6-15-bogomips=3190}
+    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=laptok-cpu=GenuineIntel-6-42-bogomips=3190}
     \caption{\label{fig:cpu-eyrie}
@@
       Values of the total user mode cpu time reported by the GNU time utility are
-        and normalised by the grid size ($nx=ny$) and the number of timesteps $nt=2^{24}/(nx \cdot ny)$.
+        and normalised by the grid size ($nx \cdot ny$) and the number of timesteps $nt=2^{24}/(nx \cdot ny)$.
       Before normalisation the time reported for an $nt=0$ run for a corresponding
@@
       Results obtained with an Intel\textsuperscript{\textregistered} 
-        Core\textsuperscript{\texttrademark} 2 Duo 1.6 GHz processor.
+        Core\textsuperscript{\texttrademark} i5 1.6 GHz processor.
     }
@@
       \begin{itemize}
-        \item{GCC g++ 4.8.0\footnote{\label{fnt:gcc-snapshot}GNU Compiler Collection trunk revision 193662 packaged in the Debian's gcc-snapshot\_20121120-1}
-          with Blitz++ 0.10}
+        \item{GCC g++ 4.8.0\footnote{\label{fnt:gcc-snapshot}GNU Compiler Collection packaged in the Debian's gcc-snapshot\_20130222-1}
+          and Blitz++ 0.10}
+        \item{LLVM Clang 3.2 and Blitz 0.10}
       \end{itemize}
@@
       \begin{itemize}
-        \item{CPython 2.7.3 with NumPy 1.6.2}
+        \item{CPython 2.7.3 and NumPy 1.7}
         \item{PyPy 1.9.0 with built-in NumPy implementation}
@@
   \end{description}
-  All performance tests were run on a Debian GNU/Linux system with the above-listed software obtained 
-    via binary packages from the Debian's ''sid'' collection (most recent package versions at the time of writing).
+  The performance tests were run on a Debian and an Ubuntu GNU/Linux systems with the above-listed software obtained 
+    via binary packages from the distributions's package repositories (most recent package versions at the time of writing).
   The tests were performed on two 64-bit machines equipped with
-    an AMD Phenom\textsuperscript{\texttrademark} II X6 1055T (800 MHz, ca. 5600 BogoMIPS)
-    and an Intel\textsuperscript{\textregistered} Core\textsuperscript{\texttrademark}~2~Duo P7500 (1.6 GHz, ca. 3200 BogoMIPS)
+    an AMD Phenom\textsuperscript{\texttrademark} II X6 1055T (800 MHz)
+    and an Intel\textsuperscript{\textregistered} Core\textsuperscript{\texttrademark}~i5-2467M (1.6 GHz)
     processors.
 
-  For both C++ and Fortran the GCC compilers were invoked with the \prog{-Ofast} and \prog{-march=native} 
-    options imposing, among other optimisations, automatic vectorisation and finite-only arithmetics 
-    (i.e.  non-standard-compliant treatment of NaN and Inf values).
-  GCC compilation did result in vectorised code for C++ only (vectorised loops within the Blitz++ headers, 
-    as reported by the messages triggered by the \prog{-ftree-vectorizer-verbose} option).
-%  At least a part of possible auto-vectorisations was not done with the Fortran implementation 
-%    due to employment of non-contiguous pointers (in particular when referring to the grid elements 
-%    excluding the halo region).
+  For both C++ and Fortran the GCC compilers were invoked with the \prog{-Ofast} and the 
+    \prog{-march=native} options.
+  The Clang compiler was invoked with the \prog{-O3}, the \prog{-mllvm -vectorize}, the \prog{-ffast-math} 
+    and the \prog{-march=native} options.
 
@@
     PyPy allows to eliminate the use of temporary matrices for storing intermediate results,
-    and to perform multiple operations on the arrays within a single array index traversal.
+    and to perform multiple operations on the arrays within a single array index traversal
+    \footnote{Lazy evaluation available in PyPy 1.9 has been temporarily removed from PyPy during a refactoring of
+      the code. It'll be reinstantiated in the codebase as soon as possible, but past PyPy 2.0 release}.
   Consequently, PyPy allows to overcome the same performance-limiting factors as those addressed by Blitz++, although 
@@
   The CPU time statistics presented in Figures \ref{fig:cpu-eyrie} and \ref{fig:cpu-skua} reveal
-    notable differences between results obtained with the two different processors,
+    some differences between results obtained with the two different processors,
     however the results agree qualitatively leading to the following observations
@@
     \item{CPython requires from around 4 to almost 10 times more CPU time than Fortran depending on the grid size;}
-    \item{PyPy execution times are in most cases closer to C++ than CPython.}
+    \item{PyPy execution times are in most cases closer to C++ than to CPython.}
   \end{itemize}
-  The support for OOP features in gfortran, the NumPy support in PyPy, and the auto-vectorisation
+  The support for OOP features in gfortran, the NumPy support in PyPy, and the relevant optimisation
     mechanisms in GCC are still in active development and hence the performance with some of the set-ups may 
@@
     \noindent
-    Authors would like to thank Tobias Burnus, Julian Cummings, Patrik Jonsson and
-      Arjen Markus for their feedback on questions posted to Blitz++ and gfortran
+    Authors would like to thank Tobias Burnus, Julian Cummings, Patrik Jonsson,
+      Arjen Markus and Janus Weil for their feedback on questions posted to Blitz++ and gfortran
       mailing lists.
@@
   \codepyt{code/pyt/listings.py}{listing18}{listing19}{1}
+
+  \subsection*{{\bf Usage example} (cf. Sect. \ref{sec:example})}
+  \ldots
+  \codepyt{code/pyt/test.py}{listing19}{listing20}{1}
+  \ldots
 
@@
 
+  \subsection*{{\bf Usage example} (cf. Sect. \ref{sec:example})}
+  \ldots
+  \codefor{code/for/test.f}{listing23}{listing24}{1}
+  \ldots
+  \codefor{code/for/test.f}{listing25}{listing26}{1}
+  \ldots
+
   \bibliographystyle{model1-num-names}

### 2013-03-12 paper/for/listings.f

@@ -200,7 +200,7 @@ module solver_m
     use arakawa_c_m
     use halo_m
     class(solver_t) :: this
-    class(bcd_t), intent(in), pointer :: bcx, bcy
+    class(bcd_t), intent(in), target :: bcx, bcy
     integer, intent(in) :: nx, ny, hlo
 
     this%n = 0
@@ -379,7 +379,7 @@ module solver_donorcell_m
   
   subroutine donorcell_ctor(this, bcx, bcy, nx, ny)
     class(donorcell_t) :: this
-    class(bcd_t), intent(in), pointer :: bcx, bcy
+    class(bcd_t), intent(in), target :: bcx, bcy
     integer, intent(in) :: nx, ny
     call solver_ctor(this, bcx,bcy, nx,ny, 1)
   end subroutine
@@ -484,7 +484,7 @@ module solver_mpdata_m
 
   subroutine mpdata_ctor(this, n_iters, bcx, bcy, nx, ny)
     class(mpdata_t) :: this
-    class(bcd_t), pointer :: bcx, bcy
+    class(bcd_t), target :: bcx, bcy
     integer, intent(in) :: n_iters, nx, ny
     integer :: c
 

### 2013-03-14 paper/for/listings.f
further paper polishing (incl. code indenting etc)

@@ -447,7 +447,7 @@ module mpdata_m
     ) / 4               
   end function
 !listing20
-  function antidiff(d, psi, i, j, C) result (return)
+  function C_adf(d, psi, i, j, C) result (return)
     integer :: d
     integer, intent(in) :: i(2), j(2)
     real(real_t) :: return(span(d, i, j), span(d, j, i))
@@ -512,7 +512,7 @@ module solver_mpdata_m
 
     associate (i => this%i, j => this%j, im => this%im,&
       jm => this%jm, psi => this%psi, n => this%n,     &
-      hlo => this%hlo, bcx => this%bcx, bcy => this%bcy &
+      hlo => this%hlo, bcx => this%bcx, bcy => this%bcy&
     )
       do step=0, this%n_iters-1
         if (step == 0) then
@@ -523,11 +523,11 @@ module solver_mpdata_m
           end block
         else
           call this%cycle()
-          call bcx%fill_halos(                    &
-            psi%at( n )%p%a, ext(j, hlo)     &
+          call bcx%fill_halos(                         &
+            psi%at( n )%p%a, ext(j, hlo)               &
           )
-          call bcy%fill_halos(                    &
-            psi%at( n )%p%a, ext(i, hlo)     &
+          call bcy%fill_halos(                         &
+            psi%at( n )%p%a, ext(i, hlo)               &
           )
 
           block
@@ -548,18 +548,18 @@ module solver_mpdata_m
 
             ! calculating the antidiffusive velo
             ptr => pi(0, C_corr%at( 0 )%p%a, im+h, j)
-            ptr = antidiff(                            &
+            ptr = C_adf(                               &
               0, psi%at( n )%p%a, im, j, C_unco        &
             )      
-            call bcy%fill_halos(                  &
+            call bcy%fill_halos(                       &
               C_corr%at(0)%p%a, ext(i, h)              &
             )
 
             ptr => pi(0, C_corr%at( 1 )%p%a, i, jm+h)
-            ptr = antidiff(                            &
+            ptr = C_adf(                               &
               1, psi%at( n )%p%a, jm, i, C_unco        &
             )
-            call bcx%fill_halos(                  &
+            call bcx%fill_halos(                       &
               C_corr%at(1)%p%a, ext(j, h)              &
             )
 

### 2013-03-15 paper/pyt/numba/listings_noslice.py
Numba: work in progress

@@ -55,17 +55,44 @@ def pi(d, idx_0, idx_1):
   idx = [idx_0, idx_1]
   return (idx[d], idx[d-1]) 
 
+#listing08                                                                                           
+@jit
+class Cyclic(object):
+  @void(int_, int_, int_, int_)
+  def __init__(self, d, i_start, i_stop, hlo):
+    i  = slice(i_start, i_stop)
+    self.d = d
+    self.left_halo = slice(i.start-hlo, i.start    )
+    self.rght_edge = slice(i.stop -hlo, i.stop     )
+    self.rght_halo = slice(i.stop,      i.stop +hlo)
+    self.left_edge = slice(i.start,     i.start+hlo)
+
+  # method invoked by the solver                                                                     
+  @void(float64_2d, int_, int_)
+  def fill_halos(self, psi, j_start, j_stop):
+    j = slice(j_start, j_stop)
+    psi[pi(self.d, self.left_halo, j)] = (
+      psi[pi(self.d, self.rght_edge, j)]
+    )
+    psi[pi(self.d, self.rght_halo, j)] = (
+      psi[pi(self.d, self.left_edge, j)]
+    )
+
+
+
 #listing07
+@jit
 class Solver(object):
   # ctor-like method
-  def __init__(self, bcx, bcy, nx, ny, hlo):
+  @void(int_, int_, int_)
+  def __init__(self, nx, ny, hlo):
     self.n = 0
     self.hlo = hlo
     self.i = slice(hlo, nx + hlo)
     self.j = slice(hlo, ny + hlo)
 
-    self.bcx = bcx(0, self.i.start, self.i.stop, hlo)
-    self.bcy = bcy(1, self.j.start, self.j.stop, hlo)
+    self.bcx = Cyclic(0, self.i.start, self.i.stop, hlo)
+    self.bcy = Cyclic(1, self.j.start, self.j.stop, hlo)
 
     self.psi = (
       numpy.empty((
@@ -90,17 +117,21 @@ class Solver(object):
     )
 
   # accessor methods
+  @float64_2d()
   def state(self):
     return self.psi[self.n][self.i, self.j]
 
   # helper methods invoked by solve()
+  @float64_2d(int_)
   def courant(self,d):
     return self.C[d][:]
 
+  @void()
   def cycle(self):
     self.n  = (self.n + 1) % 2 - 2
 
    # integration logic
+  @void(int_)
[…]

### 2013-03-15 paper/doc/paper.tex

@@
 
-%\journal{Computer Physics Communications}
-\journal{arXiv. Paper presented at the American Meteorological Society 2013 Annual Meeting in Austin, Texas, USA}
+\journal{Computer Physics Communications}
+%\journal{arXiv. Paper presented at the American Meteorological Society 2013 Annual Meeting in Austin, Texas, USA}
 
@@
           used in the literature within the program code.
-        The introduced codes serve as a basis for discussion on the tradeoffs of the programming language choice.
+        The codes serve also as a basis for discussion on the tradeoffs of the programming language choice.
         The main angles of comparison are code brevity and syntax clarity
@@
     number of dimensions assuring positive-definiteness of scalar field $\psi$ 
-    and incurring small numerical diffusion
-    \citep[see][for a recent review of MPDATA-based techniques]{Smolarkiewicz_2006}.
+    and incurring small numerical diffusion.
+  All relevant MPDATA formul\ae are given in the text but are presented without
+    derivation or detailed discussion.
+  For a recent review of MPDATA-based techniques see \citet[][and references therein]{Smolarkiewicz_2006}.
 
@@
 
-  Solution of equation (\ref{eq:adv}) using MPDATA implies discretisation of the $\psi$ and 
-    the $\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}$ (Courant number) fields onto a grid - an array of points.
+  Solution of equation (\ref{eq:adv}) using MPDATA implies discretisation onto a grid of the $\psi$ and 
+    the Courant number $\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}$ fields, where $\Delta t$ is the 
+    solver timestep and $\Delta x$ is the grid spacing.
   Among the three considered languages only Fortran is equipped with built-in
@@
 
-  Presented C++ implementation of MPDATA is built upon the Blitz++ library. %TODO: cite something
+  Presented C++ implementation of MPDATA is built upon the Blitz++ 
+    library\footnote{Blitz++ is a C++ class library for scientific computing which uses the expression templates technique to achieve high performance, see http://sf.net/projects/blitz/}.
   Blitz offers object-oriented representation of n-dimensional arrays,
@@
   In particular, it offers loop-free notation for array arithmetics
-    that does not incur creation of intermediate temporary objects, 
-    a crucial feature for achieving high-performance in array manipulation
-    \citep[][]{Veldhuizen_et_al_1997}.
-  Blitz++ is a header-only library\footnote{unless debugging mode is used} -- to use it, it is enough to include the appropriate header file,
+    that does not incur creation of intermediate temporary objects.
+  Blitz++ is a header-only library\footnote{Blitz++ requires linking with \prog{libblitz} if debugging mode is used} -- to use it, it is enough to include the appropriate header file,
     and optionally expose the required classes to the present namespace:
   \codecpp{code/cpp/listings.hpp}{listing02}{listing03}{1}
-  Here, \prog{arr\_t}, \prog{rng\_t} and \prog{idx\_t} will serve as alias identifiers for the
-    2D double precision version of the Blitz++ \prog{Array} 
-    container, for the \prog{Range} class representing a sequence of integers numbering 
-    array elements, and for the 2D version of the \prog{RectDomain} class representing 
-    a vector of \prog{Range} objects defining a 2D slab of an array, respectively.
-
-  Arguably, the power of Blitz++ comes from the possibility to express array expressions as objects.
+  Here \prog{arr\_t}, \prog{rng\_t} and \prog{idx\_t} serve as alias identifiers 
+    and are introduced in order to shorten the code.
+
+  Arguably, the power of Blitz++ comes from the ability to express array expressions as objects.
   In particular, it is possible to define a function that returns an array expression;
@@
     to be performed on the arguments.
-  As a consequence, the return types of such functions become unintelligible, and the \prog{auto} 
-    return type declaration from the C++11 standard allows to simplify the code significantly,
+  As a consequence, the return types of such functions become unintelligible.
+  Luckily, the \prog{auto} return type declaration from the C++11 standard allows to simplify the code significantly,
     even more if used through the following preprocessor macro:
   \codecpp{code/cpp/listings.hpp}{listing03}{listing04}{1}
-    in which a call to \prog{blitz::safeToReturn()} function is also included ensuring
+  The call to \prog{blitz::safeToReturn()} function is included in order to ensure that
     all arrays involved in the expression being returned continue to exist in the
@@
     and the lazy evaluation mode is turned on through the \prog{set\_invalidation(False)} call.
-  PyPy's lazy evaluation obtained with the help of a just-in-time compiler enables it to achieve
+  PyPy's lazy evaluation obtained with the help of a just-in-time compiler enables to achieve
     an analogous to Blitz++ temporary-array-free handling of array-valued expressions 
     (see discussion in section~\ref{sec:perf}).
-  Second, the NumPy package is instructed to ignore any floating-point errors, if such an option
-    is available in the interpreter (it is not supported in PyPy as of version 1.9).
+  Second, to match the settings of C++ and Fortran compilers used herein, the NumPy package is instructed 
+    to ignore any floating-point errors (only if such an option
+    is available in the interpreter\footnote{\prog{numpy.seterr()} is not supported in PyPy as of version 1.9}).
   The above lines conclude all code modifications that needed to be added in order to run
@@
     hence there is no need for using an external package as with C++ and Python.
-  Fortran array-handling features are not object-oriented, however.
-  In particular, it is not possible to alter the default behaviour through
-    inheritance.
-
-  \subsection{Array container sequences}\label{sec:sequence}
+  Fortran array-handling features are not object-oriented, though.
+
+  \subsection{Array sequence containers}\label{sec:sequence}
 
   While discretisation in space of the scalar field $\psi(x,y)$ into its $\psi_{[i,j]}$ 
-    grid representation requires floating-point array containers (cf. section \ref{sec:array}), 
-    discretisation in time requires a container class for storing
-    sequences of such arrays, i.e. \{$\psi_{[i,j]}^{[n]}$, $\psi_{[i,j]}^{[n+1]}$\}.
-  Similarly the vector field $\vec{C}$ is in fact a \{$C^{[x]}$, $C^{[y]}$\} sequence.
+    grid representation requires floating-point array containers (cf. section \ref{sec:array}).
+  Discretisation in time requires a container class for storing
+    sequences of such arrays, i.e. \{$\psi^{[n]}$, $\psi^{[n+1]}$\}.
+  Similarly the components of the vector field $\vec{C}$ are in fact a \{$C^{[x]}$, $C^{[y]}$\} 
+    array sequence.
 
@@
   \codefor{code/for/listings.f}{listing02}{listing03}{1}
-  The \prog{arr\_t} and \prog{arrptr\_t} helper types are defined solely for the purpose 
+  The \prog{arr\_t} type is defined solely for the purpose 
     of overcoming the Fortran's limitation of not allowing to declare arrays of arrays.
+  The \prog{arrptr\_t} type is defined as creating an allocatable of its instances instead
+    of multi-element pointer of \prog{arr\_t} ensures automatic memory deallocation.
   The \prog{ctor()} method initialises the container for a given number of elements \prog{n}.
@@
     an newly allocated 2D array spanning indices \prog{i(1)}:\prog{i(2)}, 
-    and \prog{j(1)}:\prog{j(2)} in the first, and last dimensions respectively\footnote{When an array
-    is passed as function argument in Fortran its base is locally set to unity, regardless of the setting
-    at the caller scope, hence the first elements of \prog{i} and \prog{j} are always \prog{i(1)} and \prog{j(1)}.}.
-  The \prog{dtor()} method serves as a destructor\footnote{as of time of writing there is no
-    open-source Fortran compiler supporting destructor-like FINAL methods defined in the Fortran 2003 standard} 
-    and deallocates all the memory allocated
-    through previous calls to \prog{init}.
+    and \prog{j(1)}:\prog{j(2)} in the first, and last dimensions respectively\footnote{In Fortran, when an array
+    is passed as a function argument its base is locally set to unity, regardless of the setting
+    at the caller scope.}.
   Access to the arrays stored in the container is provided through the \prog{at} member field.
@@
 
-%  TODO: mention why module procedures have to be prefixed with module name in Fortran
   \subsection{Staggered grid}\label{sec:grid}
@@
 
-%TODO: explain why h=(-1,0) is needed with Blitz and Python (in order not to lose memory)
-
   In the C++ implementation first a global instance \prog{h} of an empty structure 
@@
     in expressions like \prog{i+one}, \prog{i+hlf}, where \prog{i} is an instance of \prog{slice}
-    \footnote{\label{fnt:slice}one could argue that not using an own implementation of a slice-representing class
+    \footnote{\label{fnt:slice}One could argue that not using an own implementation of a slice-representing class
     in NumPy is a design flaw -- being able to modify behaviour of a hypothetical numpy.slice class 
@@
     coding of array range extensions:
-% TODO: works for int 1 and slice
   \codecpp{code/cpp/listings.hpp}{listing06}{listing07}{1}
@@
   Consequently, a range depicted by $i\pm1/2$ may be expressed in the code as \prog{ext(i, h)}.
-
-%TODO: {discuss how halo region relates to parallel programming}
+  In all three implementations the \prog{ext()} function accept the second
+    argument to be an integer or a ''half'' (cf. section \ref{sec:grid}).
 
@@
   In order to ensure compile-time evaluation, the permutation order is passed 
-    via the template parameter \prog{d}:
+    via the template parameter \prog{d}
+    (note the different order of \prog{i} and \prog{j} arguments in the two template specialisations):
   \codecpp{code/cpp/listings.hpp}{listing07}{listing08}{1}
-  (note the different order of \prog{i} and \prog{j} arguments in the two template specialisations).
-
-  NumPy uses tuples of slices for addressing multi-dimensional arrays with a
-    single object, hence the following definition of function \prog{pi()} suffices to represent $\pi$:
+
+  NumPy uses tuples of slices for addressing multi-dimensional array with a
+    single object.
+  Therefore, the following definition of function \prog{pi()} suffices to represent $\pi$:
   \codepyt{code/pyt/listings.py}{listing06}{listing07}{1}
@@
   \codefor{code/for/listings.f}{listing05}{listing06}{1}
-
-%  TODO: templates -> branchless code by design, while F relies on opti
+  \prog{span()} is used to shorten the declarations of arrays to be returned 
+    from functions in the Fortran implementation (see listings F.11 and F.17--F.20).
+
+  It is worth noting here that the C++ implementation of \prog{pi()} is
+    branchless thanks to employment of template specialisation.
+  With Fortran one needs to rely on compiler optimisations to eliminate the conditional expression
+    within the \prog{pi()} that depends on value of \prog{d} which is always known 
+    at compile time.
 
@@
      requiring its descendants to implement at least
-     the \prog{advop()} method which is expected to fill \prog{psi[n+1]} with the values of \prog{psi[n]} advected
-     within one timestep.
+     the \prog{advop()} method which is expected to fill \prog{psi[n+1]} with an updated (advected)
+     values of \prog{psi[n]}.
   The two template parameters \prog{bcx\_t} and \prog{bcy\_t} allow the solver to operate with 
@@
 
-  The sizes of the arrays representing the two time-levels $\psi^{[n]}$ and $\psi^{[n+1]}$ of $\psi$ 
+  The donor-cell and MPDATA schemes both require only the previous state of an advected field
+    in order to advance the solution.
+  Consequently, memory for two time levels ($\psi^{[n]}$ and $\psi^{[n+1]}$) is allocated in the
+    constructor.
+  The sizes of the arrays representing the two time-levels of $\psi$ 
     are defined by the domain size ({\em nx}$~\times$~{\em ny}) plus the halo region.
   The size of the halo region is an argument of the constructor.
+  The \prog{cycle()} method is used to swap the time levels without copying any data.
 
@@
     (being laid out on the Arakawa-C staggered grid).
-
-%  TODO: clarify what cycle() does
 
@@
   Consequently, while in C++ (and Fortran) the computational domain is chosen to start at (i=0, j=0)
-    and hence a part of the halo region to have negative indices, in Python the halo region starts %TODO: footnote o roznych zakresach dla MPI
-    at (0,0).
+    and hence a part of the halo region to have negative indices, in Python the halo region starts 
+    at (0,0)\footnote{The reason to allow the domain to begin at arbitrary index
+    is mainly to ease debugging in case the code would be used in parallel computations
+    using domain decomposition where each subdomain could have its own index base corresponding to
+    the location within the computational domain}.
   However, since the whole halo logic is hidden within the solver, such details are not exposed to the
@@
 
-  The C++ and Python prototype solver implementations introduced above
-    allowed the solvers to operate with any boundary condition
-    objects provided the requirements imposed by calls to these objects' methods were fulfilled 
+  The above C++ and Python prototype solvers allow to operate with any boundary condition
+    objects provided the requirements imposed by calls to these objects' methods are fulfilled 
     (at compile-time in case of C++, and at run-time in the case of Python).
@@
     an abstract type with deferred procedures having abstract interfaces 
-    \citep[see Table 2.1 in][for a summary of approximate correspondence of OOP 
+    \citep[sic!, see Table 2.1 in][for a summary of approximate correspondence of OOP 
       nomenclature between Fortran and C++]{Rouson_et_al_2012}:
@@
 
-  As a side note, it is worth mentioning that the proposed design is readily extendable 
-    to handle concurrent computations using domain decomposition.
-  The proposed prototype solver is basically ready to be applied 
-    in a distributed memory set-up (e.g. using MPI)
-    since all necessary communication logic could be placed in the boundary-condition objects
-    not requiring any changes to the solver itself.
-  Shared-memory parallelisation (e.g. OpenMP) would require minimal changes to the solver structure
-    (the \prog{advop()} method would need to be called in a concurrently executed loop),
-    leaving the boundary condition- and advection operator-representing logic basically untouched.
-
-  Finally, it has to be underlined that the proposed prototype solver is designed
-    just as a minimal example to present the concepts described in the paper.
-%  It is by no means an optimal or versatile solution.
-
   \subsection{Periodic boundaries (C++)}\label{sec:cyclic}
@@
   MPDATA is an iterative algorithm in which each iteration takes the form of the 
-    so-called donor-cell formula which itself is a first-order advection scheme.
+    so-called donor-cell formula (which itself is a first-order advection scheme).
   The set of functions related to the donor-cell scheme is encapsulated within
@@
   \codecpp{code/cpp/listings.hpp}{listing15}{listing16}{1}
+  The above definition is not used in the following MPDATA implementation -- 
+    it is given as an example only.
 
@@
   MPDATA introduces corrective steps to the algorithm defined by equation~\ref{eq:donor} and \ref{eq:donor:F}.
-  Each corrective step has the form of equation~\ref{eq:donor} with the Courant numbers replaced
-    with those defined by the so-called antidiffusive velocities, and with $\psi$ fields from 
-    the previous iterations.
-  The Courant number field corresponding to the basic MPDATA antidiffusive velocity takes the following 
+  Each corrective step is a donor-cell step (eq.~\ref{eq:donor}) with the Courant number
+    fields corresponding to the MPDATA antidiffusive velocities of the following 
     form \citep[eqs~13,~14~in][]{Smolarkiewicz_1984}:
@@
   \end{equation}
-  where
+  where $\psi$ and $C$ represent values from the previous iteration and where
   \begin{equation}\label{eq:antidiff2}
@@
   \end{equation}
-  and where for positive-definite $\psi$, the $A$ and $B$ terms take the 
+  For positive-definite $\psi$, the $A$ and $B$ terms take the 
   following form\footnote{
@@
 
-  Listings P.13-PP.17 and F.15-F.21 contain the Python and Fortran counterparts to listing C.16-C.22.
+  Listings P.13-P.17 and F.15-F.21 contain the Python and Fortran counterparts to listing C.16-C.22.
 
@@
 
-  An MPDATA solver may be now constructed with the following definition in C++:
+  An MPDATA solver may be now constructed by inheriting from \prog{solver} class
+    with the following definition in C++:
   \codecpp{code/cpp/listings.hpp}{listing23}{listing24}{1}
+  The sequence of temporary arrays \prog{tmp} allocated in the constructor
+    is used to store the antidiffusive velocities from the present and optionally 
+    previous timestep (if using more than two iterations).
+
+  The \prog{advop()} method controlls the MPDATA iterations within one timestep.
+  The first (step = 0) iteration of MPDATA is an unmodified donor-cell step (see listing C.15).
+  Subsequent iterations begin with calculation of the antidiffusive Courant fields using 
+    formula \ref{eq:antidiff1}.
+  In order to calculate values spanning an (i\mhlf~... i+\phlf) range using a formula
+    for $C_{[i+\phlf,\ldots]}$ only, the formula is evaluated using extended
+    index ranges \prog{im} and \prog{jm}.
+  In the second (step=1) iteration the uncorrected Courant field (\prog{C\_unco}) points
+    to the original \prog{C} field, and the antidiffusive Courant field is written into
+    \prog{C\_corr} which points to \prog{tmp[1]}.
+  In the third (step=2) iteration \prog{C\_unco} points to \prog{tmp[1]} while \prog{C\_corr}
+    points to \prog{tmp[0]}.
+  In subsequent iterations \prog{tmp[0]} and \prog{tmp[1]} are alternately swapped.
+
+  Listings P.18 and F.22 contain the Python and Fortran counterparts to listing C.23.
 
@@
     settings of the number of iterations used in MPDATA.
+  The velocity field is constant in time and space (although it is
+    not assumed in the presented implementations).
   The signal shape at the end of each simulation is plotted as well.
   Plotting is done with the help of the gnuplot-iostream library.
-  The resultant plot is presented herein as Figure~\ref{fig:mpdata}
+
+  The resultant plot is presented herein as Figure~\ref{fig:mpdata}.
+  The top panel depicts the initial condition. 
+  The three other panels show a snapshot of the field after 75 timesteps.
+  The donor-cell solution is characterised by strongest numerical diffusion
+    resulting in significant drop in the signal amplitude.
+  The signals advected using MPDATA show smaller numerical diffusion.
 
   \codecpp{code/cpp/plot.cpp}{listing24}{listing25}{1}
+  Listings P.19 and F.23-F.24 contain the Python and Fortran 
+    counterparts to listing C.24 (with the set-up and plotting logic omitted).
 
@@
       See section~\ref{sec:example} for details.
-      %TODO: describe the test, point out adv., numerical diffusion, periodic boundaries, not much diff between 2 and 7
     }
@@
     and the \prog{-march=native} options.
+  The CPython interpreter was invoked with the \prog{-OO} option.
 
@@
 
-  Figure \ref{fig:mem} presents a plot of the peak memory use\footnote{the resident set size (rss)
+  Figure \ref{fig:mem} presents a plot of the peak memory use\footnote{The resident set size (rss)
     as reported by GNU time (version 1.7-24)} as a function of grid size.
@@
 
-  One of the aim of this paper is to exemplify OOP features of the three
-    programming languages (or language-library pairs)  trying to represent
-    what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012} within the
-     scientific software.
+  One of the aims of this paper is to show the applicability of OOP features of the three
+    programming languages (or language-library pairs) for scientific programming.
+  The main focus is to represent
+    what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012} within the code.
   Presented benchmark tests, although quite simplistic, together with the experience gained 
-    from the development of equally structured codes in three different
-    languages provide a basis for discussion of the tradeoffs of the programming language.
+    from the development of codes in three different
+    languages provide a basis for discussion on the tradeoffs of programming language choice.
   The discussion concerns in principle the development of finite-difference solvers for 
     partial differential equations, but is likely applicable the scientific software in general.
-  A partly objective and partly subjective summary of pros and cons of C++, Python and Fortran,
-    as observed during the development of the presented codes, is presented in the four following subsections.
+  A partly objective and partly subjective summary of pros and cons of C++, Python and Fortran
+    is presented in the four following subsections.
 
@@
     \item{definitions of functions returning array-valued expressions;}
-    \item{permutations of array indices allowing dimension-independent definitions
-      of functions (see e.g. listings C.12 and C.13, P.10 and P.11, F11 and F.12);}
-    \item{fractional indexing of arrays corresponding to employment of a staggered grid;}
-    \item{associating identifiers with slabs of arrays (e.g. for representing 
-      a psi[left\_halo]=psi[rght\_edge] assignment).} (DOROTA: w zasadzie tylko przyklad rozumiem)
+    \item{permutations of array indices allowing dimension-independant definitions
+      of functions (see e.g. listings C.12 and C.13, P.10 and P.11, F.11 and F.12);}
+    \item{fractional indexing of arrays corresponding to employment of a staggered grid.}
   \end{itemize}
@@
     \item{Fortran does not allow function calls to appear on the left hand side
-      of assignment (see e.g. how the \prog{tmp} pointers were used as a workaround in the \prog{cyclic\_fill\_halos}
-      method in listing F.8)}
-    \item{Fortran's built-in array-handling features are not object-oriented,
-      and hence not easily extendable via inheritance, but also subject to
-      some inherent limitations such as lack of support for arrays of arrays
-      (DOROTA: nie podoba mi sie to 'but also', moze wystarczy tylko pierwsza czesc punktu?)
-      (cf. Sect. \ref{sec:sequence})}
+      of assignment (see e.g. how the \prog{ptr} pointers were used as a workaround in the \prog{cyclic\_fill\_halos}
+      method in listing F.8);}
+    \item{Fortran lacks support for arrays of arrays (cf. Sect. \ref{sec:sequence}).}
   \end{itemize}
@@
   Taking into account such measures as code length or coding time,
-    the Python environment clearly gains significantly.
+    the Python environment gains significantly.
   Presented Python code is shorter than the C++ and Fortran counterparts,
@@
     make Python a favourable choice even if high performance is important, especially
-    if a performance-to-development cost measure is to be considered.
+    if a cobbined measure of performance and development cost is to be considered.
 
@@
 
-(DOROTA: zastanawiam sie, czy tego nie wyrzucic)
-  Python simplifies also building and shearing of software,
-    especially if multiple architectures and operating systems are targeted.
+  Python implementations do not expose to the user the compilation or linking processes. 
+  As a result, Python-written software is easier to deploy and share, especially 
+    if multiple architectures and operating systems are targeted.
   However, CMake\footnote{CMake is a family of open-source, cross-platform
     tools automating building, testing and packaging of C/C++/Fortran software,
-    see http://cmake.org/ for details}
+    see http://cmake.org/}
     is a tool (or arguably the tool) that allows to efficiently automate 
@@
 
-  Python is also definitely easiest to debug among the three languages.
-  There exist great debugging tools for C++, but the debugging and development is 
-    often hindered by indecipherable compiler messages.
-  OOP with Fortran is a relatively new concept, and the support for the Fortran OOP
-    features among free and open source compilers, debuggers and other programming
-    aids are still not mature.
-
-  A factor contributing to the ease of use of the array-handling 
-    facilities is the clarity of rules determining their performance.
+  Python is definitely easiest to debug among the three languages.
+  Great debugging tools for C++ do exist, however the debugging and development is 
+    often hindered by indecipherable compiler messages
+    flooded with lengthy type names stemming from employment of templates.
+  Support for the OOP features of Fortran among free and open source compilers, 
+    debuggers and other programming aids remains immature.
+
   With both Fortran and Python, the memory footprint caused by employment
-    of temporary objects in array arithmetics is dependent on compiler optimisations.
-  In contrast, Blitz++ ensures temporary-array-free computations by design.
+    of temporary objects in array arithmetics is dependant on compiler choice or
+    the level of optimisations.
+  In contrast, Blitz++ ensures temporary-array-free computations by design
+    \citep{Veldhuizen_et_al_1997}.
   Similarly, the C++ template mechanism and such features as the C++11
@@
   The size of the programmers' community of a given language 
-    influence the availability of trained personnel, 
+    influences the availability of trained personnel, 
     reusable software components and information resources.
-    It also affect the maturity and quality of compilers and tools. 
+  It also affects the maturity and quality of compilers and tools. 
   Fortran is a domain-specific language while Python and C++ are general-purpose languages
     with disproportionately larger users' communities.
-  The OOP features were introduced to Fortran only recently, and have not gained
-    wide popularity among users \citep{Worth_2008}\footnote{an anecdotal yet significant
+  The OOP features of Fortran have not gained
+    wide popularity among users \citep{Worth_2008}\footnote{An anecdotal yet significant
     example being the incomplete support for syntax-highlighting of modern Fortran in Vim and Emacs editors}.
-  Fortran is almost no longer taught at the universities \citep{Kendall_et_al_2008},
+  Fortran is no longer routinely taught at the universities \citep{Kendall_et_al_2008},
     in contrast to C++ and Python.
-  An example ofdecreasing popularity of Fortran in academia 
+  An example of decreasing popularity of Fortran in academia 
     is the discontinuation of Fortran printed editions of the ''Numerical Recipes'' 
@@
 
-  In the case of C++ and Python it is also the size of the particular
-    numerical-library users' community that is of importance.
   Blitz++ is one of several packages that offer high-performance object-oriented
@@
     (the gnuplot-iostream used in listing C.24 being a much-appreciated counterexample).
- ( DOROTA: nie rozumiem nastepnego zdania)
-  However, Blitz++ allows to instantiate its arrays using preexisting data what
-    allows to interface with virtually any library, although requiring additional
-    coding.
-
-(DOROTA: a moze skrocic to i przeniesc do 4.1?)
+  However, Blitz++ allows to interface with virtually any library (including Fortran libraries), 
+    by resorting to referencing the underlying memory with raw pointers.
+
   The availability and quality of object-oriented 
@@
     algorithms and facilities for interaction with host operating system.
-%  External Fortran libraries are mostly limited to scientific applications, 
-%    while there exist collections of general-purpose C++ and Python libraries.
-%  Interfacing any of the three considered languages from
-%    another one does not pose a problem, though.
   In the authors' experience, the small popularity of OOP techniques among
-    Fortran users is reflected in the library designs.
+    Fortran users is reflected in the library designs (including the Fortran's
+    built-in library routines).
   What makes correct use of external libraries more difficult with Fortran
@@
 
-(DOROTA: nastepny akapit zaznaczylabym Fijalowi do uwaznego przejrzenia, 
-Ja nie rozumime tej czesci o numexpr itp)
   Finally, the three languages differ as well with regard to availability of 
     mechanisms for leveraging shared-memory parallelisation (e.g. with multi-core processors).
-  GCC supports OpenMP with Fortran and C++, as well as the C++11 built-in support for threads.
+  GCC supports OpenMP with Fortran and C++.
   The CPython and PyPy implementations of Python do not offer any
     built-in solution for multithreading.
-  However, there are third-party party libraries such as numexpr which can employ multiple cores from
-    a single program.
-  There exist object-oriented libraries for both C++ and Python to 
-    transparently use graphics processing units (GPU) such as Thrust \citep{Thrust}
-    and Theano \citep{Bergstra_et_al_2010}.
 
@@
 
-  Three object-oriented implementations of a prototype solver 
+  Three implementations of a prototype solver 
     for the advection equation were introduced.
-  The solvers are based on the MPDATA algorithm \citep{Smolarkiewicz_2006}.
-  All implementations follow the same structure but are implemented
-    in three different language and library pairs:
+  The solvers are based on MPDATA - an algorithm of particular applicability
+    in geophysical fluid dynamics \citep{Smolarkiewicz_2006}.
+  All implementations follow the same object-oriented structure but are implemented
+    in three different languages:
   \begin{itemize}
@@
   \end{itemize}
-(DOROTA: przelozylam to i delikatnie zmienilam, ale nadal nie rozuemi czesci 'employed tools...')
-  The fact that all considered standards are open and the employed
-    tools implementing them are free and open-source
-    is certainly a great advantage \citep{Anel_2011}.
-
-  Finite-difference partial differential equation solvers were example used to study 
-    the capabilities of these languages
-    for object-oriented scientific programming.
-  The emphasis was on the capabilities for reproduction 
-    of the mathematical abstractions used for explaining the algorithms
-    in journal papers within the program code.
 
@@
     the NumPy support in the PyPy implementation of Python.
+  The fact that all considered standards are open and the employed
+    tools implementing them are free and open-source
+    is certainly an advantage \citep{Anel_2011}.
+
   The key conclusion is that all considered language/library/compiler
@@
     represent the mathematical abstractions within the program code. 
-(OROTA: dwa nastepne zdania bym przeformulowala, ale chce sie upewnic, co wlasciwie chcemy tu powiedziec)
-  The readability and brevity of such code may contribute to its 
+  The readability and brevity of such code:
+  \begin{itemize}
+    \item{contribute to its 
       auditability, indispensable for credible and reproducible research in computational science 
-    \citep{Post_et_al_2005, Merali_et_al_2010, Stodden_et_al_2012}.
-  Employment of OOP helps as well to keep complex programs maintainable, avoiding accumulation of the technical/code 
-    debt \citep{Buschmann_2011} that besets scientific software in such domains as climate modelling
+      \citep{Post_et_al_2005, Merali_et_al_2010, Stodden_et_al_2012};
+    }
+    \item{help to keep the programs maintainable, avoiding accumulation of the code 
+      debt\footnote{See \citet{Buschmann_2011} for discussion of technical/code debt.} 
+      that besets scientific software in such domains as climate modelling
       \citep{Freeman_et_al_2010}.
+    }
+  \end{itemize}
 
@@
     allowing to reproduce the analysis and plots presented in section~\ref{sec:perf} are all 
-    available for download from the project repository\footnote{git repository at \url{http://github.com/slayoo/mpdata/}},
+    available in the CPC Program Library and at
+    the project repository\footnote{git repository at \url{http://github.com/slayoo/mpdata/}},
     and are released under the GNU GPL license \citep{GPLv3}.
-
-(DOROTA: to trzeba zmienic, nie wiem co dokladnie chciales powiedziec.)
+  The authors encourage to use the presented codes for teaching and benchmarking purposes.
+
   The OOP design enhances the possibilities to reuse and extend the presented code.
@@
     \noindent 
-    Thanks are due Piotr Smolarkiewicz and Hanna Pawłowska.
+    We thank Piotr Smolarkiewicz and Hanna Pawłowska for their help throughout the project.
+    This study was partly inspired by the lectures of Lech Łobocki.
+
     \noindent
-    Authors would like to thank Tobias Burnus, Julian Cummings, Patrik Jonsson,
-      Arjen Markus and Janus Weil for their feedback on questions posted to Blitz++ and gfortran
-      mailing lists.
+    Tobias Burnus, Ond\v{r}ej \v{C}ert\'ik, Julian Cummings, Patrik Jonsson,
+      Arjen Markus, Zbigniew Piotrowski, Davide del Vento, Janus Weil and Andrzej Wyszogrodzki 
+      provided valuable feedback to the initial version of the
+      manuscript and/or responces to questions posted to Blitz++ and gfortran mailing lists.
 

### 2013-03-16 paper/pyt/numba/listings_noslice.py
version that suspend numba... I believe it's related to bcx

@@ -17,7 +17,7 @@ except AttributeError:
   pass
 
 try:
-    from numba import jit, float64, int_, void, autojit
+    from numba import jit, float64, int_, void, autojit, object_
     float64_1d=float64[:]
     float64_2d=float64[:,:]
     print "Numba or Numbapro imported"
@@ -84,15 +84,15 @@ class Cyclic(object):
 @jit
 class Solver(object):
   # ctor-like method
-  @void(int_, int_, int_)
-  def __init__(self, nx, ny, hlo):
+  @void(object_, object_, int_, int_, int_)
+  def __init__(self, bcx, bcy, nx, ny, hlo):
     self.n = 0
     self.hlo = hlo
     self.i = slice(hlo, nx + hlo)
     self.j = slice(hlo, ny + hlo)
 
-    self.bcx = Cyclic(0, self.i.start, self.i.stop, hlo)
-    self.bcy = Cyclic(1, self.j.start, self.j.stop, hlo)
+#    self.bcx = bcx(0, self.i.start, self.i.stop, hlo)
+#    self.bcy = bcy(1, self.j.start, self.j.stop, hlo)
 
     self.psi = (
       numpy.empty((
@@ -135,16 +135,17 @@ class Solver(object):
   def solve(self, nt):
     #pdb.set_trace()
     for t in range(nt):
-      self.bcx.fill_halos(
-        self.psi[self.n], ext(self.j, self.hlo, self.hlo).start, 
-        ext(self.j, self.hlo, self.hlo).stop 
-      )
-      self.bcy.fill_halos(
-        self.psi[self.n], ext(self.i, self.hlo, self.hlo).start, 
-        ext(self.i, self.hlo, self.hlo).stop 
-      )
-      self.advop() 
-      self.cycle()
+      #self.bcx.fill_halos(
+      #  self.psi[self.n], ext(self.j, self.hlo, self.hlo).start, 
+      #  ext(self.j, self.hlo, self.hlo).stop 
+      #)
+      #self.bcy.fill_halos(
+      #  self.psi[self.n], ext(self.i, self.hlo, self.hlo).start, 
+      #  ext(self.i, self.hlo, self.hlo).stop 
+      #)
+      #self.advop() 
+      #self.cycle()
+      pass
   
 
 #listing09

### 2013-03-16 paper/pyt/numba/listings_noslice.py
coming back to the 2988b0c commit and changing clasc Cyclic with method fill_halos to function cyclic_fill_halos, since I had problems with passing bcx/bcy to the other classes and methods; it works with test_numba.py (test_pythononly doesn't work with listing_noslice.py since I had to changed the init in Mpdata)

@@ -11,13 +11,13 @@ try:
 except ImportError:
   pass
 import numpy
-try: 
+try:
   numpy.seterr(all='ignore')
 except AttributeError:
   pass
 
 try:
-    from numba import jit, float64, int_, void, autojit, object_
+    from numba import jit, float64, int_, void, autojit
     float64_1d=float64[:]
     float64_2d=float64[:,:]
     print "Numba or Numbapro imported"
@@ -46,114 +46,98 @@ def sl_add(sl, int_num):
 @autojit
 def ext(r, lft, rgh):
   return slice(
-    r.start - lft, 
+    r.start - lft,
     r.stop + rgh
   )
 #listing06
 @autojit
 def pi(d, idx_0, idx_1):
   idx = [idx_0, idx_1]
-  return (idx[d], idx[d-1]) 
-
-#listing08                                                                                           
-@jit
-class Cyclic(object):
-  @void(int_, int_, int_, int_)
-  def __init__(self, d, i_start, i_stop, hlo):
-    i  = slice(i_start, i_stop)
-    self.d = d
-    self.left_halo = slice(i.start-hlo, i.start    )
-    self.rght_edge = slice(i.stop -hlo, i.stop     )
-    self.rght_halo = slice(i.stop,      i.stop +hlo)
-    self.left_edge = slice(i.start,     i.start+hlo)
-
-  # method invoked by the solver                                                                     
-  @void(float64_2d, int_, int_)
-  def fill_halos(self, psi, j_start, j_stop):
-    j = slice(j_start, j_stop)
-    psi[pi(self.d, self.left_halo, j)] = (
-      psi[pi(self.d, self.rght_edge, j)]
-    )
-    psi[pi(self.d, self.rght_halo, j)] = (
-      psi[pi(self.d, self.left_edge, j)]
-    )
-
-
+  return (idx[d], idx[d-1])
 
 #listing07
-@jit
 class Solver(object):
   # ctor-like method
-  @void(object_, object_, int_, int_, int_)
-  def __init__(self, bcx, bcy, nx, ny, hlo):
+  def __init__(self, nx, ny, hlo):
     self.n = 0
     self.hlo = hlo
[…]

### 2013-03-16 paper/pyt/numba/test_numba.py

@@ -1,4 +1,4 @@
-from listings import *
+from listings_noslice import *
 import sys
 import time
 
@@ -30,7 +30,7 @@ def read_file(fname, nx, ny):
 # comapring array from output files with results after advection
 # dec is the accuracy of comparison, it might be needed sometimes to use smaller
 def checking_output(fin, fout, nx, ny, Cx, Cy, nt, it, dec=4):
-  slv = Mpdata(it, Cyclic, Cyclic, nx, ny)
+  slv = Mpdata(it, nx, ny)
   slv.state()[:] = read_file(fin, nx, ny)
   slv.courant(0)[:] = Cx
   slv.courant(1)[:] = Cy

### 2013-03-17 paper/pyt/numba/listings_noslice.py
trying to jit the Solver: I had to remove real_t to avoid error: NoneType object has no attribute dtype; this version still doesn't work...

@@ -55,7 +55,28 @@ def pi(d, idx_0, idx_1):
   idx = [idx_0, idx_1]
   return (idx[d], idx[d-1])
 
+#listing08                                                                                        
+#changed class to function for testing numba                                                       
+@autojit
+def cyclic_fill_halos(d, i_start, i_stop, hlo, psi, j_start, j_stop):
+    i = slice(i_start, i_stop)
+    left_halo = slice(i.start-hlo, i.start )
+    rght_edge = slice(i.stop -hlo, i.stop )
+    rght_halo = slice(i.stop, i.stop +hlo)
+    left_edge = slice(i.start, i.start+hlo)
+
+    j = slice(j_start, j_stop)
+    psi[pi(d, left_halo, j)] = (
+      psi[pi(d, rght_edge, j)]
+    )
+    psi[pi(d, rght_halo, j)] = (
+      psi[pi(d, left_edge, j)]
+    )
+
+
+
 #listing07
+@jit
 class Solver(object):
   # ctor-like method
   def __init__(self, nx, ny, hlo):
@@ -69,36 +90,40 @@ class Solver(object):
       numpy.empty((
         ext(self.i, self.hlo, self.hlo).stop,
         ext(self.j, self.hlo, self.hlo).stop
-      ), real_t),
+      )),
       numpy.empty((
         ext(self.i, self.hlo, self.hlo).stop,
         ext(self.j, self.hlo, self.hlo).stop
-      ), real_t)
+      ))
     )
 
     self.C = (
       numpy.empty((
         ext(self.i, 1, 0).stop,
         ext(self.j, self.hlo, self.hlo).stop
-      ), real_t),
+      )),
       numpy.empty((
         ext(self.i, self.hlo, self.hlo).stop,
         ext(self.j, 1, 0).stop
-      ), real_t)
+      ))
     )
 
   # accessor methods
+  @float64_2d()
   def state(self):
     return self.psi[self.n][self.i, self.j]
 
   # helper methods invoked by solve()
+  @float64_2d(int_)
   def courant(self,d):
     return self.C[d][:]
 
+  @void()
   def cycle(self):
     self.n = (self.n + 1) % 2 - 2
 
    # integration logic
+  @void(int_)
   def solve(self, nt):
     #pdb.set_trace()
     for t in range(nt):
@@ -115,23 +140,6 @@ class Solver(object):
[…]

### 2013-03-17 paper/pyt/numba/listings_noslice.py
it doesn't work - error: Cannot access attribute advop of type Solver. I believe the class Solver has to have a method advop (numba stoped complaining when I added), this kind of inheritance probably impossible with numba

@@ -122,10 +122,14 @@ class Solver(object):
   def cycle(self):
     self.n = (self.n + 1) % 2 - 2
 
+#  @void()
+#  def advop(self):
+#    pass
+
    # integration logic
   @void(int_)
   def solve(self, nt):
-    #pdb.set_trace()
+    pdb.set_trace()
     for t in range(nt):
       cyclic_fill_halos(
         0, self.i.start, self.i.stop, self.hlo,

### 2013-03-17 paper/pyt/numba/listings_noslice.py
adding @void to init method in the Solver, for some reason numba didn't complaine...

@@ -56,7 +56,7 @@ def pi(d, idx_0, idx_1):
   return (idx[d], idx[d-1])
 
 #listing08                                                                                        
-#changed class to function for testing numba                                                       
+#changed class to function for testing numba 
 @autojit
 def cyclic_fill_halos(d, i_start, i_stop, hlo, psi, j_start, j_stop):
     i = slice(i_start, i_stop)
@@ -79,6 +79,7 @@ def cyclic_fill_halos(d, i_start, i_stop, hlo, psi, j_start, j_stop):
 @jit
 class Solver(object):
   # ctor-like method
+  @void(int_, int_, int_)
   def __init__(self, nx, ny, hlo):
     self.n = 0
     self.hlo = hlo

### 2013-03-17 paper/pyt/numba/listings_noslice.py
coming back to the Cyclic method (it didn't cause yesterday problems); code still doesn't work without advop method in the Solver

@@ -17,7 +17,7 @@ except AttributeError:
   pass
 
 try:
-    from numba import jit, float64, int_, void, autojit
+    from numba import jit, float64, int_, void, autojit, object_
     float64_1d=float64[:]
     float64_2d=float64[:,:]
     print "Numba or Numbapro imported"
@@ -56,36 +56,40 @@ def pi(d, idx_0, idx_1):
   return (idx[d], idx[d-1])
 
 #listing08                                                                                        
-#changed class to function for testing numba 
-@autojit
-def cyclic_fill_halos(d, i_start, i_stop, hlo, psi, j_start, j_stop):
+@jit
+class Cyclic(object):
+  @void(int_, int_, int_, int_)
+  def __init__(self, d, i_start, i_stop, hlo):
     i = slice(i_start, i_stop)
-    left_halo = slice(i.start-hlo, i.start )
-    rght_edge = slice(i.stop -hlo, i.stop )
-    rght_halo = slice(i.stop, i.stop +hlo)
-    left_edge = slice(i.start, i.start+hlo)
+    self.d = d
+    self.left_halo = slice(i.start-hlo, i.start )
+    self.rght_edge = slice(i.stop -hlo, i.stop )
+    self.rght_halo = slice(i.stop, i.stop +hlo)
+    self.left_edge = slice(i.start, i.start+hlo)
 
+  # method invoked by the solver
+  @void(float64_2d, int_, int_)
+  def fill_halos(self, psi, j_start, j_stop):
     j = slice(j_start, j_stop)
-    psi[pi(d, left_halo, j)] = (
-      psi[pi(d, rght_edge, j)]
+    psi[pi(self.d, self.left_halo, j)] = (
+      psi[pi(self.d, self.rght_edge, j)]
     )
-    psi[pi(d, rght_halo, j)] = (
-      psi[pi(d, left_edge, j)]
+    psi[pi(self.d, self.rght_halo, j)] = (
+      psi[pi(self.d, self.left_edge, j)]
     )
-
-
 
 #listing07
 @jit
 class Solver(object):
   # ctor-like method
-  @void(int_, int_, int_)
-  def __init__(self, nx, ny, hlo):
+  @void(object_, object_, int_, int_, int_)
+  def __init__(self, bcx, bcy, nx, ny, hlo):
     self.n = 0
     self.hlo = hlo
     self.i = slice(hlo, nx + hlo)
[…]

### 2013-03-17 paper/pyt/numba/test_numba.py
test_numba is not needed anymore, after restoration od Cyclic class test_pythononly.py works

@@ -1,91 +1,0 @@
-from listings_noslice import *
-import sys
-import time
-
-Examples_san = {"nx" : [3, 3, 3], 
-                "ny" : [3, 3, 3],
-                "Cx" : [0, 1, 1],
-                "Cy" : [1, 0, 0],
-                "nt" : [3, 4, 4],
-                "it" : [1, 1, 3]}
-
-Examples_time = {"nx" : [64, 81, 512],
-                 "ny" : [64, 81, 512],
-                 "Cx" : [.2, .2, .2],
-                 "Cy" : [.5, .5, .5],
-                 "nt" : [4096, 2557, 64],
-                 "it" : [3, 3, 3]}
-
-# reading array from input/output files 
-def read_file(fname, nx, ny):
-  tmp = numpy.empty((nx, ny), real_t)
-  with open(fname, 'r') as f:
-    x = 0
-    for line in f:
-      tmp[x,:] = numpy.fromstring(line, dtype=real_t, sep='\t')
-      x += 1
-  assert(x == nx)
-  return tmp
-
-# comapring array from output files with results after advection
-# dec is the accuracy of comparison, it might be needed sometimes to use smaller
-def checking_output(fin, fout, nx, ny, Cx, Cy, nt, it, dec=4):
-  slv = Mpdata(it, nx, ny)
-  slv.state()[:] = read_file(fin, nx, ny)
-  slv.courant(0)[:] = Cx
-  slv.courant(1)[:] = Cy
-  t0 = time.time()
-  slv.solve(nt)
-  print "time", time.time() - t0
-
-  diff = numpy.amax(abs(slv.state() - read_file(fout, nx, ny)))
-  print "diff=", diff
-  if nx*ny < 100:
-    print "out", slv.state()
-  if (diff >= .5 * pow(10, -dec)): 
-    print slv.state().dtype, slv.state()
-    tmp = read_file(fout, nx, ny)
-    print tmp.dtype, tmp
-    print numpy.max(numpy.abs(slv.state() - tmp))
-    raise Exception()
-
-def main(ex = Examples_san, kat="/glade/u/home/jarecka/mpdata/paper/tests/sanity/"):
-  print "sys", sys.argv
-# progarm should work when running without any additional command-line arguments 
-# or with one additional command-line argument "sanity" or "timing"
-  if len(sys.argv) == 1:
-    pass
-  elif len(sys.argv) == 2:
-    if sys.argv[1] == "sanity":
[…]

### 2013-03-17 paper/pyt/numba/test_pythononly.py
just adding import listings_noslice, test_pythononly works again for this file

@@ -1,4 +1,4 @@
-from listings import *
+from listings_noslice import *
 import sys
 import time
 

### 2013-03-18 paper/cpp/listings.hpp
various cleanups and text polishing

@@ -138,9 +138,6 @@ struct cyclic
   }
 };
 //listing10
-namespace donorcell
-{
-//listing11
   template<class T1, class T2, class T3> 
   inline auto F(
     const T1 &psi_l, const T2 &psi_r, const T3 &C
@@ -150,7 +147,7 @@ namespace donorcell
       (C - abs(C)) * psi_r
     ) / 2
   )
-//listing12
+//listing11
   template<int d>  
   inline auto donorcell( 
     const arr_t &psi, const arr_t &C, 
@@ -167,8 +164,8 @@ namespace donorcell
         C(pi<d>(i-h, j))
     )
   )
-//listing13
-  void op(
+//listing12
+  void donorcell_op(
     const arrvec_t &psi, const int n,
     const arrvec_t &C, 
     const rng_t &i, const rng_t &j
@@ -177,9 +174,7 @@ namespace donorcell
       - donorcell<0>(psi[n], C[0], i, j)
       - donorcell<1>(psi[n], C[1], j, i); 
   }
-//listing14
-}; 
-//listing15
+//listing13
 template<class bcx_t, class bcy_t>
 struct solver_donorcell : solver<bcx_t, bcy_t> 
 {
@@ -189,73 +184,68 @@ struct solver_donorcell : solver<bcx_t, bcy_t>
 
   void advop()
   {
-    donorcell::op(
+    donorcell_op(
       this->psi, this->n, this->C, 
       this->i, this->j
     );
   }
 };
-//listing16
-namespace mpdata
-{
-//listing17
+//listing14
   template<class nom_t, class den_t>
-  inline auto frac(
+  inline auto mpdata_frac(
     const nom_t &nom, const den_t &den
   ) return_macro(
     where(den > 0, nom / den, 0)
   ) 
-//listing18
+//listing15
   template<int d>
-  inline auto A(const arr_t &psi, 
+  inline auto mpdata_A(const arr_t &psi, 
     const rng_t &i, const rng_t &j
   ) return_macro(
-    frac(
+    mpdata_frac(
       psi(pi<d>(i+1, j)) - psi(pi<d>(i,j)),
       psi(pi<d>(i+1, j)) + psi(pi<d>(i,j))
     ) 
   ) 
-//listing19
+//listing16
   template<int d>
-  inline auto B(const arr_t &psi, 
[…]

### 2013-03-18 paper/cpp/plot.cpp

@@ -1,4 +1,4 @@
-//listing24
+//listing20
 #include "listings.hpp"
 #define GNUPLOT_ENABLE_BLITZ
 #include <gnuplot-iostream/gnuplot-iostream.h>
@@ -52,7 +52,8 @@ int main()
        << "splot '-' binary" << binfmt
        << "with lines notitle\n";
     gp.sendBinary(slv.state().copy());
-  } {
+  } 
+  {
     const int it = 2;
     solver_mpdata<it, cyclic<x>, cyclic<y>> 
       slv(n[x], n[y]); 
@@ -63,7 +64,8 @@ int main()
        << "splot '-' binary" << binfmt
        << "with lines notitle\n";
     gp.sendBinary(slv.state().copy());
-  } {
+  } 
+  {
     const int it = 44;
     solver_mpdata<it, cyclic<x>, cyclic<y>> 
       slv(n[x], n[y]); 
@@ -76,4 +78,4 @@ int main()
     gp.sendBinary(slv.state().copy());
   }
 }
-//listing25
+//listing21

### 2013-03-18 paper/for/listings.f

@@ -389,7 +389,7 @@ module solver_donorcell_m
     class(arrvec_t), pointer :: C
     C => this%C
     call donorcell_op(                                 &
-      this%psi, this%n, C, this%i, this%j         &
+      this%psi, this%n, C, this%i, this%j              &
     )
   end subroutine
 end module
@@ -401,7 +401,7 @@ module mpdata_m
   implicit none
   contains 
 !listing16
-  function frac(nom, den) result (return)
+  function mpdata_frac(nom, den) result (return)
     real(real_t), intent(in) :: nom(:,:), den(:,:)
     real(real_t) :: return(size(nom, 1), size(nom, 2))
     where (den > 0)
@@ -411,23 +411,23 @@ module mpdata_m
     end where
   end function
 !listing17
-  function A(d, psi, i, j) result (return)
+  function mpdata_A(d, psi, i, j) result (return)
     integer :: d
     real(real_t), allocatable, intent(in) :: psi(:,:)
     integer, intent(in) :: i(2), j(2)
     real(real_t) :: return(span(d, i, j), span(d, j, i))
-    return = frac(                                     &
+    return = mpdata_frac(                              &
       pi(d, psi, i+1, j) - pi(d, psi, i, j),           &
       pi(d, psi, i+1, j) + pi(d, psi, i, j)            &
     )  
   end function
 !listing18
-  function B(d, psi, i, j) result (return)
+  function mpdata_B(d, psi, i, j) result (return)
     integer :: d
     real(real_t), allocatable, intent(in) :: psi(:,:) 
     integer, intent(in) :: i(2), j(2)
     real(real_t) :: return(span(d, i, j), span(d, j, i))
-    return = frac(                                     &
+    return = mpdata_frac(                              &
       pi(d, psi, i+1, j+1) + pi(d, psi, i,   j+1)      &
     - pi(d, psi, i+1, j-1) - pi(d, psi, i,   j-1),     &
       pi(d, psi, i+1, j+1) + pi(d, psi, i,   j+1)      &
@@ -435,7 +435,7 @@ module mpdata_m
     ) / 2
   end function
 !listing19
-  function C_bar(d, C, i, j) result (return)
+  function mpdata_C_bar(d, C, i, j) result (return)
     integer :: d
     real(real_t), allocatable, intent(in) :: C(:,:) 
     integer, intent(in) :: i(2), j(2)
@@ -447,7 +447,7 @@ module mpdata_m
     ) / 4               
   end function
 !listing20
[…]

### 2013-03-18 paper/pyt/listings.py

@@ -16,7 +16,7 @@ try:
 except AttributeError:
   pass
 #listing03
-class Shift():
+class shift():
   def __init__(self, plus, mnus):
     self.plus = plus
     self.mnus = mnus
@@ -31,8 +31,8 @@ class Shift():
       arg.stop  - self.mnus
     )
 #listing04
-one = Shift(1,1) 
-hlf = Shift(0,1)
+one = shift(1,1) 
+hlf = shift(0,1)
 #listing05
 def ext(r, n):
   if (type(n) == int) & (n == 1): 
@@ -45,7 +45,7 @@ def ext(r, n):
 def pi(d, *idx): 
   return (idx[d], idx[d-1])
 #listing07
-class Solver(object):
+class solver(object):
   # ctor-like method
   def __init__(self, bcx, bcy, nx, ny, hlo):
     self.n = 0
@@ -102,7 +102,7 @@ class Solver(object):
       self.cycle()
   
 #listing08
-class Cyclic(object):
+class cyclic(object):
   # ctor
   def __init__(self, d, i, hlo): 
     self.d = d
@@ -147,9 +147,9 @@ def donorcell_op(psi, n, C, i, j):
     - donorcell(1, psi[n], C[1], j, i)
   )
 #listing12
-class Donorcell(Solver):
+class solver_donorcell(solver):
   def __init__(self, bcx, bcy, nx, ny):
-    Solver.__init__(self, bcx, bcy, nx, ny, 1)
+    solver.__init__(self, bcx, bcy, nx, ny, 1)
 
   def advop(self):
     donorcell_op(
@@ -157,17 +157,17 @@ class Donorcell(Solver):
       self.C, self.i, self.j
     )
 #listing13
-def frac(nom, den):
+def mpdata_frac(nom, den):
   return numpy.where(den > 0, nom/den, 0)
 #listing14
-def a_op(d, psi, i, j):
-  return frac(
+def mpdata_A(d, psi, i, j):
+  return mpdata_frac(
     psi[pi(d, i+one, j)] - psi[pi(d, i, j)],
     psi[pi(d, i+one, j)] + psi[pi(d, i, j)]
   )
 #listing15
-def b_op(d, psi, i, j):
-  return frac( 
+def mpdata_B(d, psi, i, j):
+  return mpdata_frac( 
[…]

### 2013-03-18 paper/pyt/test.py

@@ -26,7 +26,7 @@ def main():
   dec = int(sys.argv[9])
 
 #listing19
-  slv = Mpdata(it, Cyclic, Cyclic, nx, ny)
+  slv = solver_mpdata(it, cyclic, cyclic, nx, ny)
   slv.state()[:] = read_file(fname, nx, ny)
   slv.courant(0)[:] = Cx
   slv.courant(1)[:] = Cy

### 2013-03-18 paper/cpp/listings.hpp
further text cleanups

@@ -18,7 +18,7 @@ struct arrvec_t : boost::ptr_vector<arr_t>
 {
   const arr_t &operator[](const int i) const 
   {   
-    return this->at( (i + this->size()) % this->size() ); 
+    return this->at((i + this->size()) % this->size()); 
   }
 };
 //listing05

### 2013-03-18 paper/doc/paper.tex
further text cleanups (with missing paper.tex)

@@
 \usepackage{enumitem}
-\setitemize{noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}
+\setitemize{noitemsep,itemsep=1pt,topsep=1pt,parsep=0pt,partopsep=0pt}
 
@@
 \journal{Computer Physics Communications}
-%\journal{arXiv. Paper presented at the American Meteorological Society 2013 Annual Meeting in Austin, Texas, USA}
 
@@
         Three object-oriented implementations of a prototype solver of the advection equation are introduced.
-        Presented programs are based on Blitz++ (C++), NumPy (Python), and Fortran's built-in array containers.
+        The presented programs are based on Blitz++ (C++), NumPy (Python), and Fortran's built-in array containers.
         The solvers include an implementation of the Multidimensional Positive-Definite 
@@
           used in the literature within the program code.
-        The codes serve also as a basis for discussion on the tradeoffs of the programming language choice.
+        A discussion on the tradeoffs of the programming language choice is presented.
         The main angles of comparison are code brevity and syntax clarity
           (and hence maintainability and auditability) as well as performance.
-        In case of Python, a significant performance gain is observed when switching from the standard 
+        In the case of Python, a significant performance gain is observed when switching from the standard 
           interpreter (CPython) to the PyPy implementation of Python.
@@
     paradigm for creating complex software''} \citep[][Sec.~1.3]{Press_et_al_2007}.
-  It is intriguing that while the quoted statement comes from the very book subtitled 
-   {\em The Art of Scientific Computing}, hardly any (if not none) of currently operational 
+  It is intriguing that, while the quoted statement comes from the very book subtitled 
+   {\em The Art of Scientific Computing}, hardly any (if not none) of the currently operational 
     weather and climate prediction systems - flagship examples of complex scientific software - 
-    makes extensive use of OOP techniques\footnote{
-  Fortran have been the language of choice in oceanic \citep{Griffies_et_al_2000}, 
+    make extensive use of OOP techniques.
+  Fortran has been the language of choice in oceanic \citep{Griffies_et_al_2000}, 
     weather-prediction \citep{Sundberg_2009} and Earth system \citep{Legutke_2012} modelling, 
-    and none of its 20-century editions were object-oriented languages \citep[see e.g.][for discussion]{Norton_et_al_2007}.}.
-  Application of OOP techniques in development of numerical modelling software may help to
-    {\bf (i)} maintain modularity and separation of program logic layers (e.g. separation of
-      a dynamical core, parallelisation mechanisms, data input/output, error handling and
-      the description of physical processes); and
-    {\bf (ii)} {\bf shorten and simplify the source code and improve its readability by reproducing within 
-      the program logic the mathematical abstractions and notation used in the literature}.
+    and none of its 20-century editions were object-oriented languages \citep[see e.g.][for discussion]{Norton_et_al_2007}.
+
+  Application of OOP techniques in development of numerical modelling software may help to:
+  \begin{enumerate}[label=(\roman*), leftmargin=*, widest=ii]
+    \item{maintain modularity and separation of program logic layers (e.g. separation of
+      numerical algorithms, parallelisation mechanisms, data input/output, error handling and
+      the description of physical processes); and}
+    \item{{\bf shorten and simplify the source code and improve its readability by reproducing within 
+      the program logic the mathematical notation used in the literature}.}
+  \end{enumerate}
   The first application is attainable, yet arguably cumbersome, with procedural programming.
@@
   It also enables the compiler or library authors to relieve the user (i.e. scientific programmer)
-    from hand-coding (premature) optimisations, a practice long recognised as having {\em a strong negative impact when debugging
+    from hand-coding optimisations, a practice long recognised as having {\em a strong negative impact when debugging
     and maintenance are considered} \citep{Knuth_1974}.
@@
     \citep[e.g.][respectively]{Ziemianski_et_al_2011,Abiodun_et_al_2011,Ezer_et_al_2002}.
-  MPDATA is a solver for systems of equations of the following form:
+  MPDATA is a solver for systems of advection equations of the following form:
   \begin{equation}\label{eq:adv}
@@
   \end{equation}
-  that describe transport of quantity $\psi$ by fluid flow with velocity $\vec{v}$.
+  that describe evolution of a scalar field $\psi$ transported by the fluid flow with velocity $\vec{v}$.
   Quoting Numerical Recipes once more, development of methods to numerically solve such problems 
@@
     and MPDATA is an example of the state-of-the art in this field.
-  MPDATA is designed to accurately solve equation (\ref{eq:adv}) in arbitrary
+  MPDATA is designed to accurately solve equation (\ref{eq:adv}) in an arbitrary
     number of dimensions assuring positive-definiteness of scalar field $\psi$ 
     and incurring small numerical diffusion.
-  All relevant MPDATA formul\ae are given in the text but are presented without
+  All relevant MPDATA formul\ae~are given in the text but are presented without
     derivation or detailed discussion.
@@
     \citep[][Chapter~11]{Markus_2012}, respectively.
-  The third section covers performance evaluation of the three implementations.
+  The third section of this paper covers performance evaluation of the three implementations.
   The fourth section covers discussion of the tradeoffs of the programming language choice.
@@
 
-  Throughout the paper the three implementations are presented by discussing 
-    source code listings which cover entire program code.
+  Throughout the paper we present the three implementations by discussing 
+    source code listings which cover the entire program code.
   Subsections \ref{sec:array}-\ref{sec:solver} describe all three implementations,
@@
   The relevant parts of Python and Fortran codes do not differ significantly, and for readability reasons 
-    are presented in the \ref{app:P} and the \ref{app:F}, respectively.
+    are presented in \ref{app:P} and \ref{app:F}, respectively.
 
@@
     solver timestep and $\Delta x$ is the grid spacing.
-  Among the three considered languages only Fortran is equipped with built-in
-    array handling facilities of practical use in high-performance computing.
 
   Presented C++ implementation of MPDATA is built upon the Blitz++ 
-    library\footnote{Blitz++ is a C++ class library for scientific computing which uses the expression templates technique to achieve high performance, see http://sf.net/projects/blitz/}.
+    library\footnote{Blitz++ is a C++ class library for scientific computing which uses the expression templates technique to achieve high performance, see \url{http://sf.net/projects/blitz/}}.
   Blitz offers object-oriented representation of n-dimensional arrays,
@@
 
-  Arguably, the power of Blitz++ comes from the ability to express array expressions as objects.
+  The power of Blitz++ comes from the ability to express array expressions as objects.
   In particular, it is possible to define a function that returns an array expression;
@@
     caller scope.
-  For example, definition of a function returning its array-valued argument doubled reads:
+  For example, definition of a function returning its array-valued argument doubled, reads:
     \prog{auto~f(arr\_t~x)~return\_macro(2*x)}.
@@
 
-  For the Python implementation of MPDATA the NumPy package is used.
+  For the Python implementation of MPDATA the NumPy\footnote{NumPy is a Python package
+    for scientific computing offering support for multi-dimensional arrays and a library
+    of numerical algorithms, see \url{http://numpy.org/}} package is used.
   In order to make the code compatible with both the standard CPython
     as well as the alternative PyPy implementation of Python \citep[][]{Bolz_et_al_2011},
-    the Python implementation includes the following sequence of \prog{import} statements:
+    the Python code includes the following sequence of \prog{import} statements:
   \codepyt{code/pyt/listings.py}{listing02}{listing03}{1}
@@
   Second, to match the settings of C++ and Fortran compilers used herein, the NumPy package is instructed 
-    to ignore any floating-point errors (only if such an option
-    is available in the interpreter\footnote{\prog{numpy.seterr()} is not supported in PyPy as of version 1.9}).
+    to ignore any floating-point errors, if such an option
+    is available in the interpreter\footnote{\prog{numpy.seterr()} is not supported in PyPy as of version 1.9}.
   The above lines conclude all code modifications that needed to be added in order to run
@@
 
-  Fortran features built-in multidimensional array handling mechanisms,
-    hence there is no need for using an external package as with C++ and Python.
+  Among the three considered languages only Fortran is equipped with built-in
+    array handling facilities of practical use in high-performance computing. 
+  Therefore, there is no need for using an external package as with C++ and Python.
   Fortran array-handling features are not object-oriented, though.
 
-  \subsection{Array sequence containers}\label{sec:sequence}
-
-  While discretisation in space of the scalar field $\psi(x,y)$ into its $\psi_{[i,j]}$ 
-    grid representation requires floating-point array containers (cf. section \ref{sec:array}).
-  Discretisation in time requires a container class for storing
+  \subsection{Containers for sequences of arrays}\label{sec:sequence}
+
+  As discussed above, discretisation in space of the scalar field $\psi(x,y)$ into its $\psi_{[i,j]}$ 
+    grid representation requires floating-point array containers.
+  In turn, discretisation in time requires a container class for storing
     sequences of such arrays, i.e. \{$\psi^{[n]}$, $\psi^{[n+1]}$\}.
@@
   Second, the order of dimensions would need to be different for different languages to assure that
-      the contiguous dimension is not used for time levels.
+    the contiguous dimension is used for one of the space dimensions and not for time levels.
 
@@
     and at the same time to handle automatic freeing of dynamically allocated memory.
-  For clearer ,,translation'' of mathematical formul\ae~into C++ code, the
-    \prog{ptr\_vector} class is further customised by defining a derived structure whose element-access \prog{[~]} 
-    operator is overloaded with a modulo-variant:
+  The \prog{ptr\_vector} class is further customised by defining a derived structure which element-access \prog{[~]} 
+    operator is overloaded with a modulo variant:
   \codecpp{code/cpp/listings.hpp}{listing04}{listing05}{1}
@@
   \codefor{code/for/listings.f}{listing02}{listing03}{1}
-  The \prog{arr\_t} type is defined solely for the purpose 
-    of overcoming the Fortran's limitation of not allowing to declare arrays of arrays.
-  The \prog{arrptr\_t} type is defined as creating an allocatable of its instances instead
-    of multi-element pointer of \prog{arr\_t} ensures automatic memory deallocation.
-  The \prog{ctor()} method initialises the container for a given number of elements \prog{n}.
-  The \prog{init()} method initialises the \prog{n}-th element of the container with 
-    an newly allocated 2D array spanning indices \prog{i(1)}:\prog{i(2)}, 
+  The \prog{arr\_t} type is defined solely for the purpose of overcoming the limitation 
+    of lack of an array-of-arrays construct, and its only member field is a two-dimensional array.
+  An array of \prog{arr\_t} is used hereinafter as a container for sequences of arrays.
+
+  The \prog{arrptr\_t} type is defined solely for the purpose of overcoming Fortran's limitation
+    of not supporting allocatables of pointers.
+  \prog{arrptr\_t}'s single member field is a pointer to an instance of \prog{arr\_t}.
+  Creating an allocatable of \prog{arrptr\_t}, instead of 
+    a multi-element pointer of \prog{arr\_t}, ensures automatic memory deallocation.
+
+  Type \prog{arrptr\_t} is used to implement the from-the-end addressing of elements in \prog{arrvec\_t}.
+  The array data is stored in the \prog{arrs} member field (of type \prog{arr\_t}).
+  The \prog{at} member field (of type \prog{arrptr\_t}) stores pointers to the elements of \prog{arrs}. 
+  \prog{at} has double the length of \prog{arrs} and is initialised in a cyclic manner so that
+    the \prog{-1} element of \prog{at} points to the last element of \prog{arrs}, and so on.
+  Assuming \prog{psi} is an instance of \prog{arrptr\_t}, the \prog{(i,j)} element of the \prog{n}-th array in \prog{psi}
+    may be accessed with\\ \prog{psi\%at( n )\%p\%a( i, j )}.
+
+  The \prog{ctor(n)} method initialises the container for a given number of elements \prog{n}.
+  The \prog{init(n,i,j)} method initialises the \prog{n}-th element of the container with 
+    a newly allocated 2D array spanning indices \prog{i(1)}:\prog{i(2)}, 
     and \prog{j(1)}:\prog{j(2)} in the first, and last dimensions respectively\footnote{In Fortran, when an array
@@
     at the caller scope.}.
-  Access to the arrays stored in the container is provided through the \prog{at} member field.
-  Assuming \prog{psi} is an instance of \prog{arrptr\_t}, the \prog{(i,j)} element of the \prog{n}-th array in \prog{psi}
-    may be accessed with \prog{psi\%at( n )\%p\%a( i, j )}.
 
@@
     is a natural choice for MPDATA.
-%  MPDATA is a finite-difference approximation to the flux-form of the equation~(\ref{eq:adv}) 
-%    resulting in a finite-volume scheme.
-%  Using the Arakawa-C grid allows to evaluate the fluxes at 
-%    the grid cell (finite volume) edges without interpolation.
   As a consequence, the discretised representations of the $\psi$ scalar field, and each 
@@
     employed programming languages.
-  A desired syntax would translate:
-  \begin{itemize}
-    \item{\prog{$i-\phlf$} ~~ $\leadsto$ ~~ \prog{$i-1$}} 
-    \item{\prog{$i+\phlf$} ~~ $\leadsto$ ~~ \prog{$i$}.}
-  \end{itemize}
+  A desired syntax would translate \prog{$i-\phlf$} to \prog{$i-1$} and
+    \prog{$i+\phlf$} to \prog{$i$}.
   OOP offers a convenient way to implement such notation
@@
     but also for shifting the slices by integer intervals ($i\pm1$).
-  It is implemented here by declaring a \prog{Shift} class with the adequate operator overloads:
+  It is implemented here by declaring a \prog{shift} class with the adequate operator overloads:
   \codepyt{code/pyt/listings.py}{listing03}{listing04}{1}
@@
     through inheritance would allow to implement the same behaviour as obtained in listing P.3 without the need to represent 
-    the unity as a separate class}:
+    the unity as a separate object}:
   \codepyt{code/pyt/listings.py}{listing04}{listing05}{1}
@@
   Hereinafter, the $\pi_{a,b}^{d}$ symbol is used to denote a  
-    cyclic permutation of the order $d$ of the set $\{a,b\}$.
+    cyclic permutation of an order $d$ of a set $\{a,b\}$.
   It is used to generalise the MPDATA formul\ae~into multiple dimensions
@@
   \codefor{code/for/listings.f}{listing05}{listing06}{1}
-  \prog{span()} is used to shorten the declarations of arrays to be returned 
+  The \prog{span()} function is used to shorten the declarations of arrays to be returned 
     from functions in the Fortran implementation (see listings F.11 and F.17--F.20).
@@
 
-  \subsection{A prototype solver}\label{sec:solver}
+  \subsection{Prototype solver}\label{sec:solver}
 
   The tasks to be handled by a prototype advection equation solver proposed herein are:
-  \begin{description}
-    \item[(i)]{storing arrays representing the $\psi$ and $\vec{C}$ fields and any required housekeeping data,}
-    \item[(ii)]{allocating and deallocating the required memory,}
-    \item[(iii)]{providing access to the solver state, and}
-    \item[(iv)]{performing the integration by invoking the advection-operator and
+  \begin{enumerate}[label=(\roman*), leftmargin=*, widest=iii]
+    \item{storing arrays representing the $\psi$ and $\vec{C}$ fields and any required housekeeping data,}
+    \item{allocating and deallocating the required memory,}
+    \item{providing access to the solver state,}
+    \item{performing the integration by invoking the advection-operator and
       boundary-condition handling routines.}
-  \end{description}
+  \end{enumerate}
   In the following C++ definition of the \prog{solver} structure, task (i) is represented with the definition of the structure
@@
     constructor.
-  The sizes of the arrays representing the two time-levels of $\psi$ 
+  The sizes of the arrays representing the two time levels of $\psi$ 
     are defined by the domain size ({\em nx}$~\times$~{\em ny}) plus the halo region.
@@
   The arrays representing the $C^{[x]}$ and $C^{[y]}$ components of $\vec{C}$,
-    require {\em nx+1}~$\times$~{\em ny} 
-    and {\em nx}$~\times$~{\em ny+1} elements, respectively
+    require ({\em nx+1})~$\times$~{\em ny} 
+    and {\em nx}$~\times$~({\em ny+1}) elements, respectively
     (being laid out on the Arakawa-C staggered grid).
 
-  Python definition of the \prog{Solver} class follows closely the C++ structure definition:
+  Python definition of the \prog{solver} class follows closely the C++ structure definition:
   \codepyt{code/pyt/listings.py}{listing07}{listing08}{1}
-  The key difference stems from the fact that unlike Blitz++, NumPy does not allow an array
+  The key difference stems from the fact that, unlike Blitz++, NumPy does not allow an array
     to have arbitrary index base -- in NumPy the first element is always addressed with 0.
@@
     and hence a part of the halo region to have negative indices, in Python the halo region starts 
-    at (0,0)\footnote{The reason to allow the domain to begin at arbitrary index
+    at (0,0)\footnote{The reason to allow the domain to begin at an arbitrary index
     is mainly to ease debugging in case the code would be used in parallel computations
@@
 
-  The above C++ and Python prototype solvers allow to operate with any boundary condition
-    objects provided the requirements imposed by calls to these objects' methods are fulfilled 
-    (at compile-time in case of C++, and at run-time in the case of Python).
+  The above C++ and Python prototype solvers in principle allow to operate with any boundary condition
+    objects that implement methods called from within the solver.
+  This requirement is checked at compile-time in the case of C++, and at run-time in the case of Python.
   In order to obtain an analogous behaviour with Fortran, it is required to
@@
 
-  The solver definition described in section~\ref{sec:solver} required a given
+  From this point, only C++ implementation is explained in the main text. 
+  The Python and Fortran implementations are included in appendices P and F.
+
+  The solver definition described in section~\ref{sec:solver} requires a given
     boundary condition object to implement a \prog{fill\_halos()} method.
@@
     so-called donor-cell formula (which itself is a first-order advection scheme).
-  The set of functions related to the donor-cell scheme is encapsulated within
-    the \prog{donorcell} namespace in C++:
-  \codecpp{code/cpp/listings.hpp}{listing10}{listing11}{1}
 
@@
 
-  The flux function takes the following form in C++:
-  \codecpp{code/cpp/listings.hpp}{listing11}{listing12}{1}
-
+  \noindent The flux function takes the following form in C++:
+  \codecpp{code/cpp/listings.hpp}{listing10}{listing11}{1}
+
+  \noindent 
   Equation~\ref{eq:donor} is split into the terms under the summation 
     (effectively the 1-dimensional donor-cell formula):
+  \codecpp{code/cpp/listings.hpp}{listing11}{listing12}{1}
+
+  \noindent and the actual two-dimensional donor-cell formula:
+
   \codecpp{code/cpp/listings.hpp}{listing12}{listing13}{1}
 
-  and the actual two-dimensional donor-cell formula:
-  \codecpp{code/cpp/listings.hpp}{listing13}{listing14}{1}
-
-  Finally, the C++ namespace is ended with:
-  \codecpp{code/cpp/listings.hpp}{listing14}{listing15}{1}
-
   Listings P.9-P11 and F.9-F.13 contain the Python and Fortran counterparts to listings C.12-C.15.
 
-  \subsection{A donor-cell solver (C++)}\label{sec:donorcell_solver}
+  \subsection{Donor-cell solver (C++)}\label{sec:donorcell_solver}
 
@@
     a \prog{solver\_donorcell} implementation of the abstract \prog{solver} class: 
-  \codecpp{code/cpp/listings.hpp}{listing15}{listing16}{1}
-  The above definition is not used in the following MPDATA implementation -- 
-    it is given as an example only.
+  \codecpp{code/cpp/listings.hpp}{listing13}{listing14}{1}
+  The above definition is given as an example only.
+  In the following sections an MPDATA solver of the same structure is defined.
 
@@
   \end{equation}
-  where $\psi$ and $C$ represent values from the previous iteration and where
+  where $\psi$ and $C$ represent values from the previous iteration and where:
   \begin{equation}\label{eq:antidiff2}
@@
 
-  The set of functions related to MPDATA is encapsulated within
-    the \prog{mpdata} namespace in C++:
-  \codecpp{code/cpp/listings.hpp}{listing16}{listing17}{1}
-
   If the denominator in equations \ref{eq:A} or \ref{eq:B} equals zero for a given {\em i} and {\em j}, 
@@
     represented with the \prog{where} construct (available in all three considered languages):
+  \codecpp{code/cpp/listings.hpp}{listing14}{listing15}{1}
+
+  The $A$ term defined in equation \ref{eq:A} takes the following form:
+  \codecpp{code/cpp/listings.hpp}{listing15}{listing16}{1}
+
+  The $B$ term defined in equation \ref{eq:B} takes the following form:
+  \codecpp{code/cpp/listings.hpp}{listing16}{listing17}{1}
+
+  Equation \ref{eq:antidiff2} takes the following form:
   \codecpp{code/cpp/listings.hpp}{listing17}{listing18}{1}
 
-  The $A$ term defined in equation \ref{eq:A} takes the following form:
+  Equation \ref{eq:antidiff1} take the following form:
   \codecpp{code/cpp/listings.hpp}{listing18}{listing19}{1}
 
-  The $B$ term defined in equation \ref{eq:B} takes the following form:
-  \codecpp{code/cpp/listings.hpp}{listing19}{listing20}{1}
-
-  Equation \ref{eq:antidiff2} takes the following form:
-  \codecpp{code/cpp/listings.hpp}{listing20}{listing21}{1}
-
-  Equation \ref{eq:antidiff1} take the following form:
-  \codecpp{code/cpp/listings.hpp}{listing21}{listing22}{1}
-
-  Finally, the C++ namespace is ended with:
-  \codecpp{code/cpp/listings.hpp}{listing22}{listing23}{1}
-
   Listings P.13-P.17 and F.15-F.21 contain the Python and Fortran counterparts to listing C.16-C.22.
 
-  \subsection{An MPDATA solver (C++)}\label{sec:mpdatasolver}
+  \subsection{MPDATA solver (C++)}\label{sec:mpdatasolver}
 
@@
     with the following definition in C++:
-  \codecpp{code/cpp/listings.hpp}{listing23}{listing24}{1}
-  The sequence of temporary arrays \prog{tmp} allocated in the constructor
+  \codecpp{code/cpp/listings.hpp}{listing19}{listing20}{1}
+  The array of sequences of temporary arrays \prog{tmp} allocated in the constructor
     is used to store the antidiffusive velocities from the present and optionally 
@@
   The \prog{advop()} method controlls the MPDATA iterations within one timestep.
-  The first (step = 0) iteration of MPDATA is an unmodified donor-cell step (see listing C.15).
+  The first (step = 0) iteration of MPDATA is an unmodified donor-cell step (compare listing C.15).
   Subsequent iterations begin with calculation of the antidiffusive Courant fields using 
@@
   The signal shape at the end of each simulation is plotted as well.
-  Plotting is done with the help of the gnuplot-iostream library.
+  Plotting is done with the help of the gnuplot-iostream library\footnote{gnuplot-iostream
+    is a header-only C++ library allowing gnuplot to be controlled from C++, see \url{http://stahlke.org/dan/gnuplot-iostream/}. 
+    Gnuplot is a portable command-line driven graphing utility, see \url{http://gnuplot.info/}}.
 
@@
     resulting in significant drop in the signal amplitude.
-  The signals advected using MPDATA show smaller numerical diffusion.
-
-  \codecpp{code/cpp/plot.cpp}{listing24}{listing25}{1}
+  The signals advected using MPDATA show smaller numerical diffusion with
+    the solution obtained with more iterations preserving the signal altitude 
+    more accurately.
+  In all of the simulations the signal maintains its positive definiteness.
+  The domain periodicity is apparent in the plots as the maximum of the signal 
+    after 75 timesteps is located near the domain walls.
+
   Listings P.19 and F.23-F.24 contain the Python and Fortran 
@@
 
+  \codecpp{code/cpp/plot.cpp}{listing20}{listing21}{1}
+
   \begin{figure}
-    \pgfimage[height=.95\textheight]{code/cpp/figure}
+    \pgfimage[height=.9\textheight]{code/cpp/figure}
     \caption{\label{fig:mpdata}
       Plot generated by the program given in listing~C.24.
-      See section~\ref{sec:example} for details.
+      The top panel shows initial signal shape (at time t=0).
+      The subsequent panels show snapshots of the advected field after 75 timesteps
+        from three different simulations: donorcell (or 1 MPDATA iteration), MPDATA
+        with two iterations and MPDATA with 44 iterations.
+      The colour scale and the wire-frame surface correspond to signal amplitude.
+      See section~\ref{sec:example} for discussion.
     }
@@
 
-  \begin{figure}
+  The three introduced implementations of MPDATA were tested with the following set-ups 
+    employing free and open-source tools:
+  \begin{description}
+    \item[C++:]{~
+      \begin{itemize}
+        \item{GCC g++ 4.8.0\footnote{\label{fnt:gcc-snapshot}GNU Compiler Collection packaged in the Debian's gcc-snapshot\_20130222-1} 
+          and Blitz++ 0.10}
+        \item{LLVM Clang 3.2 and Blitz 0.10}
+      \end{itemize}
+    }
+    \item[Python:]{~
+      \begin{itemize}
+        \item{CPython 2.7.3 and NumPy 1.7}
+        \item{PyPy 1.9.0 with built-in NumPy implementation}
+      \end{itemize}
+    }
+    \item[Fortran:]{~
+      \begin{itemize}
+        \item{GCC gfortran 4.8.0\textsuperscript{\ref{fnt:gcc-snapshot}}}
+      \end{itemize}
+    }
+  \end{description}
+  The performance tests were run on a Debian and an Ubuntu GNU/Linux systems with the above-listed software obtained 
+    via binary packages from the distributions' package repositories (most recent package versions at the time of writing).
+  The tests were performed on two 64-bit machines equipped with
+    an AMD Phenom\textsuperscript{\texttrademark} II X6 1055T (800 MHz)
+    and an Intel\textsuperscript{\textregistered} Core\textsuperscript{\texttrademark}~i5-2467M (1.6 GHz)
+    processors.
+
+  For both C++ and Fortran the GCC compilers were invoked with the \prog{-Ofast} and the 
+    \prog{-march=native} options.
+  The Clang compiler was invoked with the \prog{-O3}, the \prog{-mllvm -vectorize}, the \prog{-ffast-math} 
+    and the \prog{-march=native} options.
+  The CPython interpreter was invoked with the \prog{-OO} option.
+
+  In addition to the standard Python implementation CPython,
+    the Python code was tested with PyPy.
+  PyPy is an alternative implementation of Python featuring a just-in-time compiler. 
+  PyPy includes an experimental partial reimplementation of NumPy that compiles NumPy expressions into native assembler.
+  Thanks to employment of lazy evaluation of array expressions (cf. Sect. \ref{sec:array})
+    PyPy allows to eliminate the use of temporary matrices for storing intermediate results,
+    and to perform multiple operations on the arrays within a single array index traversal
+    \footnote{Lazy evaluation available in PyPy 1.9 has been temporarily removed from PyPy during a refactoring of
+      the code. It'll be reinstantiated in the codebase as soon as possible, but past PyPy 2.0 release}.
+  Consequently, PyPy allows to overcome the same performance-limiting factors as those addressed by Blitz++, although 
+    the underlying mechanisms are different.
+  In contrast to other solutions for improving performance of NumPy-based codes such as
+    Cython\footnote{see \url{http://cython.org}}, 
+    numexpr\footnote{see \url{http://code.google.com/p/numexpr/}} or 
+    Numba\footnote{see \url{http://numba.pydata.org/}}, 
+    PyPy does not require any modifications to the code.
+  Thus, PyPy may serve as a drop-in replacement for CPython ready to be used with 
+    previously-developed codes.
+
+  The same set of tests was run with all four set-ups.
+  Each test set consisted of 16 program runs.
+  The test programs are analogous to the example code presented in section~\ref{sec:example}.
+  The tests were run with different grid sizes ranging from 64$\times$64 to 2048$\times$2048.
+  The Gaussian impulse was advected for $nt=2^{24}/(nx\cdot ny)$ timesteps ($2^{24}$ chosen arbitrarily), 
+    in order to assure comparable timing accuracy for all grid sizes.
+  Three MPDATA iterations were used (i.e. two corrective steps).
+  The initial condition was loaded from a text file, and the final values were compared at the end of the test
+    with values loaded from another text file assuring the same results were obtained with all four set-ups.
+  The tests were run multiple times; program start-up, data loading, and output verification times were
+    subtracted from the reported values (see caption of Figure~\ref{fig:cpu-eyrie} for details).
+
+  \begin{figure}[t]
     \center
@@
   \end{figure}
-  \begin{figure}
-    \center
-    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=laptok-cpu=GenuineIntel-6-42-bogomips=3190}
-    \caption{\label{fig:cpu-eyrie}
-      Execution time statistics for the test runs described in Section~\ref{sec:perf}
-        plotted as a function of grid size.
-      Values of the total user mode cpu time reported by the GNU time utility are
-        and normalised by the grid size ($nx \cdot ny$) and the number of timesteps $nt=2^{24}/(nx \cdot ny)$.
-      Before normalisation the time reported for an $nt=0$ run for a corresponding
-        domain size is subtracted from the values.
-      Both the $nt=0$ and $nt=2^{24}/()nx \cdot ny$ runs are repeated three times and
-        the shortest times are taken into account only.
-      Results obtained with an Intel\textsuperscript{\textregistered} 
-        Core\textsuperscript{\texttrademark} i5 1.6 GHz processor.
-    }
-  \end{figure}
-  \begin{figure}
-    \center
-    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=skua-cpu=AuthenticAMD-16-10-bogomips=5620}
-    \caption{\label{fig:cpu-skua}
-      Same as Fig.~\ref{fig:cpu-eyrie} for an AMD Phenom\textsuperscript{\texttrademark} II 800 MHz processor.
-    }
-  \end{figure}
-
-  The three introduced implementations of MPDATA were tested with the following set-ups 
-    employing free and open-source tools:
-  \begin{description}
-    \item[C++:]{~
-      \begin{itemize}
-        \item{GCC g++ 4.8.0\footnote{\label{fnt:gcc-snapshot}GNU Compiler Collection packaged in the Debian's gcc-snapshot\_20130222-1} 
-          and Blitz++ 0.10}
-        \item{LLVM Clang 3.2 and Blitz 0.10}
-      \end{itemize}
-    }
-    \item[Python:]{~
-      \begin{itemize}
-        \item{CPython 2.7.3 and NumPy 1.7}
-        \item{PyPy 1.9.0 with built-in NumPy implementation}
-      \end{itemize}
-    }
-    \item[Fortran:]{~
-      \begin{itemize}
-        \item{GCC gfortran 4.8.0\textsuperscript{\ref{fnt:gcc-snapshot}}}
-      \end{itemize}
-    }
-  \end{description}
-  The performance tests were run on a Debian and an Ubuntu GNU/Linux systems with the above-listed software obtained 
-    via binary packages from the distributions's package repositories (most recent package versions at the time of writing).
-  The tests were performed on two 64-bit machines equipped with
-    an AMD Phenom\textsuperscript{\texttrademark} II X6 1055T (800 MHz)
-    and an Intel\textsuperscript{\textregistered} Core\textsuperscript{\texttrademark}~i5-2467M (1.6 GHz)
-    processors.
-
-  For both C++ and Fortran the GCC compilers were invoked with the \prog{-Ofast} and the 
-    \prog{-march=native} options.
-  The Clang compiler was invoked with the \prog{-O3}, the \prog{-mllvm -vectorize}, the \prog{-ffast-math} 
-    and the \prog{-march=native} options.
-  The CPython interpreter was invoked with the \prog{-OO} option.
-
-  The Python code was tested with the standard Python implementation CPython and with PyPy.
-  PyPy is an alternative implementation of Python featuring a just-in-time compiler. 
-  PyPy includes a partial reimplementation of NumPy that compiles NumPy expressions into native assembler.
-  Thanks to employment of lazy evaluation of array expressions (cf. Sect. \ref{sec:array})
-    PyPy allows to eliminate the use of temporary matrices for storing intermediate results,
-    and to perform multiple operations on the arrays within a single array index traversal
-    \footnote{Lazy evaluation available in PyPy 1.9 has been temporarily removed from PyPy during a refactoring of
-      the code. It'll be reinstantiated in the codebase as soon as possible, but past PyPy 2.0 release}.
-  Consequently, PyPy allows to overcome the same performance-limiting factors as those addressed by Blitz++, although 
-    the underlying mechanisms are different.
-  In contrast to other solutions for improving performance of NumPy-based codes such as
-    Cython or numexpr, PyPy does not require
-    any modifications to the code and allows to preserve its object-oriented structure.
-  PyPy may thus serve as a drop-in replacement for CPython ready to be used with 
-    previously-developed codes.
-
-  The same set of tests was run with all four set-ups.
-  Each test set consisted of 16 program runs.
-  The test programs are analogous to the example code presented in section~\ref{sec:example}.
-  The tests were run with different grid sizes ranging from 64$\times$64 to 2048$\times$2048.
-  The Gaussian impulse was advected for $nt=2^{24}/(nx\cdot ny)$ timesteps, 
-    in order to assure comparable timing accuracy for all grid sizes.
-  Three MPDATA iterations were used (i.e. two corrective steps).
-  The initial condition was loaded from a text file, and the final values were compared at the end of the test
-    with values loaded from another text file assuring the same results were obtained with all four set-ups.
-  The tests were run multiple times; program startup, data loading, and output verification times were
-    subtracted from the reported values (see caption of Figure~\ref{fig:cpu-eyrie} for details).
 
   Figure \ref{fig:mem} presents a plot of the peak memory use\footnote{The resident set size (rss)
-    as reported by GNU time (version 1.7-24)} as a function of grid size.
+    as reported by GNU time (version 1.7-24)} (identical for both considered CPUs)
+    as a function of grid size.
   The plotted values are normalised by the nominal size of all data arrays used in the program
@@
     are indicative of the amount of temporary memory used for array manipulation.
-  PyPy- and Blitz++-based set-ups consume notably less memory than Fortran and CPython confirming 
-    the effectiveness of the just-in-time compilation (PyPy), and the expression-templates (Blitz++) techniques
+  PyPy- and Blitz++-based set-ups consume notably less memory than Fortran and CPython.
+  This confirms the effectiveness of the just-in-time compilation (PyPy) and the expression-templates (Blitz++) techniques
     for elimination of temporary storage during array operations.
 
+  \begin{figure}[t]
+    \center
+    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=laptok-cpu=GenuineIntel-6-42-bogomips=3190}
+    \caption{\label{fig:cpu-eyrie}
+      Execution time statistics for the test runs described in Section~\ref{sec:perf}
+        plotted as a function of grid size.
+      Values of the total user mode CPU time reported by the GNU time utility are
+        normalised by the grid size ($nx \cdot ny$) and the number of timesteps $nt=2^{24}/(nx \cdot ny)$.
+      Before normalisation the time reported for an $nt=0$ run for a corresponding
+        domain size is subtracted from the values.
+      Both the $nt=0$ and $nt=2^{24}/()nx \cdot ny$ runs are repeated three times and
+        only the shortest time is taken into account.
+      Results obtained with an Intel\textsuperscript{\textregistered} 
+        Core\textsuperscript{\texttrademark} i5 1.6 GHz processor.
+    }
+  \end{figure}
+  \begin{figure}[t]
+    \center
+    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=skua-cpu=AuthenticAMD-16-10-bogomips=5620}
+    \caption{\label{fig:cpu-skua}
+      Same as Fig.~\ref{fig:cpu-eyrie} for an AMD Phenom\textsuperscript{\texttrademark} II 800 MHz processor.
+    }
+  \end{figure}
+
   The CPU time statistics presented in Figures \ref{fig:cpu-eyrie} and \ref{fig:cpu-skua} reveal
-    some differences between results obtained with the two different processors,
-    however the results agree qualitatively leading to the following observations
+    minor differences between results obtained with the two different processors.
+  Presented results lead to the following observations
     (where by referring to language names, only the results obtained with the herein considered
@@
     all tests were carried out using square grids, this should not have had any impact on the
-    performance (both Blitz++ and NumPy support Fortran's column-major ordering as well, 
-    however this feature is still missing from PyPy's built-in NumPy implementation as of PyPy 1.9)
+    performance\footnote{Both Blitz++ and NumPy support Fortran's column-major ordering as well, 
+    however this feature is still missing from PyPy's built-in NumPy implementation as of PyPy 1.9}.
 
@@
   One of the aims of this paper is to show the applicability of OOP features of the three
-    programming languages (or language-library pairs) for scientific programming.
+    programming languages (or language-library pairs) for scientific computing.
   The main focus is to represent
@@
   The discussion concerns in principle the development of finite-difference solvers for 
-    partial differential equations, but is likely applicable the scientific software in general.
+    partial differential equations, but is likely applicable to the scientific software in general.
   A partly objective and partly subjective summary of pros and cons of C++, Python and Fortran
@@
 
-  It was shown in section~\ref{sec:impl} that C++/Blitz++, Python/NumPy and Fortran
+  It was shown in section~\ref{sec:impl} that C++11/Blitz++, Python/NumPy and Fortran 2008
     provide comparable functionalities in terms of matching the blackboard abstractions
@@
     \item{definitions of functions returning array-valued expressions;}
-    \item{permutations of array indices allowing dimension-independant definitions
+    \item{permutations of array indices allowing dimension-independent definitions
       of functions (see e.g. listings C.12 and C.13, P.10 and P.11, F.11 and F.12);}
@@
     exist partially in NumPy as well (see footnote \ref{fnt:slice}).
-  The lack of a counterpart in Fortran to the C++ template mechanism was identified
+  The lack of a counterpart in Fortran to the C++ template mechanism was identified in
     \citep{Cary_et_al_1997}
@@
     corresponded to both shortest execution time and smallest memory footprint.
-  Moreover, the shortest execution times of Fortran correspond to largest memory use.
 
   One may consider performance measures addressing not only the program efficiency but also 
-    the factors influencing the development and maintenance time/cost.
+    the factors influencing the development and maintenance time/cost 
+    \citep[of particular importance in scientific computing,][]{Wilson_2006}.
   Taking into account such measures as code length or coding time,
@@
     make Python a favourable choice even if high performance is important, especially
-    if a cobbined measure of performance and development cost is to be considered.
+    if a combined measure of performance and development cost is to be considered.
 
@@
     if multiple architectures and operating systems are targeted.
-  However, CMake\footnote{CMake is a family of open-source, cross-platform
+  However, there exist tools such as CMake\footnote{CMake is a family of open-source, cross-platform
     tools automating building, testing and packaging of C/C++/Fortran software,
-    see http://cmake.org/}
-    is a tool (or arguably the tool) that allows to efficiently automate 
-    builds, tests and packaging of C++ and Fortran programs.
+    see \url{http://cmake.org/}}
+    that allow to efficiently automate 
+    building, testing and packaging of C++ and Fortran programs.
 
@@
   In contrast, Blitz++ ensures temporary-array-free computations by design
-    \citep{Veldhuizen_et_al_1997}.
-  Similarly, the C++ template mechanism and such features as the C++11
-    \prog{constexpr} construct allow the programmer to enforce evaluation
-    at compile time\footnote{Useful in context of scientific computing for example to perform
-    compile-time dimensional analysis of the physically-meaningful expressions within
-    the code with no runtime performance penalty \citet{Schabel_et_al_2008}}.
+    \citep{Veldhuizen_et_al_1997} avoiding unintentional performance loss.
 
@@
 
-  The availability and quality of object-oriented 
-    libraries differs among the three considered languages.
+  The availability and quality of libraries 
+    that offer object-oriented interfaces
+    differs among the three considered languages.
   The built-in standard libraries of Python and C++ are richer than
@@
   What makes correct use of external libraries more difficult with Fortran
-    is the lack of standard exception handling mechanisms in Fortran,
+    is the lack of standard exception handling mechanism, a feature
     long and {\em much requested by the numerical community} \citep[][Foreword]{Press_et_al_1996}.
@@
   The CPython and PyPy implementations of Python do not offer any
-    built-in solution for multithreading.
+    built-in solution for multi-threading.
 
@@
     represent the mathematical abstractions within the program code. 
-  The readability and brevity of such code:
+  This creates the potential to improve code readability and brevity,
   \begin{itemize}
-    \item{contribute to its 
+    \item{contributing to its 
       auditability, indispensable for credible and reproducible research in computational science 
-      \citep{Post_et_al_2005, Merali_et_al_2010, Stodden_et_al_2012};
-    }
-    \item{help to keep the programs maintainable, avoiding accumulation of the code 
+      \citep{Post_et_al_2005, Merali_et_al_2010, Stodden_et_al_2012}; and
+    }
+    \item{helping to keep the programs maintainable and avoiding accumulation of the code 
       debt\footnote{See \citet{Buschmann_2011} for discussion of technical/code debt.} 
@@
   \end{itemize}
-
+  \noindent 
   The performance evaluation revealed that:
@@
       \item{C++ and PyPy set-ups offered significantly smaller memory consumption 
-        than Fortran and CPython,}
+        than Fortran and CPython for larger domains,}
       \item{the PyPy set-up was roughly twice slower than C++ and up to twice faster than CPython.}
     \end{itemize}
-  The three equally-structured implementations required ca. 200, 300, and 550 lines of code %TODO update line numbers
+  The three equally-structured implementations required ca. 200, 300, and 500 lines of code 
     in Python, C++ and Fortran, respectively.
@@
     \noindent
-    Tobias Burnus, Ond\v{r}ej \v{C}ert\'ik, Julian Cummings, Patrik Jonsson,
-      Arjen Markus, Zbigniew Piotrowski, Davide del Vento, Janus Weil and Andrzej Wyszogrodzki 
+    Tobias Burnus, Julian Cummings, Ond\v{r}ej \v{C}ert\'ik, Patrik Jonsson,
+      Arjen Markus, Zbigniew Piotrowski, Davide del Vento and Janus Weil 
       provided valuable feedback to the initial version of the
-      manuscript and/or responces to questions posted to Blitz++ and gfortran mailing lists.
+      manuscript and/or responses to questions posted to Blitz++ and gfortran mailing lists.
 
@@
 
+  \clearpage
   \appendix
@@
   \subsection*{{\bf Usage example} (cf. Sect. \ref{sec:example})}
-  \ldots
   \codepyt{code/pyt/test.py}{listing19}{listing20}{1}
-  \ldots
 
@@
   \subsection*{{\bf Usage example} (cf. Sect. \ref{sec:example})}
-  \ldots
   \codefor{code/for/test.f}{listing23}{listing24}{1}
-  \ldots
   \codefor{code/for/test.f}{listing25}{listing26}{1}
-  \ldots
 

### 2013-03-19 paper/pyt/listings.py
README update, CMakeLists.txt cleanup, listings.py fix

@@ -174,7 +174,7 @@ def mpdata_B(d, psi, i, j):
     psi[pi(d, i+one, j-one)] + psi[pi(d, i, j-one)]
   ) / 2
 #listing16
-def C_bar(d, C, i, j):
+def mpdata_C_bar(d, C, i, j):
   return (
     C[pi(d, i+one, j+hlf)] + C[pi(d, i,  j+hlf)] +
     C[pi(d, i+one, j-hlf)] + C[pi(d, i,  j-hlf)] 

### 2013-03-19 paper/doc/code-cpp-listings.hpp.tex
last changes before arXiv resubmission

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{//listing00}
\PY{c+c1}{// code licensed under the terms of GNU GPL v3}
\PY{c+c1}{// copyright holder: University of Warsaw}
\PY{c+c1}{//listing01}
\PY{k}{typedef} \PY{k+kt}{double} \PY{n}{real\PYZus{}t}\PY{p}{;}
\PY{c+c1}{//listing02}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}blitz}\PY{c+cp}{/}\PY{c+cp}{array.h\PYZgt{}}
\PY{k}{using} \PY{n}{arr\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Array}\PY{o}{\PYZlt{}}\PY{n}{real\PYZus{}t}\PY{p}{,} \PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
\PY{k}{using} \PY{n}{rng\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Range}\PY{p}{;}
\PY{k}{using} \PY{n}{idx\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{RectDomain}\PY{o}{\PYZlt{}}\PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
\PY{c+c1}{//listing03}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{define return\PYZus{}macro(expr) \PYZbs{}}
\PY{c+cp}{  -\PYZgt{} decltype(safeToReturn(expr)) \PYZbs{}}
\PY{c+cp}{\PYZob{} return safeToReturn(expr); \PYZcb{} }
\PY{c+c1}{//listing04}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}boost}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}container}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}vector.hpp\PYZgt{}}
\PY{k}{struct} \PY{n}{arrvec\PYZus{}t} \PY{o}{:} \PY{n}{boost}\PY{o}{:}\PY{o}{:}\PY{n}{ptr\PYZus{}vector}\PY{o}{\PYZlt{}}\PY{n}{arr\PYZus{}t}\PY{o}{\PYZgt{}} 
\PY{p}{\PYZob{}}
  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{k}{operator}\PY{p}{[}\PY{p}{]}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{i}\PY{p}{)} \PY{k}{const} 
  \PY{p}{\PYZob{}}   
    \PY{k}{return} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{at}\PY{p}{(}\PY{p}{(}\PY{n}{i} \PY{o}{+} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{;} 
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing05}
\PY{k}{struct} \PY{n}{hlf\PYZus{}t} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}} \PY{n}{h}\PY{p}{;}

\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{+}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)} 
\PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{i}\PY{p}{;} 
\PY{p}{\PYZcb{}}

\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{-}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)} 
\PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{;} 
\PY{p}{\PYZcb{}}
\PY{c+c1}{//listing06}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{n\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{n}{ext}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{r}\PY{p}{,} \PY{k}{const} \PY{n}{n\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{n}\PY{p}{)} \PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{rng\PYZus{}t}\PY{p}{(}
    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
    \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}
  \PY{p}{)}\PY{p}{;} 
\PY{p}{\PYZcb{}} 
\PY{c+c1}{//listing07}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}} 
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}\PY{p}{;}

\PY{k}{template}\PY{o}{\PYZlt{}}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)} 
\PY{p}{\PYZob{}}
  \PY{k}{return} \PY{n}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
\PY{p}{\PYZcb{}}\PY{p}{;}

\PY{k}{template}\PY{o}{\PYZlt{}}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{)} 
\PY{p}{\PYZob{}}
  \PY{k}{return} \PY{n}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
\PY{p}{\PYZcb{}}\PY{p}{;} 
\PY{c+c1}{//listing08}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{arrvec\PYZus{}t} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{;}
  \PY{k+kt}{int} \PY{n}{n}\PY{p}{,} \PY{n}{hlo}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{;}
  \PY{n}{bcx\PYZus{}t} \PY{n}{bcx}\PY{p}{;}
  \PY{n}{bcy\PYZus{}t} \PY{n}{bcy}\PY{p}{;}

  \PY{c+c1}{// ctor}
  \PY{n}{solver}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{,} \PY{k+kt}{int} \PY{n}{hlo}\PY{p}{)} \PY{o}{:}
    \PY{n}{hlo}\PY{p}{(}\PY{n}{hlo}\PY{p}{)}\PY{p}{,}
    \PY{n}{n}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{,} 
    \PY{n}{i}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nx}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
    \PY{n}{j}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{ny}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}  
    \PY{n}{bcx}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} 
    \PY{n}{bcy}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{l} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{l} \PY{o}{\PYZlt{}} \PY{l+m+mi}{2}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{l}\PY{p}{)} 
      \PY{n}{psi}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
    \PY{n}{C}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{h}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
    \PY{n}{C}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
  \PY{p}{\PYZcb{}}

  \PY{c+c1}{// accessor methods}
  \PY{n}{arr\PYZus{}t} \PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{p}{\PYZob{}}
    \PY{k}{return} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{.}\PY{n}{reindex}\PY{p}{(}\PY{p}{\PYZob{}}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
  \PY{p}{\PYZcb{}}

  \PY{n}{arr\PYZus{}t} \PY{n}{courant}\PY{p}{(}\PY{k+kt}{int} \PY{n}{d}\PY{p}{)} 
  \PY{p}{\PYZob{}} 
    \PY{k}{return} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{;} 
  \PY{p}{\PYZcb{}}

  \PY{c+c1}{// helper methods invoked by solve()}
  \PY{k}{virtual} \PY{k+kt}{void} \PY{n}{advop}\PY{p}{(}\PY{p}{)} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;}

  \PY{k+kt}{void} \PY{n}{cycle}\PY{p}{(}\PY{p}{)} 
  \PY{p}{\PYZob{}} 
    \PY{n}{n} \PY{o}{=} \PY{p}{(}\PY{n}{n} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2} \PY{o}{-} \PY{l+m+mi}{2}\PY{p}{;} 
  \PY{p}{\PYZcb{}}

  \PY{c+c1}{// integration logic}
  \PY{k+kt}{void} \PY{n}{solve}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{nt}\PY{p}{)} 
  \PY{p}{\PYZob{}}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{t} \PY{o}{\PYZlt{}} \PY{n}{nt}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{t}\PY{p}{)} 
    \PY{p}{\PYZob{}}
      \PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{;}
      \PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{;}
      \PY{n}{advop}\PY{p}{(}\PY{p}{)}\PY{p}{;}
      \PY{n}{cycle}\PY{p}{(}\PY{p}{)}\PY{p}{;}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing09}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{cyclic}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{rght\PYZus{}halo}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{rght\PYZus{}edge}\PY{p}{;}\PY{p}{;}

  \PY{c+c1}{// ctor}
  \PY{n}{cyclic}\PY{p}{(}
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k+kt}{int} \PY{n}{hlo}
  \PY{p}{)} \PY{o}{:}
    \PY{n}{left\PYZus{}halo}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}
    \PY{n}{rght\PYZus{}edge}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{n}{hlo}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}  \PY{p}{)}\PY{p}{,}
    \PY{n}{rght\PYZus{}halo}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{n}{hlo}  \PY{p}{)}\PY{p}{,}
    \PY{n}{left\PYZus{}edge}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{n}{hlo}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}
  \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}

  \PY{c+c1}{// method invoked by the solver}
  \PY{k+kt}{void} \PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{a}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{=} \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{rght\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{;}     
    \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{rght\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{=} \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{;}     
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing10}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{T1}\PY{p}{,} \PY{k}{class} \PY{n+nc}{T2}\PY{p}{,} \PY{k}{class} \PY{n+nc}{T3}\PY{o}{\PYZgt{}} 
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{F}\PY{p}{(}
    \PY{k}{const} \PY{n}{T1} \PY{o}{\PYZam{}}\PY{n}{psi\PYZus{}l}\PY{p}{,} \PY{k}{const} \PY{n}{T2} \PY{o}{\PYZam{}}\PY{n}{psi\PYZus{}r}\PY{p}{,} \PY{k}{const} \PY{n}{T3} \PY{o}{\PYZam{}}\PY{n}{C}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{p}{(}
      \PY{p}{(}\PY{n}{C} \PY{o}{+} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}l} \PY{o}{+} 
      \PY{p}{(}\PY{n}{C} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
  \PY{p}{)}
\PY{c+c1}{//listing11}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}  
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{donorcell}\PY{p}{(} 
    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{n}{F}\PY{p}{(}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
        \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
    \PY{p}{)} \PY{o}{-}
    \PY{n}{F}\PY{p}{(}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
        \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
    \PY{p}{)}
  \PY{p}{)}
\PY{c+c1}{//listing12}
  \PY{k+kt}{void} \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{k+kt}{int} \PY{n}{n}\PY{p}{,}
    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{p}{\PYZob{}} 
    \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}
      \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
      \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}\PY{p}{;} 
  \PY{p}{\PYZcb{}}
\PY{c+c1}{//listing13}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver\PYZus{}donorcell} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}} 
\PY{p}{\PYZob{}}
  \PY{n}{solver\PYZus{}donorcell}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:}
    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
  \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}

  \PY{k+kt}{void} \PY{n}{advop}\PY{p}{(}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
      \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} 
      \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
    \PY{p}{)}\PY{p}{;}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing14}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{nom\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{den\PYZus{}t}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
    \PY{k}{const} \PY{n}{nom\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{nom}\PY{p}{,} \PY{k}{const} \PY{n}{den\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{den}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{n}{where}\PY{p}{(}\PY{n}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nom} \PY{o}{/} \PY{n}{den}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}
  \PY{p}{)} 
\PY{c+c1}{//listing15}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}
    \PY{p}{)} 
  \PY{p}{)} 
\PY{c+c1}{//listing16}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
   \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{,}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
  \PY{p}{)}
\PY{c+c1}{//listing17}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}
    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{p}{(}
      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+}
      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)} 
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}
  \PY{p}{)}
\PY{c+c1}{//listing18}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}
    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,}
    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)} 
    \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{)} 
    \PY{o}{*} \PY{n}{mpdata\PYZus{}A}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} 
    \PY{o}{-} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} 
    \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
    \PY{o}{*} \PY{n}{mpdata\PYZus{}B}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
  \PY{p}{)} 
\PY{c+c1}{//listing19}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{n\PYZus{}iters}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver\PYZus{}mpdata} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{arrvec\PYZus{}t} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{im}\PY{p}{,} \PY{n}{jm}\PY{p}{;}

  \PY{c+c1}{// ctor}
  \PY{n}{solver\PYZus{}mpdata}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:} 
    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
    \PY{n}{im}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{,}
    \PY{n}{jm}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{k+kt}{int} \PY{n}{n\PYZus{}tmp} \PY{o}{=} \PY{n}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{2} \PY{o}{?} \PY{l+m+mi}{2} \PY{o}{:} \PY{l+m+mi}{1}\PY{p}{;}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{n} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{n} \PY{o}{\PYZlt{}} \PY{n}{n\PYZus{}tmp}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{n}\PY{p}{)}
    \PY{p}{\PYZob{}}
      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
      \PY{p}{)}\PY{p}{;}
      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
      \PY{p}{)}\PY{p}{;}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}

  \PY{c+c1}{// method invoked by the solver}
  \PY{k+kt}{void} \PY{n}{advop}\PY{p}{(}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{step} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{step} \PY{o}{\PYZlt{}} \PY{n}{n\PYZus{}iters}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{step}\PY{p}{)} 
    \PY{p}{\PYZob{}}
      \PY{k}{if} \PY{p}{(}\PY{n}{step} \PY{o}{=}\PY{o}{=} \PY{l+m+mi}{0}\PY{p}{)} 
        \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
        \PY{p}{)}\PY{p}{;}
      \PY{k}{else}
      \PY{p}{\PYZob{}}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{cycle}\PY{p}{(}\PY{p}{)}\PY{p}{;}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{hlo}\PY{p}{)}
        \PY{p}{)}\PY{p}{;}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{hlo}\PY{p}{)}
        \PY{p}{)}\PY{p}{;}

        \PY{c+c1}{// choosing input/output for antidiff C}
        \PY{k}{const} \PY{n}{arrvec\PYZus{}t} 
          \PY{o}{\PYZam{}}\PY{n}{C\PYZus{}unco} \PY{o}{=} \PY{p}{(}\PY{n}{step} \PY{o}{=}\PY{o}{=} \PY{l+m+mi}{1}\PY{p}{)} 
            \PY{o}{?} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C} 
            \PY{o}{:} \PY{p}{(}\PY{n}{step} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2}\PY{p}{)} 
              \PY{o}{?} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}  \PY{c+c1}{// odd steps}
              \PY{o}{:} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{c+c1}{// even steps}
          \PY{o}{\PYZam{}}\PY{n}{C\PYZus{}corr} \PY{o}{=} \PY{p}{(}\PY{n}{step}  \PY{o}{\PYZpc{}} \PY{l+m+mi}{2}\PY{p}{)} 
            \PY{o}{?} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}    \PY{c+c1}{// odd steps}
            \PY{o}{:} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{;}   \PY{c+c1}{// even steps}

        \PY{c+c1}{// calculating the antidiffusive C }
        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{(}\PY{n}{im}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{im}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,} \PY{n}{C\PYZus{}unco}
        \PY{p}{)}\PY{p}{;}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,}\PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{;}

        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{n}{jm}\PY{o}{+}\PY{n}{h}\PY{p}{)} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{jm}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{n}{C\PYZus{}unco}
        \PY{p}{)}\PY{p}{;}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,}\PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{;}

        \PY{c+c1}{// donor-cell step }
        \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{n}{C\PYZus{}corr}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
        \PY{p}{)}\PY{p}{;}
      \PY{p}{\PYZcb{}}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing20}
\end{Verbatim}

### 2013-03-19 paper/doc/code-for-listings.f.tex
last changes before arXiv resubmission

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c}{!listing00}
\PY{c}{! code licensed under the terms of GNU GPL v3}
\PY{c}{! copyright holder: University of Warsaw}
\PY{c}{!listing01}
\PY{k}{module }\PY{n+nv}{real\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}
\PY{k}{  }\PY{k+kt}{integer}\PY{p}{,} \PY{k}{parameter} \PY{k+kd}{::} \PY{n+nv}{real\PYZus{}t} \PY{o}{=} \PY{n+nb}{kind}\PY{p}{(}\PY{l+m+mf}{0.}\PY{n+nv}{d0}\PY{p}{)} 
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing02}
\PY{k}{module }\PY{n+nv}{arrvec\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{real\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}

\PY{k}{  }\PY{k}{type} \PY{k+kd}{::} \PY{n+nv}{arr\PYZus{}t}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{a}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
  \PY{k}{end }\PY{k}{type}

\PY{k}{  }\PY{k}{type} \PY{k+kd}{::} \PY{n+nv}{arrptr\PYZus{}t}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arr\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{p}
  \PY{k}{end }\PY{k}{type}

\PY{k}{  }\PY{k}{type} \PY{k+kd}{::} \PY{n+nv}{arrvec\PYZus{}t}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arr\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{arrs}\PY{p}{(}\PY{p}{:}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrptr\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{at}\PY{p}{(}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{length}
    \PY{k}{contains}
\PY{k}{    }\PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{ctor} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{arrvec\PYZus{}ctor}
    \PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{init} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{arrvec\PYZus{}init}
  \PY{k}{end }\PY{k}{type}

\PY{k}{  }\PY{k}{contains}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{arrvec\PYZus{}ctor}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{n}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n}

    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{length} \PY{o}{=} \PY{n+nv}{n}
    \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{o}{-}\PY{n+nv}{n} \PY{p}{:} \PY{n+nv}{n}\PY{o}{-}\PY{l+m+mi}{1} \PY{p}{)}\PY{p}{)}
    \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(} \PY{l+m+mi}{0} \PY{p}{:} \PY{n+nv}{n}\PY{o}{-}\PY{l+m+mi}{1} \PY{p}{)}\PY{p}{)}
  \PY{k}{end }\PY{k}{subroutine}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{arrvec\PYZus{}init}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}

    \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{(} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{p}{:} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{p}{:} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{p}{)}\PY{p}{)}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n} \PY{o}{-} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{length}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}
  \PY{k}{end }\PY{k}{subroutine}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing03}
\PY{k}{module }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}

\PY{k}{  }\PY{k}{type} \PY{k+kd}{::} \PY{n+nv}{half\PYZus{}t}
  \PY{k}{end }\PY{k}{type}

\PY{k}{  }\PY{k}{type}\PY{p}{(}\PY{n+nv}{half\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{h}

  \PY{k}{interface }\PY{n+nv}{operator} \PY{p}{(}\PY{o}{+}\PY{p}{)}
    \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{ph}
  \PY{k}{end }\PY{k}{interface}

\PY{k}{  }\PY{k}{interface }\PY{n+nv}{operator} \PY{p}{(}\PY{o}{-}\PY{p}{)}
    \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{mh}
  \PY{k}{end }\PY{k}{interface}

\PY{k}{  }\PY{k}{contains}

\PY{k}{  }\PY{k}{elemental }\PY{k}{function }\PY{n+nv}{ph}\PY{p}{(}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}
    \PY{k}{type}\PY{p}{(}\PY{n+nv}{half\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{h}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}
\PY{k}{    }\PY{k}{return} \PY{o}{=} \PY{n+nv}{i} 
  \PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{elemental }\PY{k}{function }\PY{n+nv}{mh}\PY{p}{(}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i} 
    \PY{k}{type}\PY{p}{(}\PY{n+nv}{half\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{h}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}
\PY{k}{    }\PY{k}{return} \PY{o}{=} \PY{n+nv}{i} \PY{o}{-} \PY{l+m+mi}{1}
  \PY{k}{end }\PY{k}{function}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing04}
\PY{k}{module }\PY{n+nv}{halo\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}

\PY{k}{  }\PY{k}{interface }\PY{n+nv}{ext}
    \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{ext\PYZus{}n}
    \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{ext\PYZus{}h}
  \PY{k}{end }\PY{k}{interface }

\PY{k}{  }\PY{k}{contains}

\PY{k}{  }\PY{k}{function }\PY{n+nv}{ext\PYZus{}n}\PY{p}{(}\PY{n+nv}{r}\PY{p}{,} \PY{n+nv}{n}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}

    \PY{k}{return} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{+} \PY{n+nv}{n} \PY{o}{/}\PY{p}{)}
  \PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{function }\PY{n+nv}{ext\PYZus{}h}\PY{p}{(}\PY{n+nv}{r}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k}{type}\PY{p}{(}\PY{n+nv}{half\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{h}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}

    \PY{k}{return} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{+} \PY{n+nv}{h} \PY{o}{/}\PY{p}{)}
  \PY{k}{end }\PY{k}{function}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing05}
\PY{k}{module }\PY{n+nv}{pi\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{real\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}
\PY{k}{  }\PY{k}{contains}
\PY{k}{  }\PY{k}{function }\PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{arr}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result}\PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{arr}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k}{select }\PY{k}{case} \PY{p}{(}\PY{n+nv}{d}\PY{p}{)} 
      \PY{k}{case} \PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)} 
        \PY{k}{return} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{arr}\PY{p}{(} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{p}{:} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{p}{:} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{p}{)}   
      \PY{k}{case} \PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} 
        \PY{k}{return} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{arr}\PY{p}{(} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{p}{:} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{p}{:} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{p}{)}   
    \PY{k}{end }\PY{k}{select}
\PY{k}{  }\PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{pure }\PY{k}{function }\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result}\PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}
\PY{k}{    }\PY{k}{select }\PY{k}{case} \PY{p}{(}\PY{n+nv}{d}\PY{p}{)}
      \PY{k}{case} \PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}
        \PY{k}{return} \PY{o}{=} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{-} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
      \PY{k}{case} \PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
        \PY{k}{return} \PY{o}{=} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{-} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
    \PY{k}{end }\PY{k}{select}
\PY{k}{  }\PY{k}{end }\PY{k}{function}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing06}
\PY{k}{module }\PY{n+nv}{bcd\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{arrvec\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}

\PY{k}{  }\PY{k}{type}\PY{p}{,} \PY{k}{abstract} \PY{k+kd}{::} \PY{n+nv}{bcd\PYZus{}t}
    \PY{k}{contains}
\PY{k}{    }\PY{n+nv}{procedure}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}fill\PYZus{}halos}\PY{p}{)}\PY{p}{,} \PY{k}{deferred} \PY{k+kd}{::} \PY{n+nv}{fill\PYZus{}halos}
    \PY{n+nv}{procedure}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}init}\PY{p}{)}\PY{p}{,} \PY{k}{deferred} \PY{k+kd}{::} \PY{n+nv}{init}
  \PY{k}{end }\PY{k}{type}
\PY{k}{ }
\PY{k}{  }\PY{k}{abstract }\PY{k}{interface }
\PY{k}{    }\PY{k}{subroutine }\PY{n+nv}{bcd\PYZus{}fill\PYZus{}halos}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
      \PY{k}{import} \PY{k+kd}{::} \PY{n+nv}{bcd\PYZus{}t}\PY{p}{,} \PY{n+nv}{real\PYZus{}t}
      \PY{k}{class}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}t} \PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
      \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{a}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)} 
      \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k}{end }\PY{k}{subroutine}

\PY{k}{    }\PY{k}{subroutine }\PY{n+nv}{bcd\PYZus{}init}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}
      \PY{k}{import} \PY{k+kd}{::} \PY{n+nv}{bcd\PYZus{}t}
      \PY{k}{class}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
      \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{hlo}
    \PY{k}{end }\PY{k}{subroutine}
\PY{k}{  }\PY{k}{end }\PY{k}{interface}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing07}
\PY{k}{module }\PY{n+nv}{solver\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{arrvec\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{bcd\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{halo\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}

\PY{k}{  }\PY{k}{type}\PY{p}{,} \PY{k}{abstract} \PY{k+kd}{::} \PY{n+nv}{solver\PYZus{}t}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{C}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{hlo}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} 
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}
    \PY{k}{contains}
\PY{k}{    }\PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{solve}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{solver\PYZus{}solve}
    \PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{state}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{solver\PYZus{}state}
    \PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{courant} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{solver\PYZus{}courant}
    \PY{n+nv}{procedure} \PY{k+kd}{::} \PY{k}{cycle}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{solver\PYZus{}cycle}
    \PY{n+nv}{procedure}\PY{p}{(}\PY{n+nv}{solver\PYZus{}advop}\PY{p}{)}\PY{p}{,} \PY{k}{deferred} \PY{k+kd}{::} \PY{n+nv}{advop}
  \PY{k}{end }\PY{k}{type }

\PY{k}{  }\PY{k}{abstract }\PY{k}{interface}
\PY{k}{    }\PY{k}{subroutine }\PY{n+nv}{solver\PYZus{}advop}\PY{p}{(}\PY{n+nv}{this}\PY{p}{)}
      \PY{k}{import }\PY{n+nv}{solver\PYZus{}t}
      \PY{k}{class}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k}{end }\PY{k}{subroutine}
\PY{k}{  }\PY{k}{end }\PY{k}{interface}

\PY{k}{  }\PY{k}{contains}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{solver\PYZus{}ctor}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}
    \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
    \PY{k}{use }\PY{n+nv}{halo\PYZus{}m}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{,} \PY{n+nv}{hlo}

    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n} \PY{o}{=} \PY{l+m+mi}{0}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{hlo} \PY{o}{=} \PY{n+nv}{hlo}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{bcx} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{bcx}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{bcy} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{bcy}

    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{nx} \PY{o}{-} \PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{ny} \PY{o}{-} \PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}

    \PY{k}{call }\PY{n+nv}{bcx}\PY{p}{\PYZpc{}}\PY{n+nv}{init}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}
    \PY{k}{call }\PY{n+nv}{bcy}\PY{p}{\PYZpc{}}\PY{n+nv}{init}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}

    \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{psi}\PY{p}{)}
    \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{ctor}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k}{block}
\PY{k}{      }\PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{n}
      \PY{k}{do }\PY{n+nv}{n}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{1}
        \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{init}\PY{p}{(}                            \PY{p}{\PYZam{}}
          \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}        \PY{p}{\PYZam{}}
        \PY{p}{)}
      \PY{k}{end }\PY{k}{do}
\PY{k}{    }\PY{k}{end }\PY{k}{block}

\PY{k}{    }\PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{C}\PY{p}{)}
    \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{ctor}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{init}\PY{p}{(}                                  \PY{p}{\PYZam{}}
      \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}              \PY{p}{\PYZam{}}
    \PY{p}{)}
    \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{init}\PY{p}{(}                                  \PY{p}{\PYZam{}}
      \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)}              \PY{p}{\PYZam{}}
    \PY{p}{)}
  \PY{k}{end }\PY{k}{subroutine}

\PY{k}{  }\PY{k}{function }\PY{n+nv}{solver\PYZus{}state}\PY{p}{(}\PY{n+nv}{this}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k}{return} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{(}                 \PY{p}{\PYZam{}}
      \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{p}{:} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,}                           \PY{p}{\PYZam{}}
      \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{p}{:} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}                            \PY{p}{\PYZam{}}
    \PY{p}{)}
  \PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{function }\PY{n+nv}{solver\PYZus{}courant}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{d}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k}{return} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}
  \PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{solver\PYZus{}cycle}\PY{p}{(}\PY{n+nv}{this}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n} \PY{o}{=} \PY{n+nb}{mod}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n} \PY{o}{+} \PY{l+m+mi}{1} \PY{o}{+} \PY{l+m+mi}{2}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{2}
  \PY{k}{end }\PY{k}{subroutine}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{solver\PYZus{}solve}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{nt}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{nt}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{t}

    \PY{k}{do }\PY{n+nv}{t} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{nt}\PY{o}{-}\PY{l+m+mi}{1} 
      \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{bcx}\PY{p}{\PYZpc{}}\PY{n+nv}{fill\PYZus{}halos}\PY{p}{(}                        \PY{p}{\PYZam{}}
        \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{hlo}\PY{p}{)} \PY{p}{\PYZam{}}
      \PY{p}{)}
      \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{bcy}\PY{p}{\PYZpc{}}\PY{n+nv}{fill\PYZus{}halos}\PY{p}{(}                        \PY{p}{\PYZam{}}
        \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{hlo}\PY{p}{)} \PY{p}{\PYZam{}}
      \PY{p}{)}
      \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{advop}\PY{p}{(}\PY{p}{)}
      \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{k}{cycle}\PY{p}{(}\PY{p}{)}
    \PY{k}{end }\PY{k}{do}
\PY{k}{  }\PY{k}{end }\PY{k}{subroutine}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing08}
\PY{k}{module }\PY{n+nv}{cyclic\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{bcd\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{pi\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}
\PY{k}{  }
\PY{k}{  }\PY{k}{type}\PY{p}{,} \PY{n+nv}{extends}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{cyclic\PYZus{}t}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{left\PYZus{}halo}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{rght\PYZus{}halo}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} 
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{left\PYZus{}edge}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{rght\PYZus{}edge}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} 
    \PY{k}{contains}
\PY{k}{    }\PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{init} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{cyclic\PYZus{}init}
    \PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{fill\PYZus{}halos} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{cyclic\PYZus{}fill\PYZus{}halos}
  \PY{k}{end }\PY{k}{type}

\PY{k}{  }\PY{k}{contains}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{cyclic\PYZus{}init}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{cyclic\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{hlo}

    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{d} \PY{o}{=} \PY{n+nv}{d}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{left\PYZus{}halo} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{o}{-}\PY{n+nv}{hlo}\PY{p}{,} \PY{o}{-}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)} 
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{rght\PYZus{}halo} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{n}\PY{o}{-}\PY{l+m+mi}{1}\PY{o}{+}\PY{n+nv}{hlo} \PY{o}{/}\PY{p}{)} 
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{left\PYZus{}edge} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{hlo}\PY{o}{-}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{rght\PYZus{}edge} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{n}\PY{o}{-}\PY{n+nv}{hlo}\PY{p}{,} \PY{n+nv}{n}\PY{o}{-}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
  \PY{k}{end }\PY{k}{subroutine}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{cyclic\PYZus{}fill\PYZus{}halos}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{cyclic\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{ptr}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{a}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{left\PYZus{}halo}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} 
    \PY{n+nv}{ptr} \PY{o}{=}  \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{rght\PYZus{}edge}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{rght\PYZus{}halo}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} 
    \PY{n+nv}{ptr} \PY{o}{=}  \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{left\PYZus{}edge}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
  \PY{k}{end }\PY{k}{subroutine}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing09}
\PY{k}{module }\PY{n+nv}{donorcell\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{real\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{pi\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{arrvec\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}
\PY{k}{  }\PY{k}{contains} 
\PY{c}{!listing10}
  \PY{k}{elemental }\PY{k}{function }\PY{n+nv}{F}\PY{p}{(}\PY{n+nv}{psi\PYZus{}l}\PY{p}{,} \PY{n+nv}{psi\PYZus{}r}\PY{p}{,} \PY{n+nv}{C}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}
\PY{k}{    }\PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi\PYZus{}l}\PY{p}{,} \PY{n+nv}{psi\PYZus{}r}\PY{p}{,} \PY{n+nv}{C}
    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
      \PY{p}{(}\PY{n+nv}{C} \PY{o}{+} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n+nv}{psi\PYZus{}l} \PY{o}{+}                           \PY{p}{\PYZam{}}
      \PY{p}{(}\PY{n+nv}{C} \PY{o}{-} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n+nv}{psi\PYZus{}r}                             \PY{p}{\PYZam{}}
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
  \PY{k}{end }\PY{k}{function}
\PY{c}{!listing11}
  \PY{k}{function }\PY{n+nv}{donorcell}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} 
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}\PY{p}{,} \PY{n+nv}{C}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}           
    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
      \PY{n+nv}{F}\PY{p}{(}                                               \PY{p}{\PYZam{}}
        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,}   \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                             \PY{p}{\PYZam{}}
      \PY{p}{)} \PY{o}{-}                                              \PY{p}{\PYZam{}}
      \PY{n+nv}{F}\PY{p}{(}                                               \PY{p}{\PYZam{}}
        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,}   \PY{n+nv}{i}\PY{o}{-}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                             \PY{p}{\PYZam{}}
      \PY{p}{)}                                                \PY{p}{\PYZam{}}
    \PY{p}{)}
  \PY{k}{end }\PY{k}{function}
\PY{c}{!listing12}
  \PY{k}{subroutine }\PY{n+nv}{donorcell\PYZus{}op}\PY{p}{(}\PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}  
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{psi}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}

    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{ptr}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{pi}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
    \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{pi}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                   \PY{p}{\PYZam{}}
      \PY{o}{-} \PY{n+nv}{donorcell}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{p}{\PYZam{}}
      \PY{o}{-} \PY{n+nv}{donorcell}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}
  \PY{k}{end }\PY{k}{subroutine}
\PY{c}{!listing13}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing14}
\PY{k}{module }\PY{n+nv}{solver\PYZus{}donorcell\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{donorcell\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{solver\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}
\PY{k}{  }
\PY{k}{  }\PY{k}{type}\PY{p}{,} \PY{n+nv}{extends}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{donorcell\PYZus{}t}
    \PY{k}{contains}
\PY{k}{    }\PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{ctor} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{donorcell\PYZus{}ctor}
    \PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{advop} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{donorcell\PYZus{}advop}
  \PY{k}{end }\PY{k}{type}

\PY{k}{  }\PY{k}{contains}
\PY{k}{  }
\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{donorcell\PYZus{}ctor}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{donorcell\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}
    \PY{k}{call }\PY{n+nv}{solver\PYZus{}ctor}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{bcx}\PY{p}{,}\PY{n+nv}{bcy}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,}\PY{n+nv}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
  \PY{k}{end }\PY{k}{subroutine}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{donorcell\PYZus{}advop}\PY{p}{(}\PY{n+nv}{this}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{donorcell\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C}
    \PY{n+nv}{C} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{C}
    \PY{k}{call }\PY{n+nv}{donorcell\PYZus{}op}\PY{p}{(}                                 \PY{p}{\PYZam{}}
      \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}              \PY{p}{\PYZam{}}
    \PY{p}{)}
  \PY{k}{end }\PY{k}{subroutine}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing15}
\PY{k}{module }\PY{n+nv}{mpdata\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{arrvec\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{pi\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}
\PY{k}{  }\PY{k}{contains} 
\PY{c}{!listing16}
  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{n+nv}{den}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{nom}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}\PY{p}{,} \PY{n+nv}{den}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{size}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} \PY{n+nv}{size}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{)}
    \PY{k}{where} \PY{p}{(}\PY{n+nv}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{)}
      \PY{k}{return} \PY{o}{=} \PY{n+nv}{nom} \PY{o}{/} \PY{n+nv}{den}
    \PY{n+nv}{elsewhere}
      \PY{k}{return} \PY{o}{=} \PY{l+m+mi}{0}
    \PY{k}{end }\PY{k}{where}
\PY{k}{  }\PY{k}{end }\PY{k}{function}
\PY{c}{!listing17}
  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}A}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
    \PY{k}{return} \PY{o}{=} \PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}                              \PY{p}{\PYZam{}}
      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{o}{-} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}           \PY{p}{\PYZam{}}
      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}            \PY{p}{\PYZam{}}
    \PY{p}{)}  
  \PY{k}{end }\PY{k}{function}
\PY{c}{!listing18}
  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}B}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)} 
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
    \PY{k}{return} \PY{o}{=} \PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}                              \PY{p}{\PYZam{}}
      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
    \PY{o}{-} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}     \PY{p}{\PYZam{}}
      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
    \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
  \PY{k}{end }\PY{k}{function}
\PY{c}{!listing19}
  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{C}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)} 
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}

    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+}        \PY{p}{\PYZam{}}
      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{-}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{-}\PY{n+nv}{h}\PY{p}{)}          \PY{p}{\PYZam{}}
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}               
  \PY{k}{end }\PY{k}{function}
\PY{c}{!listing20}
  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{C}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)} 
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C}
    \PY{k}{return} \PY{o}{=}                                           \PY{p}{\PYZam{}}
      \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{)}                  \PY{p}{\PYZam{}}
      \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}          \PY{p}{\PYZam{}}
      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}A}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                         \PY{p}{\PYZam{}}
      \PY{o}{-} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                     \PY{p}{\PYZam{}}
      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}           \PY{p}{\PYZam{}}
      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}B}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
  \PY{k}{end }\PY{k}{function}
\PY{c}{!listing21}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing22}
\PY{k}{module }\PY{n+nv}{solver\PYZus{}mpdata\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{solver\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{mpdata\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{donorcell\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{halo\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}
\PY{k}{  }
\PY{k}{  }\PY{k}{type}\PY{p}{,} \PY{n+nv}{extends}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{mpdata\PYZus{}t}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{n\PYZus{}iters}\PY{p}{,} \PY{n+nv}{n\PYZus{}tmp}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{im}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{jm}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{tmp}\PY{p}{(}\PY{p}{:}\PY{p}{)} 
    \PY{k}{contains}
\PY{k}{    }\PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{ctor} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{mpdata\PYZus{}ctor}
    \PY{n+nv}{procedure} \PY{k+kd}{::} \PY{n+nv}{advop} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{mpdata\PYZus{}advop}
  \PY{k}{end }\PY{k}{type}

\PY{k}{  }\PY{k}{contains}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{mpdata\PYZus{}ctor}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{n\PYZus{}iters}\PY{p}{,} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{mpdata\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{bcd\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}
    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n\PYZus{}iters}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{c}

    \PY{k}{call }\PY{n+nv}{solver\PYZus{}ctor}\PY{p}{(}\PY{n+nv}{this}\PY{p}{,} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}

    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}iters} \PY{o}{=} \PY{n+nv}{n\PYZus{}iters}
    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}tmp} \PY{o}{=} \PY{n+nb}{min}\PY{p}{(}\PY{n+nv}{n\PYZus{}iters} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}
    \PY{k}{if} \PY{p}{(}\PY{n+nv}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{)} \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{:}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}tmp}\PY{p}{)}\PY{p}{)}

    \PY{n+nv}{associate} \PY{p}{(}\PY{n+nv}{i} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{hlo} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{hlo}\PY{p}{)}
      \PY{k}{do }\PY{n+nv}{c}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}tmp} \PY{o}{-} \PY{l+m+mi}{1}
        \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{n+nv}{c}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{ctor}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
        \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{n+nv}{c}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{init}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}\PY{p}{)}
        \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{n+nv}{c}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{init}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)}\PY{p}{)}
      \PY{k}{end }\PY{k}{do}

\PY{k}{      }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{im} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{/}\PY{p}{)}
      \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{jm} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{/}\PY{p}{)}
    \PY{k}{end }\PY{n+nv}{associate}
  \PY{k}{end }\PY{k}{subroutine}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{mpdata\PYZus{}advop}\PY{p}{(}\PY{n+nv}{this}\PY{p}{)}
    \PY{k}{class}\PY{p}{(}\PY{n+nv}{mpdata\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{this}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{step}

    \PY{n+nv}{associate} \PY{p}{(}\PY{n+nv}{i} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{im} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{im}\PY{p}{,}\PY{p}{\PYZam{}}
      \PY{n+nv}{jm} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{jm}\PY{p}{,} \PY{n+nv}{psi} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{n} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n}\PY{p}{,}     \PY{p}{\PYZam{}}
      \PY{n+nv}{hlo} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{hlo}\PY{p}{,} \PY{n+nv}{bcx} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{bcy}\PY{p}{\PYZam{}}
    \PY{p}{)}
      \PY{k}{do }\PY{n+nv}{step}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}iters}\PY{o}{-}\PY{l+m+mi}{1}
        \PY{k}{if} \PY{p}{(}\PY{n+nv}{step} \PY{o}{==} \PY{l+m+mi}{0}\PY{p}{)} \PY{k}{then}
\PY{k}{          }\PY{k}{block}
\PY{k}{            }\PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C}
            \PY{n+nv}{C} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{C}
            \PY{k}{call }\PY{n+nv}{donorcell\PYZus{}op}\PY{p}{(}\PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
          \PY{k}{end }\PY{k}{block}
\PY{k}{        }\PY{k}{else}
\PY{k}{          }\PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{k}{cycle}\PY{p}{(}\PY{p}{)}
          \PY{k}{call }\PY{n+nv}{bcx}\PY{p}{\PYZpc{}}\PY{n+nv}{fill\PYZus{}halos}\PY{p}{(}                         \PY{p}{\PYZam{}}
            \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{n+nv}{n} \PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}               \PY{p}{\PYZam{}}
          \PY{p}{)}
          \PY{k}{call }\PY{n+nv}{bcy}\PY{p}{\PYZpc{}}\PY{n+nv}{fill\PYZus{}halos}\PY{p}{(}                         \PY{p}{\PYZam{}}
            \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{n+nv}{n} \PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{hlo}\PY{p}{)}               \PY{p}{\PYZam{}}
          \PY{p}{)}

          \PY{k}{block}
\PY{k}{            }\PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C\PYZus{}corr}\PY{p}{,} \PY{n+nv}{C\PYZus{}unco}
            \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{ptr}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}

            \PY{c}{! chosing input/output for antidiff. C}
            \PY{k}{if} \PY{p}{(}\PY{n+nv}{step} \PY{o}{==} \PY{l+m+mi}{1}\PY{p}{)} \PY{k}{then}
\PY{k}{              }\PY{n+nv}{C\PYZus{}unco} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{C}
              \PY{n+nv}{C\PYZus{}corr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}
            \PY{k}{else }\PY{k}{if} \PY{p}{(}\PY{n+nb}{mod}\PY{p}{(}\PY{n+nv}{step}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)} \PY{o}{==} \PY{l+m+mi}{1}\PY{p}{)} \PY{k}{then}
\PY{k}{              }\PY{n+nv}{C\PYZus{}unco} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{c}{! odd step}
              \PY{n+nv}{C\PYZus{}corr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)} \PY{c}{! even step}
            \PY{k}{else}
\PY{k}{              }\PY{n+nv}{C\PYZus{}unco} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)} \PY{c}{! odd step}
              \PY{n+nv}{C\PYZus{}corr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{c}{! even step}
            \PY{k}{end }\PY{k}{if}

            \PY{c}{! calculating the antidiffusive velo}
            \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{pi}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{C\PYZus{}corr}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{l+m+mi}{0} \PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{im}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
            \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}                        \PY{p}{\PYZam{}}
              \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{n+nv}{n} \PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{im}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{C\PYZus{}unco}        \PY{p}{\PYZam{}}
            \PY{p}{)}      
            \PY{k}{call }\PY{n+nv}{bcy}\PY{p}{\PYZpc{}}\PY{n+nv}{fill\PYZus{}halos}\PY{p}{(}                       \PY{p}{\PYZam{}}
              \PY{n+nv}{C\PYZus{}corr}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)}              \PY{p}{\PYZam{}}
            \PY{p}{)}

            \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{pi}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{C\PYZus{}corr}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{l+m+mi}{1} \PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{jm}\PY{o}{+}\PY{n+nv}{h}\PY{p}{)}
            \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}                        \PY{p}{\PYZam{}}
              \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{n+nv}{n} \PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{jm}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{C\PYZus{}unco}        \PY{p}{\PYZam{}}
            \PY{p}{)}
            \PY{k}{call }\PY{n+nv}{bcx}\PY{p}{\PYZpc{}}\PY{n+nv}{fill\PYZus{}halos}\PY{p}{(}                       \PY{p}{\PYZam{}}
              \PY{n+nv}{C\PYZus{}corr}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{ext}\PY{p}{(}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)}              \PY{p}{\PYZam{}}
            \PY{p}{)}

            \PY{c}{! donor-cell step}
            \PY{k}{call }\PY{n+nv}{donorcell\PYZus{}op}\PY{p}{(}\PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{C\PYZus{}corr}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} 
          \PY{k}{end }\PY{k}{block}
\PY{k}{        }\PY{k}{end }\PY{k}{if}
\PY{k}{      }\PY{k}{end }\PY{k}{do}
\PY{k}{    }\PY{k}{end }\PY{n+nv}{associate}
  \PY{k}{end }\PY{k}{subroutine}
\PY{k}{end }\PY{k}{module}
\PY{c}{!listing23}
\end{Verbatim}

### 2013-03-19 paper/doc/code-for-test.f.tex
last changes before arXiv resubmission

\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{program }\PY{n+nv}{test}
  \PY{k}{use }\PY{n+nv}{solver\PYZus{}mpdata\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{cyclic\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}

\PY{k}{  }\PY{k}{if} \PY{p}{(}\PY{n+nb}{command\PYZus{}argument\PYZus{}count}\PY{p}{(}\PY{p}{)} \PY{o}{/}\PY{o}{=} \PY{p}{(}\PY{l+m+mi}{9}\PY{p}{)}\PY{p}{)} \PY{k}{then}
\PY{k}{    }\PY{k}{print}\PY{o}{*}\PY{p}{,} \PY{l+s+s2}{"expecting 9 arguments (nx, ny, Cx, Cy, nt, it, f.in, f.out, dec)"}
    \PY{k}{stop }\PY{l+m+mi}{1}
  \PY{k}{end }\PY{k}{if}

\PY{k}{  }\PY{k}{block}
\PY{c}{!listing23}
    \PY{k}{type}\PY{p}{(}\PY{n+nv}{mpdata\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{slv}
    \PY{k}{type}\PY{p}{(}\PY{n+nv}{cyclic\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{,} \PY{n+nv}{nt}\PY{p}{,} \PY{n+nv}{it}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{Cx}\PY{p}{,} \PY{n+nv}{Cy}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{ptr}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
\PY{c}{!listing24}
    \PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{l+m+mi}{666}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{fname}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{dec}

    \PY{n+nv}{nx} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
    \PY{n+nv}{ny} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{n+nv}{Cx} \PY{o}{=} \PY{n+nv}{arg2real}\PY{p}{(}\PY{l+m+mi}{3}\PY{p}{)}
    \PY{n+nv}{Cy} \PY{o}{=} \PY{n+nv}{arg2real}\PY{p}{(}\PY{l+m+mi}{4}\PY{p}{)}
    \PY{n+nv}{nt} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{5}\PY{p}{)}
    \PY{n+nv}{it} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{6}\PY{p}{)}
    \PY{n+nv}{dec} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{9}\PY{p}{)}
    \PY{k}{call }\PY{n+nb}{get\PYZus{}command\PYZus{}argument}\PY{p}{(}\PY{l+m+mi}{7}\PY{p}{,} \PY{n+nv}{fname}\PY{p}{)}

\PY{c}{!listing25}
    \PY{k}{call }\PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{ctor}\PY{p}{(}\PY{n+nv}{it}\PY{p}{,} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{)}

    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{state}\PY{p}{(}\PY{p}{)} 
    \PY{k}{call }\PY{n+nv}{read\PYZus{}file}\PY{p}{(}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{ptr}\PY{p}{)}

    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{courant}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)} 
    \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{Cx}

    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{courant}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} 
    \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{Cy}

    \PY{k}{call }\PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{solve}\PY{p}{(}\PY{n+nv}{nt}\PY{p}{)}
\PY{c}{!listing26}

    \PY{k}{block}
\PY{k}{      }\PY{k+kt}{logical} \PY{k+kd}{::} \PY{n+nv}{error} \PY{o}{=} \PY{n+nb}{.FALSE.}
      \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{tmp}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
      \PY{k+kt}{real} \PY{k+kd}{::} \PY{n+nv}{diff}
      \PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{l+m+mi}{666}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{fname}
      \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{(}\PY{n+nv}{nx}\PY{p}{,}\PY{n+nv}{ny}\PY{p}{)}\PY{p}{)}
      \PY{k}{call }\PY{n+nb}{get\PYZus{}command\PYZus{}argument}\PY{p}{(}\PY{l+m+mi}{8}\PY{p}{,} \PY{n+nv}{fname}\PY{p}{)}
      \PY{k}{call }\PY{n+nv}{read\PYZus{}file}\PY{p}{(}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{tmp}\PY{p}{)}
      \PY{n+nv}{diff} \PY{o}{=} \PY{n+nb}{maxval}\PY{p}{(}\PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{state}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{n+nv}{tmp}\PY{p}{)}\PY{p}{)}
      \PY{k}{print}\PY{o}{*}\PY{p}{,} \PY{l+s+s2}{"diff="}\PY{p}{,} \PY{n+nv}{diff}
      \PY{k}{if} \PY{p}{(}\PY{n+nv}{diff} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mf}{.5} \PY{o}{*} \PY{l+m+mi}{1}\PY{l+m+mf}{0.}\PY{o}{**}\PY{p}{(}\PY{o}{-}\PY{n+nv}{dec}\PY{p}{)}\PY{p}{)} \PY{n+nv}{error} \PY{o}{=} \PY{n+nb}{.TRUE.}
      \PY{k}{deallocate}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{)}
      \PY{k}{if} \PY{p}{(}\PY{n+nv}{error}\PY{p}{)} \PY{k}{then}
        \PY{c}{!print*, slv\PYZpc{}state()}
        \PY{k}{stop }\PY{l+m+mi}{1}
      \PY{k}{end }\PY{k}{if}
\PY{k}{    }\PY{k}{end }\PY{k}{block}
\PY{k}{  }\PY{k}{end }\PY{k}{block}

\PY{k}{  }\PY{k}{contains}

\PY{k}{  }\PY{k}{function }\PY{n+nv}{arg2int}\PY{p}{(}\PY{n+nv}{argno}\PY{p}{)} \PY{k}{result}\PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k}{implicit }\PY{k}{none}
\PY{k}{    }\PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{argno}\PY{p}{,} \PY{k}{return}
\PY{k}{    }\PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{l+m+mi}{666}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{tmp}
    \PY{k}{call }\PY{n+nb}{get\PYZus{}command\PYZus{}argument}\PY{p}{(}\PY{n+nv}{argno}\PY{p}{,} \PY{n+nv}{tmp}\PY{p}{)}
    \PY{k}{read}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{,}\PY{o}{*}\PY{p}{)}\PY{k}{return}
\PY{k}{  }\PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{function }\PY{n+nv}{arg2real}\PY{p}{(}\PY{n+nv}{argno}\PY{p}{)} \PY{k}{result}\PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k}{implicit }\PY{k}{none}
\PY{k}{    }\PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{argno}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}
\PY{k}{    }\PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{l+m+mi}{666}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{tmp}
    \PY{k}{call }\PY{n+nb}{get\PYZus{}command\PYZus{}argument}\PY{p}{(}\PY{n+nv}{argno}\PY{p}{,} \PY{n+nv}{tmp}\PY{p}{)}
    \PY{k}{read}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{,}\PY{o}{*}\PY{p}{)}\PY{k}{return}
\PY{k}{  }\PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{read\PYZus{}file}\PY{p}{(}\PY{n+nv}{fname}\PY{p}{,} \PY{k}{array}\PY{p}{)}
    \PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{o}{*}\PY{p}{)}\PY{p}{,} \PY{k}{intent} \PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{fname}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{k}{array}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{un}
    \PY{k}{open}\PY{p}{(}\PY{n+nv}{newunit}\PY{o}{=}\PY{n+nv}{un}\PY{p}{,} \PY{n+nv}{file}\PY{o}{=}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{status}\PY{o}{=}\PY{l+s+s1}{'old'}\PY{p}{,} \PY{n+nv}{action}\PY{o}{=}\PY{l+s+s1}{'read'} \PY{p}{)}
    \PY{k}{block    }
\PY{k}{      }\PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}
      \PY{k}{do }\PY{n+nv}{i}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{size}\PY{p}{(}\PY{k}{array}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
        \PY{k}{read}\PY{p}{(}\PY{n+nv}{un}\PY{p}{,} \PY{o}{*}\PY{p}{)} \PY{p}{(}\PY{k}{array} \PY{p}{(}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{size}\PY{p}{(}\PY{k}{array}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{)}
      \PY{k}{end }\PY{k}{do}
\PY{k}{    }\PY{k}{end }\PY{k}{block}
\PY{k}{    }\PY{k}{close} \PY{p}{(}\PY{n+nv}{un}\PY{p}{)}
   \PY{k}{end }\PY{k}{subroutine }\PY{n+nv}{read\PYZus{}file}
\PY{k}{end }\PY{k}{program}
\end{Verbatim}

### 2013-03-19 paper/doc/code-pyt-listings.py.tex
last changes before arXiv resubmission

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c}{\PYZsh{}listing00}
\PY{c}{\PYZsh{} code licensed under the terms of GNU GPL v3}
\PY{c}{\PYZsh{} copyright holder: University of Warsaw}
\PY{c}{\PYZsh{}listing01}
\PY{n}{real\PYZus{}t} \PY{o}{=} \PY{l+s}{'}\PY{l+s}{float64}\PY{l+s}{'}
\PY{c}{\PYZsh{}listing02}
\PY{k}{try}\PY{p}{:}
  \PY{k+kn}{import} \PY{n+nn}{numpypy}
  \PY{k+kn}{from} \PY{n+nn}{\PYZus{}numpypy.pypy} \PY{k+kn}{import} \PY{n}{set\PYZus{}invalidation}
  \PY{n}{set\PYZus{}invalidation}\PY{p}{(}\PY{n+nb+bp}{False}\PY{p}{)}
\PY{k}{except} \PY{n+ne}{ImportError}\PY{p}{:}
  \PY{k}{pass}
\PY{k+kn}{import} \PY{n+nn}{numpy}
\PY{k}{try}\PY{p}{:} 
  \PY{n}{numpy}\PY{o}{.}\PY{n}{seterr}\PY{p}{(}\PY{n+nb}{all}\PY{o}{=}\PY{l+s}{'}\PY{l+s}{ignore}\PY{l+s}{'}\PY{p}{)}
\PY{k}{except} \PY{n+ne}{AttributeError}\PY{p}{:}
  \PY{k}{pass}
\PY{c}{\PYZsh{}listing03}
\PY{k}{class} \PY{n+nc}{shift}\PY{p}{(}\PY{p}{)}\PY{p}{:}
  \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{plus}\PY{p}{,} \PY{n}{mnus}\PY{p}{)}\PY{p}{:}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{plus} \PY{o}{=} \PY{n}{plus}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mnus} \PY{o}{=} \PY{n}{mnus}
  \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}radd\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{arg}\PY{p}{)}\PY{p}{:} 
    \PY{k}{return} \PY{n+nb}{type}\PY{p}{(}\PY{n}{arg}\PY{p}{)}\PY{p}{(}
      \PY{n}{arg}\PY{o}{.}\PY{n}{start} \PY{o}{+} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{plus}\PY{p}{,} 
      \PY{n}{arg}\PY{o}{.}\PY{n}{stop}  \PY{o}{+} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{plus}
    \PY{p}{)}
  \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}rsub\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{arg}\PY{p}{)}\PY{p}{:} 
    \PY{k}{return} \PY{n+nb}{type}\PY{p}{(}\PY{n}{arg}\PY{p}{)}\PY{p}{(}
      \PY{n}{arg}\PY{o}{.}\PY{n}{start} \PY{o}{-} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mnus}\PY{p}{,} 
      \PY{n}{arg}\PY{o}{.}\PY{n}{stop}  \PY{o}{-} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mnus}
    \PY{p}{)}
\PY{c}{\PYZsh{}listing04}
\PY{n}{one} \PY{o}{=} \PY{n}{shift}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{)} 
\PY{n}{hlf} \PY{o}{=} \PY{n}{shift}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{)}
\PY{c}{\PYZsh{}listing05}
\PY{k}{def} \PY{n+nf}{ext}\PY{p}{(}\PY{n}{r}\PY{p}{,} \PY{n}{n}\PY{p}{)}\PY{p}{:}
  \PY{k}{if} \PY{p}{(}\PY{n+nb}{type}\PY{p}{(}\PY{n}{n}\PY{p}{)} \PY{o}{==} \PY{n+nb}{int}\PY{p}{)} \PY{o}{\PYZam{}} \PY{p}{(}\PY{n}{n} \PY{o}{==} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:} 
    \PY{n}{n} \PY{o}{=} \PY{n}{one}
  \PY{k}{return} \PY{n+nb}{slice}\PY{p}{(}
    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{o}{.}\PY{n}{start}\PY{p}{,} 
    \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{o}{.}\PY{n}{stop}
  \PY{p}{)}
\PY{c}{\PYZsh{}listing06}
\PY{k}{def} \PY{n+nf}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{o}{*}\PY{n}{idx}\PY{p}{)}\PY{p}{:} 
  \PY{k}{return} \PY{p}{(}\PY{n}{idx}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{,} \PY{n}{idx}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
\PY{c}{\PYZsh{}listing07}
\PY{k}{class} \PY{n+nc}{solver}\PY{p}{(}\PY{n+nb}{object}\PY{p}{)}\PY{p}{:}
  \PY{c}{\PYZsh{} ctor-like method}
  \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{bcx}\PY{p}{,} \PY{n}{bcy}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{:}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n} \PY{o}{=} \PY{l+m+mi}{0}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo} \PY{o}{=} \PY{n}{hlo}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{hlo}\PY{p}{,} \PY{n}{nx} \PY{o}{+} \PY{n}{hlo}\PY{p}{)}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{hlo}\PY{p}{,} \PY{n}{ny} \PY{o}{+} \PY{n}{hlo}\PY{p}{)}

    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{bcx} \PY{o}{=} \PY{n}{bcx}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{bcy} \PY{o}{=} \PY{n}{bcy}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}

    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi} \PY{o}{=} \PY{p}{(}
      \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{p}{(}
        \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}\PY{o}{.}\PY{n}{stop}\PY{p}{,} 
        \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}\PY{o}{.}\PY{n}{stop}
      \PY{p}{)}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}\PY{p}{,}
      \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{p}{(}
        \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}\PY{o}{.}\PY{n}{stop}\PY{p}{,} 
        \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}\PY{o}{.}\PY{n}{stop}
      \PY{p}{)}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}
    \PY{p}{)}

    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C} \PY{o}{=} \PY{p}{(}
      \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{p}{(}
        \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n}{hlf}\PY{p}{)}\PY{o}{.}\PY{n}{stop}\PY{p}{,} 
        \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}\PY{o}{.}\PY{n}{stop}
      \PY{p}{)}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}\PY{p}{,}
      \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{p}{(}
        \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}\PY{o}{.}\PY{n}{stop}\PY{p}{,} 
        \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n}{hlf}\PY{p}{)}\PY{o}{.}\PY{n}{stop}
      \PY{p}{)}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}
    \PY{p}{)}

  \PY{c}{\PYZsh{} accessor methods}
  \PY{k}{def} \PY{n+nf}{state}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
    \PY{k}{return} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{]}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{]}

  \PY{c}{\PYZsh{} helper methods invoked by solve()}
  \PY{k}{def} \PY{n+nf}{courant}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,}\PY{n}{d}\PY{p}{)}\PY{p}{:}
    \PY{k}{return} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{]}

  \PY{k}{def} \PY{n+nf}{cycle}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}  \PY{o}{=} \PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2} \PY{o}{-} \PY{l+m+mi}{2}

   \PY{c}{\PYZsh{} integration logic}
  \PY{k}{def} \PY{n+nf}{solve}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{nt}\PY{p}{)}\PY{p}{:}
    \PY{k}{for} \PY{n}{t} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{nt}\PY{p}{)}\PY{p}{:}
      \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{bcx}\PY{o}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}
      \PY{p}{)}
      \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{bcy}\PY{o}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}
      \PY{p}{)}
      \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{advop}\PY{p}{(}\PY{p}{)} 
      \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{cycle}\PY{p}{(}\PY{p}{)}

\PY{c}{\PYZsh{}listing08}
\PY{k}{class} \PY{n+nc}{cyclic}\PY{p}{(}\PY{n+nb}{object}\PY{p}{)}\PY{p}{:}
  \PY{c}{\PYZsh{} ctor}
  \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{:} 
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{d} \PY{o}{=} \PY{n}{d}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{left\PYZus{}halo} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{start}\PY{o}{-}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{o}{.}\PY{n}{start}    \PY{p}{)}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{rght\PYZus{}edge} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{stop} \PY{o}{-}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{o}{.}\PY{n}{stop}     \PY{p}{)}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{rght\PYZus{}halo} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{stop}\PY{p}{,}      \PY{n}{i}\PY{o}{.}\PY{n}{stop} \PY{o}{+}\PY{n}{hlo}\PY{p}{)}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{left\PYZus{}edge} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{start}\PY{p}{,}     \PY{n}{i}\PY{o}{.}\PY{n}{start}\PY{o}{+}\PY{n}{hlo}\PY{p}{)}

  \PY{c}{\PYZsh{} method invoked by the solver}
  \PY{k}{def} \PY{n+nf}{fill\PYZus{}halos}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{d}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} \PY{o}{=} \PY{p}{(}
      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{d}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{rght\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
    \PY{p}{)}
    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{d}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{rght\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} \PY{o}{=} \PY{p}{(}
      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{d}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
    \PY{p}{)}

\PY{c}{\PYZsh{}listing09}
\PY{k}{def} \PY{n+nf}{f}\PY{p}{(}\PY{n}{psi\PYZus{}l}\PY{p}{,} \PY{n}{psi\PYZus{}r}\PY{p}{,} \PY{n}{C}\PY{p}{)}\PY{p}{:}
  \PY{k}{return} \PY{p}{(}
    \PY{p}{(}\PY{n}{C} \PY{o}{+} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}l} \PY{o}{+} 
    \PY{p}{(}\PY{n}{C} \PY{o}{-} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
\PY{c}{\PYZsh{}listing10}
\PY{k}{def} \PY{n+nf}{donorcell}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
  \PY{k}{return} \PY{p}{(}
    \PY{n}{f}\PY{p}{(}
      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}     \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
        \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
    \PY{p}{)} \PY{o}{-} 
    \PY{n}{f}\PY{p}{(}
      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{-}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}     \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
        \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{-}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
    \PY{p}{)} 
  \PY{p}{)}
\PY{c}{\PYZsh{}listing11}
\PY{k}{def} \PY{n+nf}{donorcell\PYZus{}op}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{n}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
  \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{]} 
    \PY{o}{-} \PY{n}{donorcell}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
    \PY{o}{-} \PY{n}{donorcell}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}
  \PY{p}{)}
\PY{c}{\PYZsh{}listing12}
\PY{k}{class} \PY{n+nc}{solver\PYZus{}donorcell}\PY{p}{(}\PY{n}{solver}\PY{p}{)}\PY{p}{:}
  \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{bcx}\PY{p}{,} \PY{n}{bcy}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}\PY{p}{:}
    \PY{n}{solver}\PY{o}{.}\PY{n}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{bcx}\PY{p}{,} \PY{n}{bcy}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}

  \PY{k}{def} \PY{n+nf}{advop}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
    \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
      \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{,} 
      \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}
    \PY{p}{)}
\PY{c}{\PYZsh{}listing13}
\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}frac}\PY{p}{(}\PY{n}{nom}\PY{p}{,} \PY{n}{den}\PY{p}{)}\PY{p}{:}
  \PY{k}{return} \PY{n}{numpy}\PY{o}{.}\PY{n}{where}\PY{p}{(}\PY{n}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nom}\PY{o}{/}\PY{n}{den}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}
\PY{c}{\PYZsh{}listing14}
\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}A}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
  \PY{k}{return} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} \PY{o}{-} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,}
    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
  \PY{p}{)}
\PY{c}{\PYZsh{}listing15}
\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}B}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
  \PY{k}{return} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(} 
    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{-}
    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{-} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{one}\PY{p}{)}\PY{p}{]}\PY{p}{,}
    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+}
    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{one}\PY{p}{)}\PY{p}{]}
  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
\PY{c}{\PYZsh{}listing16}
\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
  \PY{k}{return} \PY{p}{(}
    \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}  \PY{n}{j}\PY{o}{+}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+}
    \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}  \PY{n}{j}\PY{o}{-}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} 
  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}
\PY{c}{\PYZsh{}listing17}
\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{C}\PY{p}{)}\PY{p}{:}
  \PY{k}{return} \PY{p}{(}
    \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{)} 
    \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{)}\PY{p}{)} 
    \PY{o}{*} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
    \PY{o}{-} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} 
    \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
    \PY{o}{*} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
  \PY{p}{)}
\PY{c}{\PYZsh{}listing18}
\PY{k}{class} \PY{n+nc}{solver\PYZus{}mpdata}\PY{p}{(}\PY{n}{solver}\PY{p}{)}\PY{p}{:}
  \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{n\PYZus{}iters}\PY{p}{,} \PY{n}{bcx}\PY{p}{,} \PY{n}{bcy}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}\PY{p}{:}
    \PY{n}{solver}\PY{o}{.}\PY{n}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{bcx}\PY{p}{,} \PY{n}{bcy}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{im} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{o}{.}\PY{n}{start}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{o}{.}\PY{n}{stop}\PY{p}{)}
    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{jm} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{o}{.}\PY{n}{start}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{o}{.}\PY{n}{stop}\PY{p}{)}

    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n\PYZus{}iters} \PY{o}{=} \PY{n}{n\PYZus{}iters}

    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{tmp} \PY{o}{=} \PY{p}{[}\PY{p}{(}
      \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}\PY{p}{,}
      \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}
    \PY{p}{)}\PY{p}{]}
    \PY{k}{if} \PY{n}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{2}\PY{p}{:}
      \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{tmp}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{p}{(}
        \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}\PY{p}{,}
        \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}
      \PY{p}{)}\PY{p}{)}

  \PY{k}{def} \PY{n+nf}{advop}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
    \PY{k}{for} \PY{n}{step} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n\PYZus{}iters}\PY{p}{)}\PY{p}{:}
      \PY{k}{if} \PY{n}{step} \PY{o}{==} \PY{l+m+mi}{0}\PY{p}{:}
        \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
          \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}
        \PY{p}{)}
      \PY{k}{else}\PY{p}{:}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{cycle}\PY{p}{(}\PY{p}{)}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{bcx}\PY{o}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
          \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}
        \PY{p}{)}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{bcy}\PY{o}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
          \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{hlo}\PY{p}{)}
        \PY{p}{)}
        \PY{k}{if} \PY{n}{step} \PY{o}{==} \PY{l+m+mi}{1}\PY{p}{:}
          \PY{n}{C\PYZus{}unco}\PY{p}{,} \PY{n}{C\PYZus{}corr} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{C}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
        \PY{k}{elif} \PY{n}{step} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2}\PY{p}{:}
          \PY{n}{C\PYZus{}unco}\PY{p}{,} \PY{n}{C\PYZus{}corr} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
        \PY{k}{else}\PY{p}{:}
          \PY{n}{C\PYZus{}unco}\PY{p}{,} \PY{n}{C\PYZus{}corr} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}

        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{im}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{]} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}
          \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{im}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n}{C\PYZus{}unco}
        \PY{p}{)}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{bcy}\PY{o}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n}{hlf}\PY{p}{)}\PY{p}{)}

        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{jm}\PY{o}{+}\PY{n}{hlf}\PY{p}{]} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}
          \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{[}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{jm}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n}{C\PYZus{}unco}
        \PY{p}{)}
        \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{bcx}\PY{o}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{p}{,} \PY{n}{hlf}\PY{p}{)}\PY{p}{)}

        \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
          \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{psi}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}\PY{p}{,} \PY{n}{C\PYZus{}corr}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}
        \PY{p}{)}
\PY{c}{\PYZsh{}listing19}
\end{Verbatim}

### 2013-03-19 paper/doc/code-pyt-test.py.tex
last changes before arXiv resubmission

\begin{Verbatim}[commandchars=\\\{\}]
\PY{k+kn}{from} \PY{n+nn}{listings} \PY{k+kn}{import} \PY{o}{*}
\PY{k+kn}{import} \PY{n+nn}{sys}

\PY{k}{def} \PY{n+nf}{read\PYZus{}file}\PY{p}{(}\PY{n}{fname}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}\PY{p}{:}
  \PY{n}{tmp} \PY{o}{=} \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}
  \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,} \PY{l+s}{'}\PY{l+s}{r}\PY{l+s}{'}\PY{p}{)} \PY{k}{as} \PY{n}{f}\PY{p}{:}
    \PY{n}{x} \PY{o}{=} \PY{l+m+mi}{0}
    \PY{k}{for} \PY{n}{line} \PY{o+ow}{in} \PY{n}{f}\PY{p}{:}
      \PY{n}{tmp}\PY{p}{[}\PY{n}{x}\PY{p}{,}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{numpy}\PY{o}{.}\PY{n}{fromstring}\PY{p}{(}\PY{n}{line}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{n}{real\PYZus{}t}\PY{p}{,} \PY{n}{sep}\PY{o}{=}\PY{l+s}{'}\PY{l+s+se}{\PYZbs{}t}\PY{l+s}{'}\PY{p}{)}
      \PY{n}{x} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
  \PY{k}{assert}\PY{p}{(}\PY{n}{x} \PY{o}{==} \PY{n}{nx}\PY{p}{)}
  \PY{k}{return} \PY{n}{tmp}

\PY{k}{def} \PY{n+nf}{main}\PY{p}{(}\PY{p}{)}\PY{p}{:}
  \PY{k}{if} \PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{)} \PY{o}{!=} \PY{p}{(}\PY{l+m+mi}{9} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{p}{:} 
    \PY{k}{raise} \PY{n+ne}{Exception}\PY{p}{(}\PY{l+s}{'}\PY{l+s}{expecting 9 arguments (nx, ny, Cx, Cy, nt, it, f.in, f.out, dec)}\PY{l+s}{'}\PY{p}{)}

  \PY{n}{nx} \PY{o}{=} \PY{n+nb}{int}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
  \PY{n}{ny} \PY{o}{=} \PY{n+nb}{int}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{)}
  \PY{n}{Cx} \PY{o}{=} \PY{n+nb}{float}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{)}
  \PY{n}{Cy} \PY{o}{=} \PY{n+nb}{float}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{4}\PY{p}{]}\PY{p}{)}
  \PY{n}{nt} \PY{o}{=} \PY{n+nb}{int}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{5}\PY{p}{]}\PY{p}{)}
  \PY{n}{it} \PY{o}{=} \PY{n+nb}{int}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{6}\PY{p}{]}\PY{p}{)}
  \PY{n}{fname} \PY{o}{=} \PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{7}\PY{p}{]}
  \PY{n}{fout} \PY{o}{=} \PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{8}\PY{p}{]}
  \PY{n}{dec} \PY{o}{=} \PY{n+nb}{int}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{[}\PY{l+m+mi}{9}\PY{p}{]}\PY{p}{)}

\PY{c}{\PYZsh{}listing19}
  \PY{n}{slv} \PY{o}{=} \PY{n}{solver\PYZus{}mpdata}\PY{p}{(}\PY{n}{it}\PY{p}{,} \PY{n}{cyclic}\PY{p}{,} \PY{n}{cyclic}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}
  \PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{read\PYZus{}file}\PY{p}{(}\PY{n}{fname}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}
  \PY{n}{slv}\PY{o}{.}\PY{n}{courant}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{[}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{Cx}
  \PY{n}{slv}\PY{o}{.}\PY{n}{courant}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{[}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{Cy}
  \PY{n}{slv}\PY{o}{.}\PY{n}{solve}\PY{p}{(}\PY{n}{nt}\PY{p}{)}
\PY{c}{\PYZsh{}listing20}

  \PY{n}{diff} \PY{o}{=} \PY{n}{numpy}\PY{o}{.}\PY{n}{amax}\PY{p}{(}\PY{n+nb}{abs}\PY{p}{(}\PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{n}{read\PYZus{}file}\PY{p}{(}\PY{n}{fout}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}\PY{p}{)}\PY{p}{)}
  \PY{k}{print} \PY{l+s}{"}\PY{l+s}{diff=}\PY{l+s}{"}\PY{p}{,} \PY{n}{diff}
  \PY{k}{if} \PY{p}{(}\PY{n}{diff} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{o}{.}\PY{l+m+mi}{5} \PY{o}{*} \PY{n+nb}{pow}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{,} \PY{o}{-}\PY{n}{dec}\PY{p}{)}\PY{p}{)}\PY{p}{:} 
    \PY{k}{print} \PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{dtype}\PY{p}{,} \PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)}
    \PY{n}{tmp} \PY{o}{=} \PY{n}{read\PYZus{}file}\PY{p}{(}\PY{n}{fout}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}
    \PY{k}{print} \PY{n}{tmp}\PY{o}{.}\PY{n}{dtype}\PY{p}{,} \PY{n}{tmp}
    \PY{k}{print} \PY{n}{numpy}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{n}{numpy}\PY{o}{.}\PY{n}{abs}\PY{p}{(}\PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{n}{tmp}\PY{p}{)}\PY{p}{)}
    \PY{k}{raise} \PY{n+ne}{Exception}\PY{p}{(}\PY{p}{)}

\PY{k}{if} \PY{n}{\PYZus{}\PYZus{}name\PYZus{}\PYZus{}} \PY{o}{==} \PY{l+s}{"}\PY{l+s}{\PYZus{}\PYZus{}main\PYZus{}\PYZus{}}\PY{l+s}{"}\PY{p}{:}
    \PY{n}{main}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

### 2013-03-19 paper/doc/paper.tex
last changes before arXiv resubmission

@@
+%%%%%\documentclass[review,peprint,10pt,5p,times]{elsarticle}
+%%%%%\usepackage[switch]{lineno}
+%%%%%\addtolength{\hoffset}{.5cm}
+
 \documentclass[final,5p,times,twocolumn]{elsarticle}
@@
 
-\input{formulae/preamble}
+\input{formulae_preamble}
 
@@
 
+%%%%%  \linenumbers % for the review
+
+%%%%%  \section*{Program summary}
+
+%%%%%  \noindent
+%%%%%  Program title: mpdata-oop\\
+%%%%%  Licensing provisions: GNU GPL version 3 \citep{GPLv3}\\
+%%%%%  Programming language: C++11, Python and Fortran 2008\\
+%%%%%  External routines/libraries used: Blitz++, NumPy\\
+%%%%%  Nature of problem: advective transport\\
+%%%%%  Solution method: MPDATA\\
+%%%%%  References: Smolarkiewicz 1984: A fully multidimensional positive definite advection transport algorithm with small implicit diffusion. J. Comp. Phys. 54, 325--362 \citep{Smolarkiewicz_1984}
+
   \section{Introduction}
@@
   \begin{equation}\label{eq:adv}
-    \input{formulae/adv}
+    \input{formulae_adv}
   \end{equation}
@@
     C++ (listings prefixed with C, black frame),
-  \codecpp{code/cpp/listings.hpp}{listing00}{listing01}{1}
+  \codecpp{code-cpp-listings.hpp}{listing00}{listing01}{1}
     Python (listings prefixed with P, blue frame) and
-  \codepyt{code/pyt/listings.py}{listing00}{listing01}{1}
+  \codepyt{code-pyt-listings.py}{listing00}{listing01}{1}
     Fortran (listings prefixed with F, red frame).
-  \codefor{code/for/listings.f}{listing00}{listing01}{1}
+  \codefor{code-for-listings.f}{listing00}{listing01}{1}
   Programming language constructs when inlined in the text are 
@@
   The codes begin with the following definitions:
-  \codecpp{code/cpp/listings.hpp}{listing01}{listing02}{1}
-  \codepyt{code/pyt/listings.py}{listing01}{listing02}{1}
-  \codefor{code/for/listings.f}{listing01}{listing02}{1}
+  \codecpp{code-cpp-listings.hpp}{listing01}{listing02}{1}
+  \codepyt{code-pyt-listings.py}{listing01}{listing02}{1}
+  \codefor{code-for-listings.f}{listing01}{listing02}{1}
   which provide a convenient way of switching to different precision.
@@
     and optionally expose the required classes to the present namespace:
-  \codecpp{code/cpp/listings.hpp}{listing02}{listing03}{1}
+  \codecpp{code-cpp-listings.hpp}{listing02}{listing03}{1}
   Here \prog{arr\_t}, \prog{rng\_t} and \prog{idx\_t} serve as alias identifiers 
@@
     even more if used through the following preprocessor macro:
-  \codecpp{code/cpp/listings.hpp}{listing03}{listing04}{1}
+  \codecpp{code-cpp-listings.hpp}{listing03}{listing04}{1}
   The call to \prog{blitz::safeToReturn()} function is included in order to ensure that
@@
     the Python code includes the following sequence of \prog{import} statements:
-  \codepyt{code/pyt/listings.py}{listing02}{listing03}{1}
+  \codepyt{code-pyt-listings.py}{listing02}{listing03}{1}
   First, the PyPy's built-in NumPy implementation named \prog{numpypy} is imported if applicable (i.e. if running PyPy), 
@@
     operator is overloaded with a modulo variant:
-  \codecpp{code/cpp/listings.hpp}{listing04}{listing05}{1}
+  \codecpp{code-cpp-listings.hpp}{listing04}{listing05}{1}
   Consequently the last element of any such sequence may be accessed at index \prog{-1}, the last but one at \prog{-2}, 
@@
     hence a custom \prog{arrvec\_t} type is introduced:
-  \codefor{code/for/listings.f}{listing02}{listing03}{1}
+  \codefor{code-for-listings.f}{listing02}{listing03}{1}
   The \prog{arr\_t} type is defined solely for the purpose of overcoming the limitation 
@@
     \prog{hlf\_t} is defined, and then the plus and minus operators for \prog{hlf\_t} and \prog{rng\_t} are overloaded:
-  \codecpp{code/cpp/listings.hpp}{listing05}{listing06}{1}
+  \codecpp{code-cpp-listings.hpp}{listing05}{listing06}{1}
   This way, the arrays representing vector field components can be indexed using
@@
   It is implemented here by declaring a \prog{shift} class with the adequate operator overloads:
-  \codepyt{code/pyt/listings.py}{listing03}{listing04}{1}
+  \codepyt{code-pyt-listings.py}{listing03}{listing04}{1}
     and two instances of it to represent unity and half
@@
     the unity as a separate object}:
-  \codepyt{code/pyt/listings.py}{listing04}{listing05}{1}
+  \codepyt{code-pyt-listings.py}{listing04}{listing05}{1}
 
@@
     operator overloads:
-  \codefor{code/for/listings.f}{listing03}{listing04}{1}
+  \codefor{code-for-listings.f}{listing03}{listing04}{1}
 
@@
     coding of array range extensions:
-  \codecpp{code/cpp/listings.hpp}{listing06}{listing07}{1}
-  \codepyt{code/pyt/listings.py}{listing05}{listing06}{1}
-  \codefor{code/for/listings.f}{listing04}{listing05}{1}
+  \codecpp{code-cpp-listings.hpp}{listing06}{listing07}{1}
+  \codepyt{code-pyt-listings.py}{listing05}{listing06}{1}
+  \codefor{code-for-listings.f}{listing04}{listing05}{1}
   Consequently, a range depicted by $i\pm1/2$ may be expressed in the code as \prog{ext(i, h)}.
@@
   \begin{equation*}\label{eq:pi}
-    \input{formulae/pi}
+    \input{formulae_pi}
   \end{equation*}
@@
     (note the different order of \prog{i} and \prog{j} arguments in the two template specialisations):
-  \codecpp{code/cpp/listings.hpp}{listing07}{listing08}{1}
+  \codecpp{code-cpp-listings.hpp}{listing07}{listing08}{1}
 
@@
   Therefore, the following definition of function \prog{pi()} suffices to represent $\pi$:
-  \codepyt{code/pyt/listings.py}{listing06}{listing07}{1}
+  \codepyt{code-pyt-listings.py}{listing06}{listing07}{1}
 
@@
     length of one of the vectors passed as argument is defined:
-  \codefor{code/for/listings.f}{listing05}{listing06}{1}
+  \codefor{code-for-listings.f}{listing05}{listing06}{1}
   The \prog{span()} function is used to shorten the declarations of arrays to be returned 
@@
     task (iii) is handled by the accessor methods; task (iv) is handled within the \prog{solve} method:
-  \codecpp{code/cpp/listings.hpp}{listing08}{listing09}{1}
+  \codecpp{code-cpp-listings.hpp}{listing08}{listing09}{1}
   The \prog{solver} structure is an abstract definition (containing a pure virtual method) 
@@
   Python definition of the \prog{solver} class follows closely the C++ structure definition:
-  \codepyt{code/pyt/listings.py}{listing07}{listing08}{1}
+  \codepyt{code-pyt-listings.py}{listing07}{listing08}{1}
   The key difference stems from the fact that, unlike Blitz++, NumPy does not allow an array
@@
       nomenclature between Fortran and C++]{Rouson_et_al_2012}:
-  \codefor{code/for/listings.f}{listing06}{listing07}{1}
+  \codefor{code-for-listings.f}{listing06}{listing07}{1}
   Having defined the abstract type for boundary-condition objects, 
     a definition of a solver class following closely the C++ and Python counterparts may be provided:
-  \codefor{code/for/listings.f}{listing07}{listing08}{1}
+  \codefor{code-for-listings.f}{listing07}{listing08}{1}
 
@@
   An implementation of periodic boundary conditions in C++ is provided in the following listing:
-  \codecpp{code/cpp/listings.hpp}{listing09}{listing10}{1}
+  \codecpp{code-cpp-listings.hpp}{listing09}{listing10}{1}
 
@@
   \begin{equation}\label{eq:donor}
-    \input{formulae/donor}
+    \input{formulae_donor}
   \end{equation}
@@
   \begin{equation}\label{eq:donor:F}
-    \input{formulae/donor-F}
+    \input{formulae_donor-F}
   \end{equation}
@@
   \noindent The flux function takes the following form in C++:
-  \codecpp{code/cpp/listings.hpp}{listing10}{listing11}{1}
+  \codecpp{code-cpp-listings.hpp}{listing10}{listing11}{1}
 
@@
     (effectively the 1-dimensional donor-cell formula):
-  \codecpp{code/cpp/listings.hpp}{listing11}{listing12}{1}
+  \codecpp{code-cpp-listings.hpp}{listing11}{listing12}{1}
 
@@
 
-  \codecpp{code/cpp/listings.hpp}{listing12}{listing13}{1}
+  \codecpp{code-cpp-listings.hpp}{listing12}{listing13}{1}
 
@@
     a \prog{solver\_donorcell} implementation of the abstract \prog{solver} class: 
-  \codecpp{code/cpp/listings.hpp}{listing13}{listing14}{1}
+  \codecpp{code-cpp-listings.hpp}{listing13}{listing14}{1}
   The above definition is given as an example only.
@@
   \begin{equation}\label{eq:antidiff1}
-    \input{formulae/antidiff1}
+    \input{formulae_antidiff1}
   \end{equation}
@@
   \begin{equation}\label{eq:antidiff2}
-    \input{formulae/antidiff2}
+    \input{formulae_antidiff2}
   \end{equation}
@@
   \begin{equation}\label{eq:A}
-    \input{formulae/A}
+    \input{formulae_A}
   \end{equation}
   \begin{equation}\label{eq:B}
-    \input{formulae/B}
+    \input{formulae_B}
   \end{equation}
@@
     represented with the \prog{where} construct (available in all three considered languages):
-  \codecpp{code/cpp/listings.hpp}{listing14}{listing15}{1}
+  \codecpp{code-cpp-listings.hpp}{listing14}{listing15}{1}
 
   The $A$ term defined in equation \ref{eq:A} takes the following form:
-  \codecpp{code/cpp/listings.hpp}{listing15}{listing16}{1}
+  \codecpp{code-cpp-listings.hpp}{listing15}{listing16}{1}
 
   The $B$ term defined in equation \ref{eq:B} takes the following form:
-  \codecpp{code/cpp/listings.hpp}{listing16}{listing17}{1}
+  \codecpp{code-cpp-listings.hpp}{listing16}{listing17}{1}
 
   Equation \ref{eq:antidiff2} takes the following form:
-  \codecpp{code/cpp/listings.hpp}{listing17}{listing18}{1}
+  \codecpp{code-cpp-listings.hpp}{listing17}{listing18}{1}
 
   Equation \ref{eq:antidiff1} take the following form:
-  \codecpp{code/cpp/listings.hpp}{listing18}{listing19}{1}
+  \codecpp{code-cpp-listings.hpp}{listing18}{listing19}{1}
 
@@
     with the following definition in C++:
-  \codecpp{code/cpp/listings.hpp}{listing19}{listing20}{1}
+  \codecpp{code-cpp-listings.hpp}{listing19}{listing20}{1}
   The array of sequences of temporary arrays \prog{tmp} allocated in the constructor
@@
 
-  \codecpp{code/cpp/plot.cpp}{listing20}{listing21}{1}
+  \codecpp{code-cpp-plot.cpp}{listing20}{listing21}{1}
 
   \begin{figure}
-    \pgfimage[height=.9\textheight]{code/cpp/figure}
+    \includegraphics[height=.9\textheight]{demo}
     \caption{\label{fig:mpdata}
@@
     \center
-    \includegraphics[height=.45\textwidth,angle=-90]{plot-mem-host=laptok-cpu=GenuineIntel-6-42-bogomips=3190}
+    \includegraphics[height=.45\textwidth,angle=-90]{skua-mem}
     \caption{\label{fig:mem}
@@
     \center
-    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=laptok-cpu=GenuineIntel-6-42-bogomips=3190}
+    \includegraphics[height=.45\textwidth,angle=-90]{laptok-cpu}
     \caption{\label{fig:cpu-eyrie}
@@
     \center
-    \includegraphics[height=.45\textwidth,angle=-90]{plot-cpu-host=skua-cpu=AuthenticAMD-16-10-bogomips=5620}
+    \includegraphics[height=.45\textwidth,angle=-90]{skua-cpu}
     \caption{\label{fig:cpu-skua}
@@
 
+%%%%%%
   \clearpage
+%%%%%%
   \appendix
@@
   \noindent 
-  \codepyt{code/pyt/listings.py}{listing08}{listing09}{1}
+  \codepyt{code-pyt-listings.py}{listing08}{listing09}{1}
 
   \subsection*{{\bf Donor-cell formul\ae}~(cf. Sect. \ref{sec:donor})}
-  \codepyt{code/pyt/listings.py}{listing09}{listing10}{1}
-  \codepyt{code/pyt/listings.py}{listing10}{listing11}{1}
-  \codepyt{code/pyt/listings.py}{listing11}{listing12}{1}
+  \codepyt{code-pyt-listings.py}{listing09}{listing10}{1}
+  \codepyt{code-pyt-listings.py}{listing10}{listing11}{1}
+  \codepyt{code-pyt-listings.py}{listing11}{listing12}{1}
 
   \subsection*{{\bf Donor-cell solver} (cf. Sect. \ref{sec:donorcell_solver})}
-  \codepyt{code/pyt/listings.py}{listing12}{listing13}{1}
+  \codepyt{code-pyt-listings.py}{listing12}{listing13}{1}
 
   \subsection*{{\bf MPDATA formul\ae}~(cf. Sect. \ref{sec:mpdata})}
-  \codepyt{code/pyt/listings.py}{listing13}{listing14}{1}
-  \codepyt{code/pyt/listings.py}{listing14}{listing15}{1}
-  \codepyt{code/pyt/listings.py}{listing15}{listing16}{1}
-  \codepyt{code/pyt/listings.py}{listing16}{listing17}{1}
-  \codepyt{code/pyt/listings.py}{listing17}{listing18}{1}
+  \codepyt{code-pyt-listings.py}{listing13}{listing14}{1}
+  \codepyt{code-pyt-listings.py}{listing14}{listing15}{1}
+  \codepyt{code-pyt-listings.py}{listing15}{listing16}{1}
+  \codepyt{code-pyt-listings.py}{listing16}{listing17}{1}
+  \codepyt{code-pyt-listings.py}{listing17}{listing18}{1}
 
   \subsection*{{\bf An MPDATA solver} (cf. Sect. \ref{sec:mpdatasolver})}
-  \codepyt{code/pyt/listings.py}{listing18}{listing19}{1}
+  \codepyt{code-pyt-listings.py}{listing18}{listing19}{1}
 
   \subsection*{{\bf Usage example} (cf. Sect. \ref{sec:example})}
-  \codepyt{code/pyt/test.py}{listing19}{listing20}{1}
+  \codepyt{code-pyt-test.py}{listing19}{listing20}{1}
 
@@
   \subsection*{{\bf Periodic boundaries} (cf. Sect. \ref{sec:cyclic})}
-  \codefor{code/for/listings.f}{listing08}{listing09}{1}
+  \codefor{code-for-listings.f}{listing08}{listing09}{1}
 
   \subsection*{{\bf Donor-cell formul\ae}~(cf. Sect. \ref{sec:donor})}
-  \codefor{code/for/listings.f}{listing09}{listing10}{1}
-  \codefor{code/for/listings.f}{listing10}{listing11}{1}
-  \codefor{code/for/listings.f}{listing11}{listing12}{1}
-  \codefor{code/for/listings.f}{listing12}{listing13}{1}
-  \codefor{code/for/listings.f}{listing13}{listing14}{1}
+  \codefor{code-for-listings.f}{listing09}{listing10}{1}
+  \codefor{code-for-listings.f}{listing10}{listing11}{1}
+  \codefor{code-for-listings.f}{listing11}{listing12}{1}
+  \codefor{code-for-listings.f}{listing12}{listing13}{1}
+  \codefor{code-for-listings.f}{listing13}{listing14}{1}
 
   \subsection*{{\bf Donor-cell solver} (cf. Sect. \ref{sec:donorcell_solver})}
-  \codefor{code/for/listings.f}{listing14}{listing15}{1}
+  \codefor{code-for-listings.f}{listing14}{listing15}{1}
 
   \subsection*{{\bf MPDATA formul\ae}~(cf. Sect. \ref{sec:mpdata})}
-  \codefor{code/for/listings.f}{listing15}{listing16}{1}
-  \codefor{code/for/listings.f}{listing16}{listing17}{1}
-  \codefor{code/for/listings.f}{listing17}{listing18}{1}
-  \codefor{code/for/listings.f}{listing18}{listing19}{1}
-  \codefor{code/for/listings.f}{listing19}{listing20}{1}
-  \codefor{code/for/listings.f}{listing20}{listing21}{1}
-  \codefor{code/for/listings.f}{listing21}{listing22}{1}
+  \codefor{code-for-listings.f}{listing15}{listing16}{1}
+  \codefor{code-for-listings.f}{listing16}{listing17}{1}
+  \codefor{code-for-listings.f}{listing17}{listing18}{1}
+  \codefor{code-for-listings.f}{listing18}{listing19}{1}
+  \codefor{code-for-listings.f}{listing19}{listing20}{1}
+  \codefor{code-for-listings.f}{listing20}{listing21}{1}
+  \codefor{code-for-listings.f}{listing21}{listing22}{1}
 
   \subsection*{{\bf An MPDATA solver} (cf. Sect. \ref{sec:mpdatasolver})}
-  \codefor{code/for/listings.f}{listing22}{listing23}{1}
+  \codefor{code-for-listings.f}{listing22}{listing23}{1}
 
   \subsection*{{\bf Usage example} (cf. Sect. \ref{sec:example})}
-  \codefor{code/for/test.f}{listing23}{listing24}{1}
-  \codefor{code/for/test.f}{listing25}{listing26}{1}
+  \codefor{code-for-test.f}{listing23}{listing24}{1}
+  \codefor{code-for-test.f}{listing25}{listing26}{1}
 

### 2013-03-19 paper/tests/timing/inout.sh

@@ -1,18 +1,0 @@
-#!/bin/bash
-Cx=.2
-Cy=.5
-it=3
-
-#for it in {1,2,3}; do
-  for nxy in {0064,0081,0102,0128,0161,0203,0256,0323,0406,0512,0645,0813,1024,1290,1626,2048}; do
-    nt=`echo "2^24/$nxy/$nxy"|bc`
-    echo nx=ny=$nxy nt=$nt
-    time ../../cpp/inout-cpp $nxy $nxy $Cx $Cy $nt $it \
-      1>"nx=${nxy}_ny=${nxy}_Cx=${Cx}_Cy=${Cy}_nt=${nt}_it=${it}:in" \
-      2>"nx=${nxy}_ny=${nxy}_Cx=${Cx}_Cy=${Cy}_nt=${nt}_it=${it}:out"
-    if [ $? -ne 0 ]; then 
-      echo "calling ../../cpp/inout-cpp failed!"
-      exit 1
-    fi
-  done
-#done

### 2013-03-20 paper/doc/paper.tex
adding CPC Program Summary

@@
-%%%%%\documentclass[review,peprint,10pt,5p,times]{elsarticle}
-%%%%%\usepackage[switch]{lineno}
-%%%%%\addtolength{\hoffset}{.5cm}
-
-\documentclass[final,5p,times,twocolumn]{elsarticle}
+%%%%% <cpc>
+\documentclass[review,peprint,10pt,5p,times]{elsarticle}
+\usepackage[switch]{lineno}
+\addtolength{\hoffset}{.5cm}
+%%%%% </cpc>
+
+%%%%% <arXiv>
+%\documentclass[final,5p,times,twocolumn]{elsarticle}
+%%%%% </arXiv>
 
@@
 
+%%%%% <cpc>
+\linenumbers % for the review
+\section*{Program summary}
+\noindent
+{\it Program title:} mpdata-oop\\
+{\it Licensing provisions:} GNU GPL version 3\\
+{\it Programming languages:} C++, Python and Fortran\\
+{\it Computer:} hardware-independent (tested on x86\_64)\\
+{\it Operating system:} OS-independent (developed on GNU/Linux)\\
+{\it RAM/CPU time requirements:} problem-dependent\\
+{\it Has the code been vectorised or parallelised?:}\\ 
+  GCC 4.8 auto-vectorises parts of the Fortran and the C++ code.\\
+  The code is readily adaptable for concurrent computation using
+  domain decomposition.\\
+{\it External libraries used:} Blitz++, NumPy\\
+{\it Classification:} 70 (Mathematical and Numerical Methods)\\
+{\it Nature of problem:} advective transport\\
+{\it Solution method:} Multidimensional Positive Definite Advective Transport Algorithm (MPDATA)\\
+{\it Reference:} Smolarkiewicz 1984, J. Comp. Phys. 54, 325--362
+%%%%% </cpc>
+
   \tableofcontents
-
-%%%%%  \linenumbers % for the review
-
-%%%%%  \section*{Program summary}
-
-%%%%%  \noindent
-%%%%%  Program title: mpdata-oop\\
-%%%%%  Licensing provisions: GNU GPL version 3 \citep{GPLv3}\\
-%%%%%  Programming language: C++11, Python and Fortran 2008\\
-%%%%%  External routines/libraries used: Blitz++, NumPy\\
-%%%%%  Nature of problem: advective transport\\
-%%%%%  Solution method: MPDATA\\
-%%%%%  References: Smolarkiewicz 1984: A fully multidimensional positive definite advection transport algorithm with small implicit diffusion. J. Comp. Phys. 54, 325--362 \citep{Smolarkiewicz_1984}
 
@@
 
-%%%%%%
-  \clearpage
-%%%%%%
+%%%%%% <arXiv>
+%  \clearpage
+%%%%%% </arXiv>
   \appendix

### 2013-03-20 paper/cpp/listings.hpp
cpp: one language keyword less :)

@@ -2,7 +2,7 @@
 // code licensed under the terms of GNU GPL v3
 // copyright holder: University of Warsaw
 //listing01
-typedef double real_t;
+using real_t = double;
 //listing02
 #include <blitz/array.h>
 using arr_t = blitz::Array<real_t, 2>;

### 2013-03-21 paper/doc/paper.tex
corecting Davide's surname spelling

@@
     Tobias Burnus, Julian Cummings, Ond\v{r}ej \v{C}ert\'ik, Patrik Jonsson,
-      Arjen Markus, Zbigniew Piotrowski, Davide del Vento and Janus Weil 
+      Arjen Markus, Zbigniew Piotrowski, Davide Del Vento and Janus Weil 
       provided valuable feedback to the initial version of the

### 2013-03-24 paper/cpp/plot.cpp
fixing plot.cpp

@@ -11,11 +11,26 @@ void setup(T &solver, int n[2])
   blitz::firstIndex i;
   blitz::secondIndex j;
   solver.state() = exp(
-    -sqr(i-n[x]/2.) / (2*pow(n[x]/10., 2))
-    -sqr(j-n[y]/2.) / (2*pow(n[y]/10., 2))
+    -sqr((.5+i)-n[x]/2.) / (2*pow(n[x]/10., 2))
+    -sqr((.5+j)-n[y]/2.) / (2*pow(n[y]/10., 2))
   );  
   solver.courant(x) = -.5; 
   solver.courant(y) = -.25;
+}
+
+template <class T>
+void plot(T &solver, Gnuplot &gp)
+{
+  gp << "splot '-' binary" 
+     << gp.binfmt(solver.state())
+     << " origin=(.5,.5,-1)"
+     << " with image notitle"
+     << ", '-' binary" 
+     << gp.binfmt(solver.state())
+     << " origin=(.5,.5,0)"
+     << " with lines notitle\n";
+  gp.sendBinary(solver.state().copy());
+  gp.sendBinary(solver.state().copy());
 }
 
 int main() 
@@ -31,27 +46,20 @@ int main()
      << "unset ztics\n"    
      << "set xlabel 'X'\n"
      << "set ylabel 'Y'\n"
-     << "set xrange [0:" << n[x]-1 << "]\n"   
-     << "set yrange [0:" << n[y]-1 << "]\n"   
-     << "set zrange [-.666:1]\n"   
+     << "set xrange [0:" << n[x] << "]\n"   
+     << "set yrange [0:" << n[y] << "]\n"   
+     << "set zrange [-1:1]\n"   
      << "set cbrange [-.025:1.025]\n"     
-     << "set palette maxcolors 42\n"
-     << "set pm3d at b\n";
-  std::string binfmt;
+     << "set palette maxcolors 42\n";
   {
     solver_donorcell<cyclic<x>, cyclic<y>> 
       slv(n[x], n[y]);
     setup(slv, n);
-    binfmt = gp.binfmt(slv.state());
-    gp << "set title 't=0'\n"
-       << "splot '-' binary" << binfmt
-       << "with lines notitle\n";
-    gp.sendBinary(slv.state().copy());
+    gp << "set title 't=0'\n";
+    plot(slv, gp);
     slv.solve(nt);
-    gp << "set title 'donorcell t="<<nt<<"'\n"
-       << "splot '-' binary" << binfmt
[…]

### 2013-05-02 paper/doc/paper.tex
hint on the CMake 2.8.10 bug

@@
   \center
-  \begin{tikzpicture}
-    \coordinate (Origin)   at (0,0);
-    \coordinate (XAxisMin) at (-2.5,0);
-    \coordinate (XAxisMax) at (3,0);
-    \coordinate (YAxisMin) at (0,-2.5);
-    \coordinate (YAxisMax) at (0,3);
-    \draw [thin, gray,-latex] (XAxisMin) -- (XAxisMax);% Draw x axis
-    \draw [thin, gray,-latex] (YAxisMin) -- (YAxisMax);% Draw y axis
-
-    \clip (-2.5,-2.5) rectangle (2.5cm,2.5cm); % Clips the picture...
-    \draw[style=help lines,dashed] (-14,-14) grid[step=2cm] (14,14);
-          % Draws a grid in the new coordinates.
-          %\filldraw[fill=gray, fill opacity=0.3, draw=black] (0,0) rectangle (2,2);
-              % Puts the shaded rectangle
-    \foreach \x in {-7,-6,...,7}{% Two indices running over each
-      \foreach \y in {-7,-6,...,7}{% node on the grid we have drawn 
-        \node[draw,circle,inner sep=2pt,fill] at (2*\x,2*\y) {};
-            % Places a dot at those points
-      }
-    }
-    \draw [black] (0,0) -- (0,0) node [below right] {$\!\psi_{[i,j]}$};
-    \draw [black] (-2,0) -- (-2,0) node [below right] {$\!\psi_{[i-1,j]}$};
-    \draw [black] (0,2) -- (0,2) node [below right] {$\!\psi_{[i,j+1]}$};
-
-    \draw [ultra thick,-latex,red] (.8,0) -- (1.2,0) node [above right] {$\!\!\!\!\!\!\!\!\!\!C^{[x]}_{[i+\phlf,j]}$};
-    \draw [ultra thick,-latex,red] (-1.2,0) -- (-.8,0) node [above right] {$\!\!\!\!\!\!\!\!\!\!C^{[x]}_{[i\mhlf,j]}$};
-    \draw [ultra thick,-latex,red] (.8,2) -- (1.2,2) node [above] {};
-    \draw [ultra thick,-latex,red] (-1.2,2) -- (-.8,2) node [above] {};
-    \draw [ultra thick,-latex,red] (.8,-2) -- (1.2,-2) node [above] {};
-    \draw [ultra thick,-latex,red] (-1.2,-2) -- (-.8,-2) node [above] {};
-
-    \draw [ultra thick,-latex,red] (0,.8) -- (0,1.2) node [above] {};
-    \draw [ultra thick,-latex,red] (0,-1.2) -- (0,-.8) node [below right] {$C^{[y]}_{[i,j\mhlf]}$};
-    \draw [ultra thick,-latex,red] (-2,.8) -- (-2,1.2) node [above] {};
-    \draw [ultra thick,-latex,red] (-2,-1.2) -- (-2,-.8) node [above] {};
-    \draw [ultra thick,-latex,red] (2,.8) -- (2,1.2) node [above] {};
-    \draw [ultra thick,-latex,red] (2,-1.2) -- (2,-.8) node [above] {};
-  \end{tikzpicture}
+  \input{grid}
   \caption{\label{fig:grid}
@@
         domain size is subtracted from the values.
-      Both the $nt=0$ and $nt=2^{24}/()nx \cdot ny$ runs are repeated three times and
+      Both the $nt=0$ and $nt=2^{24}/(nx \cdot ny)$ runs are repeated three times and
         only the shortest time is taken into account.

### 2013-05-02 paper/cpp/plot.cpp

@@ -33,19 +33,19 @@ void plot(T &solver, Gnuplot &gp)
   gp.sendBinary(solver.state().copy());
 }
 
-int main() 
+int main(int argc, char **argv) 
 {
-  int n[] = {24, 24}, nt = 75;
+  int n[] = {24, 24}, nt = atof(argv[1]);
   Gnuplot gp;
-  gp << "set term pdf size 10cm, 30cm\n" 
+  gp << "set term pdf size 8cm, 14cm\n" 
      << "set output 'figure.pdf'\n"     
-     << "set multiplot layout 4,1\n" 
+     << "set multiplot layout 2,1\n" 
      << "set border 4095\n"
      << "set xtics out\n"
      << "set ytics out\n"
      << "unset ztics\n"    
-     << "set xlabel 'X'\n"
-     << "set ylabel 'Y'\n"
+     << "set xlabel 'x/dx'\n"
+     << "set ylabel 'y/dy'\n"
      << "set xrange [0:" << n[x] << "]\n"   
      << "set yrange [0:" << n[y] << "]\n"   
      << "set zrange [-1:1]\n"   
@@ -55,22 +55,25 @@ int main()
     solver_donorcell<cyclic<x>, cyclic<y>> 
       slv(n[x], n[y]);
     setup(slv, n);
-    gp << "set title 't=0'\n";
+/*
+    gp << "set title 't/dt=0'\n";
     plot(slv, gp);
+*/
     slv.solve(nt);
-    gp << "set title 'donorcell t="<<nt<<"'\n";
+    gp << "set title 'donorcell t/dt="<<nt<<"'\n";
     plot(slv, gp);
   } 
   {
-    const int it = 2;
+    const int it = 3;
     solver_mpdata<it, cyclic<x>, cyclic<y>> 
       slv(n[x], n[y]); 
     setup(slv, n); 
     slv.solve(nt);
     gp << "set title 'mpdata<" << it << "> "
-       << "t=" << nt << "'\n";
+       << "t/dt=" << nt << "'\n";
     plot(slv, gp);
   } 
+/*
   {
     const int it = 44;
     solver_mpdata<it, cyclic<x>, cyclic<y>> 
@@ -78,8 +81,9 @@ int main()
     setup(slv, n); 
     slv.solve(nt); 
     gp << "set title 'mpdata<" << it << "> "
-       << "t=" << nt << "'\n";
+       << "t/dt=" << nt << "'\n";
     plot(slv, gp);
   }
+*/
 }
 //listing21

### 2013-05-02 paper/cpp/listings.hpp
Revert "hint on the CMake 2.8.10 bug"
This reverts commit d0d49d3acb5ecb007858dad7d618d065668bf578.

@@ -250,8 +250,7 @@ template<int n_iters, class bcx_t, class bcy_t>
 struct solver_mpdata : solver<bcx_t, bcy_t>
 {
   // member fields
-  static const int n_tmp = n_iters > 2 ? 2 : 1;
-  arrvec_t tmp[n_tmp];
+  arrvec_t tmp[2];
   rng_t im, jm;
 
   // ctor
@@ -260,6 +259,7 @@ struct solver_mpdata : solver<bcx_t, bcy_t>
     im(this->i.first() - 1, this->i.last()),
     jm(this->j.first() - 1, this->j.last())
   {
+    int n_tmp = n_iters > 2 ? 2 : 1;
     for (int n = 0; n < n_tmp; ++n)
     {
       tmp[n].push_back(new arr_t(

### 2013-05-02 paper/cpp/plot.cpp

@@ -33,19 +33,19 @@ void plot(T &solver, Gnuplot &gp)
   gp.sendBinary(solver.state().copy());
 }
 
-int main(int argc, char **argv) 
+int main() 
 {
-  int n[] = {24, 24}, nt = atof(argv[1]);
+  int n[] = {24, 24}, nt = 75;
   Gnuplot gp;
-  gp << "set term pdf size 8cm, 14cm\n" 
+  gp << "set term pdf size 10cm, 30cm\n" 
      << "set output 'figure.pdf'\n"     
-     << "set multiplot layout 2,1\n" 
+     << "set multiplot layout 4,1\n" 
      << "set border 4095\n"
      << "set xtics out\n"
      << "set ytics out\n"
      << "unset ztics\n"    
-     << "set xlabel 'x/dx'\n"
-     << "set ylabel 'y/dy'\n"
+     << "set xlabel 'X'\n"
+     << "set ylabel 'Y'\n"
      << "set xrange [0:" << n[x] << "]\n"   
      << "set yrange [0:" << n[y] << "]\n"   
      << "set zrange [-1:1]\n"   
@@ -55,25 +55,22 @@ int main(int argc, char **argv)
     solver_donorcell<cyclic<x>, cyclic<y>> 
       slv(n[x], n[y]);
     setup(slv, n);
-/*
-    gp << "set title 't/dt=0'\n";
+    gp << "set title 't=0'\n";
     plot(slv, gp);
-*/
     slv.solve(nt);
-    gp << "set title 'donorcell t/dt="<<nt<<"'\n";
+    gp << "set title 'donorcell t="<<nt<<"'\n";
     plot(slv, gp);
   } 
   {
-    const int it = 3;
+    const int it = 2;
     solver_mpdata<it, cyclic<x>, cyclic<y>> 
       slv(n[x], n[y]); 
     setup(slv, n); 
     slv.solve(nt);
     gp << "set title 'mpdata<" << it << "> "
-       << "t/dt=" << nt << "'\n";
+       << "t=" << nt << "'\n";
     plot(slv, gp);
   } 
-/*
   {
     const int it = 44;
     solver_mpdata<it, cyclic<x>, cyclic<y>> 
@@ -81,9 +78,8 @@ int main(int argc, char **argv)
     setup(slv, n); 
     slv.solve(nt); 
     gp << "set title 'mpdata<" << it << "> "
-       << "t/dt=" << nt << "'\n";
[…]

### 2013-05-02 paper/cpp/listings.hpp
changing tmp[2] into tmp[n_tmp] in listings.cpp

@@ -250,7 +250,8 @@ template<int n_iters, class bcx_t, class bcy_t>
 struct solver_mpdata : solver<bcx_t, bcy_t>
 {
   // member fields
-  arrvec_t tmp[2];
+  static const int n_tmp = n_iters > 2 ? 2 : 1;
+  arrvec_t tmp[n_tmp];
   rng_t im, jm;
 
   // ctor
@@ -259,7 +260,6 @@ struct solver_mpdata : solver<bcx_t, bcy_t>
     im(this->i.first() - 1, this->i.last()),
     jm(this->j.first() - 1, this->j.last())
   {
-    int n_tmp = n_iters > 2 ? 2 : 1;
     for (int n = 0; n < n_tmp; ++n)
     {
       tmp[n].push_back(new arr_t(

### 2013-05-02 paper/cpp/plot.cpp
t -> t/dt, X -> x/dx, Y -> y/dy in plot.cpp

@@ -44,8 +44,8 @@ int main()
      << "set xtics out\n"
      << "set ytics out\n"
      << "unset ztics\n"    
-     << "set xlabel 'X'\n"
-     << "set ylabel 'Y'\n"
+     << "set xlabel 'x/dx'\n"
+     << "set ylabel 'y/dy'\n"
      << "set xrange [0:" << n[x] << "]\n"   
      << "set yrange [0:" << n[y] << "]\n"   
      << "set zrange [-1:1]\n"   
@@ -55,10 +55,10 @@ int main()
     solver_donorcell<cyclic<x>, cyclic<y>> 
       slv(n[x], n[y]);
     setup(slv, n);
-    gp << "set title 't=0'\n";
+    gp << "set title 't/dt=0'\n";
     plot(slv, gp);
     slv.solve(nt);
-    gp << "set title 'donorcell t="<<nt<<"'\n";
+    gp << "set title 'donorcell t/dt="<<nt<<"'\n";
     plot(slv, gp);
   } 
   {
@@ -68,7 +68,7 @@ int main()
     setup(slv, n); 
     slv.solve(nt);
     gp << "set title 'mpdata<" << it << "> "
-       << "t=" << nt << "'\n";
+       << "t/dt=" << nt << "'\n";
     plot(slv, gp);
   } 
   {
@@ -78,7 +78,7 @@ int main()
     setup(slv, n); 
     slv.solve(nt); 
     gp << "set title 'mpdata<" << it << "> "
-       << "t=" << nt << "'\n";
+       << "t/dt=" << nt << "'\n";
     plot(slv, gp);
   }
 }

### 2013-07-01 paper/doc/paper.tex
Revert "hint on the CMake 2.8.10 bug"
This reverts commit d0d49d3acb5ecb007858dad7d618d065668bf578.
initial changes before planned submission to GMD

--- 2013-03-21 paper/doc/paper.tex
@@
-%%%%% <cpc>
-\documentclass[review,peprint,10pt,5p,times]{elsarticle}
-\usepackage[switch]{lineno}
-\addtolength{\hoffset}{.5cm}
-%%%%% </cpc>
-
 %%%%% <arXiv>
@@
 %%%%% </arXiv>
+
+%%%%% <gmd>
+\documentclass[gmd]{copernicus}%_discussions}
+%%%%% </gmd>
 
@@
 \newcommand{\prog}[1]{{\rm\bf#1}}
-\newcommand{\url}[1]{{#1}}
 
@@
 \begin{document}
-  \begin{frontmatter}
+    \linenumbers
 
     \title{
-      Object-oriented implementations of the MPDATA advection equation solver in~C++,~Python~and~Fortran
+      Formula translation in C++, Python and modern Fortran: a~case~study~of~the~language~choice tradeoffs
     }
@@
 
-    \address[1]{Institute of Geophysics, Faculty of Physics, University of Warsaw}
-    \address[2]{PyPy Team}
+    \affil[1]{Institute of Geophysics, Faculty of Physics, University of Warsaw}
+    \affil[2]{PyPy Team}
+
+    \correspondence{Sylwester Arabas (sarabas@igf.fuw.edu.pl)}
+    \runningauthor{Arabas, Jarecka et al.}
+    \runningtitle{Formula translation in C++, Python \& Fortran}
+
+    \maketitle
 
@@
         The presented programs are based on Blitz++ (C++), NumPy (Python), and Fortran's built-in array containers.
-        The solvers include an implementation of the Multidimensional Positive-Definite 
+        The solvers constitute implementations of the Multidimensional Positive-Definite 
           Advective Transport Algorithm (MPDATA).
-        The introduced codes exemplify how the application of 
+        The introduced codes serve as exemples for how the application of 
           object-oriented programming (OOP) techniques allows to reproduce the mathematical notation 
@@
           (and hence maintainability and auditability) as well as performance.
+        All performance tests are performed using free and open-source compilers.
         In the case of Python, a significant performance gain is observed when switching from the standard 
@@
 
-    % up to six in Comp. Phys. Comm.
-    \begin{keyword}
-      object-oriented programming, advection equation, MPDATA, C++, Fortran, Python
-    \end{keyword}
-
-  \end{frontmatter}
-
-%%%%% <cpc>
-\linenumbers % for the review
-\section*{Program summary}
-\noindent
-{\it Program title:} mpdata-oop\\
-{\it Licensing provisions:} GNU GPL version 3\\
-{\it Programming languages:} C++, Python and Fortran\\
-{\it Computer:} hardware-independent (tested on x86\_64)\\
-{\it Operating system:} OS-independent (developed on GNU/Linux)\\
-{\it RAM/CPU time requirements:} problem-dependent\\
-{\it Has the code been vectorised or parallelised?:}\\ 
-  GCC 4.8 auto-vectorises parts of the Fortran and the C++ code.\\
-  The code is readily adaptable for concurrent computation using
-  domain decomposition.\\
-{\it External libraries used:} Blitz++, NumPy\\
-{\it Classification:} 70 (Mathematical and Numerical Methods)\\
-{\it Nature of problem:} advective transport\\
-{\it Solution method:} Multidimensional Positive Definite Advective Transport Algorithm (MPDATA)\\
-{\it Reference:} Smolarkiewicz 1984, J. Comp. Phys. 54, 325--362
-%%%%% </cpc>
-
-  \tableofcontents
-
-  \section{Introduction}
+%    \begin{keyword}
+%      object-oriented programming, advection equation, MPDATA, C++, Fortran, Python
+%    \end{keyword}
+
+  \introduction
 
@@
   The latter, virtually impossible to obtain with procedural programming, is the focus of this paper.
-  It also enables the compiler or library authors to relieve the user (i.e. scientific programmer)
-    from hand-coding optimisations, a practice long recognised as having {\em a strong negative impact when debugging
-    and maintenance are considered} \citep{Knuth_1974}.
-
-  MPDATA \citep{Smolarkiewicz_1984} stands for Multidimensional Positive Definite Advective Transport Algorithm and is
-    an example of a numerical procedure used in weather, climate and ocean simulation systems
+
+  The aim of this paper is to show how OOP techniques can be used to represent mathematical
+    notation within the program code.
+  For discussion purposes, three sample implementations of a numerical scheme for solving
+    the advection equation are introduced.
+  The three programs are written in C++, Python and Fortran. 
+  They implement a two-dimensional version of the 
+    Multidimensional Positive Definite Advective Transport Algorithm \citep[MPDATA][]{Smolarkiewicz_1984}.
+  MPDATA is an example of a numerical procedure used in weather, climate and ocean simulation systems
     \citep[e.g.][respectively]{Ziemianski_et_al_2011,Abiodun_et_al_2011,Ezer_et_al_2002}.
-  MPDATA is a solver for systems of advection equations of the following form:
-  \begin{equation}\label{eq:adv}
-    \input{formulae_adv}
-  \end{equation}
-  that describe evolution of a scalar field $\psi$ transported by the fluid flow with velocity $\vec{v}$.
-  Quoting Numerical Recipes once more, development of methods to numerically solve such problems 
-    {\em ''is an art as much as a science''} \citep[][Sec.~20.1]{Press_et_al_2007},
-    and MPDATA is an example of the state-of-the art in this field.
-  MPDATA is designed to accurately solve equation (\ref{eq:adv}) in an arbitrary
-    number of dimensions assuring positive-definiteness of scalar field $\psi$ 
-    and incurring small numerical diffusion.
-  All relevant MPDATA formul\ae~are given in the text but are presented without
-    derivation or detailed discussion.
-  For a recent review of MPDATA-based techniques see \citet[][and references therein]{Smolarkiewicz_2006}.
-
-  In this paper we introduce and discuss object-oriented implementations of an MPDATA-based 
-    two-dimensional (2D) advection equation solver written in C++11 (\citeauthor{ISO_14882} 14882:2011), 
-    Python \citep{Rossum_and_Drake_2011} and Fortran 2008 (\citeauthor{ISO_1539-1} 1539-1:2010). 
+  Even in its most basic form presented herein, MPDATA is complex enough to contain
+    a wide range of mathematical abstractions that can be represented using OOP constructs, 
+    yet it is simple enough to allow inclusion of the entire source code within the paper text.
+  All relevant MPDATA formul\ae~are given in the text alongside corresponding code fragments.
+  In compliance with the article focus, these formul\ae~are presented without derivation or detailed discussion.
+  For a recent review of MPDATA-based techniques including an introductory description of the algorithm
+    and a thorough list of references, see \citet[][]{Smolarkiewicz_2006}.
+
   In the following section we introduce the three implementations
@@
     discussing where and how the OOP techniques may be applied in its implementation.
-  The syntax and nomenclature of OOP techniques are used without introduction,
-    for an overview of OOP in context of C++, Python and Fortran, consult for example
-    \citep[][Part~II]{Stroustrup_2000}, \citep[][Chapter~5]{Pilgrim_2004} and
-    \citep[][Chapter~11]{Markus_2012}, respectively.
+  Subsections \ref{sec:array}-\ref{sec:solver} describe all three implementations,
+    while subsequent sections \ref{sec:cyclic}-\ref{sec:example} cover discussion of C++ code only.
+  The relevant parts of Python and Fortran codes do not differ significantly, and for readability reasons 
+    are presented in \ref{app:P} and \ref{app:F}, respectively.
   The third section of this paper covers performance evaluation of the three implementations.
@@
   The fifth section closes the article with a brief summary.
-
-  Throughout the paper we present the three implementations by discussing 
-    source code listings which cover the entire program code.
-  Subsections \ref{sec:array}-\ref{sec:solver} describe all three implementations,
-    while subsequent sections \ref{sec:cyclic}-\ref{sec:example} cover discussion of C++ code 
-    only.
-  The relevant parts of Python and Fortran codes do not differ significantly, and for readability reasons 
-    are presented in \ref{app:P} and \ref{app:F}, respectively.
 
@@
     hence every Fortran listing contains definition of a separate module.
+  The language syntax and OOP nomenclature are used without introduction in the paper,
+    for an overview of OOP in context of C++, Python and Fortran, consult for example
+    \citep[][Part~II]{Stroustrup_2000}, \citep[][Chapter~5]{Pilgrim_2004} and
+    \citep[][Chapter~11]{Markus_2012}, respectively.
 
@@
 
+  MPDATA is a solver for systems of advection equations of the following form:
+  \begin{equation}\label{eq:adv}
+    \input{formulae_adv}
+  \end{equation}
+  that describe evolution of a scalar field $\psi$ transported by the fluid flow with velocity $\vec{v}$.
   Solution of equation (\ref{eq:adv}) using MPDATA implies discretisation onto a grid of the $\psi$ and 
@@
 
-  The authors do expect some performance gain could 
-    be obtained by introducing into the codes some ''manual'' optimisations -- 
-    code rearrangements aimed solely at the purpose of increasing performance.
-  These were avoided intentionally as they degrade code readability,
-    should in principle be handled by the compilers,
-    and are generally advised to be avoided \citep[e.g.][section 3.12]{bib_CERNcpp}.
-
   \section{Discussion on the tradeoffs of language choice}
@@
   \end{itemize}
+
+  Making use of features such as loop-free arithmetics not only shortens the code,
+    but also enables the compiler or library authors to relieve the user (i.e. scientific programmer)
+    from hand-coding optimisations (e.g. loop order choice).
+  Hand-coded optimisations -- code rearrangements aimed solely at the purpose of increasing performance --
+    were long recognised as having {\em a strong negative impact when debugging
+    and maintenance are considered} \citep{Knuth_1974},
+    and are generally advised to be avoided \citep[e.g.][section 3.12]{bib_CERNcpp}.
+
   Three issues specific to Fortran that 
@@
 
-  \section{Summary and outlook}
+  \conclusions[Summary and outlook]
 
@@
 
-  \section*{Acknowledgements}
-  {\small
-    \noindent 
-    We thank Piotr Smolarkiewicz and Hanna Pawłowska for their help throughout the project.
-    This study was partly inspired by the lectures of Lech Łobocki.
-
-    \noindent
-    Tobias Burnus, Julian Cummings, Ond\v{r}ej \v{C}ert\'ik, Patrik Jonsson,
-      Arjen Markus, Zbigniew Piotrowski, Davide Del Vento and Janus Weil 
-      provided valuable feedback to the initial version of the
-      manuscript and/or responses to questions posted to Blitz++ and gfortran mailing lists.
-
-    \noindent
-    SA, AJ and DJ acknowledge funding from the Polish National Science Centre
-      (project no. 2011/01/N/ST10/01483).
-
-    \noindent
-    Part of the work was carried out during a visit of SA to the National
-      Center for Atmospheric Research (NCAR) in Boulder, Colorado, USA.
-    NCAR is operated by the University Corporation for Atmospheric Research.
-    The visit was funded by the Foundation for Polish Science (START programme).
-
-    \noindent
-    Development of NumPy support in PyPy was led by Alex Gaynor, Matti Picus and MF.
-  }
-
 %%%%%% <arXiv>
@@
 
-  \setcounter{section}{15} 
+  %\setcounter{section}{15} 
   \section{Python code for sections \ref{sec:cyclic}--\ref{sec:mpdatasolver}}\label{app:P}
@@
 
-  \setcounter{section}{5} 
+  %\setcounter{section}{5} 
   \section{Fortran code for sections \ref{sec:cyclic}--\ref{sec:mpdatasolver}}\label{app:F}
@@
 
-  \bibliographystyle{model1-num-names}
+  \begin{acknowledgements}
+    \noindent 
+    We thank Piotr Smolarkiewicz and Hanna Pawłowska for their help throughout the project.
+    This study was partly inspired by the lectures of Lech Łobocki.
+
+    \noindent
+    Tobias Burnus, Julian Cummings, Ond\v{r}ej \v{C}ert\'ik, Patrik Jonsson,
+      Arjen Markus, Zbigniew Piotrowski, Davide Del Vento and Janus Weil 
+      provided valuable feedback to the initial version of the
+      manuscript and/or responses to questions posted to Blitz++ and gfortran mailing lists.
+
+    \noindent
+    SA, AJ and DJ acknowledge funding from the Polish National Science Centre
+      (project no. 2011/01/N/ST10/01483).
+
+    \noindent
+    Part of the work was carried out during a visit of SA to the National
+      Center for Atmospheric Research (NCAR) in Boulder, Colorado, USA.
+    NCAR is operated by the University Corporation for Atmospheric Research.
+    The visit was funded by the Foundation for Polish Science (START programme).
+
+    \noindent
+    Development of NumPy support in PyPy was led by Alex Gaynor, Matti Picus and MF.
+  \end{acknowledgements}
+
+  \bibliographystyle{copernicus}
   \bibliography{paper}

### 2013-07-09 paper/doc/paper.tex
changes before gmd submission

@@
 \usepackage[T1]{fontenc}
+
+\usepackage{type1cm}
+\usepackage{relsize}
 
@@
   frame=single,
-  fontsize=\scriptsize,
+  fontsize=\relscale{.666},
   fontfamily=courier,
@@
           Advective Transport Algorithm (MPDATA).
-        The introduced codes serve as exemples for how the application of 
+        The introduced codes serve as examples for how the application of 
           object-oriented programming (OOP) techniques allows to reproduce the mathematical notation 
@@
           (and hence maintainability and auditability) as well as performance.
-        All performance tests are performed using free and open-source compilers.
+        All performance tests are carried out using free and open-source compilers.
         In the case of Python, a significant performance gain is observed when switching from the standard 
@@
     weather and climate prediction systems - flagship examples of complex scientific software - 
-    make extensive use of OOP techniques.
+    make extensive use of OOP techniques\footnote{
   Fortran has been the language of choice in oceanic \citep{Griffies_et_al_2000}, 
     weather-prediction \citep{Sundberg_2009} and Earth system \citep{Legutke_2012} modelling, 
-    and none of its 20-century editions were object-oriented languages \citep[see e.g.][for discussion]{Norton_et_al_2007}.
+    and none of its 20-century editions were object-oriented languages \citep[see e.g.][for discussion]{Norton_et_al_2007}}.
 
@@
   The latter, virtually impossible to obtain with procedural programming, is the focus of this paper.
-
-  The aim of this paper is to show how OOP techniques can be used to represent mathematical
-    notation within the program code.
-  For discussion purposes, three sample implementations of a numerical scheme for solving
-    the advection equation are introduced.
-  The three programs are written in C++, Python and Fortran. 
-  They implement a two-dimensional version of the 
-    Multidimensional Positive Definite Advective Transport Algorithm \citep[MPDATA][]{Smolarkiewicz_1984}.
-  MPDATA is an example of a numerical procedure used in weather, climate and ocean simulation systems
-    \citep[e.g.][respectively]{Ziemianski_et_al_2011,Abiodun_et_al_2011,Ezer_et_al_2002}.
+  The importance of reproducing the mathematical notation in the code lays primarily
+    in the fact that code readability and brevity \citep[and not the language choice itself,][]{Seeley_2004}
+    significantly contribute to code maintainability \citep{Wilson_et_al_2012}.
+  Code maintainability, in turn, plays a vital role in community programming efforts 
+    and hence geoscientific model development in particular.
+
+  The key aim of this paper is to show how OOP techniques can be used to faithfully
+    reproduce within the code what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012}.
+  For this purpose, a sample implementation of a numerical scheme for solving
+    the advection equation is introduced in in C++, Python and modern Fortran --
+    OOP languages commonly used in scientific computing \citep[see e.g.][chapt.~8]{Einarsson_2005}. 
+  Presented benchmark tests, although quite simplistic, together with the experience gained 
+    from the development of codes in three different
+    languages provide a basis for discussion on the tradeoffs of programming language choice.
+  The discussion concerns in principle the development of finite-difference solvers for 
+    partial differential equations, but is likely applicable to some extent to the scientific software in general.
+
+  All three programs include an equally structured implementation of the two-dimensional version of the 
+    Multidimensional Positive Definite Advective Transport Algorithm \citep[MPDATA;][]{Smolarkiewicz_1984}.
+  MPDATA is an example of a numerical procedure used in weather, climate, ocean and solar simulation systems
+    \citep[e.g.][respectively]{Ziemianski_et_al_2011,Abiodun_et_al_2011,Ezer_et_al_2002,Charbonneau_and_Smolarkiewicz_2013}.
   Even in its most basic form presented herein, MPDATA is complex enough to contain
@@
     yet it is simple enough to allow inclusion of the entire source code within the paper text.
-  All relevant MPDATA formul\ae~are given in the text alongside corresponding code fragments.
-  In compliance with the article focus, these formul\ae~are presented without derivation or detailed discussion.
-  For a recent review of MPDATA-based techniques including an introductory description of the algorithm
-    and a thorough list of references, see \citet[][]{Smolarkiewicz_2006}.
-
-  In the following section we introduce the three implementations
+  All relevant MPDATA formul\ae~are given in the text alongside corresponding code fragments
+    allowing comparison of the relevant syntax with the mathematical notation.
+  In compliance with the article focus, these formul\ae~are presented without derivation or detailed discussion
+    \citep[see][for a recent review of MPDATA-based techniques including an introductory description 
+    of the algorithm and an exhaustive list of references]{Smolarkiewicz_2006}.
+
+  The paper is structured as follows.
+  In section~\ref{sec:impl} we introduce the three implementations
     briefly describing the algorithm itself and
@@
   The relevant parts of Python and Fortran codes do not differ significantly, and for readability reasons 
-    are presented in \ref{app:P} and \ref{app:F}, respectively.
-  The third section of this paper covers performance evaluation of the three implementations.
-  The fourth section covers discussion of the tradeoffs of the programming language choice.
-  The fifth section closes the article with a brief summary.
-
-  The entire code is licensed under the terms of the GNU General Public License license version 3 \citep{GPLv3}.
-
+    are presented in appendices \ref{app:P} and \ref{app:F}, respectively.
+  Section~\ref{sec:perf} of this paper covers performance evaluation of the three implementations.
+  Section~\ref{sec:tradeoffs} covers discussion of the tradeoffs of the programming language choice.
+  Section~\ref{sec:concl} section closes the article with a brief summary.
+  Those readers not interested in technical details may wish to skip to section~\ref{sec:perf} or
+    \ref{sec:tradeoffs}.
+
+  The entire code is licensed under the terms of the GNU General Public License version 3 \citep{GPLv3}.
   All listings include line numbers printed to the left of the source code, with separate numbering for
@@
     hence every Fortran listing contains definition of a separate module.
+
   The language syntax and OOP nomenclature are used without introduction in the paper,
@@
 
-  In Fortran fractional array indexing is obtained through
+  In the Fortran implementation fractional array indexing is obtained through
     definition and instantiation of an object representing the half, and having appropriate
@@
   With Fortran one needs to rely on compiler optimisations to eliminate the conditional expression
-    within the \prog{pi()} that depends on value of \prog{d} which is always known 
+    within \prog{pi()} that depends on value of \prog{d} which is always known 
     at compile time.
@@
   From this point, only C++ implementation is explained in the main text. 
-  The Python and Fortran implementations are included in appendices P and F.
+  The Python and Fortran implementations are included in appendices A and B.
 
@@
 
-  \section{Discussion on the tradeoffs of language choice}
-
-  One of the aims of this paper is to show the applicability of OOP features of the three
-    programming languages (or language-library pairs) for scientific computing.
-  The main focus is to represent
-    what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012} within the code.
-  Presented benchmark tests, although quite simplistic, together with the experience gained 
-    from the development of codes in three different
-    languages provide a basis for discussion on the tradeoffs of programming language choice.
-  The discussion concerns in principle the development of finite-difference solvers for 
-    partial differential equations, but is likely applicable to the scientific software in general.
-  A partly objective and partly subjective summary of pros and cons of C++, Python and Fortran
-    is presented in the four following subsections.
+  \section{Discussion on the tradeoffs of language choice}\label{sec:tradeoffs}
 
@@
   Taking into account such measures as code length or coding time,
-    the Python environment gains significantly.
+    the Python environment gains significantly \citep[see also discussion in][]{Lin_2012}.
   Presented Python code is shorter than the C++ and Fortran counterparts,
@@
 
-  Python implementations do not expose to the user the compilation or linking processes. 
+  Python implementations do not expose users to compilation or linking processes. 
   As a result, Python-written software is easier to deploy and share, especially 
@@
 
-  \conclusions[Summary and outlook]
+  \conclusions[Summary and outlook]\label{sec:concl}
 
@@
   All implementations follow the same object-oriented structure but are implemented
-    in three different languages:
+    in three different languages (or language-library pairs):
   \begin{itemize}
@@
     tools implementing them are free and open-source
-    is certainly an advantage \citep{Anel_2011}.
+    is certainly an advantage \citep[][sect.~28.2.5]{Anel_2011,Syvitski_et_al_2013}.
 
@@
     allowing to reproduce the analysis and plots presented in section~\ref{sec:perf} are all 
-    available in the CPC Program Library and at
+    available in the electronic supplement and at
     the project repository\footnote{git repository at \url{http://github.com/slayoo/mpdata/}},
@@
   \begin{acknowledgements}
-    \noindent 
     We thank Piotr Smolarkiewicz and Hanna Pawłowska for their help throughout the project.
@@
 
-    \noindent
     Tobias Burnus, Julian Cummings, Ond\v{r}ej \v{C}ert\'ik, Patrik Jonsson,
@@
 
-    \noindent
     SA, AJ and DJ acknowledge funding from the Polish National Science Centre
@@
 
-    \noindent
     Part of the work was carried out during a visit of SA to the National
@@
 
-    \noindent
     Development of NumPy support in PyPy was led by Alex Gaynor, Matti Picus and MF.

### 2013-07-15 paper/doc/code-cpp-listings.hpp.tex
the new order

@@
 \PY{c+c1}{//listing01}
-\PY{k}{typedef} \PY{k+kt}{double} \PY{n}{real\PYZus{}t}\PY{p}{;}
+\PY{k}{using} \PY{n}{real\PYZus{}t} \PY{o}{=} \PY{k+kt}{double}\PY{p}{;}
 \PY{c+c1}{//listing02}
@@
 \PY{c+c1}{//listing06}
-\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{n\PYZus{}t}\PY{o}{\PYZgt{}}
-\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{n}{ext}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{r}\PY{p}{,} \PY{k}{const} \PY{n}{n\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{n}\PY{p}{)} \PY{p}{\PYZob{}}
-  \PY{k}{return} \PY{n}{rng\PYZus{}t}\PY{p}{(}
-    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
-    \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}
-  \PY{p}{)}\PY{p}{;} 
-\PY{p}{\PYZcb{}} 
-\PY{c+c1}{//listing07}
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}} 
@@
 \PY{p}{\PYZcb{}}\PY{p}{;} 
+\PY{c+c1}{//listing07}
+\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{T1}\PY{p}{,} \PY{k}{class} \PY{n+nc}{T2}\PY{p}{,} \PY{k}{class} \PY{n+nc}{T3}\PY{o}{\PYZgt{}} 
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{F}\PY{p}{(}
+  \PY{k}{const} \PY{n}{T1} \PY{o}{\PYZam{}}\PY{n}{psi\PYZus{}l}\PY{p}{,} \PY{k}{const} \PY{n}{T2} \PY{o}{\PYZam{}}\PY{n}{psi\PYZus{}r}\PY{p}{,} \PY{k}{const} \PY{n}{T3} \PY{o}{\PYZam{}}\PY{n}{C}
+\PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
+  \PY{p}{(}
+    \PY{p}{(}\PY{n}{C} \PY{o}{+} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}l} \PY{o}{+} 
+    \PY{p}{(}\PY{n}{C} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
+  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
+\PY{p}{)}
 \PY{c+c1}{//listing08}
+\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}  
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{donorcell}\PY{p}{(} 
+  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
+  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+\PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
+  \PY{n}{F}\PY{p}{(}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
+      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
+  \PY{p}{)} \PY{o}{-}
+  \PY{n}{F}\PY{p}{(}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
+      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
+  \PY{p}{)}
+\PY{p}{)}
+\PY{c+c1}{//listing09}
+\PY{k+kt}{void} \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
+  \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{k+kt}{int} \PY{n}{n}\PY{p}{,}
+  \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
+  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+\PY{p}{)} \PY{p}{\PYZob{}} 
+  \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}
+    \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
+    \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}\PY{p}{;}
+\PY{p}{\PYZcb{}}
+\PY{c+c1}{//listing10}
+\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{nom\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{den\PYZus{}t}\PY{o}{\PYZgt{}}
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
+  \PY{k}{const} \PY{n}{nom\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{nom}\PY{p}{,} \PY{k}{const} \PY{n}{den\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{den}
+\PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
+  \PY{n}{where}\PY{p}{(}\PY{n}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nom} \PY{o}{/} \PY{n}{den}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}
+\PY{p}{)} 
+\PY{c+c1}{//listing11}
+\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
+  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+\PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
+  \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}
+  \PY{p}{)} 
+\PY{p}{)} 
+\PY{c+c1}{//listing12}
+\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
+  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+\PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
+ \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{,}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}
+  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
+\PY{p}{)}
+\PY{c+c1}{//listing13}
+\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}
+  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
+  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} 
+  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+\PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
+  \PY{p}{(}
+    \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+}
+    \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)}
+  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}
+\PY{p}{)}
+\PY{c+c1}{//listing14}
+\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}
+  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
+  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,}
+  \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}
+\PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
+  \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}
+  \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{)}
+  \PY{o}{*} \PY{n}{mpdata\PYZus{}A}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} 
+  \PY{o}{-} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} 
+  \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
+  \PY{o}{*} \PY{n}{mpdata\PYZus{}B}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
+\PY{p}{)} 
+\PY{c+c1}{//listing15}
+\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{n\PYZus{}t}\PY{o}{\PYZgt{}}
+\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{n}{ext}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{r}\PY{p}{,} \PY{k}{const} \PY{n}{n\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{n}\PY{p}{)} \PY{p}{\PYZob{}}
+  \PY{k}{return} \PY{n}{rng\PYZus{}t}\PY{p}{(}
+    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
+    \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}
+  \PY{p}{)}\PY{p}{;} 
+\PY{p}{\PYZcb{}} 
+\PY{c+c1}{//listing16}
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
@@
 \PY{p}{\PYZcb{}}\PY{p}{;}
-\PY{c+c1}{//listing09}
+\PY{c+c1}{//listing17}
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
@@
 \PY{p}{\PYZcb{}}\PY{p}{;}
-\PY{c+c1}{//listing10}
-  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{T1}\PY{p}{,} \PY{k}{class} \PY{n+nc}{T2}\PY{p}{,} \PY{k}{class} \PY{n+nc}{T3}\PY{o}{\PYZgt{}} 
-  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{F}\PY{p}{(}
-    \PY{k}{const} \PY{n}{T1} \PY{o}{\PYZam{}}\PY{n}{psi\PYZus{}l}\PY{p}{,} \PY{k}{const} \PY{n}{T2} \PY{o}{\PYZam{}}\PY{n}{psi\PYZus{}r}\PY{p}{,} \PY{k}{const} \PY{n}{T3} \PY{o}{\PYZam{}}\PY{n}{C}
-  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
-    \PY{p}{(}
-      \PY{p}{(}\PY{n}{C} \PY{o}{+} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}l} \PY{o}{+} 
-      \PY{p}{(}\PY{n}{C} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
-    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
-  \PY{p}{)}
-\PY{c+c1}{//listing11}
-  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}  
-  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{donorcell}\PY{p}{(} 
-    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
-    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
-  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
-    \PY{n}{F}\PY{p}{(}
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
-        \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
-    \PY{p}{)} \PY{o}{-}
-    \PY{n}{F}\PY{p}{(}
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
-        \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
-    \PY{p}{)}
-  \PY{p}{)}
-\PY{c+c1}{//listing12}
-  \PY{k+kt}{void} \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
-    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{k+kt}{int} \PY{n}{n}\PY{p}{,}
-    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
-    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
-  \PY{p}{)} \PY{p}{\PYZob{}} 
-    \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}
-      \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
-      \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}\PY{p}{;}
-  \PY{p}{\PYZcb{}}
-\PY{c+c1}{//listing13}
+\PY{c+c1}{//listing18}
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
@@
 \PY{p}{\PYZcb{}}\PY{p}{;}
-\PY{c+c1}{//listing14}
-  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{nom\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{den\PYZus{}t}\PY{o}{\PYZgt{}}
-  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
-    \PY{k}{const} \PY{n}{nom\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{nom}\PY{p}{,} \PY{k}{const} \PY{n}{den\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{den}
-  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
-    \PY{n}{where}\PY{p}{(}\PY{n}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nom} \PY{o}{/} \PY{n}{den}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}
-  \PY{p}{)} 
-\PY{c+c1}{//listing15}
-  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
-  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
-    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
-  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
-    \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,}
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}
-    \PY{p}{)} 
-  \PY{p}{)} 
-\PY{c+c1}{//listing16}
-  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
-  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
-    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
-  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
-   \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-}
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{,}
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+}
-      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}
-    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
-  \PY{p}{)}
-\PY{c+c1}{//listing17}
-  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
-  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}
-    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
-    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} 
-    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
-  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
-    \PY{p}{(}
-      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+}
-      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)}
-    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}
-  \PY{p}{)}
-\PY{c+c1}{//listing18}
-  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
-  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}
-    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
-    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,}
-    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}
-  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
-    \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}
-    \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{)}
-    \PY{o}{*} \PY{n}{mpdata\PYZus{}A}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} 
-    \PY{o}{-} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} 
-    \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
-    \PY{o}{*} \PY{n}{mpdata\PYZus{}B}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
-  \PY{p}{)} 
 \PY{c+c1}{//listing19}
@@
   \PY{c+c1}{// member fields}
-  \PY{n}{arrvec\PYZus{}t} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{;}
+  \PY{k}{static} \PY{k}{const} \PY{k+kt}{int} \PY{n}{n\PYZus{}tmp} \PY{o}{=} \PY{n}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{2} \PY{o}{?} \PY{l+m+mi}{2} \PY{o}{:} \PY{l+m+mi}{1}\PY{p}{;}
+  \PY{n}{arrvec\PYZus{}t} \PY{n}{tmp}\PY{p}{[}\PY{n}{n\PYZus{}tmp}\PY{p}{]}\PY{p}{;}
   \PY{n}{rng\PYZus{}t} \PY{n}{im}\PY{p}{,} \PY{n}{jm}\PY{p}{;}
@@
   \PY{p}{\PYZob{}}
-    \PY{k+kt}{int} \PY{n}{n\PYZus{}tmp} \PY{o}{=} \PY{n}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{2} \PY{o}{?} \PY{l+m+mi}{2} \PY{o}{:} \PY{l+m+mi}{1}\PY{p}{;}
     \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{n} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{n} \PY{o}{\PYZlt{}} \PY{n}{n\PYZus{}tmp}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{n}\PY{p}{)}

### 2013-07-15 paper/doc/paper.tex
the new order

@@
 
-%%%%% <gmd>
-\documentclass[gmd]{copernicus}%_discussions}
-%%%%% </gmd>
+%%%%% <sciprog>
+\documentclass[twocolumn]{article}
+\usepackage{lineno,authblk,latexsym,amsmath,float}
+\usepackage[colorlinks=true]{hyperref}
+\usepackage[hmargin=.55in,vmargin=.75in]{geometry}
+\usepackage[numbers]{natbib}
+\setlength{\columnsep}{1cm}
+%%%%% </sciprog>
 
@@
 \setitemize{noitemsep,itemsep=1pt,topsep=1pt,parsep=0pt,partopsep=0pt}
+
+% fr compact section headers
+\usepackage[compact]{titlesec}
+\titlespacing*{\section}{0pt}{*0}{0pt}
 
@@
   frame=single,
-  fontsize=\relscale{.666},
+  fontsize=\relscale{.6},
   fontfamily=courier,
@@
 \newcommand{\codecpp}[4]{%
+  \linespread{.8}
   \addtocounter{lstnocpp}{1}%
@@
   \setcounter{linenocpp}{\value{FancyVerbLine}}%
+  \vspace{-1.5ex}
+  \linespread{2}
 }
 \newcommand{\codepyt}[4]{%
+  \linespread{.8}
   \addtocounter{lstnopyt}{1}%
@@
   \setcounter{linenopyt}{\value{FancyVerbLine}}%
+  \vspace{-1.5ex}
+  \linespread{2}
 }
 \newcommand{\codefor}[4]{%
+  \linespread{.8}
   \addtocounter{lstnofor}{1}%
@@
   \setcounter{linenofor}{\value{FancyVerbLine}}%
+  \vspace{-1.5ex}
+  \linespread{2}
 }
@@
 
-\journal{Computer Physics Communications}
+\linespread{2}
 
@@
 
-    \author[1]{Sylwester Arabas}
+    \author[1]{Sylwester Arabas\footnote{Corresponence to: S. Arabas -- sarabas@igf.fuw.edu.pl -- 0048225546827 -- University of Warsaw, Pasteura 7, 02-093 Warszawa, Poland}}
     \author[1]{Dorota Jarecka}
@@
 
-    \correspondence{Sylwester Arabas (sarabas@igf.fuw.edu.pl)}
-    \runningauthor{Arabas, Jarecka et al.}
-    \runningtitle{Formula translation in C++, Python \& Fortran}
-
+    \onecolumn
     \maketitle
@@
           under the terms of the GNU GPL license.
+
+        ~\\
+        {\bf Keywords:} object-oriented programming, advection equation, MPDATA, C++, Fortran, Python
     \end{abstract}
 
-%    \begin{keyword}
-%      object-oriented programming, advection equation, MPDATA, C++, Fortran, Python
-%    \end{keyword}
-
-  \introduction
+  \twocolumn
+
+  \tableofcontents
+
+  \section{Introduction}
 
@@
     significantly contribute to code maintainability \citep{Wilson_et_al_2012}.
-  Code maintainability, in turn, plays a vital role in community programming efforts 
-    and hence geoscientific model development in particular.
 
@@
   For this purpose, a sample implementation of a numerical scheme for solving
-    the advection equation is introduced in in C++, Python and modern Fortran --
+    the advection equation is introduced in C++, Python and modern Fortran --
     OOP languages commonly used in scientific computing \citep[see e.g.][chapt.~8]{Einarsson_2005}. 
-  Presented benchmark tests, although quite simplistic, together with the experience gained 
-    from the development of codes in three different
-    languages provide a basis for discussion on the tradeoffs of programming language choice.
+  Presented implementations and the results of benchmark tests
+    provide a basis for discussion on the tradeoffs of programming language choice.
   The discussion concerns in principle the development of finite-difference solvers for 
-    partial differential equations, but is likely applicable to some extent to the scientific software in general.
+    partial differential equations, but is likely applicable to some extent to the scientific programming in general.
 
@@
     \citep[e.g.][respectively]{Ziemianski_et_al_2011,Abiodun_et_al_2011,Ezer_et_al_2002,Charbonneau_and_Smolarkiewicz_2013}.
-  Even in its most basic form presented herein, MPDATA is complex enough to contain
+  The basic MPDATA scheme presented herein is complex enough to contain
     a wide range of mathematical abstractions that can be represented using OOP constructs, 
@@
   In compliance with the article focus, these formul\ae~are presented without derivation or detailed discussion
-    \citep[see][for a recent review of MPDATA-based techniques including an introductory description 
-    of the algorithm and an exhaustive list of references]{Smolarkiewicz_2006}.
+    (see~\citep{Smolarkiewicz_2006} for a recent review of MPDATA-based techniques including an introductory description 
+    of the algorithm and an exhaustive list of references).
 
   The paper is structured as follows.
-  In section~\ref{sec:impl} we introduce the three implementations
+  In section~\ref{sec:impl} we introduce the ''formula translation'' part of the three implementations
     briefly describing the algorithm itself and
-    discussing where and how the OOP techniques may be applied in its implementation.
-  Subsections \ref{sec:array}-\ref{sec:solver} describe all three implementations,
-    while subsequent sections \ref{sec:cyclic}-\ref{sec:example} cover discussion of C++ code only.
-  The relevant parts of Python and Fortran codes do not differ significantly, and for readability reasons 
-    are presented in appendices \ref{app:P} and \ref{app:F}, respectively.
-  Section~\ref{sec:perf} of this paper covers performance evaluation of the three implementations.
+    discussing where and how the OOP techniques were applied in its implementation.
+  The remaining part of the implementations -- the solver logic -- is
+    presented in appendix~\ref{sec:app}.
+  Section~\ref{sec:perf} covers performance evaluation of the three implementations.
   Section~\ref{sec:tradeoffs} covers discussion of the tradeoffs of the programming language choice.
-  Section~\ref{sec:concl} section closes the article with a brief summary.
-  Those readers not interested in technical details may wish to skip to section~\ref{sec:perf} or
-    \ref{sec:tradeoffs}.
+  Section~\ref{sec:concl} closes the article with a brief summary.
 
@@
 
-  \section{Implementation}\label{sec:impl}
+  \section{Implementation: the formul\ae}\label{sec:impl}
 
@@
 
-  MPDATA is a solver for systems of advection equations of the following form:
+  MPDATA is, in its most basic form presented herein, a solver for systems of advection equations of the following form:
+
+  \vspace{-5ex}
   \begin{equation}\label{eq:adv}
@@
   that describe evolution of a scalar field $\psi$ transported by the fluid flow with velocity $\vec{v}$.
-  Solution of equation (\ref{eq:adv}) using MPDATA implies discretisation onto a grid of the $\psi$ and 
-    the Courant number $\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}$ fields, where $\Delta t$ is the 
+  Solution of equation (\ref{eq:adv}) using MPDATA implies discretisation onto a grid of~the~$\psi$~and~the~Courant 
+    number $\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}$ fields, where $\Delta t$ is the 
     solver timestep and $\Delta x$ is the grid spacing.
@@
     and two instances of it to represent unity and half
-    in expressions like \prog{i+one}, \prog{i+hlf}, where \prog{i} is an instance of \prog{slice}
+    in expressions like \prog{i+one}, \prog{i+hlf}, where \prog{i} is an instance of \prog{slice}%
     \footnote{\label{fnt:slice}One could argue that not using an own implementation of a slice-representing class
@@
 
-  \subsection{Halo regions}
-
-  The MPDATA formul\ae~defining $\psi^{[n+1]}_{[i,j]}$ as a function of $\psi^{[n]}_{[i,j]}$
-    (discussed in the following sections) feature terms such as $\psi_{[i-1,j-1]}$.
-  One way of assuring validity of these formul\ae~on the edges of the domain (e.g. for i=0) 
-    is to introduce the so-called halo region surrounding the domain.
-  The method of populating the halo region with data depends on the boundary condition type.
-  Employment of the halo-region logic implies repeated usage of array range 
-    extensions in the code such as $i \leadsto i \pm halo$.
-
-  An \prog{ext()} function is defined in all three implementation, in order to simplify 
-    coding of array range extensions:
-  \codecpp{code-cpp-listings.hpp}{listing06}{listing07}{1}
-  \codepyt{code-pyt-listings.py}{listing05}{listing06}{1}
-  \codefor{code-for-listings.f}{listing04}{listing05}{1}
-  Consequently, a range depicted by $i\pm1/2$ may be expressed in the code as \prog{ext(i, h)}.
-  In all three implementations the \prog{ext()} function accept the second
-    argument to be an integer or a ''half'' (cf. section \ref{sec:grid}).
-
   \subsection{Array index permutations}\label{sec:pi}
@@
     using the following notation:
+
+  \vspace{-5ex}
   \begin{equation*}\label{eq:pi}
@@
     (note the different order of \prog{i} and \prog{j} arguments in the two template specialisations):
-  \codecpp{code-cpp-listings.hpp}{listing07}{listing08}{1}
+  \codecpp{code-cpp-listings.hpp}{listing06}{listing07}{1}
 
@@
     single object.
-  Therefore, the following definition of function \prog{pi()} suffices to represent $\pi$:
-  \codepyt{code-pyt-listings.py}{listing06}{listing07}{1}
+  Therefore, the following definition of function \prog{pi()} suffices to represent~$\pi$:
+  \codepyt{code-pyt-listings.py}{listing05}{listing06}{1}
 
@@
     length of one of the vectors passed as argument is defined:
-  \codefor{code-for-listings.f}{listing05}{listing06}{1}
+  \codefor{code-for-listings.f}{listing04}{listing05}{1}
   The \prog{span()} function is used to shorten the declarations of arrays to be returned 
-    from functions in the Fortran implementation (see listings F.11 and F.17--F.20).
+    from functions in the Fortran implementation (see listings F.7 and F.12--F.15).
 
@@
 
-  \subsection{Prototype solver}\label{sec:solver}
-
-  The tasks to be handled by a prototype advection equation solver proposed herein are:
-  \begin{enumerate}[label=(\roman*), leftmargin=*, widest=iii]
-    \item{storing arrays representing the $\psi$ and $\vec{C}$ fields and any required housekeeping data,}
-    \item{allocating and deallocating the required memory,}
-    \item{providing access to the solver state,}
-    \item{performing the integration by invoking the advection-operator and
-      boundary-condition handling routines.}
-  \end{enumerate}
-  In the following C++ definition of the \prog{solver} structure, task (i) is represented with the definition of the structure
-    member fields; task (ii) is split between the \prog{solver}'s constructor and the destructors of \prog{arrvec\_t};
-    task (iii) is handled by the accessor methods; task (iv) is handled within the \prog{solve} method:
-  \codecpp{code-cpp-listings.hpp}{listing08}{listing09}{1}
-  The \prog{solver} structure is an abstract definition (containing a pure virtual method) 
-     requiring its descendants to implement at least
-     the \prog{advop()} method which is expected to fill \prog{psi[n+1]} with an updated (advected)
-     values of \prog{psi[n]}.
-  The two template parameters \prog{bcx\_t} and \prog{bcy\_t} allow the solver to operate with 
-    any kind of boundary condition structures that fulfil the requirements implied by the calls to the 
-    methods of \prog{bcx} and \prog{bcy}, respectively.
-
-  The donor-cell and MPDATA schemes both require only the previous state of an advected field
-    in order to advance the solution.
-  Consequently, memory for two time levels ($\psi^{[n]}$ and $\psi^{[n+1]}$) is allocated in the
-    constructor.
-  The sizes of the arrays representing the two time levels of $\psi$ 
-    are defined by the domain size ({\em nx}$~\times$~{\em ny}) plus the halo region.
-  The size of the halo region is an argument of the constructor.
-  The \prog{cycle()} method is used to swap the time levels without copying any data.
-
-  The arrays representing the $C^{[x]}$ and $C^{[y]}$ components of $\vec{C}$,
-    require ({\em nx+1})~$\times$~{\em ny} 
-    and {\em nx}$~\times$~({\em ny+1}) elements, respectively
-    (being laid out on the Arakawa-C staggered grid).
-
-  Python definition of the \prog{solver} class follows closely the C++ structure definition:
-  \codepyt{code-pyt-listings.py}{listing07}{listing08}{1}
-  The key difference stems from the fact that, unlike Blitz++, NumPy does not allow an array
-    to have arbitrary index base -- in NumPy the first element is always addressed with 0.
-  Consequently, while in C++ (and Fortran) the computational domain is chosen to start at (i=0, j=0)
-    and hence a part of the halo region to have negative indices, in Python the halo region starts 
-    at (0,0)\footnote{The reason to allow the domain to begin at an arbitrary index
-    is mainly to ease debugging in case the code would be used in parallel computations
-    using domain decomposition where each subdomain could have its own index base corresponding to
-    the location within the computational domain}.
-  However, since the whole halo logic is hidden within the solver, such details are not exposed to the
-    user.
-  The \prog{bcx} and \prog{bcy} boundary-condition specifications are passed to the solver through
-    constructor-like \prog{\_\_init\_\_()} method as opposed to template parameters in C++.
-
-  The above C++ and Python prototype solvers in principle allow to operate with any boundary condition
-    objects that implement methods called from within the solver.
-  This requirement is checked at compile-time in the case of C++, and at run-time in the case of Python.
-  In order to obtain an analogous behaviour with Fortran, it is required to
-    define, prior to definition of a solver type, 
-    an abstract type with deferred procedures having abstract interfaces 
-    \citep[sic!, see Table 2.1 in][for a summary of approximate correspondence of OOP 
-      nomenclature between Fortran and C++]{Rouson_et_al_2012}:
-  \codefor{code-for-listings.f}{listing06}{listing07}{1}
-  Having defined the abstract type for boundary-condition objects, 
-    a definition of a solver class following closely the C++ and Python counterparts may be provided:
-  \codefor{code-for-listings.f}{listing07}{listing08}{1}
-
-  \subsection{Periodic boundaries (C++)}\label{sec:cyclic}
-
-  From this point, only C++ implementation is explained in the main text. 
-  The Python and Fortran implementations are included in appendices A and B.
-
-  The solver definition described in section~\ref{sec:solver} requires a given
-    boundary condition object to implement a \prog{fill\_halos()} method.
-  An implementation of periodic boundary conditions in C++ is provided in the following listing:
-  \codecpp{code-cpp-listings.hpp}{listing09}{listing10}{1}
-
-  As hinted by the member field names, the \prog{fill\_halos()} methods
-    fill the left/right halo regions with data from the right/left edges of the domain.
-  Thanks to employment of the function \prog{pi()} described in section~\ref{sec:pi}
-    the same code may be applied in any dimension (here being a template parameter).
-
-  Listings P.8 and F.8 contain the Python and Fortran counterparts to listing C.9.
-
-  \subsection{Donor-cell formul\ae~(C++)}\label{sec:donor}
+  \subsection{Donor-cell formul\ae}\label{sec:donor}
+
   MPDATA is an iterative algorithm in which each iteration takes the form of the 
@@
   The donor-cell scheme may be written as \citep[eq. 2 in][]{Smolarkiewicz_1984}:
+
+  \vspace{-5ex}
   \begin{equation}\label{eq:donor}
@@
     and F is the so-called flux function \citep[][eq.~3]{Smolarkiewicz_1984}:
+
+  \vspace{-5ex}
   \begin{equation}\label{eq:donor:F}
@@
   \noindent The flux function takes the following form in C++:
-  \codecpp{code-cpp-listings.hpp}{listing10}{listing11}{1}
-
-  \noindent 
+  \codecpp{code-cpp-listings.hpp}{listing07}{listing08}{1}
+  \vspace{-3.5ex}
   Equation~\ref{eq:donor} is split into the terms under the summation 
     (effectively the 1-dimensional donor-cell formula):
-  \codecpp{code-cpp-listings.hpp}{listing11}{listing12}{1}
-
-  \noindent and the actual two-dimensional donor-cell formula:
-
-  \codecpp{code-cpp-listings.hpp}{listing12}{listing13}{1}
-
-  Listings P.9-P11 and F.9-F.13 contain the Python and Fortran counterparts to listings C.12-C.15.
-
-  \subsection{Donor-cell solver (C++)}\label{sec:donorcell_solver}
-
-  As mentioned in the previous section, the donor-cell formula
-    constitutes an advection scheme, hence we may use it to create
-    a \prog{solver\_donorcell} implementation of the abstract \prog{solver} class: 
-  \codecpp{code-cpp-listings.hpp}{listing13}{listing14}{1}
-  The above definition is given as an example only.
-  In the following sections an MPDATA solver of the same structure is defined.
-
-  Listings P.12 and F.14 contain the Python and Fortran counterparts to listing C.16.
-
-  \subsection{MPDATA formul\ae~(C++)}\label{sec:mpdata}
+  \codecpp{code-cpp-listings.hpp}{listing08}{listing09}{1}
+  \vspace{-3.5ex}
+  and the actual two-dimensional donor-cell formula:
+  \codecpp{code-cpp-listings.hpp}{listing09}{listing10}{1}
+  \vspace{-3.5ex}
+  The same formul\ae~are expressed as follows in Python:
+  \codepyt{code-pyt-listings.py}{listing06}{listing07}{1}
+  \codepyt{code-pyt-listings.py}{listing07}{listing08}{1}
+  \codepyt{code-pyt-listings.py}{listing08}{listing09}{1}
+  \vspace{-3.5ex}
+  The Fortran counterparts are: 
+  \codefor{code-for-listings.f}{listing05}{listing06}{1}
+  \codefor{code-for-listings.f}{listing06}{listing07}{1}
+  \codefor{code-for-listings.f}{listing07}{listing08}{1}
+  \codefor{code-for-listings.f}{listing08}{listing09}{1}
+  \codefor{code-for-listings.f}{listing09}{listing10}{1}
+
+  \subsection{MPDATA formul\ae}\label{sec:mpdata}
 
@@
     form \citep[eqs~13,~14~in][]{Smolarkiewicz_1984}:
+
+  \vspace{-5ex}
   \begin{equation}\label{eq:antidiff1}
@@
   where $\psi$ and $C$ represent values from the previous iteration and where:
+
+  \vspace{-5ex}
   \begin{equation}\label{eq:antidiff2}
@@
   }:
+
+  \vspace{-5ex}
   \begin{equation}\label{eq:A}
@@
     the corresponding $A_{[i,j]}$ and $B_{[i,j]}$ are set to zero what may be conveniently 
-    represented with the \prog{where} construct (available in all three considered languages):
+    represented with the \prog{where} construct in all three considered languages:
+  \codecpp{code-cpp-listings.hpp}{listing10}{listing11}{1}
+  \codepyt{code-pyt-listings.py}{listing09}{listing10}{1}
+  \codefor{code-for-listings.f}{listing10}{listing11}{1}
+  \codefor{code-for-listings.f}{listing11}{listing12}{1}
+
+  The $A$ term defined in equation \ref{eq:A} takes the following form:
+  \codecpp{code-cpp-listings.hpp}{listing11}{listing12}{1}
+  \codepyt{code-pyt-listings.py}{listing10}{listing11}{1}
+  \codefor{code-for-listings.f}{listing12}{listing13}{1}
+
+  The $B$ term defined in equation \ref{eq:B} takes the following form:
+  \codecpp{code-cpp-listings.hpp}{listing12}{listing13}{1}
+  \codepyt{code-pyt-listings.py}{listing11}{listing12}{1}
+  \codefor{code-for-listings.f}{listing13}{listing14}{1}
+
+  Equation \ref{eq:antidiff2} takes the following form:
+  \codecpp{code-cpp-listings.hpp}{listing13}{listing14}{1}
+  \codepyt{code-pyt-listings.py}{listing12}{listing13}{1}
+  \codefor{code-for-listings.f}{listing14}{listing15}{1}
+
+  Equation \ref{eq:antidiff1} takes the following form:
   \codecpp{code-cpp-listings.hpp}{listing14}{listing15}{1}
-
-  The $A$ term defined in equation \ref{eq:A} takes the following form:
-  \codecpp{code-cpp-listings.hpp}{listing15}{listing16}{1}
-
-  The $B$ term defined in equation \ref{eq:B} takes the following form:
-  \codecpp{code-cpp-listings.hpp}{listing16}{listing17}{1}
-
-  Equation \ref{eq:antidiff2} takes the following form:
-  \codecpp{code-cpp-listings.hpp}{listing17}{listing18}{1}
-
-  Equation \ref{eq:antidiff1} take the following form:
-  \codecpp{code-cpp-listings.hpp}{listing18}{listing19}{1}
-
-  Listings P.13-P.17 and F.15-F.21 contain the Python and Fortran counterparts to listing C.16-C.22.
-
-  \subsection{MPDATA solver (C++)}\label{sec:mpdatasolver}
-
-  An MPDATA solver may be now constructed by inheriting from \prog{solver} class
-    with the following definition in C++:
-  \codecpp{code-cpp-listings.hpp}{listing19}{listing20}{1}
-  The array of sequences of temporary arrays \prog{tmp} allocated in the constructor
-    is used to store the antidiffusive velocities from the present and optionally 
-    previous timestep (if using more than two iterations).
-
-  The \prog{advop()} method controlls the MPDATA iterations within one timestep.
-  The first (step = 0) iteration of MPDATA is an unmodified donor-cell step (compare listing C.15).
-  Subsequent iterations begin with calculation of the antidiffusive Courant fields using 
-    formula \ref{eq:antidiff1}.
-  In order to calculate values spanning an (i\mhlf~... i+\phlf) range using a formula
-    for $C_{[i+\phlf,\ldots]}$ only, the formula is evaluated using extended
-    index ranges \prog{im} and \prog{jm}.
-  In the second (step=1) iteration the uncorrected Courant field (\prog{C\_unco}) points
-    to the original \prog{C} field, and the antidiffusive Courant field is written into
-    \prog{C\_corr} which points to \prog{tmp[1]}.
-  In the third (step=2) iteration \prog{C\_unco} points to \prog{tmp[1]} while \prog{C\_corr}
-    points to \prog{tmp[0]}.
-  In subsequent iterations \prog{tmp[0]} and \prog{tmp[1]} are alternately swapped.
-
-  Listings P.18 and F.22 contain the Python and Fortran counterparts to listing C.23.
-
-  \subsection{Usage example (C++)}\label{sec:example}
-
-  The following listing provides an example of how the MPDATA solver
-    defined in section~\ref{sec:mpdatasolver} may be used together
-    with the cyclic boundary conditions defined in section~\ref{sec:cyclic}.
-  In the example a Gaussian signal is advected in a 2D domain 
-    defined over a grid of 24$\times$24 cells.
-  The program first plots the initial condition, then performs
-    the integration for 75 timesteps with three different 
-    settings of the number of iterations used in MPDATA.
-  The velocity field is constant in time and space (although it is
-    not assumed in the presented implementations).
-  The signal shape at the end of each simulation is plotted as well.
-  Plotting is done with the help of the gnuplot-iostream library\footnote{gnuplot-iostream
-    is a header-only C++ library allowing gnuplot to be controlled from C++, see \url{http://stahlke.org/dan/gnuplot-iostream/}. 
-    Gnuplot is a portable command-line driven graphing utility, see \url{http://gnuplot.info/}}.
-
-  The resultant plot is presented herein as Figure~\ref{fig:mpdata}.
-  The top panel depicts the initial condition. 
-  The three other panels show a snapshot of the field after 75 timesteps.
-  The donor-cell solution is characterised by strongest numerical diffusion
-    resulting in significant drop in the signal amplitude.
-  The signals advected using MPDATA show smaller numerical diffusion with
-    the solution obtained with more iterations preserving the signal altitude 
-    more accurately.
-  In all of the simulations the signal maintains its positive definiteness.
-  The domain periodicity is apparent in the plots as the maximum of the signal 
-    after 75 timesteps is located near the domain walls.
-
-  Listings P.19 and F.23-F.24 contain the Python and Fortran 
-    counterparts to listing C.24 (with the set-up and plotting logic omitted).
-
-  \codecpp{code-cpp-plot.cpp}{listing20}{listing21}{1}
-
-  \begin{figure}
-    \includegraphics[height=.9\textheight]{demo}
-    \caption{\label{fig:mpdata}
-      Plot generated by the program given in listing~C.24.
-      The top panel shows initial signal shape (at time t=0).
-      The subsequent panels show snapshots of the advected field after 75 timesteps
-        from three different simulations: donorcell (or 1 MPDATA iteration), MPDATA
-        with two iterations and MPDATA with 44 iterations.
-      The colour scale and the wire-frame surface correspond to signal amplitude.
-      See section~\ref{sec:example} for discussion.
-    }
-  \end{figure}
+  \codepyt{code-pyt-listings.py}{listing13}{listing14}{1}
+  \codefor{code-for-listings.f}{listing15}{listing16}{1}
+  \codefor{code-for-listings.f}{listing16}{listing17}{1}
 
   \section{Performance evaluation}\label{sec:perf}
+
+  \subsection{Set up}
 
@@
 
-  \begin{figure}[t]
+  \begin{figure}[h!]
     \center
@@
   \end{figure}
+
+  \subsection{Results}
 
@@
   \end{figure}
-  \begin{figure}[t]
+  \begin{figure}[th!]
     \center
@@
 
-  \section{Discussion on the tradeoffs of language choice}\label{sec:tradeoffs}
-
-  \subsection{OOP for blackboard abstractions}
+  \section{The tradeoffs of language choice}\label{sec:tradeoffs}
+
+  \subsection{Representability of blackboard abstractions}
 
@@
   Taking into account solely the part of code representing particular formul\ae~
-    (e.g. listings C.21, P.17, F.20 and equation \ref{eq:antidiff1}) all three
+    (e.g. listings C.14, P.13, F.15 and equation \ref{eq:antidiff1}) all three
     languages allow to match (or surpass) \LaTeX~in its brevity of formula translation syntax.
@@
     \item{loop-free array arithmetics;}
-    \item{definitions of functions returning array-valued expressions;}
+    \item{functions returning array-valued expressions;}
     \item{permutations of array indices allowing dimension-independent definitions
-      of functions (see e.g. listings C.12 and C.13, P.10 and P.11, F.11 and F.12);}
+      of functions (see e.g. listings C.8 and C.9, P.7 and P.8, F.7 and F.8);}
     \item{fractional indexing of arrays corresponding to employment of a staggered grid.}
@@
     \item{Fortran does not feature a mechanism allowing to reuse a single piece of code (algorithm)
-      with different data types (compare e.g. listings C.6, P.5 and F.4) such as
+      with different data types (compare e.g. listings C.15, P.14 and F.17) such as
       templates in C++ and the so-called duck typing in Python;}
@@
       of assignment (see e.g. how the \prog{ptr} pointers were used as a workaround in the \prog{cyclic\_fill\_halos}
-      method in listing F.8);}
+      method in listing F.20);}
     \item{Fortran lacks support for arrays of arrays (cf. Sect. \ref{sec:sequence}).}
@@
 
-  \subsection{Added values}
+  \subsection{User- and code-base}
 
@@
     there are apparently very few C++ libraries offering Blitz++ support out of the box
-    (the gnuplot-iostream used in listing C.24 being a much-appreciated counterexample).
+    (the gnuplot-iostream used in listing C.20 being a much-appreciated counterexample).
   However, Blitz++ allows to interface with virtually any library (including Fortran libraries), 
@@
 
-  \conclusions[Summary and outlook]\label{sec:concl}
+  \section{Summary and outlook}\label{sec:concl}
 
@@
 
-  %\setcounter{section}{15} 
-  \section{Python code for sections \ref{sec:cyclic}--\ref{sec:mpdatasolver}}\label{app:P}
-
-  \subsection*{{\bf Periodic Boundaries} (cf. Sect. \ref{sec:cyclic})}
-  \noindent 
-  \codepyt{code-pyt-listings.py}{listing08}{listing09}{1}
-
-  \subsection*{{\bf Donor-cell formul\ae}~(cf. Sect. \ref{sec:donor})}
-  \codepyt{code-pyt-listings.py}{listing09}{listing10}{1}
-  \codepyt{code-pyt-listings.py}{listing10}{listing11}{1}
-  \codepyt{code-pyt-listings.py}{listing11}{listing12}{1}
-
-  \subsection*{{\bf Donor-cell solver} (cf. Sect. \ref{sec:donorcell_solver})}
-  \codepyt{code-pyt-listings.py}{listing12}{listing13}{1}
-
-  \subsection*{{\bf MPDATA formul\ae}~(cf. Sect. \ref{sec:mpdata})}
-  \codepyt{code-pyt-listings.py}{listing13}{listing14}{1}
+  \section{Implementation: the solvers}\label{sec:app}
+  \subsection{Halo regions}
+
+  The MPDATA formul\ae~defining $\psi^{[n+1]}_{[i,j]}$ as a function of $\psi^{[n]}_{[i,j]}$
+    (discussed in the following sections) feature terms such as $\psi_{[i-1,j-1]}$.
+  One way of assuring validity of these formul\ae~on the edges of the domain (e.g. for i=0) 
+    is to introduce the so-called halo region surrounding the domain.
+  The method of populating the halo region with data depends on the boundary condition type.
+  Employment of the halo-region logic implies repeated usage of array range 
+    extensions in the code such as $i \leadsto i \pm halo$.
+
+  An \prog{ext()} function is defined in all three implementation, in order to simplify 
+    coding of array range extensions:
+  \codecpp{code-cpp-listings.hpp}{listing15}{listing16}{1}
   \codepyt{code-pyt-listings.py}{listing14}{listing15}{1}
+  \codefor{code-for-listings.f}{listing17}{listing18}{1}
+  Consequently, a range depicted by $i\pm1/2$ may be expressed in the code as \prog{ext(i, h)}.
+  In all three implementations the \prog{ext()} function accept the second
+    argument to be an integer or a ''half'' (cf. section \ref{sec:grid}).
+
+  \subsection{Prototype solver}\label{sec:solver}
+
+  The tasks to be handled by a prototype advection equation solver proposed herein are:
+  \begin{enumerate}[label=(\roman*), leftmargin=*, widest=iii]
+    \item{storing arrays representing the $\psi$ and $\vec{C}$ fields and any required housekeeping data,}
+    \item{allocating and deallocating the required memory,}
+    \item{providing access to the solver state,}
+    \item{performing the integration by invoking the advection-operator and
+      boundary-condition handling routines.}
+  \end{enumerate}
+  In the following C++ definition of the \prog{solver} structure, task (i) is represented with the definition of the structure
+    member fields; task (ii) is split between the \prog{solver}'s constructor and the destructors of \prog{arrvec\_t};
+    task (iii) is handled by the accessor methods; task (iv) is handled within the \prog{solve} method:
+  \codecpp{code-cpp-listings.hpp}{listing16}{listing17}{1}
+  The \prog{solver} structure is an abstract definition (containing a pure virtual method) 
+     requiring its descendants to implement at least
+     the \prog{advop()} method which is expected to fill \prog{psi[n+1]} with an updated (advected)
+     values of \prog{psi[n]}.
+  The two template parameters \prog{bcx\_t} and \prog{bcy\_t} allow the solver to operate with 
+    any kind of boundary condition structures that fulfil the requirements implied by the calls to the 
+    methods of \prog{bcx} and \prog{bcy}, respectively.
+
+  The donor-cell and MPDATA schemes both require only the previous state of an advected field
+    in order to advance the solution.
+  Consequently, memory for two time levels ($\psi^{[n]}$ and $\psi^{[n+1]}$) is allocated in the
+    constructor.
+  The sizes of the arrays representing the two time levels of $\psi$ 
+    are defined by the domain size ({\em nx}$~\times$~{\em ny}) plus the halo region.
+  The size of the halo region is an argument of the constructor.
+  The \prog{cycle()} method is used to swap the time levels without copying any data.
+
+  The arrays representing the $C^{[x]}$ and $C^{[y]}$ components of $\vec{C}$,
+    require ({\em nx+1})~$\times$~{\em ny} 
+    and {\em nx}$~\times$~({\em ny+1}) elements, respectively
+    (being laid out on the Arakawa-C staggered grid).
+
+  Python definition of the \prog{solver} class follows closely the C++ structure definition:
   \codepyt{code-pyt-listings.py}{listing15}{listing16}{1}
+  The key difference stems from the fact that, unlike Blitz++, NumPy does not allow an array
+    to have arbitrary index base -- in NumPy the first element is always addressed with 0.
+  Consequently, while in C++ (and Fortran) the computational domain is chosen to start at (i=0, j=0)
+    and hence a part of the halo region to have negative indices, in Python the halo region starts 
+    at (0,0)\footnote{The reason to allow the domain to begin at an arbitrary index
+    is mainly to ease debugging in case the code would be used in parallel computations
+    using domain decomposition where each subdomain could have its own index base corresponding to
+    the location within the computational domain}.
+  However, since the whole halo logic is hidden within the solver, such details are not exposed to the
+    user.
+  The \prog{bcx} and \prog{bcy} boundary-condition specifications are passed to the solver through
+    constructor-like \prog{\_\_init\_\_()} method as opposed to template parameters in C++.
+
+  The above C++ and Python prototype solvers in principle allow to operate with any boundary condition
+    objects that implement methods called from within the solver.
+  This requirement is checked at compile-time in the case of C++, and at run-time in the case of Python.
+  In order to obtain an analogous behaviour with Fortran, it is required to
+    define, prior to definition of a solver type, 
+    an abstract type with deferred procedures having abstract interfaces 
+    \citep[sic!, see Table 2.1 in][for a summary of approximate correspondence of OOP 
+      nomenclature between Fortran and C++]{Rouson_et_al_2012}:
+  \codefor{code-for-listings.f}{listing18}{listing19}{1}
+  Having defined the abstract type for boundary-condition objects, 
+    a definition of a solver class following closely the C++ and Python counterparts may be provided:
+  \codefor{code-for-listings.f}{listing19}{listing20}{1}
+
+  \subsection{Periodic boundaries}\label{sec:cyclic}
+
+  The solver definition described in section~\ref{sec:solver} requires a given
+    boundary condition object to implement a \prog{fill\_halos()} method.
+  An implementation of periodic boundary conditions in C++ is provided in the following listing:
+  \codecpp{code-cpp-listings.hpp}{listing17}{listing18}{1}
+
+  As hinted by the member field names, the \prog{fill\_halos()} methods
+    fill the left/right halo regions with data from the right/left edges of the domain.
+  Thanks to employment of the function \prog{pi()} described in section~\ref{sec:pi}
+    the same code may be applied in any dimension (here being a template parameter).
+
+  The following listings contain the Python and Fortran counterparts to listing C.17.
   \codepyt{code-pyt-listings.py}{listing16}{listing17}{1}
+  \codefor{code-for-listings.f}{listing20}{listing21}{1}
+
+  \subsection{Donor-cell solver}\label{sec:donorcell_solver}
+
+  As mentioned in the previous section, the donor-cell formula
+    constitutes an advection scheme, hence we may use it to create
+    a \prog{solver\_donorcell} implementation of the abstract \prog{solver} class: 
+  \codecpp{code-cpp-listings.hpp}{listing18}{listing19}{1}
+  The above definition is given as an example only.
+  In the following sections an MPDATA solver of the same structure is defined.
+
+  The following listings contain the Python and Fortran counterparts to listing C.18.
   \codepyt{code-pyt-listings.py}{listing17}{listing18}{1}
-
-  \subsection*{{\bf An MPDATA solver} (cf. Sect. \ref{sec:mpdatasolver})}
+  \codefor{code-for-listings.f}{listing21}{listing2}{1}
+
+  \subsection{MPDATA solver}\label{sec:mpdatasolver}
+
+  An MPDATA solver may be now constructed by inheriting from \prog{solver} class
+    with the following definition in C++:
+  \codecpp{code-cpp-listings.hpp}{listing19}{listing20}{1}
+  The array of sequences of temporary arrays \prog{tmp} allocated in the constructor
+    is used to store the antidiffusive velocities from the present and optionally 
+    previous timestep (if using more than two iterations).
+
+  The \prog{advop()} method controlls the MPDATA iterations within one timestep.
+  The first (step = 0) iteration of MPDATA is an unmodified donor-cell step.
+  Subsequent iterations begin with calculation of the antidiffusive Courant fields using 
+    formula \ref{eq:antidiff1}.
+  In order to calculate values spanning an (i\mhlf~... i+\phlf) range using a formula
+    for $C_{[i+\phlf,\ldots]}$ only, the formula is evaluated using extended
+    index ranges \prog{im} and \prog{jm}.
+  In the second (step=1) iteration the uncorrected Courant field (\prog{C\_unco}) points
+    to the original \prog{C} field, and the antidiffusive Courant field is written into
+    \prog{C\_corr} which points to \prog{tmp[1]}.
+  In the third (step=2) iteration \prog{C\_unco} points to \prog{tmp[1]} while \prog{C\_corr}
+    points to \prog{tmp[0]}.
+  In subsequent iterations \prog{tmp[0]} and \prog{tmp[1]} are alternately swapped.
+
+  The following listings contain the Python and Fortran counterparts to listing C.19.
   \codepyt{code-pyt-listings.py}{listing18}{listing19}{1}
-
-  \subsection*{{\bf Usage example} (cf. Sect. \ref{sec:example})}
+  \codefor{code-for-listings.f}{listing22}{listing23}{1}
+
+  \subsection{Usage example}\label{sec:example}
+
+  The following listing provides an example of how the MPDATA solver
+    defined in section~\ref{sec:mpdatasolver} may be used together
+    with the cyclic boundary conditions defined in section~\ref{sec:cyclic}.
+  In the example a Gaussian signal is advected in a 2D domain 
+    defined over a grid of 24$\times$24 cells.
+  The program first plots the initial condition, then performs
+    the integration for 75 timesteps with three different 
+    settings of the number of iterations used in MPDATA.
+  The velocity field is constant in time and space (although it is
+    not assumed in the presented implementations).
+  The signal shape at the end of each simulation is plotted as well.
+  Plotting is done with the help of the gnuplot-iostream library\footnote{gnuplot-iostream
+    is a header-only C++ library allowing gnuplot to be controlled from C++, see \url{http://stahlke.org/dan/gnuplot-iostream/}. 
+    Gnuplot is a portable command-line driven graphing utility, see \url{http://gnuplot.info/}}.
+
+  The resultant plot is presented herein as Figure~\ref{fig:mpdata}.
+  The top panel depicts the initial condition. 
+  The three other panels show a snapshot of the field after 75 timesteps.
+  The donor-cell solution is characterised by strongest numerical diffusion
+    resulting in significant drop in the signal amplitude.
+  The signals advected using MPDATA show smaller numerical diffusion with
+    the solution obtained with more iterations preserving the signal altitude 
+    more accurately.
+  In all of the simulations the signal maintains its positive definiteness.
+  The domain periodicity is apparent in the plots as the maximum of the signal 
+    after 75 timesteps is located near the domain walls.
+  \codecpp{code-cpp-plot.cpp}{listing20}{listing21}{1}
+
+  The following listings contain the Python and Fortran 
+    counterparts to listing C.20 (with the set-up and plotting logic omitted).
   \codepyt{code-pyt-test.py}{listing19}{listing20}{1}
-
-  %\setcounter{section}{5} 
-  \section{Fortran code for sections \ref{sec:cyclic}--\ref{sec:mpdatasolver}}\label{app:F}
-
-  \subsection*{{\bf Periodic boundaries} (cf. Sect. \ref{sec:cyclic})}
-  \codefor{code-for-listings.f}{listing08}{listing09}{1}
-
-  \subsection*{{\bf Donor-cell formul\ae}~(cf. Sect. \ref{sec:donor})}
-  \codefor{code-for-listings.f}{listing09}{listing10}{1}
-  \codefor{code-for-listings.f}{listing10}{listing11}{1}
-  \codefor{code-for-listings.f}{listing11}{listing12}{1}
-  \codefor{code-for-listings.f}{listing12}{listing13}{1}
-  \codefor{code-for-listings.f}{listing13}{listing14}{1}
-
-  \subsection*{{\bf Donor-cell solver} (cf. Sect. \ref{sec:donorcell_solver})}
-  \codefor{code-for-listings.f}{listing14}{listing15}{1}
-
-  \subsection*{{\bf MPDATA formul\ae}~(cf. Sect. \ref{sec:mpdata})}
-  \codefor{code-for-listings.f}{listing15}{listing16}{1}
-  \codefor{code-for-listings.f}{listing16}{listing17}{1}
-  \codefor{code-for-listings.f}{listing17}{listing18}{1}
-  \codefor{code-for-listings.f}{listing18}{listing19}{1}
-  \codefor{code-for-listings.f}{listing19}{listing20}{1}
-  \codefor{code-for-listings.f}{listing20}{listing21}{1}
-  \codefor{code-for-listings.f}{listing21}{listing22}{1}
-
-  \subsection*{{\bf An MPDATA solver} (cf. Sect. \ref{sec:mpdatasolver})}
-  \codefor{code-for-listings.f}{listing22}{listing23}{1}
-
-  \subsection*{{\bf Usage example} (cf. Sect. \ref{sec:example})}
   \codefor{code-for-test.f}{listing23}{listing24}{1}
@@
 
-  \begin{acknowledgements}
+  \small
+
+  \section*{Acknowledgements}
+  {
     We thank Piotr Smolarkiewicz and Hanna Pawłowska for their help throughout the project.
@@
     Development of NumPy support in PyPy was led by Alex Gaynor, Matti Picus and MF.
-  \end{acknowledgements}
-
-  \bibliographystyle{copernicus}
+  }
+
+  \bibliographystyle{plainnat}
   \bibliography{paper}
 
+  \begin{figure}[bh!]
+    \caption{\label{fig:mpdata}
+      Plot generated by the program given in listing~C.20.
+      The top panel shows initial signal shape (at time t=0).
+      The subsequent panels show snapshots of the advected field after 75 timesteps
+        from three different simulations: donorcell (or 1 MPDATA iteration), MPDATA
+        with two iterations and MPDATA with 44 iterations.
+      The colour scale and the wire-frame surface correspond to signal amplitude.
+      See section~\ref{sec:example} for discussion.
+    }
+  \end{figure}
+  \includegraphics[height=\textheight]{demo}
+
 \end{document}

### 2013-07-15 paper/cpp/listings.hpp

@@ -34,14 +34,6 @@ inline rng_t operator-(const rng_t &i, const hlf_t &)
   return i-1; 
 }
 //listing06
-template<class n_t>
-inline rng_t ext(const rng_t &r, const n_t &n) { 
-  return rng_t(
-    (r - n).first(), 
-    (r + n).last()
-  ); 
-} 
-//listing07
 template<int d> 
 inline idx_t pi(const rng_t &i, const rng_t &j);
 
@@ -56,7 +48,107 @@ inline idx_t pi<1>(const rng_t &j, const rng_t &i)
 {
   return idx_t({i,j});
 }; 
+//listing07
+template<class T1, class T2, class T3> 
+inline auto F(
+  const T1 &psi_l, const T2 &psi_r, const T3 &C
+) return_macro(
+  (
+    (C + abs(C)) * psi_l + 
+    (C - abs(C)) * psi_r
+  ) / 2
+)
 //listing08
+template<int d>  
+inline auto donorcell( 
+  const arr_t &psi, const arr_t &C, 
+  const rng_t &i, const rng_t &j
+) return_macro(
+  F(
+    psi(pi<d>(i,   j)), 
+    psi(pi<d>(i+1, j)), 
+      C(pi<d>(i+h, j))
+  ) -
+  F(
+    psi(pi<d>(i-1, j)), 
+    psi(pi<d>(i,   j)), 
+      C(pi<d>(i-h, j))
+  )
+)
+//listing09
+void donorcell_op(
+  const arrvec_t &psi, const int n,
+  const arrvec_t &C, 
+  const rng_t &i, const rng_t &j
+) { 
+  psi[n+1](i,j) = psi[n](i,j)
+    - donorcell<0>(psi[n], C[0], i, j)
+    - donorcell<1>(psi[n], C[1], j, i); 
+}
+//listing10
+template<class nom_t, class den_t>
+inline auto mpdata_frac(
+  const nom_t &nom, const den_t &den
+) return_macro(
+  where(den > 0, nom / den, 0)
+) 
+//listing11
+template<int d>
+inline auto mpdata_A(const arr_t &psi, 
+  const rng_t &i, const rng_t &j
+) return_macro(
+  mpdata_frac(
+    psi(pi<d>(i+1, j)) - psi(pi<d>(i,j)),
[…]

### 2013-07-15 paper/for/listings.f

@@ -83,34 +83,6 @@ module arakawa_c_m
   end function
 end module
 !listing04
-module halo_m
-  use arakawa_c_m
-  implicit none
-
-  interface ext
-    module procedure ext_n
-    module procedure ext_h
-  end interface 
-
-  contains
-
-  function ext_n(r, n) result (return)
-    integer, intent(in) :: r(2)
-    integer, intent(in) :: n
-    integer :: return(2)
-    
-    return = (/ r(1) - n, r(2) + n /)
-  end function
-
-  function ext_h(r, h) result (return)
-    integer, intent(in) :: r(2)
-    type(half_t), intent(in) :: h
-    integer :: return(2)
-    
-    return = (/ r(1) - h, r(2) + h /)
-  end function
-end module
-!listing05
 module pi_m
   use real_m
   implicit none
@@ -140,7 +112,156 @@ module pi_m
     end select
   end function
 end module
+!listing05
+module donorcell_m
+  use real_m
+  use arakawa_c_m
+  use pi_m
+  use arrvec_m
+  implicit none
+  contains 
 !listing06
+  elemental function F(psi_l, psi_r, C) result (return)
+    real(real_t) :: return
+    real(real_t), intent(in) :: psi_l, psi_r, C
+    return = (                                         &
+      (C + abs(C)) * psi_l +                           &
+      (C - abs(C)) * psi_r                             &
+    ) / 2
+  end function
+!listing07
+  function donorcell(d, psi, C, i, j) result (return)
+    integer :: d
+    integer, intent(in) :: i(2), j(2) 
+    real(real_t) :: return(span(d, i, j), span(d, j, i))
+    real(real_t), allocatable, intent(in) :: psi(:,:), C(:,:)           
+    return = (                                         &
+      F(                                               &
+        pi(d, psi, i,   j),                            &
+        pi(d, psi, i+1, j),                            &
+        pi(d,   C, i+h, j)                             &
+      ) -                                              &
+      F(                                               &
+        pi(d, psi, i-1, j),                            &
+        pi(d, psi, i,   j),                            &
+        pi(d, C,   i-h, j)                             &
[…]

### 2013-07-15 paper/pyt/listings.py

@@ -34,6 +34,68 @@ class shift():
 one = shift(1,1) 
 hlf = shift(0,1)
 #listing05
+def pi(d, *idx): 
+  return (idx[d], idx[d-1])
+#listing06
+def f(psi_l, psi_r, C):
+  return (
+    (C + abs(C)) * psi_l + 
+    (C - abs(C)) * psi_r
+  ) / 2
+#listing07
+def donorcell(d, psi, C, i, j):
+  return (
+    f(
+      psi[pi(d, i,     j)], 
+      psi[pi(d, i+one, j)], 
+        C[pi(d, i+hlf, j)]
+    ) - 
+    f(
+      psi[pi(d, i-one, j)], 
+      psi[pi(d, i,     j)], 
+        C[pi(d, i-hlf, j)]
+    ) 
+  )
+#listing08
+def donorcell_op(psi, n, C, i, j):
+  psi[n+1][i,j] = (psi[n][i,j] 
+    - donorcell(0, psi[n], C[0], i, j)
+    - donorcell(1, psi[n], C[1], j, i)
+  )
+#listing09
+def mpdata_frac(nom, den):
+  return numpy.where(den > 0, nom/den, 0)
+#listing10
+def mpdata_A(d, psi, i, j):
+  return mpdata_frac(
+    psi[pi(d, i+one, j)] - psi[pi(d, i, j)],
+    psi[pi(d, i+one, j)] + psi[pi(d, i, j)]
+  )
+#listing11
+def mpdata_B(d, psi, i, j):
+  return mpdata_frac( 
+    psi[pi(d, i+one, j+one)] + psi[pi(d, i, j+one)] -
+    psi[pi(d, i+one, j-one)] - psi[pi(d, i, j-one)],
+    psi[pi(d, i+one, j+one)] + psi[pi(d, i, j+one)] +
+    psi[pi(d, i+one, j-one)] + psi[pi(d, i, j-one)]
+  ) / 2
+#listing12
+def mpdata_C_bar(d, C, i, j):
+  return (
+    C[pi(d, i+one, j+hlf)] + C[pi(d, i,  j+hlf)] +
+    C[pi(d, i+one, j-hlf)] + C[pi(d, i,  j-hlf)] 
+  ) / 4
+#listing13
+def mpdata_C_adf(d, psi, i, j, C):
+  return (
+    abs(C[d][pi(d, i+hlf, j)]) 
[…]

### 2013-07-17 paper/doc/paper.tex
adding mention of C++11 and F2008 in the abstract

@@
         The introduced codes serve as examples for how the application of 
-          object-oriented programming (OOP) techniques allows to reproduce the mathematical notation 
+          object-oriented programming (OOP) techniques and new language constructs from
+          C++11 and Fortran 2008 allow to reproduce the mathematical notation 
           used in the literature within the program code.

### 2013-07-26 paper/doc/paper.tex
fixing the donor-cell parenthesis issue; paper.tex updated befor Sci. Prog. submission

@@
 
-    \author[1]{Sylwester Arabas\footnote{Corresponence to: S. Arabas -- sarabas@igf.fuw.edu.pl -- 0048225546827 -- University of Warsaw, Pasteura 7, 02-093 Warszawa, Poland}}
-    \author[1]{Dorota Jarecka}
-    \author[1]{Anna Jaruga}
-    \author[2]{Maciej Fijałkowski}
-
-    \affil[1]{Institute of Geophysics, Faculty of Physics, University of Warsaw}
-    \affil[2]{PyPy Team}
+    \author{~}
+%    \author[1]{Sylwester Arabas\footnote{Correspondence to: S. Arabas -- sarabas@igf.fuw.edu.pl -- 0048225546827 -- University of Warsaw, Pasteura 7, 02-093 Warszawa, Poland}}
+%    \author[1]{Dorota Jarecka}
+%    \author[1]{Anna Jaruga}
+%    \author[2]{Maciej Fijałkowski}
+
+%    \affil[1]{Institute of Geophysics, Faculty of Physics, University of Warsaw}
+%    \affil[2]{PyPy Team}
 
@@
     weather-prediction \citep{Sundberg_2009} and Earth system \citep{Legutke_2012} modelling, 
-    and none of its 20-century editions were object-oriented languages \citep[see e.g.][for discussion]{Norton_et_al_2007}}.
+    and none of its 20\textsuperscript{th}-century editions were object-oriented languages \citep[see e.g.][for discussion]{Norton_et_al_2007}}.
 
@@
   The importance of reproducing the mathematical notation in the code lays primarily
-    in the fact that code readability and brevity \citep[and not the language choice itself,][]{Seeley_2004}
+    in the fact that code readability and brevity 
+    %\citep[and not the language choice itself,][]{Seeley_2004}
     significantly contribute to code maintainability \citep{Wilson_et_al_2012}.
@@
     allowing comparison of the relevant syntax with the mathematical notation.
-  In compliance with the article focus, these formul\ae~are presented without derivation or detailed discussion
+  These formul\ae~are presented without derivation or detailed discussion
     (see~\citep{Smolarkiewicz_2006} for a recent review of MPDATA-based techniques including an introductory description 
@@
   The remaining part of the implementations -- the solver logic -- is
-    presented in appendix~\ref{sec:app}.
+    presented in appendix~\ref{app:solver}.
+  A usage example is given in appendix~\ref{app:usage}
   Section~\ref{sec:perf} covers performance evaluation of the three implementations.
@@
 
-  \section{Implementation: the formul\ae}\label{sec:impl}
+  \section{Implementation of the formul\ae}\label{sec:impl}
 
@@
 
-  In NumPy in order to prevent copying of array data during slicing one needs to operate on the
+  In NumPy, in order to prevent copying of array data during slicing, one needs to operate on the
     so-called array views.
@@
 
-  In the Fortran implementation \prog{pi()} returns a pointer to the array elements specified
+  Fortran does not feature an analogous mechanism for specifying array ranges in multiple
+    dimensions using a single entity.
+  As a workaround, in in the Fortran implementation, \prog{pi()} returns a pointer to the array elements specified
     by \prog{i} and \prog{j} interpreted as (i,j) or
     (j,i) depending on the value of the argument \prog{d}.
+% TODO: disucss creation of temporary arrays and disabling of vectorisation due to employment of POINTERS
   In addition to \prog{pi()}, a helper \prog{span()} function returning the 
@@
 
+  The brevity of the code in the above listings as well as its similarity
+    to the mathematical notation is the main point of this paper.
+  The ''formula translation'' features include:
+  \begin{itemize}
+    \item{loop-free notation;}
+    \item{array-valued functions enabling reuse of sub-expressions;}
+    \item{fractional indexing obtained with the help of operator overloading;}
+    \item{dimension-independent indexing with the help of \prog{pi} permutation functions.}
+  \end{itemize}
+  The same features are applied to translation of more complex formul\ae~in the
+    following section.
+
   \subsection{MPDATA formul\ae}\label{sec:mpdata}
@@
 
-  If the denominator in equations \ref{eq:A} or \ref{eq:B} equals zero for a given {\em i} and {\em j}, 
+  If the (positive-defined) denominator in equations \ref{eq:A} or \ref{eq:B} 
+    equals zero for a given {\em i} and {\em j}, %TODO: >0 in the code
     the corresponding $A_{[i,j]}$ and $B_{[i,j]}$ are set to zero what may be conveniently 
@@
   \codefor{code-for-listings.f}{listing16}{listing17}{1}
+
+  The above listings conclude the formula-translation part of this paper.
+  Implementation of a prototype MPDATA solver using the above code
+    is presented in appendix~\ref{app:solver}.
 
@@
   PyPy is an alternative implementation of Python featuring a just-in-time compiler. 
-  PyPy includes an experimental partial reimplementation of NumPy that compiles NumPy expressions into native assembler.
+  PyPy includes an experimental partial re-implementation of NumPy that compiles NumPy expressions into native assembler.
   Thanks to employment of lazy evaluation of array expressions (cf. Sect. \ref{sec:array})
@@
   Each test set consisted of 16 program runs.
-  The test programs are analogous to the example code presented in section~\ref{sec:example}.
+  The test programs are analogous to the example code presented in appendix~\ref{app:usage}.
   The tests were run with different grid sizes ranging from 64$\times$64 to 2048$\times$2048.
@@
   Three MPDATA iterations were used (i.e. two corrective steps).
-  The initial condition was loaded from a text file, and the final values were compared at the end of the test
-    with values loaded from another text file assuring the same results were obtained with all four set-ups.
+  %The initial condition was loaded from a text file, and the final values were compared at the end of the test
+  %  with values loaded from another text file assuring the same results were obtained with all four set-ups.
   The tests were run multiple times; program start-up, data loading, and output verification times were
@@
 
-  The CPU time statistics presented in Figures \ref{fig:cpu-eyrie} and \ref{fig:cpu-skua} reveal
+  The CPU time statistics presented in Figures \ref{fig:cpu-eyrie} and~\ref{fig:cpu-skua} reveal
     minor differences between results obtained with the two different processors.
@@
   \begin{itemize}
+    \item{Fortran lacks support for specifying array ranges in multiple dimensions
+      with a single entity (cf. tuples of slices in Python and blitz::RectDomain in C++);
+    }
     \item{Fortran does not feature a mechanism allowing to reuse a single piece of code (algorithm)
@@
 
-  With both Fortran and Python, the memory footprint caused by employment
+  Blitz++ ensures temporary-array-free computations by design
+    \citep{Veldhuizen_et_al_1997} avoiding unintentional performance loss.
+  In contrast, with both Fortran and Python, the memory footprint caused by employment
     of temporary objects in array arithmetics is dependant on compiler choice or
     the level of optimisations.
-  In contrast, Blitz++ ensures temporary-array-free computations by design
-    \citep{Veldhuizen_et_al_1997} avoiding unintentional performance loss.
-
-  \subsection{User- and code-base}
+
+  \subsection{Developers' community and libraries}%User- and code-base}
 
@@
 
-  Finally, the three languages differ as well with regard to availability of 
+  The three languages differ as well with regard to availability of 
     mechanisms for leveraging shared-memory parallelisation (e.g. with multi-core processors).
@@
     in Python, C++ and Fortran, respectively.  
+  It is the authors' impression that these figures are somehow indicative of the programming 
+    effort.
 
@@
 
-  \section{Implementation: the solvers}\label{sec:app}
+  \section{Prototype solvers}\label{app:solver}
+
+  The following sections provide a complete description of a minimal example of application
+    of the formul\ae~''translated'' into C++, Python and Fortran in the main body of the paper.
+
   \subsection{Halo regions}
@@
 
-  The \prog{advop()} method controlls the MPDATA iterations within one timestep.
+  The \prog{advop()} method controls the MPDATA iterations within one timestep.
   The first (step = 0) iteration of MPDATA is an unmodified donor-cell step.
@@
 
-  \subsection{Usage example}\label{sec:example}
+  \section{Usage example}\label{app:usage}
 
@@
 
-  \small
+  \footnotesize
 
@@
       The colour scale and the wire-frame surface correspond to signal amplitude.
-      See section~\ref{sec:example} for discussion.
+      See appendix~\ref{app:usage} for discussion.
     }

### 2013-07-26 paper/cpp/listings.hpp

@@ -81,9 +81,10 @@ void donorcell_op(
   const arrvec_t &C, 
   const rng_t &i, const rng_t &j
 ) { 
-  psi[n+1](i,j) = psi[n](i,j)
-    - donorcell<0>(psi[n], C[0], i, j)
-    - donorcell<1>(psi[n], C[1], j, i); 
+  psi[n+1](i,j) = psi[n](i,j) - (
+    donorcell<0>(psi[n], C[0], i, j) +
+    donorcell<1>(psi[n], C[1], j, i)
+  ); 
 }
 //listing10
 template<class nom_t, class den_t>

### 2013-07-26 paper/for/listings.f

@@ -157,9 +157,10 @@ module donorcell_m
     
     real(real_t), pointer :: ptr(:,:)
     ptr => pi(0, psi%at(n+1)%p%a, i, j)
-    ptr = pi(0, psi%at(n)%p%a, i, j)                   &
-      - donorcell(0, psi%at(n)%p%a, C%at(0)%p%a, i, j) &
-      - donorcell(1, psi%at(n)%p%a, C%at(1)%p%a, j, i)
+    ptr = pi(0, psi%at(n)%p%a, i, j) - (               &
+      donorcell(0, psi%at(n)%p%a, C%at(0)%p%a, i, j) + &
+      donorcell(1, psi%at(n)%p%a, C%at(1)%p%a, j, i)   &
+    )
   end subroutine
 !listing09
 end module

### 2013-07-26 paper/pyt/listings.py

@@ -58,9 +58,9 @@ def donorcell(d, psi, C, i, j):
   )
 #listing08
 def donorcell_op(psi, n, C, i, j):
-  psi[n+1][i,j] = (psi[n][i,j] 
-    - donorcell(0, psi[n], C[0], i, j)
-    - donorcell(1, psi[n], C[1], j, i)
+  psi[n+1][i,j] = psi[n][i,j] - (
+    donorcell(0, psi[n], C[0], i, j) +
+    donorcell(1, psi[n], C[1], j, i)
   )
 #listing09
 def mpdata_frac(nom, den):

### 2014-02-19 paper/doc/code-cpp-listings.hpp.tex
changes before Sci Prog resubmission

@@
 \PY{c+c1}{//listing01}
-\PY{k}{using} \PY{n}{real\PYZus{}t} \PY{o}{=} \PY{k+kt}{double}\PY{p}{;}
+\PY{k}{using} \PY{k+kt}{real\PYZus{}t} \PY{o}{=} \PY{k+kt}{double}\PY{p}{;}
 \PY{c+c1}{//listing02}
 \PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}blitz}\PY{c+cp}{/}\PY{c+cp}{array.h\PYZgt{}}
-\PY{k}{using} \PY{n}{arr\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Array}\PY{o}{\PYZlt{}}\PY{n}{real\PYZus{}t}\PY{p}{,} \PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
-\PY{k}{using} \PY{n}{rng\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Range}\PY{p}{;}
-\PY{k}{using} \PY{n}{idx\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{RectDomain}\PY{o}{\PYZlt{}}\PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
+\PY{k}{using} \PY{k+kt}{arr\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Array}\PY{o}{\PYZlt{}}\PY{k+kt}{real\PYZus{}t}\PY{p}{,} \PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
+\PY{k}{using} \PY{k+kt}{rng\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Range}\PY{p}{;}
+\PY{k}{using} \PY{k+kt}{idx\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{RectDomain}\PY{o}{\PYZlt{}}\PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
 \PY{c+c1}{//listing03}
 \PY{c+cp}{\PYZsh{}}\PY{c+cp}{define return\PYZus{}macro(expr) \PYZbs{}}
-\PY{c+cp}{  -\PYZgt{} decltype(safeToReturn(expr)) \PYZbs{}}
+\PY{c+cp}{  \PYZhy{}\PYZgt{} decltype(safeToReturn(expr)) \PYZbs{}}
 \PY{c+cp}{\PYZob{} return safeToReturn(expr); \PYZcb{} }
@@
 \PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}boost}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}container}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}vector.hpp\PYZgt{}}
-\PY{k}{struct} \PY{n}{arrvec\PYZus{}t} \PY{o}{:} \PY{n}{boost}\PY{o}{:}\PY{o}{:}\PY{n}{ptr\PYZus{}vector}\PY{o}{\PYZlt{}}\PY{n}{arr\PYZus{}t}\PY{o}{\PYZgt{}} 
+\PY{k}{struct} \PY{k+kt}{arrvec\PYZus{}t} \PY{o}{:} \PY{n}{boost}\PY{o}{:}\PY{o}{:}\PY{n}{ptr\PYZus{}vector}\PY{o}{\PYZlt{}}\PY{k+kt}{arr\PYZus{}t}\PY{o}{\PYZgt{}} 
 \PY{p}{\PYZob{}}
-  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{k}{operator}\PY{p}{[}\PY{p}{]}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{i}\PY{p}{)} \PY{k}{const} 
+  \PY{k}{const} \PY{k+kt}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{k}{operator}\PY{p}{[}\PY{p}{]}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{i}\PY{p}{)} \PY{k}{const} 
   \PY{p}{\PYZob{}}   
-    \PY{k}{return} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{at}\PY{p}{(}\PY{p}{(}\PY{n}{i} \PY{o}{+} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{;}
+    \PY{k}{return} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{at}\PY{p}{(}\PY{p}{(}\PY{n}{i} \PY{o}{+} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{;}
   \PY{p}{\PYZcb{}}
@@
 \PY{c+c1}{//listing05}
-\PY{k}{struct} \PY{n}{hlf\PYZus{}t} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}} \PY{n}{h}\PY{p}{;}
+\PY{k}{struct} \PY{k+kt}{hlf\PYZus{}t} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}} \PY{n}{h}\PY{p}{;}
 
-\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{+}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)}
+\PY{k+kr}{inline} \PY{k+kt}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{+}\PY{p}{(}\PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)}
 \PY{p}{\PYZob{}} 
@@
 
-\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{-}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)}
+\PY{k+kr}{inline} \PY{k+kt}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{\PYZhy{}}\PY{p}{(}\PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)}
 \PY{p}{\PYZob{}} 
-  \PY{k}{return} \PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{;} 
+  \PY{k}{return} \PY{n}{i}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{;} 
 \PY{p}{\PYZcb{}}
@@
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}} 
-\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}\PY{p}{;}
+\PY{k+kr}{inline} \PY{k+kt}{idx\PYZus{}t} \PY{n}{pi}\PY{p}{(}\PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}\PY{p}{;}
 
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{o}{\PYZgt{}}
-\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}
+\PY{k+kr}{inline} \PY{k+kt}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}
 \PY{p}{\PYZob{}}
-  \PY{k}{return} \PY{n}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
+  \PY{k}{return} \PY{k+kt}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
 \PY{p}{\PYZcb{}}\PY{p}{;}
@@
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{o}{\PYZgt{}}
-\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{)}
+\PY{k+kr}{inline} \PY{k+kt}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{)}
 \PY{p}{\PYZob{}}
-  \PY{k}{return} \PY{n}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
+  \PY{k}{return} \PY{k+kt}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
 \PY{p}{\PYZcb{}}\PY{p}{;} 
@@
     \PY{p}{(}\PY{n}{C} \PY{o}{+} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}l} \PY{o}{+} 
-    \PY{p}{(}\PY{n}{C} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
+    \PY{p}{(}\PY{n}{C} \PY{o}{\PYZhy{}} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
   \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
@@
 \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{donorcell}\PY{p}{(} 
-  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
-  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+  \PY{k}{const} \PY{k+kt}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{k+kt}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
+  \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
 \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
@@
       \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
-  \PY{p}{)} \PY{o}{-}
+  \PY{p}{)} \PY{o}{\PYZhy{}}
   \PY{n}{F}\PY{p}{(}
-    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
     \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
-      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
+      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{\PYZhy{}}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
   \PY{p}{)}
@@
 \PY{k+kt}{void} \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
-  \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{k+kt}{int} \PY{n}{n}\PY{p}{,}
-  \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
-  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+  \PY{k}{const} \PY{k+kt}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{k+kt}{int} \PY{n}{n}\PY{p}{,}
+  \PY{k}{const} \PY{k+kt}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
+  \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
 \PY{p}{)} \PY{p}{\PYZob{}} 
-  \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}
-    \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
-    \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}\PY{p}{;}
+  \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{p}{(}
+    \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} \PY{o}{+}
+    \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}
+  \PY{p}{)}\PY{p}{;} 
 \PY{p}{\PYZcb{}}
@@
 \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
-  \PY{k}{const} \PY{n}{nom\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{nom}\PY{p}{,} \PY{k}{const} \PY{n}{den\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{den}
+  \PY{k}{const} \PY{k+kt}{nom\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{nom}\PY{p}{,} \PY{k}{const} \PY{k+kt}{den\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{den}
 \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
@@
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
-\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
-  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{k}{const} \PY{k+kt}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
+  \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
 \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
   \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
-    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,}
     \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}
@@
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
-\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
-  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+\PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{k}{const} \PY{k+kt}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
+  \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
 \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
  \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
-    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-}
-    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{,}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZhy{}}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{,}
     \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+}
-    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}
+    \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}
   \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
@@
 \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}
-  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
-  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} 
-  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
+  \PY{k}{const} \PY{k+kt}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
+  \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} 
+  \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
 \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
@@
     \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+}
-    \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)}
+    \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{n}{h}\PY{p}{)}\PY{p}{)}
   \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}
@@
 \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}
-  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
-  \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,}
-  \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}
+  \PY{k}{const} \PY{k+kt}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
+  \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,}
+  \PY{k}{const} \PY{k+kt}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}
 \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
   \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}
-  \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{)}
+  \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{\PYZhy{}} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{)}
   \PY{o}{*} \PY{n}{mpdata\PYZus{}A}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} 
-  \PY{o}{-} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} 
-  \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
+  \PY{o}{\PYZhy{}} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} 
+  \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
   \PY{o}{*} \PY{n}{mpdata\PYZus{}B}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
@@
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{n\PYZus{}t}\PY{o}{\PYZgt{}}
-\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{n}{ext}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{r}\PY{p}{,} \PY{k}{const} \PY{n}{n\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{n}\PY{p}{)} \PY{p}{\PYZob{}}
-  \PY{k}{return} \PY{n}{rng\PYZus{}t}\PY{p}{(}
-    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
+\PY{k+kr}{inline} \PY{k+kt}{rng\PYZus{}t} \PY{n}{ext}\PY{p}{(}\PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{r}\PY{p}{,} \PY{k}{const} \PY{k+kt}{n\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{n}\PY{p}{)} \PY{p}{\PYZob{}}
+  \PY{k}{return} \PY{k+kt}{rng\PYZus{}t}\PY{p}{(}
+    \PY{p}{(}\PY{n}{r} \PY{o}{\PYZhy{}} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
     \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}
@@
   \PY{c+c1}{// member fields}
-  \PY{n}{arrvec\PYZus{}t} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{;}
+  \PY{k+kt}{arrvec\PYZus{}t} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{;}
   \PY{k+kt}{int} \PY{n}{n}\PY{p}{,} \PY{n}{hlo}\PY{p}{;}
-  \PY{n}{rng\PYZus{}t} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{;}
-  \PY{n}{bcx\PYZus{}t} \PY{n}{bcx}\PY{p}{;}
-  \PY{n}{bcy\PYZus{}t} \PY{n}{bcy}\PY{p}{;}
+  \PY{k+kt}{rng\PYZus{}t} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{;}
+  \PY{k+kt}{bcx\PYZus{}t} \PY{n}{bcx}\PY{p}{;}
+  \PY{k+kt}{bcy\PYZus{}t} \PY{n}{bcy}\PY{p}{;}
 
@@
     \PY{n}{n}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{,} 
-    \PY{n}{i}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nx}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
-    \PY{n}{j}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{ny}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}  
+    \PY{n}{i}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nx}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
+    \PY{n}{j}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{ny}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}  
     \PY{n}{bcx}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} 
@@
     \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{l} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{l} \PY{o}{\PYZlt{}} \PY{l+m+mi}{2}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{l}\PY{p}{)} 
-      \PY{n}{psi}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
-    \PY{n}{C}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{h}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
-    \PY{n}{C}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
+      \PY{n}{psi}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{k+kt}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
+    \PY{n}{C}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{k+kt}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{h}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
+    \PY{n}{C}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{k+kt}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
   \PY{p}{\PYZcb{}}
@@
   \PY{c+c1}{// accessor methods}
-  \PY{n}{arr\PYZus{}t} \PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{p}{\PYZob{}}
+  \PY{k+kt}{arr\PYZus{}t} \PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{p}{\PYZob{}}
     \PY{k}{return} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{.}\PY{n}{reindex}\PY{p}{(}\PY{p}{\PYZob{}}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
@@
 
-  \PY{n}{arr\PYZus{}t} \PY{n}{courant}\PY{p}{(}\PY{k+kt}{int} \PY{n}{d}\PY{p}{)} 
+  \PY{k+kt}{arr\PYZus{}t} \PY{n}{courant}\PY{p}{(}\PY{k+kt}{int} \PY{n}{d}\PY{p}{)} 
   \PY{p}{\PYZob{}} 
@@
 
-  \PY{k+kt}{void} \PY{n}{cycle}\PY{p}{(}\PY{p}{)} 
+  \PY{k+kt}{void} \PY{n+nf}{cycle}\PY{p}{(}\PY{p}{)} 
   \PY{p}{\PYZob{}} 
-    \PY{n}{n} \PY{o}{=} \PY{p}{(}\PY{n}{n} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2} \PY{o}{-} \PY{l+m+mi}{2}\PY{p}{;} 
+    \PY{n}{n} \PY{o}{=} \PY{p}{(}\PY{n}{n} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2} \PY{o}{\PYZhy{}} \PY{l+m+mi}{2}\PY{p}{;} 
   \PY{p}{\PYZcb{}}
@@
   \PY{c+c1}{// integration logic}
-  \PY{k+kt}{void} \PY{n}{solve}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{nt}\PY{p}{)} 
+  \PY{k+kt}{void} \PY{n+nf}{solve}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{nt}\PY{p}{)} 
   \PY{p}{\PYZob{}}
@@
   \PY{c+c1}{// member fields}
-  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{rght\PYZus{}halo}\PY{p}{;}
-  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{rght\PYZus{}edge}\PY{p}{;}\PY{p}{;}
+  \PY{k+kt}{rng\PYZus{}t} \PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{rght\PYZus{}halo}\PY{p}{;}
+  \PY{k+kt}{rng\PYZus{}t} \PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{rght\PYZus{}edge}\PY{p}{;}\PY{p}{;}
 
@@
   \PY{n}{cyclic}\PY{p}{(}
-    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k+kt}{int} \PY{n}{hlo}
+    \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k+kt}{int} \PY{n}{hlo}
   \PY{p}{)} \PY{o}{:}
-    \PY{n}{left\PYZus{}halo}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}
-    \PY{n}{rght\PYZus{}edge}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{n}{hlo}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}  \PY{p}{)}\PY{p}{,}
+    \PY{n}{left\PYZus{}halo}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}
+    \PY{n}{rght\PYZus{}edge}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{n}{hlo}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}  \PY{p}{)}\PY{p}{,}
     \PY{n}{rght\PYZus{}halo}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{n}{hlo}  \PY{p}{)}\PY{p}{,}
-    \PY{n}{left\PYZus{}edge}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{n}{hlo}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}
+    \PY{n}{left\PYZus{}edge}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{n}{hlo}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}
   \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
@@
   \PY{c+c1}{// method invoked by the solver}
-  \PY{k+kt}{void} \PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{a}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}
+  \PY{k+kt}{void} \PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{k}{const} \PY{k+kt}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{a}\PY{p}{,} \PY{k}{const} \PY{k+kt}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}
   \PY{p}{\PYZob{}}
@@
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
-\PY{k}{struct} \PY{n}{solver\PYZus{}donorcell} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}} 
+\PY{k}{struct} \PY{n}{solver\PYZus{}donorcell} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{k+kt}{bcx\PYZus{}t}\PY{p}{,} \PY{k+kt}{bcy\PYZus{}t}\PY{o}{\PYZgt{}} 
 \PY{p}{\PYZob{}}
   \PY{n}{solver\PYZus{}donorcell}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:}
-    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
+    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{k+kt}{bcx\PYZus{}t}\PY{p}{,} \PY{k+kt}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
   \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
@@
     \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
-      \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} 
-      \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
+      \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} 
+      \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}
     \PY{p}{)}\PY{p}{;}
@@
 \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{n\PYZus{}iters}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
-\PY{k}{struct} \PY{n}{solver\PYZus{}mpdata} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
+\PY{k}{struct} \PY{n}{solver\PYZus{}mpdata} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{k+kt}{bcx\PYZus{}t}\PY{p}{,} \PY{k+kt}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
 \PY{p}{\PYZob{}}
@@
   \PY{k}{static} \PY{k}{const} \PY{k+kt}{int} \PY{n}{n\PYZus{}tmp} \PY{o}{=} \PY{n}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{2} \PY{o}{?} \PY{l+m+mi}{2} \PY{o}{:} \PY{l+m+mi}{1}\PY{p}{;}
-  \PY{n}{arrvec\PYZus{}t} \PY{n}{tmp}\PY{p}{[}\PY{n}{n\PYZus{}tmp}\PY{p}{]}\PY{p}{;}
-  \PY{n}{rng\PYZus{}t} \PY{n}{im}\PY{p}{,} \PY{n}{jm}\PY{p}{;}
+  \PY{k+kt}{arrvec\PYZus{}t} \PY{n}{tmp}\PY{p}{[}\PY{n}{n\PYZus{}tmp}\PY{p}{]}\PY{p}{;}
+  \PY{k+kt}{rng\PYZus{}t} \PY{n}{im}\PY{p}{,} \PY{n}{jm}\PY{p}{;}
 
@@
   \PY{n}{solver\PYZus{}mpdata}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:} 
-    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
-    \PY{n}{im}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{,}
-    \PY{n}{jm}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}
+    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{k+kt}{bcx\PYZus{}t}\PY{p}{,} \PY{k+kt}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
+    \PY{n}{im}\PY{p}{(}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{,}
+    \PY{n}{jm}\PY{p}{(}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}
   \PY{p}{\PYZob{}}
@@
     \PY{p}{\PYZob{}}
-      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}
-        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
+      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{k+kt}{arr\PYZus{}t}\PY{p}{(}
+        \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
       \PY{p}{)}\PY{p}{;}
-      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}
-        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
+      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{k+kt}{arr\PYZus{}t}\PY{p}{(}
+        \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
       \PY{p}{)}\PY{p}{;}
@@
         \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
-          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
+          \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}
         \PY{p}{)}\PY{p}{;}
@@
       \PY{p}{\PYZob{}}
-        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{cycle}\PY{p}{(}\PY{p}{)}\PY{p}{;}
-        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
-          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{hlo}\PY{p}{)}
+        \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{cycle}\PY{p}{(}\PY{p}{)}\PY{p}{;}
+        \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
+          \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{hlo}\PY{p}{)}
         \PY{p}{)}\PY{p}{;}
-        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
-          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{hlo}\PY{p}{)}
+        \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
+          \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{hlo}\PY{p}{)}
         \PY{p}{)}\PY{p}{;}
@@
         \PY{c+c1}{// choosing input/output for antidiff C}
-        \PY{k}{const} \PY{n}{arrvec\PYZus{}t} 
+        \PY{k}{const} \PY{k+kt}{arrvec\PYZus{}t} 
           \PY{o}{\PYZam{}}\PY{n}{C\PYZus{}unco} \PY{o}{=} \PY{p}{(}\PY{n}{step} \PY{o}{=}\PY{o}{=} \PY{l+m+mi}{1}\PY{p}{)} 
-            \PY{o}{?} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C} 
+            \PY{o}{?} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{C} 
             \PY{o}{:} \PY{p}{(}\PY{n}{step} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2}\PY{p}{)} 
@@
         \PY{c+c1}{// calculating the antidiffusive C }
-        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{(}\PY{n}{im}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}
-          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{im}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,} \PY{n}{C\PYZus{}unco}
+        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{(}\PY{n}{im}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}
+          \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{im}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,} \PY{n}{C\PYZus{}unco}
         \PY{p}{)}\PY{p}{;}
-        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,}\PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{;}
+        \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,}\PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{;}
 
-        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{n}{jm}\PY{o}{+}\PY{n}{h}\PY{p}{)} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}
-          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{jm}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{n}{C\PYZus{}unco}
+        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{n}{jm}\PY{o}{+}\PY{n}{h}\PY{p}{)} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}
+          \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{jm}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{n}{C\PYZus{}unco}
         \PY{p}{)}\PY{p}{;}
-        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,}\PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{;}
+        \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,}\PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{;}
 
-        \PY{c+c1}{// donor-cell step }
+        \PY{c+c1}{// donor\PYZhy{}cell step }
         \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
-          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{n}{C\PYZus{}corr}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
+          \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{n}{C\PYZus{}corr}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{\PYZhy{}}\PY{o}{\PYZgt{}}\PY{n}{j}
         \PY{p}{)}\PY{p}{;}

### 2014-02-19 paper/doc/code-for-listings.f.tex
the new order
changes before Sci Prog resubmission

@@
     \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{length} \PY{o}{=} \PY{n+nv}{n}
-    \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{o}{-}\PY{n+nv}{n} \PY{p}{:} \PY{n+nv}{n}\PY{o}{-}\PY{l+m+mi}{1} \PY{p}{)}\PY{p}{)}
-    \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(} \PY{l+m+mi}{0} \PY{p}{:} \PY{n+nv}{n}\PY{o}{-}\PY{l+m+mi}{1} \PY{p}{)}\PY{p}{)}
+    \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(} \PY{o}{\PYZhy{}}\PY{n+nv}{n} \PY{p}{:} \PY{n+nv}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1} \PY{p}{)}\PY{p}{)}
+    \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(} \PY{l+m+mi}{0} \PY{p}{:} \PY{n+nv}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1} \PY{p}{)}\PY{p}{)}
   \PY{k}{end }\PY{k}{subroutine}
@@
     \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n} \PY{o}{-} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{length}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n} \PY{o}{\PYZhy{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{length}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{arrs}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}
   \PY{k}{end }\PY{k}{subroutine}
@@
 
-\PY{k}{  }\PY{k}{interface }\PY{n+nv}{operator} \PY{p}{(}\PY{o}{-}\PY{p}{)}
+\PY{k}{  }\PY{k}{interface }\PY{n+nv}{operator} \PY{p}{(}\PY{o}{\PYZhy{}}\PY{p}{)}
     \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{mh}
@@
     \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}
-\PY{k}{    }\PY{k}{return} \PY{o}{=} \PY{n+nv}{i} \PY{o}{-} \PY{l+m+mi}{1}
+\PY{k}{    }\PY{k}{return} \PY{o}{=} \PY{n+nv}{i} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1}
   \PY{k}{end }\PY{k}{function}
@@
 \PY{c}{!listing04}
-\PY{k}{module }\PY{n+nv}{halo\PYZus{}m}
-  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
-  \PY{k}{implicit }\PY{k}{none}
-
-\PY{k}{  }\PY{k}{interface }\PY{n+nv}{ext}
-    \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{ext\PYZus{}n}
-    \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{ext\PYZus{}h}
-  \PY{k}{end }\PY{k}{interface }
-
-\PY{k}{  }\PY{k}{contains}
-
-\PY{k}{  }\PY{k}{function }\PY{n+nv}{ext\PYZus{}n}\PY{p}{(}\PY{n+nv}{r}\PY{p}{,} \PY{n+nv}{n}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n}
-    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-
-    \PY{k}{return} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{+} \PY{n+nv}{n} \PY{o}{/}\PY{p}{)}
-  \PY{k}{end }\PY{k}{function}
-
-\PY{k}{  }\PY{k}{function }\PY{n+nv}{ext\PYZus{}h}\PY{p}{(}\PY{n+nv}{r}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-    \PY{k}{type}\PY{p}{(}\PY{n+nv}{half\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{h}
-    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-
-    \PY{k}{return} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{+} \PY{n+nv}{h} \PY{o}{/}\PY{p}{)}
-  \PY{k}{end }\PY{k}{function}
-\PY{k}{end }\PY{k}{module}
-\PY{c}{!listing05}
 \PY{k}{module }\PY{n+nv}{pi\PYZus{}m}
@@
       \PY{k}{case} \PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}
-        \PY{k}{return} \PY{o}{=} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{-} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
+        \PY{k}{return} \PY{o}{=} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
       \PY{k}{case} \PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
-        \PY{k}{return} \PY{o}{=} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{-} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
+        \PY{k}{return} \PY{o}{=} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
     \PY{k}{end }\PY{k}{select}
@@
 \PY{k}{end }\PY{k}{module}
+\PY{c}{!listing05}
+\PY{k}{module }\PY{n+nv}{donorcell\PYZus{}m}
+  \PY{k}{use }\PY{n+nv}{real\PYZus{}m}
+  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
+  \PY{k}{use }\PY{n+nv}{pi\PYZus{}m}
+  \PY{k}{use }\PY{n+nv}{arrvec\PYZus{}m}
+  \PY{k}{implicit }\PY{k}{none}
+\PY{k}{  }\PY{k}{contains} 
 \PY{c}{!listing06}
+  \PY{k}{elemental }\PY{k}{function }\PY{n+nv}{F}\PY{p}{(}\PY{n+nv}{psi\PYZus{}l}\PY{p}{,} \PY{n+nv}{psi\PYZus{}r}\PY{p}{,} \PY{n+nv}{C}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}
+\PY{k}{    }\PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi\PYZus{}l}\PY{p}{,} \PY{n+nv}{psi\PYZus{}r}\PY{p}{,} \PY{n+nv}{C}
+    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
+      \PY{p}{(}\PY{n+nv}{C} \PY{o}{+} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n+nv}{psi\PYZus{}l} \PY{o}{+}                           \PY{p}{\PYZam{}}
+      \PY{p}{(}\PY{n+nv}{C} \PY{o}{\PYZhy{}} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n+nv}{psi\PYZus{}r}                             \PY{p}{\PYZam{}}
+    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
+  \PY{k}{end }\PY{k}{function}
+\PY{c}{!listing07}
+  \PY{k}{function }\PY{n+nv}{donorcell}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} 
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}\PY{p}{,} \PY{n+nv}{C}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
+    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
+      \PY{n+nv}{F}\PY{p}{(}                                               \PY{p}{\PYZam{}}
+        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
+        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
+        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,}   \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                             \PY{p}{\PYZam{}}
+      \PY{p}{)} \PY{o}{\PYZhy{}}                                              \PY{p}{\PYZam{}}
+      \PY{n+nv}{F}\PY{p}{(}                                               \PY{p}{\PYZam{}}
+        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
+        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
+        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,}   \PY{n+nv}{i}\PY{o}{\PYZhy{}}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                             \PY{p}{\PYZam{}}
+      \PY{p}{)}                                                \PY{p}{\PYZam{}}
+    \PY{p}{)}
+  \PY{k}{end }\PY{k}{function}
+\PY{c}{!listing08}
+  \PY{k}{subroutine }\PY{n+nv}{donorcell\PYZus{}op}\PY{p}{(}\PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}  
+    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{psi}
+    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{ptr}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
+    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{pi}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
+    \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{pi}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{p}{(}               \PY{p}{\PYZam{}}
+      \PY{n+nv}{donorcell}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{o}{+} \PY{p}{\PYZam{}}
+      \PY{n+nv}{donorcell}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}   \PY{p}{\PYZam{}}
+    \PY{p}{)}
+  \PY{k}{end }\PY{k}{subroutine}
+\PY{c}{!listing09}
+\PY{k}{end }\PY{k}{module}
+\PY{c}{!listing10}
+\PY{k}{module }\PY{n+nv}{mpdata\PYZus{}m}
+  \PY{k}{use }\PY{n+nv}{arrvec\PYZus{}m}
+  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
+  \PY{k}{use }\PY{n+nv}{pi\PYZus{}m}
+  \PY{k}{implicit }\PY{k}{none}
+\PY{k}{  }\PY{k}{contains} 
+\PY{c}{!listing11}
+  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{n+nv}{den}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{nom}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}\PY{p}{,} \PY{n+nv}{den}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{size}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} \PY{n+nv}{size}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{)}
+    \PY{k}{where} \PY{p}{(}\PY{n+nv}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{)}
+      \PY{k}{return} \PY{o}{=} \PY{n+nv}{nom} \PY{o}{/} \PY{n+nv}{den}
+    \PY{n+nv}{elsewhere}
+      \PY{k}{return} \PY{o}{=} \PY{l+m+mi}{0}
+    \PY{k}{end }\PY{k}{where}
+\PY{k}{  }\PY{k}{end }\PY{k}{function}
+\PY{c}{!listing12}
+  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}A}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
+    \PY{k}{return} \PY{o}{=} \PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}                              \PY{p}{\PYZam{}}
+      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}           \PY{p}{\PYZam{}}
+      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}            \PY{p}{\PYZam{}}
+    \PY{p}{)}  
+  \PY{k}{end }\PY{k}{function}
+\PY{c}{!listing13}
+  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}B}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
+    \PY{k}{return} \PY{o}{=} \PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}                              \PY{p}{\PYZam{}}
+      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
+    \PY{o}{\PYZhy{}} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}     \PY{p}{\PYZam{}}
+      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
+    \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
+    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
+  \PY{k}{end }\PY{k}{function}
+\PY{c}{!listing14}
+  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{C}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
+
+    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
+      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+}        \PY{p}{\PYZam{}}
+      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{\PYZhy{}}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{\PYZhy{}}\PY{n+nv}{h}\PY{p}{)}          \PY{p}{\PYZam{}}
+    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}               
+  \PY{k}{end }\PY{k}{function}
+\PY{c}{!listing15}
+  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{C}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
+    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
+    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C}
+    \PY{k}{return} \PY{o}{=}                                           \PY{p}{\PYZam{}}
+      \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{)}                  \PY{p}{\PYZam{}}
+      \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{\PYZhy{}} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}          \PY{p}{\PYZam{}}
+      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}A}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                         \PY{p}{\PYZam{}}
+      \PY{o}{\PYZhy{}} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                     \PY{p}{\PYZam{}}
+      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}           \PY{p}{\PYZam{}}
+      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}B}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
+  \PY{k}{end }\PY{k}{function}
+\PY{c}{!listing16}
+\PY{k}{end }\PY{k}{module}
+\PY{c}{!listing17}
+\PY{k}{module }\PY{n+nv}{halo\PYZus{}m}
+  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
+  \PY{k}{implicit }\PY{k}{none}
+
+\PY{k}{  }\PY{k}{interface }\PY{n+nv}{ext}
+    \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{ext\PYZus{}n}
+    \PY{k}{module }\PY{n+nv}{procedure} \PY{n+nv}{ext\PYZus{}h}
+  \PY{k}{end }\PY{k}{interface }
+
+\PY{k}{  }\PY{k}{contains}
+
+\PY{k}{  }\PY{k}{function }\PY{n+nv}{ext\PYZus{}n}\PY{p}{(}\PY{n+nv}{r}\PY{p}{,} \PY{n+nv}{n}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n}
+    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+
+    \PY{k}{return} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{+} \PY{n+nv}{n} \PY{o}{/}\PY{p}{)}
+  \PY{k}{end }\PY{k}{function}
+
+\PY{k}{  }\PY{k}{function }\PY{n+nv}{ext\PYZus{}h}\PY{p}{(}\PY{n+nv}{r}\PY{p}{,} \PY{n+nv}{h}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
+    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+    \PY{k}{type}\PY{p}{(}\PY{n+nv}{half\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{h}
+    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
+
+    \PY{k}{return} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{r}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{+} \PY{n+nv}{h} \PY{o}{/}\PY{p}{)}
+  \PY{k}{end }\PY{k}{function}
+\PY{k}{end }\PY{k}{module}
+\PY{c}{!listing18}
 \PY{k}{module }\PY{n+nv}{bcd\PYZus{}m}
@@
 \PY{k}{end }\PY{k}{module}
-\PY{c}{!listing07}
+\PY{c}{!listing19}
 \PY{k}{module }\PY{n+nv}{solver\PYZus{}m}
@@
 
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{nx} \PY{o}{-} \PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{ny} \PY{o}{-} \PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{nx} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{ny} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
 
@@
     \PY{k}{class}\PY{p}{(}\PY{n+nv}{solver\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{this}
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n} \PY{o}{=} \PY{n+nb}{mod}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n} \PY{o}{+} \PY{l+m+mi}{1} \PY{o}{+} \PY{l+m+mi}{2}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{2}
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n} \PY{o}{=} \PY{n+nb}{mod}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n} \PY{o}{+} \PY{l+m+mi}{1} \PY{o}{+} \PY{l+m+mi}{2}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{l+m+mi}{2}
   \PY{k}{end }\PY{k}{subroutine}
@@
 
-    \PY{k}{do }\PY{n+nv}{t} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{nt}\PY{o}{-}\PY{l+m+mi}{1} 
+    \PY{k}{do }\PY{n+nv}{t} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{nt}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1} 
       \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{bcx}\PY{p}{\PYZpc{}}\PY{n+nv}{fill\PYZus{}halos}\PY{p}{(}                        \PY{p}{\PYZam{}}
@@
 \PY{k}{end }\PY{k}{module}
-\PY{c}{!listing08}
+\PY{c}{!listing20}
 \PY{k}{module }\PY{n+nv}{cyclic\PYZus{}m}
@@
     \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{d} \PY{o}{=} \PY{n+nv}{d}
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{left\PYZus{}halo} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{o}{-}\PY{n+nv}{hlo}\PY{p}{,} \PY{o}{-}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)} 
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{rght\PYZus{}halo} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{n}\PY{o}{-}\PY{l+m+mi}{1}\PY{o}{+}\PY{n+nv}{hlo} \PY{o}{/}\PY{p}{)} 
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{left\PYZus{}edge} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{hlo}\PY{o}{-}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{rght\PYZus{}edge} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{n}\PY{o}{-}\PY{n+nv}{hlo}\PY{p}{,} \PY{n+nv}{n}\PY{o}{-}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{left\PYZus{}halo} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{o}{\PYZhy{}}\PY{n+nv}{hlo}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)} 
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{rght\PYZus{}halo} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{o}{+}\PY{n+nv}{hlo} \PY{o}{/}\PY{p}{)} 
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{left\PYZus{}edge} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{hlo}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{rght\PYZus{}edge} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{n}\PY{o}{\PYZhy{}}\PY{n+nv}{hlo}\PY{p}{,} \PY{n+nv}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1} \PY{o}{/}\PY{p}{)}
   \PY{k}{end }\PY{k}{subroutine}
@@
 \PY{k}{end }\PY{k}{module}
-\PY{c}{!listing09}
-\PY{k}{module }\PY{n+nv}{donorcell\PYZus{}m}
-  \PY{k}{use }\PY{n+nv}{real\PYZus{}m}
-  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
-  \PY{k}{use }\PY{n+nv}{pi\PYZus{}m}
-  \PY{k}{use }\PY{n+nv}{arrvec\PYZus{}m}
-  \PY{k}{implicit }\PY{k}{none}
-\PY{k}{  }\PY{k}{contains} 
-\PY{c}{!listing10}
-  \PY{k}{elemental }\PY{k}{function }\PY{n+nv}{F}\PY{p}{(}\PY{n+nv}{psi\PYZus{}l}\PY{p}{,} \PY{n+nv}{psi\PYZus{}r}\PY{p}{,} \PY{n+nv}{C}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}
-\PY{k}{    }\PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi\PYZus{}l}\PY{p}{,} \PY{n+nv}{psi\PYZus{}r}\PY{p}{,} \PY{n+nv}{C}
-    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
-      \PY{p}{(}\PY{n+nv}{C} \PY{o}{+} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n+nv}{psi\PYZus{}l} \PY{o}{+}                           \PY{p}{\PYZam{}}
-      \PY{p}{(}\PY{n+nv}{C} \PY{o}{-} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n+nv}{psi\PYZus{}r}                             \PY{p}{\PYZam{}}
-    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
-  \PY{k}{end }\PY{k}{function}
-\PY{c}{!listing11}
-  \PY{k}{function }\PY{n+nv}{donorcell}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} 
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}\PY{p}{,} \PY{n+nv}{C}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
-    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
-      \PY{n+nv}{F}\PY{p}{(}                                               \PY{p}{\PYZam{}}
-        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
-        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
-        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,}   \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                             \PY{p}{\PYZam{}}
-      \PY{p}{)} \PY{o}{-}                                              \PY{p}{\PYZam{}}
-      \PY{n+nv}{F}\PY{p}{(}                                               \PY{p}{\PYZam{}}
-        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
-        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}                            \PY{p}{\PYZam{}}
-        \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,}   \PY{n+nv}{i}\PY{o}{-}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                             \PY{p}{\PYZam{}}
-      \PY{p}{)}                                                \PY{p}{\PYZam{}}
-    \PY{p}{)}
-  \PY{k}{end }\PY{k}{function}
-\PY{c}{!listing12}
-  \PY{k}{subroutine }\PY{n+nv}{donorcell\PYZus{}op}\PY{p}{(}\PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}  
-    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable} \PY{k+kd}{::} \PY{n+nv}{psi}
-    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{n}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{ptr}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
-    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{pi}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
-    \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{pi}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                   \PY{p}{\PYZam{}}
-      \PY{o}{-} \PY{n+nv}{donorcell}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{p}{\PYZam{}}
-      \PY{o}{-} \PY{n+nv}{donorcell}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{n}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}
-  \PY{k}{end }\PY{k}{subroutine}
-\PY{c}{!listing13}
-\PY{k}{end }\PY{k}{module}
-\PY{c}{!listing14}
+\PY{c}{!listing21}
 \PY{k}{module }\PY{n+nv}{solver\PYZus{}donorcell\PYZus{}m}
@@
 \PY{k}{end }\PY{k}{module}
-\PY{c}{!listing15}
-\PY{k}{module }\PY{n+nv}{mpdata\PYZus{}m}
-  \PY{k}{use }\PY{n+nv}{arrvec\PYZus{}m}
-  \PY{k}{use }\PY{n+nv}{arakawa\PYZus{}c\PYZus{}m}
-  \PY{k}{use }\PY{n+nv}{pi\PYZus{}m}
-  \PY{k}{implicit }\PY{k}{none}
-\PY{k}{  }\PY{k}{contains} 
-\PY{c}{!listing16}
-  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{n+nv}{den}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{nom}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}\PY{p}{,} \PY{n+nv}{den}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{size}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} \PY{n+nv}{size}\PY{p}{(}\PY{n+nv}{nom}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{)}
-    \PY{k}{where} \PY{p}{(}\PY{n+nv}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{)}
-      \PY{k}{return} \PY{o}{=} \PY{n+nv}{nom} \PY{o}{/} \PY{n+nv}{den}
-    \PY{n+nv}{elsewhere}
-      \PY{k}{return} \PY{o}{=} \PY{l+m+mi}{0}
-    \PY{k}{end }\PY{k}{where}
-\PY{k}{  }\PY{k}{end }\PY{k}{function}
-\PY{c}{!listing17}
-  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}A}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
-    \PY{k}{return} \PY{o}{=} \PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}                              \PY{p}{\PYZam{}}
-      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{o}{-} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,}           \PY{p}{\PYZam{}}
-      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}            \PY{p}{\PYZam{}}
-    \PY{p}{)}  
-  \PY{k}{end }\PY{k}{function}
-\PY{c}{!listing18}
-  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}B}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
-    \PY{k}{return} \PY{o}{=} \PY{n+nv}{mpdata\PYZus{}frac}\PY{p}{(}                              \PY{p}{\PYZam{}}
-      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
-    \PY{o}{-} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}     \PY{p}{\PYZam{}}
-      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
-    \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}      \PY{p}{\PYZam{}}
-    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
-  \PY{k}{end }\PY{k}{function}
-\PY{c}{!listing19}
-  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{C}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
-
-    \PY{k}{return} \PY{o}{=} \PY{p}{(}                                         \PY{p}{\PYZam{}}
-      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{+}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{+}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+}        \PY{p}{\PYZam{}}
-      \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{o}{-}\PY{n+nv}{h}\PY{p}{)} \PY{o}{+} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,}   \PY{n+nv}{j}\PY{o}{-}\PY{n+nv}{h}\PY{p}{)}          \PY{p}{\PYZam{}}
-    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}               
-  \PY{k}{end }\PY{k}{function}
-\PY{c}{!listing20}
-  \PY{k}{function }\PY{n+nv}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{C}\PY{p}{)} \PY{k}{result} \PY{p}{(}\PY{k}{return}\PY{p}{)}
-    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{d}
-    \PY{k+kt}{integer}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}\PY{p}{(}\PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{span}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{i}\PY{p}{)}\PY{p}{)}
-    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{allocatable}\PY{p}{,} \PY{k}{intent}\PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{psi}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
-    \PY{k}{class}\PY{p}{(}\PY{n+nv}{arrvec\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{C}
-    \PY{k}{return} \PY{o}{=}                                           \PY{p}{\PYZam{}}
-      \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{)}                  \PY{p}{\PYZam{}}
-      \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}          \PY{p}{\PYZam{}}
-      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}A}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                         \PY{p}{\PYZam{}}
-      \PY{o}{-} \PY{n+nv}{pi}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{o}{+}\PY{n+nv}{h}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}                     \PY{p}{\PYZam{}}
-      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{C}\PY{p}{\PYZpc{}}\PY{n+nv}{at}\PY{p}{(}\PY{n+nv}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{p}\PY{p}{\PYZpc{}}\PY{n+nv}{a}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}           \PY{p}{\PYZam{}}
-      \PY{o}{*} \PY{n+nv}{mpdata\PYZus{}B}\PY{p}{(}\PY{n+nv}{d}\PY{p}{,} \PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}
-  \PY{k}{end }\PY{k}{function}
-\PY{c}{!listing21}
-\PY{k}{end }\PY{k}{module}
 \PY{c}{!listing22}
@@
     \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}iters} \PY{o}{=} \PY{n+nv}{n\PYZus{}iters}
-    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}tmp} \PY{o}{=} \PY{n+nb}{min}\PY{p}{(}\PY{n+nv}{n\PYZus{}iters} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}
+    \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}tmp} \PY{o}{=} \PY{n+nb}{min}\PY{p}{(}\PY{n+nv}{n\PYZus{}iters} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}
     \PY{k}{if} \PY{p}{(}\PY{n+nv}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{)} \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{:}\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}tmp}\PY{p}{)}\PY{p}{)}
@@
     \PY{n+nv}{associate} \PY{p}{(}\PY{n+nv}{i} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{j}\PY{p}{,} \PY{n+nv}{hlo} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{hlo}\PY{p}{)}
-      \PY{k}{do }\PY{n+nv}{c}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}tmp} \PY{o}{-} \PY{l+m+mi}{1}
+      \PY{k}{do }\PY{n+nv}{c}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}tmp} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1}
         \PY{k}{call }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{tmp}\PY{p}{(}\PY{n+nv}{c}\PY{p}{)}\PY{p}{\PYZpc{}}\PY{n+nv}{ctor}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
@@
 
-\PY{k}{      }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{im} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{/}\PY{p}{)}
-      \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{jm} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{/}\PY{p}{)}
+\PY{k}{      }\PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{im} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{i}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{/}\PY{p}{)}
+      \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{jm} \PY{o}{=} \PY{p}{(}\PY{o}{/} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{j}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)} \PY{o}{/}\PY{p}{)}
     \PY{k}{end }\PY{n+nv}{associate}
@@
     \PY{p}{)}
-      \PY{k}{do }\PY{n+nv}{step}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}iters}\PY{o}{-}\PY{l+m+mi}{1}
+      \PY{k}{do }\PY{n+nv}{step}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nv}{this}\PY{p}{\PYZpc{}}\PY{n+nv}{n\PYZus{}iters}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}
         \PY{k}{if} \PY{p}{(}\PY{n+nv}{step} \PY{o}{==} \PY{l+m+mi}{0}\PY{p}{)} \PY{k}{then}
@@
 
-            \PY{c}{! donor-cell step}
+            \PY{c}{! donor\PYZhy{}cell step}
             \PY{k}{call }\PY{n+nv}{donorcell\PYZus{}op}\PY{p}{(}\PY{n+nv}{psi}\PY{p}{,} \PY{n+nv}{n}\PY{p}{,} \PY{n+nv}{C\PYZus{}corr}\PY{p}{,} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)} 

### 2014-02-19 paper/doc/code-for-test.f.tex
changes before Sci Prog resubmission

@@
 \PY{k}{  }\PY{k}{if} \PY{p}{(}\PY{n+nb}{command\PYZus{}argument\PYZus{}count}\PY{p}{(}\PY{p}{)} \PY{o}{/}\PY{o}{=} \PY{p}{(}\PY{l+m+mi}{9}\PY{p}{)}\PY{p}{)} \PY{k}{then}
-\PY{k}{    }\PY{k}{print}\PY{o}{*}\PY{p}{,} \PY{l+s+s2}{"expecting 9 arguments (nx, ny, Cx, Cy, nt, it, f.in, f.out, dec)"}
+\PY{k}{    }\PY{k}{print}\PY{o}{*}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}expecting 9 arguments (nx, ny, Cx, Cy, nt, it, f.in, f.out, dec)\PYZdq{}}
     \PY{k}{stop }\PY{l+m+mi}{1}
@@
       \PY{k}{call }\PY{n+nv}{read\PYZus{}file}\PY{p}{(}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{tmp}\PY{p}{)}
-      \PY{n+nv}{diff} \PY{o}{=} \PY{n+nb}{maxval}\PY{p}{(}\PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{state}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{n+nv}{tmp}\PY{p}{)}\PY{p}{)}
-      \PY{k}{print}\PY{o}{*}\PY{p}{,} \PY{l+s+s2}{"diff="}\PY{p}{,} \PY{n+nv}{diff}
-      \PY{k}{if} \PY{p}{(}\PY{n+nv}{diff} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mf}{.5} \PY{o}{*} \PY{l+m+mi}{1}\PY{l+m+mf}{0.}\PY{o}{**}\PY{p}{(}\PY{o}{-}\PY{n+nv}{dec}\PY{p}{)}\PY{p}{)} \PY{n+nv}{error} \PY{o}{=} \PY{n+nb}{.TRUE.}
+      \PY{n+nv}{diff} \PY{o}{=} \PY{n+nb}{maxval}\PY{p}{(}\PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{state}\PY{p}{(}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n+nv}{tmp}\PY{p}{)}\PY{p}{)}
+      \PY{k}{print}\PY{o}{*}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}diff=\PYZdq{}}\PY{p}{,} \PY{n+nv}{diff}
+      \PY{k}{if} \PY{p}{(}\PY{n+nv}{diff} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mf}{.5} \PY{o}{*} \PY{l+m+mi}{1}\PY{l+m+mf}{0.}\PY{o}{**}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{n+nv}{dec}\PY{p}{)}\PY{p}{)} \PY{n+nv}{error} \PY{o}{=} \PY{n+nb}{.TRUE.}
       \PY{k}{deallocate}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{)}
@@
     \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{un}
-    \PY{k}{open}\PY{p}{(}\PY{n+nv}{newunit}\PY{o}{=}\PY{n+nv}{un}\PY{p}{,} \PY{n+nv}{file}\PY{o}{=}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{status}\PY{o}{=}\PY{l+s+s1}{'old'}\PY{p}{,} \PY{n+nv}{action}\PY{o}{=}\PY{l+s+s1}{'read'} \PY{p}{)}
+    \PY{k}{open}\PY{p}{(}\PY{n+nv}{newunit}\PY{o}{=}\PY{n+nv}{un}\PY{p}{,} \PY{n+nv}{file}\PY{o}{=}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{status}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}old\PYZsq{}}\PY{p}{,} \PY{n+nv}{action}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}read\PYZsq{}} \PY{p}{)}
     \PY{k}{block    }

### 2014-02-19 paper/doc/code-pyt-listings.py.tex
the new order
changes before Sci Prog resubmission

@@
 \PY{c}{\PYZsh{}listing01}
-\PY{n}{real\PYZus{}t} \PY{o}{=} \PY{l+s}{'}\PY{l+s}{float64}\PY{l+s}{'}
+\PY{n}{real\PYZus{}t} \PY{o}{=} \PY{l+s}{\PYZsq{}}\PY{l+s}{float64}\PY{l+s}{\PYZsq{}}
 \PY{c}{\PYZsh{}listing02}
@@
 \PY{k}{try}\PY{p}{:} 
-  \PY{n}{numpy}\PY{o}{.}\PY{n}{seterr}\PY{p}{(}\PY{n+nb}{all}\PY{o}{=}\PY{l+s}{'}\PY{l+s}{ignore}\PY{l+s}{'}\PY{p}{)}
+  \PY{n}{numpy}\PY{o}{.}\PY{n}{seterr}\PY{p}{(}\PY{n+nb}{all}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{ignore}\PY{l+s}{\PYZsq{}}\PY{p}{)}
 \PY{k}{except} \PY{n+ne}{AttributeError}\PY{p}{:}
@@
     \PY{k}{return} \PY{n+nb}{type}\PY{p}{(}\PY{n}{arg}\PY{p}{)}\PY{p}{(}
-      \PY{n}{arg}\PY{o}{.}\PY{n}{start} \PY{o}{-} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mnus}\PY{p}{,} 
-      \PY{n}{arg}\PY{o}{.}\PY{n}{stop}  \PY{o}{-} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mnus}
+      \PY{n}{arg}\PY{o}{.}\PY{n}{start} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mnus}\PY{p}{,} 
+      \PY{n}{arg}\PY{o}{.}\PY{n}{stop}  \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mnus}
     \PY{p}{)}
@@
 \PY{c}{\PYZsh{}listing05}
+\PY{k}{def} \PY{n+nf}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{o}{*}\PY{n}{idx}\PY{p}{)}\PY{p}{:} 
+  \PY{k}{return} \PY{p}{(}\PY{n}{idx}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{,} \PY{n}{idx}\PY{p}{[}\PY{n}{d}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
+\PY{c}{\PYZsh{}listing06}
+\PY{k}{def} \PY{n+nf}{f}\PY{p}{(}\PY{n}{psi\PYZus{}l}\PY{p}{,} \PY{n}{psi\PYZus{}r}\PY{p}{,} \PY{n}{C}\PY{p}{)}\PY{p}{:}
+  \PY{k}{return} \PY{p}{(}
+    \PY{p}{(}\PY{n}{C} \PY{o}{+} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}l} \PY{o}{+} 
+    \PY{p}{(}\PY{n}{C} \PY{o}{\PYZhy{}} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
+  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
+\PY{c}{\PYZsh{}listing07}
+\PY{k}{def} \PY{n+nf}{donorcell}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
+  \PY{k}{return} \PY{p}{(}
+    \PY{n}{f}\PY{p}{(}
+      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}     \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
+      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
+        \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
+    \PY{p}{)} \PY{o}{\PYZhy{}} 
+    \PY{n}{f}\PY{p}{(}
+      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{\PYZhy{}}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
+      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}     \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
+        \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{\PYZhy{}}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
+    \PY{p}{)} 
+  \PY{p}{)}
+\PY{c}{\PYZsh{}listing08}
+\PY{k}{def} \PY{n+nf}{donorcell\PYZus{}op}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{n}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
+  \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{]} \PY{o}{=} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{]} \PY{o}{\PYZhy{}} \PY{p}{(}
+    \PY{n}{donorcell}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} \PY{o}{+}
+    \PY{n}{donorcell}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}
+  \PY{p}{)}
+\PY{c}{\PYZsh{}listing09}
+\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}frac}\PY{p}{(}\PY{n}{nom}\PY{p}{,} \PY{n}{den}\PY{p}{)}\PY{p}{:}
+  \PY{k}{return} \PY{n}{numpy}\PY{o}{.}\PY{n}{where}\PY{p}{(}\PY{n}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nom}\PY{o}{/}\PY{n}{den}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}
+\PY{c}{\PYZsh{}listing10}
+\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}A}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
+  \PY{k}{return} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
+    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} \PY{o}{\PYZhy{}} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,}
+    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
+  \PY{p}{)}
+\PY{c}{\PYZsh{}listing11}
+\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}B}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
+  \PY{k}{return} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(} 
+    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{\PYZhy{}}
+    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{\PYZhy{}} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{n}{one}\PY{p}{)}\PY{p}{]}\PY{p}{,}
+    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+}
+    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{n}{one}\PY{p}{)}\PY{p}{]}
+  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
+\PY{c}{\PYZsh{}listing12}
+\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
+  \PY{k}{return} \PY{p}{(}
+    \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}  \PY{n}{j}\PY{o}{+}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+}
+    \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{\PYZhy{}}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}  \PY{n}{j}\PY{o}{\PYZhy{}}\PY{n}{hlf}\PY{p}{)}\PY{p}{]}
+  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}
+\PY{c}{\PYZsh{}listing13}
+\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{C}\PY{p}{)}\PY{p}{:}
+  \PY{k}{return} \PY{p}{(}
+    \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{)} 
+    \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{\PYZhy{}} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{)}\PY{p}{)}
+    \PY{o}{*} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
+    \PY{o}{\PYZhy{}} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} 
+    \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
+    \PY{o}{*} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
+  \PY{p}{)}
+\PY{c}{\PYZsh{}listing14}
 \PY{k}{def} \PY{n+nf}{ext}\PY{p}{(}\PY{n}{r}\PY{p}{,} \PY{n}{n}\PY{p}{)}\PY{p}{:}
@@
   \PY{k}{return} \PY{n+nb}{slice}\PY{p}{(}
-    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{o}{.}\PY{n}{start}\PY{p}{,} 
+    \PY{p}{(}\PY{n}{r} \PY{o}{\PYZhy{}} \PY{n}{n}\PY{p}{)}\PY{o}{.}\PY{n}{start}\PY{p}{,} 
     \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{o}{.}\PY{n}{stop}
   \PY{p}{)}
-\PY{c}{\PYZsh{}listing06}
-\PY{k}{def} \PY{n+nf}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{o}{*}\PY{n}{idx}\PY{p}{)}\PY{p}{:} 
-  \PY{k}{return} \PY{p}{(}\PY{n}{idx}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{,} \PY{n}{idx}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
-\PY{c}{\PYZsh{}listing07}
+\PY{c}{\PYZsh{}listing15}
 \PY{k}{class} \PY{n+nc}{solver}\PY{p}{(}\PY{n+nb}{object}\PY{p}{)}\PY{p}{:}
-  \PY{c}{\PYZsh{} ctor-like method}
+  \PY{c}{\PYZsh{} ctor\PYZhy{}like method}
   \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{bcx}\PY{p}{,} \PY{n}{bcy}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{:}
@@
   \PY{k}{def} \PY{n+nf}{cycle}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{)}\PY{p}{:}
-    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}  \PY{o}{=} \PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2} \PY{o}{-} \PY{l+m+mi}{2}
+    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n}  \PY{o}{=} \PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{n} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2} \PY{o}{\PYZhy{}} \PY{l+m+mi}{2}
 
@@
 
-\PY{c}{\PYZsh{}listing08}
+\PY{c}{\PYZsh{}listing16}
 \PY{k}{class} \PY{n+nc}{cyclic}\PY{p}{(}\PY{n+nb}{object}\PY{p}{)}\PY{p}{:}
@@
     \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{d} \PY{o}{=} \PY{n}{d}
-    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{left\PYZus{}halo} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{start}\PY{o}{-}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{o}{.}\PY{n}{start}    \PY{p}{)}
-    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{rght\PYZus{}edge} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{stop} \PY{o}{-}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{o}{.}\PY{n}{stop}     \PY{p}{)}
+    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{left\PYZus{}halo} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{start}\PY{o}{\PYZhy{}}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{o}{.}\PY{n}{start}    \PY{p}{)}
+    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{rght\PYZus{}edge} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{stop} \PY{o}{\PYZhy{}}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{o}{.}\PY{n}{stop}     \PY{p}{)}
     \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{rght\PYZus{}halo} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n}{i}\PY{o}{.}\PY{n}{stop}\PY{p}{,}      \PY{n}{i}\PY{o}{.}\PY{n}{stop} \PY{o}{+}\PY{n}{hlo}\PY{p}{)}
@@
 
-\PY{c}{\PYZsh{}listing09}
-\PY{k}{def} \PY{n+nf}{f}\PY{p}{(}\PY{n}{psi\PYZus{}l}\PY{p}{,} \PY{n}{psi\PYZus{}r}\PY{p}{,} \PY{n}{C}\PY{p}{)}\PY{p}{:}
-  \PY{k}{return} \PY{p}{(}
-    \PY{p}{(}\PY{n}{C} \PY{o}{+} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}l} \PY{o}{+} 
-    \PY{p}{(}\PY{n}{C} \PY{o}{-} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
-  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
-\PY{c}{\PYZsh{}listing10}
-\PY{k}{def} \PY{n+nf}{donorcell}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
-  \PY{k}{return} \PY{p}{(}
-    \PY{n}{f}\PY{p}{(}
-      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}     \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
-      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
-        \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
-    \PY{p}{)} \PY{o}{-} 
-    \PY{n}{f}\PY{p}{(}
-      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{-}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
-      \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}     \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,} 
-        \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{-}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
-    \PY{p}{)} 
-  \PY{p}{)}
-\PY{c}{\PYZsh{}listing11}
-\PY{k}{def} \PY{n+nf}{donorcell\PYZus{}op}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{n}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
-  \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{]}
-    \PY{o}{-} \PY{n}{donorcell}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
-    \PY{o}{-} \PY{n}{donorcell}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}
-  \PY{p}{)}
-\PY{c}{\PYZsh{}listing12}
+\PY{c}{\PYZsh{}listing17}
 \PY{k}{class} \PY{n+nc}{solver\PYZus{}donorcell}\PY{p}{(}\PY{n}{solver}\PY{p}{)}\PY{p}{:}
@@
     \PY{p}{)}
-\PY{c}{\PYZsh{}listing13}
-\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}frac}\PY{p}{(}\PY{n}{nom}\PY{p}{,} \PY{n}{den}\PY{p}{)}\PY{p}{:}
-  \PY{k}{return} \PY{n}{numpy}\PY{o}{.}\PY{n}{where}\PY{p}{(}\PY{n}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nom}\PY{o}{/}\PY{n}{den}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}
-\PY{c}{\PYZsh{}listing14}
-\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}A}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
-  \PY{k}{return} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
-    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} \PY{o}{-} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{,}
-    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}
-  \PY{p}{)}
-\PY{c}{\PYZsh{}listing15}
-\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}B}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
-  \PY{k}{return} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(} 
-    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{-}
-    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{-} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{one}\PY{p}{)}\PY{p}{]}\PY{p}{,}
-    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+}
-    \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{one}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{psi}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{one}\PY{p}{)}\PY{p}{]}
-  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
-\PY{c}{\PYZsh{}listing16}
-\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{C}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{:}
-  \PY{k}{return} \PY{p}{(}
-    \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}  \PY{n}{j}\PY{o}{+}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+}
-    \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{one}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{hlf}\PY{p}{)}\PY{p}{]} \PY{o}{+} \PY{n}{C}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{p}{,}  \PY{n}{j}\PY{o}{-}\PY{n}{hlf}\PY{p}{)}\PY{p}{]}
-  \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}
-\PY{c}{\PYZsh{}listing17}
-\PY{k}{def} \PY{n+nf}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{C}\PY{p}{)}\PY{p}{:}
-  \PY{k}{return} \PY{p}{(}
-    \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{)} 
-    \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n+nb}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]}\PY{p}{)}\PY{p}{)}
-    \PY{o}{*} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
-    \PY{o}{-} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{[}\PY{n}{pi}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{i}\PY{o}{+}\PY{n}{hlf}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{]} 
-    \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
-    \PY{o}{*} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{n}{d}\PY{p}{,} \PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
-  \PY{p}{)}
 \PY{c}{\PYZsh{}listing18}
@@
     \PY{n}{solver}\PY{o}{.}\PY{n}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{bcx}\PY{p}{,} \PY{n}{bcy}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
-    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{im} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{o}{.}\PY{n}{start}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{o}{.}\PY{n}{stop}\PY{p}{)}
-    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{jm} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{o}{.}\PY{n}{start}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{o}{.}\PY{n}{stop}\PY{p}{)}
+    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{im} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{o}{.}\PY{n}{start}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{i}\PY{o}{.}\PY{n}{stop}\PY{p}{)}
+    \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{jm} \PY{o}{=} \PY{n+nb}{slice}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{o}{.}\PY{n}{start}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{j}\PY{o}{.}\PY{n}{stop}\PY{p}{)}
 

### 2014-02-19 paper/doc/code-pyt-test.py.tex
changes before Sci Prog resubmission

@@
   \PY{n}{tmp} \PY{o}{=} \PY{n}{numpy}\PY{o}{.}\PY{n}{empty}\PY{p}{(}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}\PY{p}{,} \PY{n}{real\PYZus{}t}\PY{p}{)}
-  \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,} \PY{l+s}{'}\PY{l+s}{r}\PY{l+s}{'}\PY{p}{)} \PY{k}{as} \PY{n}{f}\PY{p}{:}
+  \PY{k}{with} \PY{n+nb}{open}\PY{p}{(}\PY{n}{fname}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{)} \PY{k}{as} \PY{n}{f}\PY{p}{:}
     \PY{n}{x} \PY{o}{=} \PY{l+m+mi}{0}
     \PY{k}{for} \PY{n}{line} \PY{o+ow}{in} \PY{n}{f}\PY{p}{:}
-      \PY{n}{tmp}\PY{p}{[}\PY{n}{x}\PY{p}{,}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{numpy}\PY{o}{.}\PY{n}{fromstring}\PY{p}{(}\PY{n}{line}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{n}{real\PYZus{}t}\PY{p}{,} \PY{n}{sep}\PY{o}{=}\PY{l+s}{'}\PY{l+s+se}{\PYZbs{}t}\PY{l+s}{'}\PY{p}{)}
+      \PY{n}{tmp}\PY{p}{[}\PY{n}{x}\PY{p}{,}\PY{p}{:}\PY{p}{]} \PY{o}{=} \PY{n}{numpy}\PY{o}{.}\PY{n}{fromstring}\PY{p}{(}\PY{n}{line}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{n}{real\PYZus{}t}\PY{p}{,} \PY{n}{sep}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s+se}{\PYZbs{}t}\PY{l+s}{\PYZsq{}}\PY{p}{)}
       \PY{n}{x} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
@@
   \PY{k}{if} \PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{sys}\PY{o}{.}\PY{n}{argv}\PY{p}{)} \PY{o}{!=} \PY{p}{(}\PY{l+m+mi}{9} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{p}{:} 
-    \PY{k}{raise} \PY{n+ne}{Exception}\PY{p}{(}\PY{l+s}{'}\PY{l+s}{expecting 9 arguments (nx, ny, Cx, Cy, nt, it, f.in, f.out, dec)}\PY{l+s}{'}\PY{p}{)}
+    \PY{k}{raise} \PY{n+ne}{Exception}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{expecting 9 arguments (nx, ny, Cx, Cy, nt, it, f.in, f.out, dec)}\PY{l+s}{\PYZsq{}}\PY{p}{)}
 
@@
 
-  \PY{n}{diff} \PY{o}{=} \PY{n}{numpy}\PY{o}{.}\PY{n}{amax}\PY{p}{(}\PY{n+nb}{abs}\PY{p}{(}\PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{n}{read\PYZus{}file}\PY{p}{(}\PY{n}{fout}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}\PY{p}{)}\PY{p}{)}
-  \PY{k}{print} \PY{l+s}{"}\PY{l+s}{diff=}\PY{l+s}{"}\PY{p}{,} \PY{n}{diff}
-  \PY{k}{if} \PY{p}{(}\PY{n}{diff} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{o}{.}\PY{l+m+mi}{5} \PY{o}{*} \PY{n+nb}{pow}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{,} \PY{o}{-}\PY{n}{dec}\PY{p}{)}\PY{p}{)}\PY{p}{:} 
+  \PY{n}{diff} \PY{o}{=} \PY{n}{numpy}\PY{o}{.}\PY{n}{amax}\PY{p}{(}\PY{n+nb}{abs}\PY{p}{(}\PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{read\PYZus{}file}\PY{p}{(}\PY{n}{fout}\PY{p}{,} \PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{)}\PY{p}{)}\PY{p}{)}
+  \PY{k}{print} \PY{l+s}{\PYZdq{}}\PY{l+s}{diff=}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{diff}
+  \PY{k}{if} \PY{p}{(}\PY{n}{diff} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{o}{.}\PY{l+m+mi}{5} \PY{o}{*} \PY{n+nb}{pow}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{n}{dec}\PY{p}{)}\PY{p}{)}\PY{p}{:} 
     \PY{k}{print} \PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{dtype}\PY{p}{,} \PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)}
@@
     \PY{k}{print} \PY{n}{tmp}\PY{o}{.}\PY{n}{dtype}\PY{p}{,} \PY{n}{tmp}
-    \PY{k}{print} \PY{n}{numpy}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{n}{numpy}\PY{o}{.}\PY{n}{abs}\PY{p}{(}\PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{n}{tmp}\PY{p}{)}\PY{p}{)}
+    \PY{k}{print} \PY{n}{numpy}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{n}{numpy}\PY{o}{.}\PY{n}{abs}\PY{p}{(}\PY{n}{slv}\PY{o}{.}\PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{tmp}\PY{p}{)}\PY{p}{)}
     \PY{k}{raise} \PY{n+ne}{Exception}\PY{p}{(}\PY{p}{)}
 
-\PY{k}{if} \PY{n}{\PYZus{}\PYZus{}name\PYZus{}\PYZus{}} \PY{o}{==} \PY{l+s}{"}\PY{l+s}{\PYZus{}\PYZus{}main\PYZus{}\PYZus{}}\PY{l+s}{"}\PY{p}{:}
+\PY{k}{if} \PY{n}{\PYZus{}\PYZus{}name\PYZus{}\PYZus{}} \PY{o}{==} \PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZus{}\PYZus{}main\PYZus{}\PYZus{}}\PY{l+s}{\PYZdq{}}\PY{p}{:}
     \PY{n}{main}\PY{p}{(}\PY{p}{)}

### 2014-02-19 paper/doc/paper.tex
changes before Sci Prog resubmission

@@
 
+% highlighting (or not) the modified portions of the text
+\newcommand{\changed}[1]{{\color{blue}#1}}
+%\newcommand{\changed}[1]{#1}
+
+\usepackage{soul}
+\newcommand{\removed}[1]{{\color{blue}\st{#1}}}
+
 % input encoding assuring proper handling of Polish characters (surnames!)
 \usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
+\usepackage{fontenc}
 
@@
     \title{
-      Formula translation in C++, Python and modern Fortran: a~case~study~of~the~language~choice tradeoffs
+      Formula translation in \changed{Blitz++, NumPy} and modern Fortran: a~case~study~of~the~language~choice tradeoffs
     }
@@
 
-  Object oriented programming (OOP) {\em ''has become recognised as the almost unique successful 
+  Object oriented programming (OOP) {\em ``has become recognised as the almost unique successful 
     paradigm for creating complex software''} \citep[][Sec.~1.3]{Press_et_al_2007}.
@@
     reproduce within the code what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012}.
+  \changed{These may relate to several levels of mathematical abstraction. 
+    Object-oriented logic can be used to make the code resemble analytical formul\ae~\citep[e.g.][]{Weller_et_al_1998}
+    and/or numerical algorithms, the latter being exemplified in this paper.
+    }
   For this purpose, a sample implementation of a numerical scheme for solving
@@
   The paper is structured as follows.
-  In section~\ref{sec:impl} we introduce the ''formula translation'' part of the three implementations
+  In section~\ref{sec:impl} we introduce the ``formula translation'' part of the three implementations
     briefly describing the algorithm itself and
@@
   \codefor{code-for-listings.f}{listing01}{listing02}{1}
-  which provide a convenient way of switching to different precision.
+  which provide a convenient way of switching to different precision\footnote{Fortran's
+    \prog{selected\_real\_kind()}" intrinsic function may be used instead to improve portability.}.
 
@@
     for an overview of OOP in context of C++, Python and Fortran, consult for example
-    \citep[][Part~II]{Stroustrup_2000}, \citep[][Chapter~5]{Pilgrim_2004} and
+    \citep[][Part~III]{Stroustrup_2013}, \citep[][Chapter~5]{Pilgrim_2004} and
     \citep[][Chapter~11]{Markus_2012}, respectively.
@@
   In particular, it is possible to define a function that returns an array expression;
-    i.e. not the resultant array, but an object representing a ,,recipe'' defining the operations
+    i.e. not the resultant array, but an object representing a ``recipe'' defining the operations
     to be performed on the arguments.
@@
 
-  For the Python implementation of MPDATA the NumPy\footnote{NumPy is a Python package
+  For the Python implementation of MPDATA, the NumPy\footnote{NumPy is a Python package
     for scientific computing offering support for multi-dimensional arrays and a library
@@
   Therefore, there is no need for using an external package as with C++ and Python.  
-  Fortran array-handling features are not object-oriented, though.
+  Fortran array-handling features are not object-oriented, though \changed{(e.g. it is impossible
+    to overload array operators or to provide constructor-like initialisation logic)}.
 
@@
 
-  In the C++ implementation the Boost\footnote{
+  In the C++ implementation, the Boost\footnote{
     Boost is a free and open-source collection of peer-reviewed C++ libraries available at \url{http://www.boost.org/}.
@@
 
-  In the C++ implementation first a global instance \prog{h} of an empty structure 
+  In the C++ implementation, first a global instance \prog{h} of an empty structure 
     \prog{hlf\_t} is defined, and then the plus and minus operators for \prog{hlf\_t} and \prog{rng\_t} are overloaded:
@@
     dimensions using a single entity.
-  As a workaround, in in the Fortran implementation, \prog{pi()} returns a pointer to the array elements specified
+  As a workaround, in the Fortran implementation, \prog{pi()} returns a pointer to the array elements specified
     by \prog{i} and \prog{j} interpreted as (i,j) or
     (j,i) depending on the value of the argument \prog{d}.
-% TODO: disucss creation of temporary arrays and disabling of vectorisation due to employment of POINTERS
   In addition to \prog{pi()}, a helper \prog{span()} function returning the 
@@
     to the mathematical notation is the main point of this paper.
-  The ''formula translation'' features include:
+  The ``formula translation'' features include:
   \begin{itemize}
@@
   If the (positive-defined) denominator in equations \ref{eq:A} or \ref{eq:B} 
-    equals zero for a given {\em i} and {\em j}, %TODO: >0 in the code
+    equals zero for a given {\em i} and {\em j},
     the corresponding $A_{[i,j]}$ and $B_{[i,j]}$ are set to zero what may be conveniently 
@@
   The Clang compiler was invoked with the \prog{-O3}, the \prog{-mllvm -vectorize}, the \prog{-ffast-math} 
-    and the \prog{-march=native} options.
+    and the \prog{-march=native} options\footnote{Clang 3.4 enables vectorisation by default and supports the \prog{-Ofast} option}.
   The CPython interpreter was invoked with the \prog{-OO} option.
@@
   Three MPDATA iterations were used (i.e. two corrective steps).
-  %The initial condition was loaded from a text file, and the final values were compared at the end of the test
-  %  with values loaded from another text file assuring the same results were obtained with all four set-ups.
   The tests were run multiple times; program start-up, data loading, and output verification times were
@@
     not intended for use in production binaries, NumPy performs them always.
-  Additionally, the C++ and Fortran set-ups may, in principle, benefit from GCC's auto-vectorisation
+  Additionally, the C++ and Fortran set-ups may, in principle, benefit from auto-vectorisation
     features which do not have yet counterparts in CPython or PyPy.
@@
 
+  \changed{
+  The timing and memory usage statistics presented in figures \ref{fig:mem}-\ref{fig:cpu-skua}
+    reveal that, in the presented case, no single language/library/compiler set-up 
+    corresponds to both shortest execution time and smallest memory footprint.
+  Yet, performance is not the only criterion for the selection of a given language.
+  Presented case study allows as well to assess other language characteristics 
+    that define the tradeoffs of language choice.
+  }
+
   \subsection{Representability of blackboard abstractions}
@@
   It was shown in section~\ref{sec:impl} that C++11/Blitz++, Python/NumPy and Fortran 2008
-    provide comparable functionalities in terms of matching the blackboard abstractions
+    provide comparable functionality in terms of matching the blackboard abstractions
     within the program code.
@@
     \item{Fortran does not feature a mechanism allowing to reuse a single piece of code (algorithm)
-      with different data types (compare e.g. listings C.15, P.14 and F.17) such as
+      with different data \removed{types} \changed{structures} (compare e.g. listings C.15, P.14 and F.17) such as
       templates in C++ and the so-called duck typing in Python;} 
@@
 
-  \subsection{Performance}
-
-  The timing and memory usage statistics presented in figures \ref{fig:mem}-\ref{fig:cpu-skua}
-    reveal that no single language/library/compiler set-up 
-    corresponded to both shortest execution time and smallest memory footprint.
-
-  One may consider performance measures addressing not only the program efficiency but also 
-    the factors influencing the development and maintenance time/cost 
-    \citep[of particular importance in scientific computing,][]{Wilson_2006}.
-  Taking into account such measures as code length or coding time,
-    the Python environment gains significantly \citep[see also discussion in][]{Lin_2012}.
-  Presented Python code is shorter than the C++ and Fortran counterparts,
-    and is simpler in terms of syntax and usage (see discussion below).
-
-  Employment of the PyPy drop-in replacement for the standard Python implementation brings 
-    Python's performance significantly closer to those of C++ and Fortran, in some
-    cases making it the least memory consuming set-up.
-  Python has already been the language of choice for scientific software projects having code clarity 
-    or ease of use as the first requirement \citep[see e.g.][]{Barnes_and_Jones_2011}.
-  PyPy's capability to improve performance of unmodified Python code may 
-    make Python a favourable choice even if high performance is important, especially
-    if a combined measure of performance and development cost is to be considered.
-
-  \subsection{Ease of use and abuse}
-
-  Using the number of lines of code or the number of distinct language keywords
-    needed to implement the MPDATA-based
-    solver presented in section \ref{sec:impl} as measures of syntax 
-    brevity, Python clearly surpasses its rivals.
-  Python was developed with emphasis on code readability and object-orientation.
-  Arguably, taking it to the extreme - Python uses line indentation to define 
-    blocks of code and treats even single integers as objects.
-  As a consequence Python is easy to learn and easy to teach.
-  It is also much harder to abuse Python than C++ or Fortran
-    (for instance with \prog{goto} statements, employment of the preprocessor,
-    or the implicit typing in Fortran).
-
-  Python implementations do not expose users to compilation or linking processes. 
-  As a result, Python-written software is easier to deploy and share, especially 
-    if multiple architectures and operating systems are targeted.
-  However, there exist tools such as CMake\footnote{CMake is a family of open-source, cross-platform
-    tools automating building, testing and packaging of C/C++/Fortran software,
-    see \url{http://cmake.org/}}
-    that allow to efficiently automate 
-    building, testing and packaging of C++ and Fortran programs.
-
-  Python is definitely easiest to debug among the three languages.
-  Great debugging tools for C++ do exist, however the debugging and development is 
-    often hindered by indecipherable compiler messages
-    flooded with lengthy type names stemming from employment of templates.
-  Support for the OOP features of Fortran among free and open source compilers, 
-    debuggers and other programming aids remains immature.
-
-  Blitz++ ensures temporary-array-free computations by design
-    \citep{Veldhuizen_et_al_1997} avoiding unintentional performance loss.
-  In contrast, with both Fortran and Python, the memory footprint caused by employment
-    of temporary objects in array arithmetics is dependant on compiler choice or
-    the level of optimisations.
-
   \subsection{Developers' community and libraries}%User- and code-base}
@@
     example being the incomplete support for syntax-highlighting of modern Fortran in Vim and Emacs editors}.
-  Fortran is no longer routinely taught at the universities \citep{Kendall_et_al_2008},
-    in contrast to C++ and Python.
+  Fortran is no longer routinely taught at the \changed{university computer science
+    departments} \citep{Kendall_et_al_2008}, in contrast to C++ and Python.
   An example of decreasing popularity of Fortran in academia 
-    is the discontinuation of Fortran printed editions of the ''Numerical Recipes'' 
-    series of Press et al.
+    is the discontinuation of Fortran printed editions of the ``Numerical Recipes'' 
+    series of Press et al. \changed{(as of the third edition, the C++ version is the only one)}.
 
@@
     purpose \citep{Iglberger_et_al_2012}).
-  In contrast, the NumPy package became a de facto standard solution for Python.
+  In contrast, the NumPy package became a de-facto standard solution for Python.
   Consequently, numerous Python libraries adopted NumPy but
@@
 
-  The availability and quality of libraries 
-    that offer object-oriented interfaces
+  The availability and quality of libraries that offer object-oriented interfaces
     differs among the three considered languages.
@@
     built-in library routines).
-  What makes correct use of external libraries more difficult with Fortran
+  What makes correct use of external libraries \changed{less convenient} with Fortran 
     is the lack of standard exception handling mechanism, a feature
@@
 
+  \changed{
   The three languages differ as well with regard to availability of 
-    mechanisms for leveraging shared-memory parallelisation (e.g. with multi-core processors).
-  GCC supports OpenMP with Fortran and C++.
-  The CPython and PyPy implementations of Python do not offer any
-    built-in solution for multi-threading.
+    mechanisms \changed{(either built-in or available in external libraries)}
+    for handling concurrent computations.
+  For instance, GCC supports OpenMP with Fortran and C++ what allows to easily
+    leverage shared-memory parallelisation possibilities of multi-core processors.
+  There is no equivalent built-in solution for multi-threading in
+    CPython or PyPy.
+  Fortan 2008 standard includes the ``coarray'' built-in parallel programming model for
+    which counterparts are available as external libraries in case of C++ and Python.
+  Implementations of the Message Passing Interface (MPI) for handling communication 
+    in distributed-memory set-ups are available for all three languages.
+  }
+
+  \subsection{\changed{Productivity, ease of use and misuse}}
+
+  \changed{
+  The factors influencing the development and maintenance time/cost \changed{are}
+    of particular importance in scientific computing \citep{Wilson_2006}.
+  Among the three compared environments, 
+    Python gains significantly if code length or coding time
+    is prioritised \citep[see also discussion in][]{Lin_2012}.
+  Python has already been the language of choice for scientific software projects having code clarity 
+    or ease of use as the first requirement \citep[see e.g.][]{Barnes_and_Jones_2011}.
+  PyPy's capability to improve performance of unmodified Python code may 
+    make Python a favourable choice even if high performance is important, especially
+    if a combined measure of performance and development cost is to be considered.
+
+  Using the number of lines of code or the number of distinct language keywords
+    needed to implement a given logic as measures of syntax 
+    brevity, Python clearly surpasses its rivals.
+  Python was developed with emphasis on code readability and object-orientation.
+  Arguably, taking it to the extreme - Python uses line indentation to define 
+    blocks of code and treats even single integers as objects.
+  As a consequence Python is \changed{relatively} easy to learn and to teach.
+
+  Fortran's lack of exception mechanism poses a misuse risk when using 
+    both internal and external library calls.
+  The lack of exceptions results in a default policy to ignore 
+    recoverable errors. 
+  With no additional error-handling code, a Fortran program may silently continue after an  
+    error - additional code is needed to detect the error. 
+  In C++ and Python such program will stop by default, while additional code may be introduced to recover from the error condition.
+  Python does not feature such notorious mechanisms as
+    the preprocessor in C++ and the implicit typing in Fortran,
+    making it less prone to misuse.
+
+  Python implementations do not expose users to compilation or linking processes. 
+  As a result, Python-written software is easier to deploy and share, especially 
+    if multiple architectures and operating systems are targeted.
+  However, there exist tools such as CMake\footnote{CMake is a family of open-source, cross-platform
+    tools automating building, testing and packaging of C/C++/Fortran software,
+    see \url{http://cmake.org/}}
+    that allow to efficiently automate 
+    building, testing and packaging of C++ and Fortran programs.
+
+  It is worth noting one advantage of the C++/Blitz++ set-up.
+  Blitz++ ensures temporary-array-free computations by design
+    \citep{Veldhuizen_et_al_1997} avoiding unintentional performance loss.
+  In contrast, with both Fortran and Python, the memory footprint caused by employment
+    of temporary objects in array arithmetics is dependant on compiler choice or
+    the level of optimisations.
+
+  Finally, Python is definitely easiest to debug among the three languages.
+  Great debugging tools for C++ do exist, however the debugging and development is 
+    often hindered by indecipherable compiler messages
+    flooded with lengthy type names stemming from employment of templates.
+  Support for the OOP features of Fortran among compilers, 
+    debuggers and other programming aids remains immature at the time of writing.
+  }
 
@@
   The following sections provide a complete description of a minimal example of application
-    of the formul\ae~''translated'' into C++, Python and Fortran in the main body of the paper.
+    of the formul\ae~``translated'' into C++, Python and Fortran in the main body of the paper.
 
@@
   In all three implementations the \prog{ext()} function accept the second
-    argument to be an integer or a ''half'' (cf. section \ref{sec:grid}).
+    argument to be an integer or a ``half'' (cf. section \ref{sec:grid}).
 
@@
 
+    \changed{The final version of the manuscript benefited from suggestions of two
+      anonymous reviewers.}
+
     SA, AJ and DJ acknowledge funding from the Polish National Science Centre
@@
   \end{figure}
-  \includegraphics[height=\textheight]{demo}
+  \includegraphics[height=.7\textheight]{demo}
 

### 2014-02-20 paper/doc/paper.tex
changes before sending LaTeX files to Sci Prog

@@
-%%%%% <arXiv>
-%\documentclass[final,5p,times,twocolumn]{elsarticle}
-%%%%% </arXiv>
-
-%%%%% <sciprog>
 \documentclass[twocolumn]{article}
-\usepackage{lineno,authblk,latexsym,amsmath,float}
-\usepackage[colorlinks=true]{hyperref}
+\usepackage{authblk,latexsym,amsmath,float}
+\usepackage[runin]{abstract}
+\usepackage[colorlinks=true,allcolors=blue]{hyperref}
 \usepackage[hmargin=.55in,vmargin=.75in]{geometry}
 \usepackage[numbers]{natbib}
-\setlength{\columnsep}{1cm}
-%%%%% </sciprog>
-
-% highlighting (or not) the modified portions of the text
-\newcommand{\changed}[1]{{\color{blue}#1}}
-%\newcommand{\changed}[1]{#1}
-
-\usepackage{soul}
-\newcommand{\removed}[1]{{\color{blue}\st{#1}}}
+
+% do-nothing commands
+\newcommand{\changed}[1]{#1}
+\newcommand{\removed}[1]{#1}
 
@@
 
-% fr compact section headers
+% for compact section headers
 \usepackage[compact]{titlesec}
-\titlespacing*{\section}{0pt}{*0}{0pt}
 
@@
   frame=single,
-  fontsize=\relscale{.6},
+  fontsize=\relscale{.66},
   fontfamily=courier,
@@
 \newcommand{\codecpp}[4]{%
-  \linespread{.8}
   \addtocounter{lstnocpp}{1}%
@@
   \setcounter{linenocpp}{\value{FancyVerbLine}}%
-  \vspace{-1.5ex}
-  \linespread{2}
 }
 \newcommand{\codepyt}[4]{%
-  \linespread{.8}
   \addtocounter{lstnopyt}{1}%
@@
   \setcounter{linenopyt}{\value{FancyVerbLine}}%
-  \vspace{-1.5ex}
-  \linespread{2}
 }
 \newcommand{\codefor}[4]{%
-  \linespread{.8}
   \addtocounter{lstnofor}{1}%
@@
   \setcounter{linenofor}{\value{FancyVerbLine}}%
-  \vspace{-1.5ex}
-  \linespread{2}
 }
@@
 
-\linespread{2}
-
 \begin{document}
-    \linenumbers
 
@@
 
-    \author{~}
-%    \author[1]{Sylwester Arabas\footnote{Correspondence to: S. Arabas -- sarabas@igf.fuw.edu.pl -- 0048225546827 -- University of Warsaw, Pasteura 7, 02-093 Warszawa, Poland}}
-%    \author[1]{Dorota Jarecka}
-%    \author[1]{Anna Jaruga}
-%    \author[2]{Maciej Fijałkowski}
-
-%    \affil[1]{Institute of Geophysics, Faculty of Physics, University of Warsaw}
-%    \affil[2]{PyPy Team}
-
-    \onecolumn
+    \date{}
+
+    \author[1]{Sylwester Arabas}%\footnote{Correspondence to: S. Arabas -- sarabas@igf.fuw.edu.pl -- University of Warsaw, Pasteura 7, 02-093 Warszawa, Poland}}
+    \author[1]{Dorota Jarecka}
+    \author[1]{Anna Jaruga}
+    \author[2]{Maciej Fijałkowski}
+
+    \affil[1]{Institute of Geophysics, Faculty of Physics, University of Warsaw}
+    \affil[2]{PyPy Team}
+
+    \twocolumn[
       \maketitle
 
-    \begin{abstract}
+      \begin{onecolabstract}
 	  Three object-oriented implementations of a prototype solver of the advection equation are introduced.
@@
 	  {\bf Keywords:} object-oriented programming, advection equation, MPDATA, C++, Fortran, Python
-    \end{abstract}
-
-  \twocolumn
+	  ~\\
+      \end{onecolabstract}
+  ]
 
   \tableofcontents
+  \newpage
 
@@
 
-  \vspace{-5ex}
   \begin{equation}\label{eq:adv}
@@
 
-  \vspace{-5ex}
   \begin{equation*}\label{eq:pi}
@@
 
-  \vspace{-5ex}
   \begin{equation}\label{eq:donor}
@@
 
-  \vspace{-5ex}
   \begin{equation}\label{eq:donor:F}
@@
   \codecpp{code-cpp-listings.hpp}{listing07}{listing08}{1}
-  \vspace{-3.5ex}
   Equation~\ref{eq:donor} is split into the terms under the summation 
@@
   \codecpp{code-cpp-listings.hpp}{listing08}{listing09}{1}
-  \vspace{-3.5ex}
   and the actual two-dimensional donor-cell formula:
   \codecpp{code-cpp-listings.hpp}{listing09}{listing10}{1}
-  \vspace{-3.5ex}
   The same formul\ae~are expressed as follows in Python:
@@
   \codepyt{code-pyt-listings.py}{listing08}{listing09}{1}
-  \vspace{-3.5ex}
   The Fortran counterparts are: 
@@
 
-  \vspace{-5ex}
   \begin{equation}\label{eq:antidiff1}
@@
 
-  \vspace{-5ex}
   \begin{equation}\label{eq:antidiff2}
@@
 
-  \vspace{-5ex}
   \begin{equation}\label{eq:A}
@@
 
-%%%%%% <arXiv>
-%  \clearpage
-%%%%%% </arXiv>
+  \newpage
   \appendix
@@
   \end{figure}
-  \includegraphics[height=.7\textheight]{demo}
+  \begin{center}
+    \includegraphics[height=.83\textheight]{demo}
+  \end{center}
 

### 2014-07-01 paper/doc/grid.tex
adding grid.tex

--- 2013-07-26 paper/doc/paper.tex
@@
-%%%%% <arXiv>
-%\documentclass[final,5p,times,twocolumn]{elsarticle}
-%%%%% </arXiv>
-
-%%%%% <sciprog>
-\documentclass[twocolumn]{article}
-\usepackage{lineno,authblk,latexsym,amsmath,float}
-\usepackage[colorlinks=true]{hyperref}
-\usepackage[hmargin=.55in,vmargin=.75in]{geometry}
-\usepackage[numbers]{natbib}
-\setlength{\columnsep}{1cm}
-%%%%% </sciprog>
-
-% input encoding assuring proper handling of Polish characters (surnames!)
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-
-\usepackage{type1cm}
-\usepackage{relsize}
-
-\input{formulae_preamble}
-
-% for Arakawa-C figure
-\usepackage{tikz}
-\usetikzlibrary{calc}
-
-% for conpact itemize etc
-\usepackage{enumitem}
-\setitemize{noitemsep,itemsep=1pt,topsep=1pt,parsep=0pt,partopsep=0pt}
-
-% fr compact section headers
-\usepackage[compact]{titlesec}
-\titlespacing*{\section}{0pt}{*0}{0pt}
-
-% for source code listings
-\usepackage{fancyvrb}
-\fvset{
-  frame=single,
-  fontsize=\relscale{.6},
-  fontfamily=courier,
-  framerule=.2mm,
-  framesep=.7mm,
-  numbers=left,
-  xleftmargin=4mm,
-  numbersep=.25mm,
-  firstnumber=last
-}
-\input{pygments}
-
-\newcounter{lstnocpp}
-\newcounter{lstnopyt}
-\newcounter{lstnofor}
-
-\setcounter{lstnocpp}{-1}
-\setcounter{lstnopyt}{-1}
-\setcounter{lstnofor}{-1}
-
-\newcounter{linenocpp}
-\newcounter{linenopyt}
-\newcounter{linenofor}
-
-\setcounter{linenocpp}{0}
-\setcounter{linenopyt}{0}
-\setcounter{linenofor}{0}
-
-\newcommand*\FancyVerbStartString{}
-\newcommand*\FancyVerbStopString{}
-
-\newcommand{\codecpp}[4]{%
-  \linespread{.8}
-  \addtocounter{lstnocpp}{1}%
-  \renewcommand*\FancyVerbStartString{\PY{c+c1}{//#2}}%
-  \renewcommand*\FancyVerbStopString{\PY{c+c1}{//#3}}%
-  \setcounter{FancyVerbLine}{\thelinenocpp}%
-  \fvset{label={listing~C.\thelstnocpp~(C++)},rulecolor=\color{black},stepnumber=#4}%
-  \input{#1}%
-  \setcounter{linenocpp}{\value{FancyVerbLine}}%
-  \vspace{-1.5ex}
-  \linespread{2}
-}
-\newcommand{\codepyt}[4]{%
-  \linespread{.8}
-  \addtocounter{lstnopyt}{1}%
-  \renewcommand*\FancyVerbStartString{\PY{c}{\PYZsh{}#2}}
-  \renewcommand*\FancyVerbStopString{\PY{c}{\PYZsh{}#3}}
-  \setcounter{FancyVerbLine}{\thelinenopyt}%
-  \fvset{label={listing~P.\thelstnopyt~(Python)},rulecolor=\color{blue},stepnumber=#4}%
-  \input{#1}%
-  \setcounter{linenopyt}{\value{FancyVerbLine}}%
-  \vspace{-1.5ex}
-  \linespread{2}
-}
-\newcommand{\codefor}[4]{%
-  \linespread{.8}
-  \addtocounter{lstnofor}{1}%
-  \renewcommand*\FancyVerbStartString{\PY{c}{!#2}}
-  \renewcommand*\FancyVerbStopString{\PY{c}{!#3}}
-  \setcounter{FancyVerbLine}{\thelinenofor}%
-  \fvset{label={listing~F.\thelstnofor~(Fortran)},rulecolor=\color{red},stepnumber=#4}%
-  \input{#1}%
-  \setcounter{linenofor}{\value{FancyVerbLine}}%
-  \vspace{-1.5ex}
-  \linespread{2}
-}
-
-\newcommand{\prog}[1]{{\rm\bf#1}}
-
-\linespread{2}
-
-\begin{document}
-    \linenumbers
-
-    \title{
-      Formula translation in C++, Python and modern Fortran: a~case~study~of~the~language~choice tradeoffs
-    }
-
-    \author{~}
-%    \author[1]{Sylwester Arabas\footnote{Correspondence to: S. Arabas -- sarabas@igf.fuw.edu.pl -- 0048225546827 -- University of Warsaw, Pasteura 7, 02-093 Warszawa, Poland}}
-%    \author[1]{Dorota Jarecka}
-%    \author[1]{Anna Jaruga}
-%    \author[2]{Maciej Fijałkowski}
-
-%    \affil[1]{Institute of Geophysics, Faculty of Physics, University of Warsaw}
-%    \affil[2]{PyPy Team}
-
-    \onecolumn
-    \maketitle
-
-    \begin{abstract}
-        Three object-oriented implementations of a prototype solver of the advection equation are introduced.
-        The presented programs are based on Blitz++ (C++), NumPy (Python), and Fortran's built-in array containers.
-        The solvers constitute implementations of the Multidimensional Positive-Definite 
-          Advective Transport Algorithm (MPDATA).
-        The introduced codes serve as examples for how the application of 
-          object-oriented programming (OOP) techniques and new language constructs from
-          C++11 and Fortran 2008 allow to reproduce the mathematical notation 
-          used in the literature within the program code.
-        A discussion on the tradeoffs of the programming language choice is presented.
-        The main angles of comparison are code brevity and syntax clarity
-          (and hence maintainability and auditability) as well as performance.
-        All performance tests are carried out using free and open-source compilers.
-        In the case of Python, a significant performance gain is observed when switching from the standard 
-          interpreter (CPython) to the PyPy implementation of Python.
-        Entire source code of all three implementations is embedded in the text and is licensed
-          under the terms of the GNU GPL license.
-
-        ~\\
-        {\bf Keywords:} object-oriented programming, advection equation, MPDATA, C++, Fortran, Python
-    \end{abstract}
-
-  \twocolumn
-
-  \tableofcontents
-
-  \section{Introduction}
-
-  Object oriented programming (OOP) {\em ''has become recognised as the almost unique successful 
-    paradigm for creating complex software''} \citep[][Sec.~1.3]{Press_et_al_2007}.
-  It is intriguing that, while the quoted statement comes from the very book subtitled 
-   {\em The Art of Scientific Computing}, hardly any (if not none) of the currently operational 
-    weather and climate prediction systems - flagship examples of complex scientific software - 
-    make extensive use of OOP techniques\footnote{
-  Fortran has been the language of choice in oceanic \citep{Griffies_et_al_2000}, 
-    weather-prediction \citep{Sundberg_2009} and Earth system \citep{Legutke_2012} modelling, 
-    and none of its 20\textsuperscript{th}-century editions were object-oriented languages \citep[see e.g.][for discussion]{Norton_et_al_2007}}.
-
-  Application of OOP techniques in development of numerical modelling software may help to:
-  \begin{enumerate}[label=(\roman*), leftmargin=*, widest=ii]
-    \item{maintain modularity and separation of program logic layers (e.g. separation of
-      numerical algorithms, parallelisation mechanisms, data input/output, error handling and
-      the description of physical processes); and}
-    \item{{\bf shorten and simplify the source code and improve its readability by reproducing within 
-      the program logic the mathematical notation used in the literature}.}
-  \end{enumerate}
-  The first application is attainable, yet arguably cumbersome, with procedural programming.
-  The latter, virtually impossible to obtain with procedural programming, is the focus of this paper.
-  The importance of reproducing the mathematical notation in the code lays primarily
-    in the fact that code readability and brevity 
-    %\citep[and not the language choice itself,][]{Seeley_2004}
-    significantly contribute to code maintainability \citep{Wilson_et_al_2012}.
-
-  The key aim of this paper is to show how OOP techniques can be used to faithfully
-    reproduce within the code what can be referred to as {\em blackboard abstractions} \citep{Rouson_et_al_2012}.
-  For this purpose, a sample implementation of a numerical scheme for solving
-    the advection equation is introduced in C++, Python and modern Fortran --
-    OOP languages commonly used in scientific computing \citep[see e.g.][chapt.~8]{Einarsson_2005}. 
-  Presented implementations and the results of benchmark tests
-    provide a basis for discussion on the tradeoffs of programming language choice.
-  The discussion concerns in principle the development of finite-difference solvers for 
-    partial differential equations, but is likely applicable to some extent to the scientific programming in general.
-
-  All three programs include an equally structured implementation of the two-dimensional version of the 
-    Multidimensional Positive Definite Advective Transport Algorithm \citep[MPDATA;][]{Smolarkiewicz_1984}.
-  MPDATA is an example of a numerical procedure used in weather, climate, ocean and solar simulation systems
-    \citep[e.g.][respectively]{Ziemianski_et_al_2011,Abiodun_et_al_2011,Ezer_et_al_2002,Charbonneau_and_Smolarkiewicz_2013}.
-  The basic MPDATA scheme presented herein is complex enough to contain
-    a wide range of mathematical abstractions that can be represented using OOP constructs, 
-    yet it is simple enough to allow inclusion of the entire source code within the paper text.
-  All relevant MPDATA formul\ae~are given in the text alongside corresponding code fragments
-    allowing comparison of the relevant syntax with the mathematical notation.
-  These formul\ae~are presented without derivation or detailed discussion
-    (see~\citep{Smolarkiewicz_2006} for a recent review of MPDATA-based techniques including an introductory description 
-    of the algorithm and an exhaustive list of references).
-
-  The paper is structured as follows.
-  In section~\ref{sec:impl} we introduce the ''formula translation'' part of the three implementations
-    briefly describing the algorithm itself and
-    discussing where and how the OOP techniques were applied in its implementation.
-  The remaining part of the implementations -- the solver logic -- is
-    presented in appendix~\ref{app:solver}.
-  A usage example is given in appendix~\ref{app:usage}
-  Section~\ref{sec:perf} covers performance evaluation of the three implementations.
-  Section~\ref{sec:tradeoffs} covers discussion of the tradeoffs of the programming language choice.
-  Section~\ref{sec:concl} closes the article with a brief summary.
-
-  The entire code is licensed under the terms of the GNU General Public License version 3 \citep{GPLv3}.
-  All listings include line numbers printed to the left of the source code, with separate numbering for
-    C++ (listings prefixed with C, black frame),
-  \codecpp{code-cpp-listings.hpp}{listing00}{listing01}{1}
-    Python (listings prefixed with P, blue frame) and
-  \codepyt{code-pyt-listings.py}{listing00}{listing01}{1}
-    Fortran (listings prefixed with F, red frame).
-  \codefor{code-for-listings.f}{listing00}{listing01}{1}
-  Programming language constructs when inlined in the text are 
-    typeset in bold, e.g. \prog{GOTO 2}.
-
-  \section{Implementation of the formul\ae}\label{sec:impl}
-
-  Double precision floating-point format is used in all three implementations.
-  The codes begin with the following definitions:
-  \codecpp{code-cpp-listings.hpp}{listing01}{listing02}{1}
-  \codepyt{code-pyt-listings.py}{listing01}{listing02}{1}
-  \codefor{code-for-listings.f}{listing01}{listing02}{1}
-  which provide a convenient way of switching to different precision.
-
-  All codes are structured in a way allowing compilation of the code 
-    in exactly the same order as presented in the text within one source file,
-    hence every Fortran listing contains definition of a separate module.
-
-  The language syntax and OOP nomenclature are used without introduction in the paper,
-    for an overview of OOP in context of C++, Python and Fortran, consult for example
-    \citep[][Part~II]{Stroustrup_2000}, \citep[][Chapter~5]{Pilgrim_2004} and
-    \citep[][Chapter~11]{Markus_2012}, respectively.
-
-  \subsection{Array containers}\label{sec:array}
-
-  MPDATA is, in its most basic form presented herein, a solver for systems of advection equations of the following form:
-
-  \vspace{-5ex}
-  \begin{equation}\label{eq:adv}
-    \input{formulae_adv}
-  \end{equation}
-  that describe evolution of a scalar field $\psi$ transported by the fluid flow with velocity $\vec{v}$.
-  Solution of equation (\ref{eq:adv}) using MPDATA implies discretisation onto a grid of~the~$\psi$~and~the~Courant 
-    number $\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}$ fields, where $\Delta t$ is the 
-    solver timestep and $\Delta x$ is the grid spacing.
-
-  Presented C++ implementation of MPDATA is built upon the Blitz++ 
-    library\footnote{Blitz++ is a C++ class library for scientific computing which uses the expression templates technique to achieve high performance, see \url{http://sf.net/projects/blitz/}}.
-  Blitz offers object-oriented representation of n-dimensional arrays,
-    and array-valued mathematical expressions.
-  In particular, it offers loop-free notation for array arithmetics
-    that does not incur creation of intermediate temporary objects.
-  Blitz++ is a header-only library\footnote{Blitz++ requires linking with \prog{libblitz} if debugging mode is used} -- to use it, it is enough to include the appropriate header file,
-    and optionally expose the required classes to the present namespace:
-  \codecpp{code-cpp-listings.hpp}{listing02}{listing03}{1}
-  Here \prog{arr\_t}, \prog{rng\_t} and \prog{idx\_t} serve as alias identifiers 
-    and are introduced in order to shorten the code.
-
-  The power of Blitz++ comes from the ability to express array expressions as objects.
-  In particular, it is possible to define a function that returns an array expression;
-    i.e. not the resultant array, but an object representing a ,,recipe'' defining the operations
-    to be performed on the arguments.
-  As a consequence, the return types of such functions become unintelligible.
-  Luckily, the \prog{auto} return type declaration from the C++11 standard allows to simplify the code significantly,
-    even more if used through the following preprocessor macro:
-  \codecpp{code-cpp-listings.hpp}{listing03}{listing04}{1}
-  The call to \prog{blitz::safeToReturn()} function is included in order to ensure that
-    all arrays involved in the expression being returned continue to exist in the
-    caller scope.
-  For example, definition of a function returning its array-valued argument doubled, reads:
-    \prog{auto~f(arr\_t~x)~return\_macro(2*x)}.
-  This is the only preprocessor macro defined herein.
-
-  For the Python implementation of MPDATA the NumPy\footnote{NumPy is a Python package
-    for scientific computing offering support for multi-dimensional arrays and a library
-    of numerical algorithms, see \url{http://numpy.org/}} package is used.
-  In order to make the code compatible with both the standard CPython
-    as well as the alternative PyPy implementation of Python \citep[][]{Bolz_et_al_2011},
-    the Python code includes the following sequence of \prog{import} statements:
-  \codepyt{code-pyt-listings.py}{listing02}{listing03}{1}
-  First, the PyPy's built-in NumPy implementation named \prog{numpypy} is imported if applicable (i.e. if running PyPy), 
-    and the lazy evaluation mode is turned on through the \prog{set\_invalidation(False)} call.
-  PyPy's lazy evaluation obtained with the help of a just-in-time compiler enables to achieve
-    an analogous to Blitz++ temporary-array-free handling of array-valued expressions 
-    (see discussion in section~\ref{sec:perf}).
-  Second, to match the settings of C++ and Fortran compilers used herein, the NumPy package is instructed 
-    to ignore any floating-point errors, if such an option
-    is available in the interpreter\footnote{\prog{numpy.seterr()} is not supported in PyPy as of version 1.9}.
-  The above lines conclude all code modifications that needed to be added in order to run
-    the code with PyPy.
-
-  Among the three considered languages only Fortran is equipped with built-in
-    array handling facilities of practical use in high-performance computing. 
-  Therefore, there is no need for using an external package as with C++ and Python.
-  Fortran array-handling features are not object-oriented, though.
-
-  \subsection{Containers for sequences of arrays}\label{sec:sequence}
-
-  As discussed above, discretisation in space of the scalar field $\psi(x,y)$ into its $\psi_{[i,j]}$ 
-    grid representation requires floating-point array containers.
-  In turn, discretisation in time requires a container class for storing
-    sequences of such arrays, i.e. \{$\psi^{[n]}$, $\psi^{[n+1]}$\}.
-  Similarly the components of the vector field $\vec{C}$ are in fact a \{$C^{[x]}$, $C^{[y]}$\} 
-    array sequence.
-
-  Using an additional array dimension to represent the sequence elements is not considered for two reasons.
-  First, the $C^{[x]}$ and $C^{[y]}$ arrays constituting the sequence have different sizes
-    (see discussion of the Arakawa-C grid in section~\ref{sec:grid}).
-  Second, the order of dimensions would need to be different for different languages to assure that
-    the contiguous dimension is used for one of the space dimensions and not for time levels.
-
-  In the C++ implementation the Boost\footnote{
-    Boost is a free and open-source collection of peer-reviewed C++ libraries available at \url{http://www.boost.org/}.
-    Several parts of Boost have been integrated into or inspired new additions to the C++ standard.
-  } \prog{ptr\_vector} class is used to represent sequences of Blitz++ arrays 
-    and at the same time to handle automatic freeing of dynamically allocated memory.
-  The \prog{ptr\_vector} class is further customised by defining a derived structure which element-access \prog{[~]} 
-    operator is overloaded with a modulo variant:
-  \codecpp{code-cpp-listings.hpp}{listing04}{listing05}{1}
-  Consequently the last element of any such sequence may be accessed at index \prog{-1}, the last but one at \prog{-2}, 
-    and so on.
-
-  In the Python implementation the built-in \prog{tuple} type is used to store sequences of NumPy arrays.
-  Employment of negative indices for handling from-the-end addressing of elements
-    is a built-in feature of all sequence containers in Python.
-
-  Fortran does not feature any built-in sequence container capable of storing arrays, 
-    hence a custom \prog{arrvec\_t} type is introduced:
-  \codefor{code-for-listings.f}{listing02}{listing03}{1}
-  The \prog{arr\_t} type is defined solely for the purpose of overcoming the limitation 
-    of lack of an array-of-arrays construct, and its only member field is a two-dimensional array.
-  An array of \prog{arr\_t} is used hereinafter as a container for sequences of arrays.
-
-  The \prog{arrptr\_t} type is defined solely for the purpose of overcoming Fortran's limitation
-    of not supporting allocatables of pointers.
-  \prog{arrptr\_t}'s single member field is a pointer to an instance of \prog{arr\_t}.
-  Creating an allocatable of \prog{arrptr\_t}, instead of 
-    a multi-element pointer of \prog{arr\_t}, ensures automatic memory deallocation.
-
-  Type \prog{arrptr\_t} is used to implement the from-the-end addressing of elements in \prog{arrvec\_t}.
-  The array data is stored in the \prog{arrs} member field (of type \prog{arr\_t}).
-  The \prog{at} member field (of type \prog{arrptr\_t}) stores pointers to the elements of \prog{arrs}. 
-  \prog{at} has double the length of \prog{arrs} and is initialised in a cyclic manner so that
-    the \prog{-1} element of \prog{at} points to the last element of \prog{arrs}, and so on.
-  Assuming \prog{psi} is an instance of \prog{arrptr\_t}, the \prog{(i,j)} element of the \prog{n}-th array in \prog{psi}
-    may be accessed with\\ \prog{psi\%at( n )\%p\%a( i, j )}.
-
-  The \prog{ctor(n)} method initialises the container for a given number of elements \prog{n}.
-  The \prog{init(n,i,j)} method initialises the \prog{n}-th element of the container with 
-    a newly allocated 2D array spanning indices \prog{i(1)}:\prog{i(2)}, 
-    and \prog{j(1)}:\prog{j(2)} in the first, and last dimensions respectively\footnote{In Fortran, when an array
-    is passed as a function argument its base is locally set to unity, regardless of the setting
-    at the caller scope.}.
-
-  \subsection{Staggered grid}\label{sec:grid}
-
-  \begin{figure}[h!]
-  \center
   \begin{tikzpicture}
@@
   \end{tikzpicture}
-  \caption{\label{fig:grid}
-    A schematic of the Arakawa-C grid.
-  }
-  \end{figure}
-  The so-called Arakawa-C staggered grid \citep{Arakawa_and_Lamb_1977} depicted in Figure~\ref{fig:grid}
-    is a natural choice for MPDATA.
-  As a consequence, the discretised representations of the $\psi$ scalar field, and each 
-    component of the ${\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}}$ vector field 
-    in eq.~(\ref{eq:adv}) are defined over different grid point locations.
-  In mathematical notation this can be indicated by usage of fractional indices, e.g.
-    $C^{[x]}_{[i\mhlf,j]}$, $C^{[x]}_{[i+\phlf,j]}$, $C^{[y]}_{[i,j\mhlf]}$ and $C^{[y]}_{[i,j+\phlf]}$
-    to depict the grid values of the ${\vec{C}}$ vector components surrounding $\psi_{[i,j]}$.
-  However, fractional indexing does not have a built-in counterpart in any of the
-    employed programming languages.
-  A desired syntax would translate \prog{$i-\phlf$} to \prog{$i-1$} and
-    \prog{$i+\phlf$} to \prog{$i$}.
-  OOP offers a convenient way to implement such notation
-    by overloading the \prog{+} and \prog{-} operators for objects representing array indices.
-
-  In the C++ implementation first a global instance \prog{h} of an empty structure 
-    \prog{hlf\_t} is defined, and then the plus and minus operators for \prog{hlf\_t} and \prog{rng\_t} are overloaded:
-  \codecpp{code-cpp-listings.hpp}{listing05}{listing06}{1}
-  This way, the arrays representing vector field components can be indexed using
-    \prog{(i+h,j)}, \prog{(i-h,j)} etc. where \prog{h}~represents the half.
-
-  In NumPy, in order to prevent copying of array data during slicing, one needs to operate on the
-    so-called array views.
-  Array views are obtained when indexing the arrays with objects of the Python's
-    built-it \prog{slice} type (or tuples of such objects in case of multi-dimensional arrays).
-  Python forbids overloading of operators of built-in types such as \prog{slices}, 
-    and does not define addition/subtraction operators for \prog{slice} and \prog{int} pairs.
-  Consequently, a custom logic has to be defined not only for fractional indexing,
-    but also for shifting the slices by integer intervals ($i\pm1$).
-  It is implemented here by declaring a \prog{shift} class with the adequate operator overloads:
-  \codepyt{code-pyt-listings.py}{listing03}{listing04}{1}
-    and two instances of it to represent unity and half
-    in expressions like \prog{i+one}, \prog{i+hlf}, where \prog{i} is an instance of \prog{slice}%
-    \footnote{\label{fnt:slice}One could argue that not using an own implementation of a slice-representing class
-    in NumPy is a design flaw -- being able to modify behaviour of a hypothetical numpy.slice class 
-    through inheritance would allow to implement the same behaviour as obtained in listing P.3 without the need to represent 
-    the unity as a separate object}:
-  \codepyt{code-pyt-listings.py}{listing04}{listing05}{1}
-
-  In the Fortran implementation fractional array indexing is obtained through
-    definition and instantiation of an object representing the half, and having appropriate
-    operator overloads:
-  \codefor{code-for-listings.f}{listing03}{listing04}{1}
-
-  \subsection{Array index permutations}\label{sec:pi}
-  Hereinafter, the $\pi_{a,b}^{d}$ symbol is used to denote a  
-    cyclic permutation of an order $d$ of a set $\{a,b\}$.
-  It is used to generalise the MPDATA formul\ae~into multiple dimensions
-    using the following notation:
-
-  \vspace{-5ex}
-  \begin{equation*}\label{eq:pi}
-    \input{formulae_pi}
-  \end{equation*}
-
-  Blitz++ ships with the \prog{RectDomain} class (aliased here as \prog{idx\_t}) 
-    for specifying array ranges in multiple dimensions.
-  The $\pi$ permutation is implemented in C++ as a function \prog{pi()}
-    returning an instance of \prog{idx\_t}.
-  In order to ensure compile-time evaluation, the permutation order is passed 
-    via the template parameter \prog{d}
-    (note the different order of \prog{i} and \prog{j} arguments in the two template specialisations):
-  \codecpp{code-cpp-listings.hpp}{listing06}{listing07}{1}
-
-  NumPy uses tuples of slices for addressing multi-dimensional array with a
-    single object.
-  Therefore, the following definition of function \prog{pi()} suffices to represent~$\pi$:
-  \codepyt{code-pyt-listings.py}{listing05}{listing06}{1}
-
-  Fortran does not feature an analogous mechanism for specifying array ranges in multiple
-    dimensions using a single entity.
-  As a workaround, in in the Fortran implementation, \prog{pi()} returns a pointer to the array elements specified
-    by \prog{i} and \prog{j} interpreted as (i,j) or
-    (j,i) depending on the value of the argument \prog{d}.
-% TODO: disucss creation of temporary arrays and disabling of vectorisation due to employment of POINTERS
-  In addition to \prog{pi()}, a helper \prog{span()} function returning the 
-    length of one of the vectors passed as argument is defined:
-  \codefor{code-for-listings.f}{listing04}{listing05}{1}
-  The \prog{span()} function is used to shorten the declarations of arrays to be returned 
-    from functions in the Fortran implementation (see listings F.7 and F.12--F.15).
-
-  It is worth noting here that the C++ implementation of \prog{pi()} is
-    branchless thanks to employment of template specialisation.
-  With Fortran one needs to rely on compiler optimisations to eliminate the conditional expression
-    within \prog{pi()} that depends on value of \prog{d} which is always known 
-    at compile time.
-
-  \subsection{Donor-cell formul\ae}\label{sec:donor}
-
-  MPDATA is an iterative algorithm in which each iteration takes the form of the 
-    so-called donor-cell formula (which itself is a first-order advection scheme).
-
-  MPDATA and donor-cell are explicit forward-in-time algorithms -- they allow to predict $\psi^{[n+1]}$ as a
-    function of $\psi^{[n]}$ where $n$ and $n+1$ denote two adjacent time levels.
-  The donor-cell scheme may be written as \citep[eq. 2 in][]{Smolarkiewicz_1984}:
-
-  \vspace{-5ex}
-  \begin{equation}\label{eq:donor}
-    \input{formulae_donor}
-  \end{equation}
-  where $N$ is the number of dimensions, 
-    and F is the so-called flux function \citep[][eq.~3]{Smolarkiewicz_1984}:
-
-  \vspace{-5ex}
-  \begin{equation}\label{eq:donor:F}
-    \input{formulae_donor-F}
-  \end{equation}
-
-  \noindent The flux function takes the following form in C++:
-  \codecpp{code-cpp-listings.hpp}{listing07}{listing08}{1}
-  \vspace{-3.5ex}
-  Equation~\ref{eq:donor} is split into the terms under the summation 
-    (effectively the 1-dimensional donor-cell formula):
-  \codecpp{code-cpp-listings.hpp}{listing08}{listing09}{1}
-  \vspace{-3.5ex}
-  and the actual two-dimensional donor-cell formula:
-  \codecpp{code-cpp-listings.hpp}{listing09}{listing10}{1}
-  \vspace{-3.5ex}
-  The same formul\ae~are expressed as follows in Python:
-  \codepyt{code-pyt-listings.py}{listing06}{listing07}{1}
-  \codepyt{code-pyt-listings.py}{listing07}{listing08}{1}
-  \codepyt{code-pyt-listings.py}{listing08}{listing09}{1}
-  \vspace{-3.5ex}
-  The Fortran counterparts are: 
-  \codefor{code-for-listings.f}{listing05}{listing06}{1}
-  \codefor{code-for-listings.f}{listing06}{listing07}{1}
-  \codefor{code-for-listings.f}{listing07}{listing08}{1}
-  \codefor{code-for-listings.f}{listing08}{listing09}{1}
-  \codefor{code-for-listings.f}{listing09}{listing10}{1}
-
-  The brevity of the code in the above listings as well as its similarity
-    to the mathematical notation is the main point of this paper.
-  The ''formula translation'' features include:
-  \begin{itemize}
-    \item{loop-free notation;}
-    \item{array-valued functions enabling reuse of sub-expressions;}
-    \item{fractional indexing obtained with the help of operator overloading;}
-    \item{dimension-independent indexing with the help of \prog{pi} permutation functions.}
-  \end{itemize}
-  The same features are applied to translation of more complex formul\ae~in the
-    following section.
-
-  \subsection{MPDATA formul\ae}\label{sec:mpdata}
-
-  MPDATA introduces corrective steps to the algorithm defined by equation~\ref{eq:donor} and \ref{eq:donor:F}.
-  Each corrective step is a donor-cell step (eq.~\ref{eq:donor}) with the Courant number
-    fields corresponding to the MPDATA antidiffusive velocities of the following 
-    form \citep[eqs~13,~14~in][]{Smolarkiewicz_1984}:
-
-  \vspace{-5ex}
-  \begin{equation}\label{eq:antidiff1}
-    \input{formulae_antidiff1}
-  \end{equation}
-  where $\psi$ and $C$ represent values from the previous iteration and where:
-
-  \vspace{-5ex}
-  \begin{equation}\label{eq:antidiff2}
-    \input{formulae_antidiff2}
-  \end{equation}
-  For positive-definite $\psi$, the $A$ and $B$ terms take the 
-  following form\footnote{
-    Since $\psi\ge0$, $|A|\le1$ and $|B|\le1$. 
-    See \citet[Sec. 4.2]{Smolarkiewicz_2006} for description of adaptation of the 
-    formul\ae~for advection of fields of variable sign
-  }:
-
-  \vspace{-5ex}
-  \begin{equation}\label{eq:A}
-    \input{formulae_A}
-  \end{equation}
-  \begin{equation}\label{eq:B}
-    \input{formulae_B}
-  \end{equation}
-
-  If the (positive-defined) denominator in equations \ref{eq:A} or \ref{eq:B} 
-    equals zero for a given {\em i} and {\em j}, %TODO: >0 in the code
-    the corresponding $A_{[i,j]}$ and $B_{[i,j]}$ are set to zero what may be conveniently 
-    represented with the \prog{where} construct in all three considered languages:
-  \codecpp{code-cpp-listings.hpp}{listing10}{listing11}{1}
-  \codepyt{code-pyt-listings.py}{listing09}{listing10}{1}
-  \codefor{code-for-listings.f}{listing10}{listing11}{1}
-  \codefor{code-for-listings.f}{listing11}{listing12}{1}
-
-  The $A$ term defined in equation \ref{eq:A} takes the following form:
-  \codecpp{code-cpp-listings.hpp}{listing11}{listing12}{1}
-  \codepyt{code-pyt-listings.py}{listing10}{listing11}{1}
-  \codefor{code-for-listings.f}{listing12}{listing13}{1}
-
-  The $B$ term defined in equation \ref{eq:B} takes the following form:
-  \codecpp{code-cpp-listings.hpp}{listing12}{listing13}{1}
-  \codepyt{code-pyt-listings.py}{listing11}{listing12}{1}
-  \codefor{code-for-listings.f}{listing13}{listing14}{1}
-
-  Equation \ref{eq:antidiff2} takes the following form:
-  \codecpp{code-cpp-listings.hpp}{listing13}{listing14}{1}
-  \codepyt{code-pyt-listings.py}{listing12}{listing13}{1}
-  \codefor{code-for-listings.f}{listing14}{listing15}{1}
-
-  Equation \ref{eq:antidiff1} takes the following form:
-  \codecpp{code-cpp-listings.hpp}{listing14}{listing15}{1}
-  \codepyt{code-pyt-listings.py}{listing13}{listing14}{1}
-  \codefor{code-for-listings.f}{listing15}{listing16}{1}
-  \codefor{code-for-listings.f}{listing16}{listing17}{1}
-
-  The above listings conclude the formula-translation part of this paper.
-  Implementation of a prototype MPDATA solver using the above code
-    is presented in appendix~\ref{app:solver}.
-
-  \section{Performance evaluation}\label{sec:perf}
-
-  \subsection{Set up}
-
-  The three introduced implementations of MPDATA were tested with the following set-ups 
-    employing free and open-source tools:
-  \begin{description}
-    \item[C++:]{~
-      \begin{itemize}
-        \item{GCC g++ 4.8.0\footnote{\label{fnt:gcc-snapshot}GNU Compiler Collection packaged in the Debian's gcc-snapshot\_20130222-1} 
-          and Blitz++ 0.10}
-        \item{LLVM Clang 3.2 and Blitz 0.10}
-      \end{itemize}
-    }
-    \item[Python:]{~
-      \begin{itemize}
-        \item{CPython 2.7.3 and NumPy 1.7}
-        \item{PyPy 1.9.0 with built-in NumPy implementation}
-      \end{itemize}
-    }
-    \item[Fortran:]{~
-      \begin{itemize}
-        \item{GCC gfortran 4.8.0\textsuperscript{\ref{fnt:gcc-snapshot}}}
-      \end{itemize}
-    }
-  \end{description}
-  The performance tests were run on a Debian and an Ubuntu GNU/Linux systems with the above-listed software obtained 
-    via binary packages from the distributions' package repositories (most recent package versions at the time of writing).
-  The tests were performed on two 64-bit machines equipped with
-    an AMD Phenom\textsuperscript{\texttrademark} II X6 1055T (800 MHz)
-    and an Intel\textsuperscript{\textregistered} Core\textsuperscript{\texttrademark}~i5-2467M (1.6 GHz)
-    processors.
-
-  For both C++ and Fortran the GCC compilers were invoked with the \prog{-Ofast} and the 
-    \prog{-march=native} options.
-  The Clang compiler was invoked with the \prog{-O3}, the \prog{-mllvm -vectorize}, the \prog{-ffast-math} 
-    and the \prog{-march=native} options.
-  The CPython interpreter was invoked with the \prog{-OO} option.
-
-  In addition to the standard Python implementation CPython,
-    the Python code was tested with PyPy.
-  PyPy is an alternative implementation of Python featuring a just-in-time compiler. 
-  PyPy includes an experimental partial re-implementation of NumPy that compiles NumPy expressions into native assembler.
-  Thanks to employment of lazy evaluation of array expressions (cf. Sect. \ref{sec:array})
-    PyPy allows to eliminate the use of temporary matrices for storing intermediate results,
-    and to perform multiple operations on the arrays within a single array index traversal
-    \footnote{Lazy evaluation available in PyPy 1.9 has been temporarily removed from PyPy during a refactoring of
-      the code. It'll be reinstantiated in the codebase as soon as possible, but past PyPy 2.0 release}.
-  Consequently, PyPy allows to overcome the same performance-limiting factors as those addressed by Blitz++, although 
-    the underlying mechanisms are different.
-  In contrast to other solutions for improving performance of NumPy-based codes such as
-    Cython\footnote{see \url{http://cython.org}}, 
-    numexpr\footnote{see \url{http://code.google.com/p/numexpr/}} or 
-    Numba\footnote{see \url{http://numba.pydata.org/}}, 
-    PyPy does not require any modifications to the code.
-  Thus, PyPy may serve as a drop-in replacement for CPython ready to be used with 
-    previously-developed codes.
-
-  The same set of tests was run with all four set-ups.
-  Each test set consisted of 16 program runs.
-  The test programs are analogous to the example code presented in appendix~\ref{app:usage}.
-  The tests were run with different grid sizes ranging from 64$\times$64 to 2048$\times$2048.
-  The Gaussian impulse was advected for $nt=2^{24}/(nx\cdot ny)$ timesteps ($2^{24}$ chosen arbitrarily), 
-    in order to assure comparable timing accuracy for all grid sizes.
-  Three MPDATA iterations were used (i.e. two corrective steps).
-  %The initial condition was loaded from a text file, and the final values were compared at the end of the test
-  %  with values loaded from another text file assuring the same results were obtained with all four set-ups.
-  The tests were run multiple times; program start-up, data loading, and output verification times were
-    subtracted from the reported values (see caption of Figure~\ref{fig:cpu-eyrie} for details).
-
-  \begin{figure}[h!]
-    \center
-    \includegraphics[height=.45\textwidth,angle=-90]{skua-mem}
-    \caption{\label{fig:mem}
-      Memory consumption statistics for the test runs described in Section~\ref{sec:perf}
-        plotted as a function of grid size.
-      Peak resident set size (rss) values reported by the GNU time utility are normalised by the size of
-        data that needs to be allocated in the program to store all declared grid-sized arrays.
-      Asymptotic values reached at the largest grid sizes are indicative 
-        of temporary storage requirements.
-    }
-  \end{figure}
-
-  \subsection{Results}
-
-  Figure \ref{fig:mem} presents a plot of the peak memory use\footnote{The resident set size (rss)
-    as reported by GNU time (version 1.7-24)} (identical for both considered CPUs)
-    as a function of grid size.
-  The plotted values are normalised by the nominal size of all data arrays used in the program
-    (i.e. two (nx+2)$\times$(ny+2) arrays representing the two time levels of $\psi$, 
-     a (nx+1)$\times$(ny+2) array representing the $C^{[x]}$ component of the Courant number field,
-     a (nx+2)$\times$(ny+1) array representing the $C^{[y]}$ component, 
-     and two pairs of arrays of the size of $C^{[x]}$ and $C^{[y]}$ for storing the 
-     antidiffusive velocities, all composed of 8-byte double-precision floating point numbers).
-  Plotted statistics reveal a notable memory footprint of the Python interpreter itself
-    for both CPython and PyPy, losing its significance for domains larger than 1024$\times$1024.
-  The roughly asymptotic values reached in all four set-ups for grid sizes larger that 1024$\times$1024
-    are indicative of the amount of temporary memory used for array manipulation.
-  PyPy- and Blitz++-based set-ups consume notably less memory than Fortran and CPython.
-  This confirms the effectiveness of the just-in-time compilation (PyPy) and the expression-templates (Blitz++) techniques
-    for elimination of temporary storage during array operations.
-
-  \begin{figure}[h!]
-    \center
-    \includegraphics[height=.45\textwidth,angle=-90]{laptok-cpu}
-    \caption{\label{fig:cpu-eyrie}
-      Execution time statistics for the test runs described in Section~\ref{sec:perf}
-        plotted as a function of grid size.
-      Values of the total user mode CPU time reported by the GNU time utility are
-        normalised by the grid size ($nx \cdot ny$) and the number of timesteps $nt=2^{24}/(nx \cdot ny)$.
-      Before normalisation the time reported for an $nt=0$ run for a corresponding
-        domain size is subtracted from the values.
-      Both the $nt=0$ and $nt=2^{24}/(nx \cdot ny)$ runs are repeated three times and
-        only the shortest time is taken into account.
-      Results obtained with an Intel\textsuperscript{\textregistered} 
-        Core\textsuperscript{\texttrademark} i5 1.6 GHz processor.
-    }
-  \end{figure}
-  \begin{figure}[th!]
-    \center
-    \includegraphics[height=.45\textwidth,angle=-90]{skua-cpu}
-    \caption{\label{fig:cpu-skua}
-      Same as Fig.~\ref{fig:cpu-eyrie} for an AMD Phenom\textsuperscript{\texttrademark} II 800 MHz processor.
-    }
-  \end{figure}
-
-  The CPU time statistics presented in Figures \ref{fig:cpu-eyrie} and~\ref{fig:cpu-skua} reveal
-    minor differences between results obtained with the two different processors.
-  Presented results lead to the following observations
-    (where by referring to language names, only the results obtained with the herein considered
-     program codes, and software/hardware configurations are meant):
-  \begin{itemize}
-    \item{Fortran gives shortest execution times for any domain size;}
-    \item{C++ execution times are less than twice those of Fortran for grids larger than 
-      256$\times$256;}
-    \item{CPython requires from around 4 to almost 10 times more CPU time than Fortran depending on the grid size;}
-    \item{PyPy execution times are in most cases closer to C++ than to CPython.}
-  \end{itemize}
-  The support for OOP features in gfortran, the NumPy support in PyPy, and the relevant optimisation
-    mechanisms in GCC are still in active development and hence the performance with some of the set-ups may 
-    likely change with newer versions of these packages.
-
-  It is worth mentioning, that even though the three implementations are equally structured,
-    the three considered languages have some inherent differences influencing the execution times.
-  Notably, while Fortran and Blitz++ offer runtime array-bounds and array-shape checks as options
-    not intended for use in production binaries, NumPy performs them always.
-  Additionally, the C++ and Fortran set-ups may, in principle, benefit from GCC's auto-vectorisation
-    features which do not have yet counterparts in CPython or PyPy.
-  Finally, Fortran uses different ordering for storing array elements in memory, but since
-    all tests were carried out using square grids, this should not have had any impact on the
-    performance\footnote{Both Blitz++ and NumPy support Fortran's column-major ordering as well, 
-    however this feature is still missing from PyPy's built-in NumPy implementation as of PyPy 1.9}.
-
-  \section{The tradeoffs of language choice}\label{sec:tradeoffs}
-
-  \subsection{Representability of blackboard abstractions}
-
-  It was shown in section~\ref{sec:impl} that C++11/Blitz++, Python/NumPy and Fortran 2008
-    provide comparable functionalities in terms of matching the blackboard abstractions
-    within the program code.
-  Taking into account solely the part of code representing particular formul\ae~
-    (e.g. listings C.14, P.13, F.15 and equation \ref{eq:antidiff1}) all three
-    languages allow to match (or surpass) \LaTeX~in its brevity of formula translation syntax.
-  All three languages were shown to be capable of providing mechanisms to compactly represent such abstractions as:
-  \begin{itemize}
-    \item{loop-free array arithmetics;}
-    \item{functions returning array-valued expressions;}
-    \item{permutations of array indices allowing dimension-independent definitions
-      of functions (see e.g. listings C.8 and C.9, P.7 and P.8, F.7 and F.8);}
-    \item{fractional indexing of arrays corresponding to employment of a staggered grid.}
-  \end{itemize}
-
-  Making use of features such as loop-free arithmetics not only shortens the code,
-    but also enables the compiler or library authors to relieve the user (i.e. scientific programmer)
-    from hand-coding optimisations (e.g. loop order choice).
-  Hand-coded optimisations -- code rearrangements aimed solely at the purpose of increasing performance --
-    were long recognised as having {\em a strong negative impact when debugging
-    and maintenance are considered} \citep{Knuth_1974},
-    and are generally advised to be avoided \citep[e.g.][section 3.12]{bib_CERNcpp}.
-
-  Three issues specific to Fortran that 
-    resulted in employment of a more repetitive or cumbersome syntax than in C++ or Python
-    were observed:
-  \begin{itemize}
-    \item{Fortran lacks support for specifying array ranges in multiple dimensions
-      with a single entity (cf. tuples of slices in Python and blitz::RectDomain in C++);
-    }
-    \item{Fortran does not feature a mechanism allowing to reuse a single piece of code (algorithm)
-      with different data types (compare e.g. listings C.15, P.14 and F.17) such as
-      templates in C++ and the so-called duck typing in Python;}
-    \item{Fortran does not allow function calls to appear on the left hand side
-      of assignment (see e.g. how the \prog{ptr} pointers were used as a workaround in the \prog{cyclic\_fill\_halos}
-      method in listing F.20);}
-    \item{Fortran lacks support for arrays of arrays (cf. Sect. \ref{sec:sequence}).}
-  \end{itemize}
-  Interestingly, the limitation in extendability via inheritance was found to
-    exist partially in NumPy as well (see footnote \ref{fnt:slice}).
-  The lack of a counterpart in Fortran to the C++ template mechanism was identified in
-    \citep{Cary_et_al_1997}
-    as one of the key deficiencies of Fortran when compared with C++ in context 
-    of applicability to object-oriented scientific programming.
-
-  \subsection{Performance}
-
-  The timing and memory usage statistics presented in figures \ref{fig:mem}-\ref{fig:cpu-skua}
-    reveal that no single language/library/compiler set-up 
-    corresponded to both shortest execution time and smallest memory footprint.
-
-  One may consider performance measures addressing not only the program efficiency but also 
-    the factors influencing the development and maintenance time/cost 
-    \citep[of particular importance in scientific computing,][]{Wilson_2006}.
-  Taking into account such measures as code length or coding time,
-    the Python environment gains significantly \citep[see also discussion in][]{Lin_2012}.
-  Presented Python code is shorter than the C++ and Fortran counterparts,
-    and is simpler in terms of syntax and usage (see discussion below).
-
-  Employment of the PyPy drop-in replacement for the standard Python implementation brings 
-    Python's performance significantly closer to those of C++ and Fortran, in some
-    cases making it the least memory consuming set-up.
-  Python has already been the language of choice for scientific software projects having code clarity 
-    or ease of use as the first requirement \citep[see e.g.][]{Barnes_and_Jones_2011}.
-  PyPy's capability to improve performance of unmodified Python code may 
-    make Python a favourable choice even if high performance is important, especially
-    if a combined measure of performance and development cost is to be considered.
-
-  \subsection{Ease of use and abuse}
-
-  Using the number of lines of code or the number of distinct language keywords
-    needed to implement the MPDATA-based
-    solver presented in section \ref{sec:impl} as measures of syntax 
-    brevity, Python clearly surpasses its rivals.
-  Python was developed with emphasis on code readability and object-orientation.
-  Arguably, taking it to the extreme - Python uses line indentation to define 
-    blocks of code and treats even single integers as objects.
-  As a consequence Python is easy to learn and easy to teach.
-  It is also much harder to abuse Python than C++ or Fortran
-    (for instance with \prog{goto} statements, employment of the preprocessor,
-    or the implicit typing in Fortran).
-
-  Python implementations do not expose users to compilation or linking processes. 
-  As a result, Python-written software is easier to deploy and share, especially 
-    if multiple architectures and operating systems are targeted.
-  However, there exist tools such as CMake\footnote{CMake is a family of open-source, cross-platform
-    tools automating building, testing and packaging of C/C++/Fortran software,
-    see \url{http://cmake.org/}}
-    that allow to efficiently automate 
-    building, testing and packaging of C++ and Fortran programs.
-
-  Python is definitely easiest to debug among the three languages.
-  Great debugging tools for C++ do exist, however the debugging and development is 
-    often hindered by indecipherable compiler messages
-    flooded with lengthy type names stemming from employment of templates.
-  Support for the OOP features of Fortran among free and open source compilers, 
-    debuggers and other programming aids remains immature.
-
-  Blitz++ ensures temporary-array-free computations by design
-    \citep{Veldhuizen_et_al_1997} avoiding unintentional performance loss.
-  In contrast, with both Fortran and Python, the memory footprint caused by employment
-    of temporary objects in array arithmetics is dependant on compiler choice or
-    the level of optimisations.
-
-  \subsection{Developers' community and libraries}%User- and code-base}
-
-  The size of the programmers' community of a given language 
-    influences the availability of trained personnel, 
-    reusable software components and information resources.
-  It also affects the maturity and quality of compilers and tools. 
-  Fortran is a domain-specific language while Python and C++ are general-purpose languages
-    with disproportionately larger users' communities.
-  The OOP features of Fortran have not gained
-    wide popularity among users \citep{Worth_2008}\footnote{An anecdotal yet significant
-    example being the incomplete support for syntax-highlighting of modern Fortran in Vim and Emacs editors}.
-  Fortran is no longer routinely taught at the universities \citep{Kendall_et_al_2008},
-    in contrast to C++ and Python.
-  An example of decreasing popularity of Fortran in academia 
-    is the discontinuation of Fortran printed editions of the ''Numerical Recipes'' 
-    series of Press et al.
-
-  Blitz++ is one of several packages that offer high-performance object-oriented
-    array manipulation functionality with C++ (and is not necessarily optimal for every
-    purpose \citep{Iglberger_et_al_2012}).
-  In contrast, the NumPy package became a de facto standard solution for Python.
-  Consequently, numerous Python libraries adopted NumPy but
-    there are apparently very few C++ libraries offering Blitz++ support out of the box
-    (the gnuplot-iostream used in listing C.20 being a much-appreciated counterexample).
-  However, Blitz++ allows to interface with virtually any library (including Fortran libraries), 
-    by resorting to referencing the underlying memory with raw pointers.
-
-  The availability and quality of libraries 
-    that offer object-oriented interfaces
-    differs among the three considered languages.
-  The built-in standard libraries of Python and C++ are richer than
-    those of Fortran and offer versatile data types, collections of
-    algorithms and facilities for interaction with host operating system.
-  In the authors' experience, the small popularity of OOP techniques among
-    Fortran users is reflected in the library designs (including the Fortran's
-    built-in library routines).
-  What makes correct use of external libraries more difficult with Fortran
-    is the lack of standard exception handling mechanism, a feature
-    long and {\em much requested by the numerical community} \citep[][Foreword]{Press_et_al_1996}.
-
-  The three languages differ as well with regard to availability of 
-    mechanisms for leveraging shared-memory parallelisation (e.g. with multi-core processors).
-  GCC supports OpenMP with Fortran and C++.
-  The CPython and PyPy implementations of Python do not offer any
-    built-in solution for multi-threading.
-
-  \section{Summary and outlook}\label{sec:concl}
-
-  Three implementations of a prototype solver 
-    for the advection equation were introduced.
-  The solvers are based on MPDATA - an algorithm of particular applicability
-    in geophysical fluid dynamics \citep{Smolarkiewicz_2006}.
-  All implementations follow the same object-oriented structure but are implemented
-    in three different languages (or language-library pairs):
-  \begin{itemize}
-    \item{C++ with Blitz++;}
-    \item{Python with NumPy;}
-    \item{Fortran.}
-  \end{itemize}
-
-  Presented programs were developed making use of such recent
-    developments as support for C++11 and Fortran 2008 in GCC, and
-    the NumPy support in the PyPy implementation of Python.
-  The fact that all considered standards are open and the employed
-    tools implementing them are free and open-source
-    is certainly an advantage \citep[][sect.~28.2.5]{Anel_2011,Syvitski_et_al_2013}.
-
-  The key conclusion is that all considered language/library/compiler
-    set-ups offer possibilities for using OOP to compactly 
-    represent the mathematical abstractions within the program code. 
-  This creates the potential to improve code readability and brevity,
-  \begin{itemize}
-    \item{contributing to its 
-      auditability, indispensable for credible and reproducible research in computational science 
-      \citep{Post_et_al_2005, Merali_et_al_2010, Stodden_et_al_2012}; and
-    }
-    \item{helping to keep the programs maintainable and avoiding accumulation of the code 
-      debt\footnote{See \citet{Buschmann_2011} for discussion of technical/code debt.} 
-      that besets scientific software in such domains as climate modelling
-      \citep{Freeman_et_al_2010}.
-    }
-  \end{itemize}
-  \noindent 
-  The performance evaluation revealed that:
-    \begin{itemize}
-      \item{the Fortran set-up offered shortest execution times,}
-      \item{it took the C++ set-up less than twice longer to compute than Fortran,}
-      \item{C++ and PyPy set-ups offered significantly smaller memory consumption 
-        than Fortran and CPython for larger domains,}
-      \item{the PyPy set-up was roughly twice slower than C++ and up to twice faster than CPython.}
-    \end{itemize}
-  The three equally-structured implementations required ca. 200, 300, and 500 lines of code 
-    in Python, C++ and Fortran, respectively.  
-  It is the authors' impression that these figures are somehow indicative of the programming 
-    effort.
-
-  In addition to the source code presented within the text,
-    a set of tests and build-/test-automation scripts
-    allowing to reproduce the analysis and plots presented in section~\ref{sec:perf} are all 
-    available in the electronic supplement and at
-    the project repository\footnote{git repository at \url{http://github.com/slayoo/mpdata/}},
-    and are released under the GNU GPL license \citep{GPLv3}.
-  The authors encourage to use the presented codes for teaching and benchmarking purposes.
-
-  The OOP design enhances the possibilities to reuse and extend the presented code.
-  Development is underway of an object-oriented C++ library featuring concepts presented herein,
-    supporting integration in one to three dimensions, handling systems of equations with source terms, 
-    providing miscellaneous options of MPDATA and several parallel processing approaches.
-
-%%%%%% <arXiv>
-%  \clearpage
-%%%%%% </arXiv>
-  \appendix
-
-  \section{Prototype solvers}\label{app:solver}
-
-  The following sections provide a complete description of a minimal example of application
-    of the formul\ae~''translated'' into C++, Python and Fortran in the main body of the paper.
-
-  \subsection{Halo regions}
-
-  The MPDATA formul\ae~defining $\psi^{[n+1]}_{[i,j]}$ as a function of $\psi^{[n]}_{[i,j]}$
-    (discussed in the following sections) feature terms such as $\psi_{[i-1,j-1]}$.
-  One way of assuring validity of these formul\ae~on the edges of the domain (e.g. for i=0) 
-    is to introduce the so-called halo region surrounding the domain.
-  The method of populating the halo region with data depends on the boundary condition type.
-  Employment of the halo-region logic implies repeated usage of array range 
-    extensions in the code such as $i \leadsto i \pm halo$.
-
-  An \prog{ext()} function is defined in all three implementation, in order to simplify 
-    coding of array range extensions:
-  \codecpp{code-cpp-listings.hpp}{listing15}{listing16}{1}
-  \codepyt{code-pyt-listings.py}{listing14}{listing15}{1}
-  \codefor{code-for-listings.f}{listing17}{listing18}{1}
-  Consequently, a range depicted by $i\pm1/2$ may be expressed in the code as \prog{ext(i, h)}.
-  In all three implementations the \prog{ext()} function accept the second
-    argument to be an integer or a ''half'' (cf. section \ref{sec:grid}).
-
-  \subsection{Prototype solver}\label{sec:solver}
-
-  The tasks to be handled by a prototype advection equation solver proposed herein are:
-  \begin{enumerate}[label=(\roman*), leftmargin=*, widest=iii]
-    \item{storing arrays representing the $\psi$ and $\vec{C}$ fields and any required housekeeping data,}
-    \item{allocating and deallocating the required memory,}
-    \item{providing access to the solver state,}
-    \item{performing the integration by invoking the advection-operator and
-      boundary-condition handling routines.}
-  \end{enumerate}
-  In the following C++ definition of the \prog{solver} structure, task (i) is represented with the definition of the structure
-    member fields; task (ii) is split between the \prog{solver}'s constructor and the destructors of \prog{arrvec\_t};
-    task (iii) is handled by the accessor methods; task (iv) is handled within the \prog{solve} method:
-  \codecpp{code-cpp-listings.hpp}{listing16}{listing17}{1}
-  The \prog{solver} structure is an abstract definition (containing a pure virtual method) 
-     requiring its descendants to implement at least
-     the \prog{advop()} method which is expected to fill \prog{psi[n+1]} with an updated (advected)
-     values of \prog{psi[n]}.
-  The two template parameters \prog{bcx\_t} and \prog{bcy\_t} allow the solver to operate with 
-    any kind of boundary condition structures that fulfil the requirements implied by the calls to the 
-    methods of \prog{bcx} and \prog{bcy}, respectively.
-
-  The donor-cell and MPDATA schemes both require only the previous state of an advected field
-    in order to advance the solution.
-  Consequently, memory for two time levels ($\psi^{[n]}$ and $\psi^{[n+1]}$) is allocated in the
-    constructor.
-  The sizes of the arrays representing the two time levels of $\psi$ 
-    are defined by the domain size ({\em nx}$~\times$~{\em ny}) plus the halo region.
-  The size of the halo region is an argument of the constructor.
-  The \prog{cycle()} method is used to swap the time levels without copying any data.
-
-  The arrays representing the $C^{[x]}$ and $C^{[y]}$ components of $\vec{C}$,
-    require ({\em nx+1})~$\times$~{\em ny} 
-    and {\em nx}$~\times$~({\em ny+1}) elements, respectively
-    (being laid out on the Arakawa-C staggered grid).
-
-  Python definition of the \prog{solver} class follows closely the C++ structure definition:
-  \codepyt{code-pyt-listings.py}{listing15}{listing16}{1}
-  The key difference stems from the fact that, unlike Blitz++, NumPy does not allow an array
-    to have arbitrary index base -- in NumPy the first element is always addressed with 0.
-  Consequently, while in C++ (and Fortran) the computational domain is chosen to start at (i=0, j=0)
-    and hence a part of the halo region to have negative indices, in Python the halo region starts 
-    at (0,0)\footnote{The reason to allow the domain to begin at an arbitrary index
-    is mainly to ease debugging in case the code would be used in parallel computations
-    using domain decomposition where each subdomain could have its own index base corresponding to
-    the location within the computational domain}.
-  However, since the whole halo logic is hidden within the solver, such details are not exposed to the
-    user.
-  The \prog{bcx} and \prog{bcy} boundary-condition specifications are passed to the solver through
-    constructor-like \prog{\_\_init\_\_()} method as opposed to template parameters in C++.
-
-  The above C++ and Python prototype solvers in principle allow to operate with any boundary condition
-    objects that implement methods called from within the solver.
-  This requirement is checked at compile-time in the case of C++, and at run-time in the case of Python.
-  In order to obtain an analogous behaviour with Fortran, it is required to
-    define, prior to definition of a solver type, 
-    an abstract type with deferred procedures having abstract interfaces 
-    \citep[sic!, see Table 2.1 in][for a summary of approximate correspondence of OOP 
-      nomenclature between Fortran and C++]{Rouson_et_al_2012}:
-  \codefor{code-for-listings.f}{listing18}{listing19}{1}
-  Having defined the abstract type for boundary-condition objects, 
-    a definition of a solver class following closely the C++ and Python counterparts may be provided:
-  \codefor{code-for-listings.f}{listing19}{listing20}{1}
-
-  \subsection{Periodic boundaries}\label{sec:cyclic}
-
-  The solver definition described in section~\ref{sec:solver} requires a given
-    boundary condition object to implement a \prog{fill\_halos()} method.
-  An implementation of periodic boundary conditions in C++ is provided in the following listing:
-  \codecpp{code-cpp-listings.hpp}{listing17}{listing18}{1}
-
-  As hinted by the member field names, the \prog{fill\_halos()} methods
-    fill the left/right halo regions with data from the right/left edges of the domain.
-  Thanks to employment of the function \prog{pi()} described in section~\ref{sec:pi}
-    the same code may be applied in any dimension (here being a template parameter).
-
-  The following listings contain the Python and Fortran counterparts to listing C.17.
-  \codepyt{code-pyt-listings.py}{listing16}{listing17}{1}
-  \codefor{code-for-listings.f}{listing20}{listing21}{1}
-
-  \subsection{Donor-cell solver}\label{sec:donorcell_solver}
-
-  As mentioned in the previous section, the donor-cell formula
-    constitutes an advection scheme, hence we may use it to create
-    a \prog{solver\_donorcell} implementation of the abstract \prog{solver} class: 
-  \codecpp{code-cpp-listings.hpp}{listing18}{listing19}{1}
-  The above definition is given as an example only.
-  In the following sections an MPDATA solver of the same structure is defined.
-
-  The following listings contain the Python and Fortran counterparts to listing C.18.
-  \codepyt{code-pyt-listings.py}{listing17}{listing18}{1}
-  \codefor{code-for-listings.f}{listing21}{listing2}{1}
-
-  \subsection{MPDATA solver}\label{sec:mpdatasolver}
-
-  An MPDATA solver may be now constructed by inheriting from \prog{solver} class
-    with the following definition in C++:
-  \codecpp{code-cpp-listings.hpp}{listing19}{listing20}{1}
-  The array of sequences of temporary arrays \prog{tmp} allocated in the constructor
-    is used to store the antidiffusive velocities from the present and optionally 
-    previous timestep (if using more than two iterations).
-
-  The \prog{advop()} method controls the MPDATA iterations within one timestep.
-  The first (step = 0) iteration of MPDATA is an unmodified donor-cell step.
-  Subsequent iterations begin with calculation of the antidiffusive Courant fields using 
-    formula \ref{eq:antidiff1}.
-  In order to calculate values spanning an (i\mhlf~... i+\phlf) range using a formula
-    for $C_{[i+\phlf,\ldots]}$ only, the formula is evaluated using extended
-    index ranges \prog{im} and \prog{jm}.
-  In the second (step=1) iteration the uncorrected Courant field (\prog{C\_unco}) points
-    to the original \prog{C} field, and the antidiffusive Courant field is written into
-    \prog{C\_corr} which points to \prog{tmp[1]}.
-  In the third (step=2) iteration \prog{C\_unco} points to \prog{tmp[1]} while \prog{C\_corr}
-    points to \prog{tmp[0]}.
-  In subsequent iterations \prog{tmp[0]} and \prog{tmp[1]} are alternately swapped.
-
-  The following listings contain the Python and Fortran counterparts to listing C.19.
-  \codepyt{code-pyt-listings.py}{listing18}{listing19}{1}
-  \codefor{code-for-listings.f}{listing22}{listing23}{1}
-
-  \section{Usage example}\label{app:usage}
-
-  The following listing provides an example of how the MPDATA solver
-    defined in section~\ref{sec:mpdatasolver} may be used together
-    with the cyclic boundary conditions defined in section~\ref{sec:cyclic}.
-  In the example a Gaussian signal is advected in a 2D domain 
-    defined over a grid of 24$\times$24 cells.
-  The program first plots the initial condition, then performs
-    the integration for 75 timesteps with three different 
-    settings of the number of iterations used in MPDATA.
-  The velocity field is constant in time and space (although it is
-    not assumed in the presented implementations).
-  The signal shape at the end of each simulation is plotted as well.
-  Plotting is done with the help of the gnuplot-iostream library\footnote{gnuplot-iostream
-    is a header-only C++ library allowing gnuplot to be controlled from C++, see \url{http://stahlke.org/dan/gnuplot-iostream/}. 
-    Gnuplot is a portable command-line driven graphing utility, see \url{http://gnuplot.info/}}.
-
-  The resultant plot is presented herein as Figure~\ref{fig:mpdata}.
-  The top panel depicts the initial condition. 
-  The three other panels show a snapshot of the field after 75 timesteps.
-  The donor-cell solution is characterised by strongest numerical diffusion
-    resulting in significant drop in the signal amplitude.
-  The signals advected using MPDATA show smaller numerical diffusion with
-    the solution obtained with more iterations preserving the signal altitude 
-    more accurately.
-  In all of the simulations the signal maintains its positive definiteness.
-  The domain periodicity is apparent in the plots as the maximum of the signal 
-    after 75 timesteps is located near the domain walls.
-  \codecpp{code-cpp-plot.cpp}{listing20}{listing21}{1}
-
-  The following listings contain the Python and Fortran 
-    counterparts to listing C.20 (with the set-up and plotting logic omitted).
-  \codepyt{code-pyt-test.py}{listing19}{listing20}{1}
-  \codefor{code-for-test.f}{listing23}{listing24}{1}
-  \codefor{code-for-test.f}{listing25}{listing26}{1}
-
-  \footnotesize
-
-  \section*{Acknowledgements}
-  {
-    We thank Piotr Smolarkiewicz and Hanna Pawłowska for their help throughout the project.
-    This study was partly inspired by the lectures of Lech Łobocki.
-
-    Tobias Burnus, Julian Cummings, Ond\v{r}ej \v{C}ert\'ik, Patrik Jonsson,
-      Arjen Markus, Zbigniew Piotrowski, Davide Del Vento and Janus Weil 
-      provided valuable feedback to the initial version of the
-      manuscript and/or responses to questions posted to Blitz++ and gfortran mailing lists.
-
-    SA, AJ and DJ acknowledge funding from the Polish National Science Centre
-      (project no. 2011/01/N/ST10/01483).
-
-    Part of the work was carried out during a visit of SA to the National
-      Center for Atmospheric Research (NCAR) in Boulder, Colorado, USA.
-    NCAR is operated by the University Corporation for Atmospheric Research.
-    The visit was funded by the Foundation for Polish Science (START programme).
-
-    Development of NumPy support in PyPy was led by Alex Gaynor, Matti Picus and MF.
-  }
-
-  \bibliographystyle{plainnat}
-  \bibliography{paper}
-
-  \begin{figure}[bh!]
-    \caption{\label{fig:mpdata}
-      Plot generated by the program given in listing~C.20.
-      The top panel shows initial signal shape (at time t=0).
-      The subsequent panels show snapshots of the advected field after 75 timesteps
-        from three different simulations: donorcell (or 1 MPDATA iteration), MPDATA
-        with two iterations and MPDATA with 44 iterations.
-      The colour scale and the wire-frame surface correspond to signal amplitude.
-      See appendix~\ref{app:usage} for discussion.
-    }
-  \end{figure}
-  \includegraphics[height=\textheight]{demo}
-
-\end{document}
